### 2021-11-01 writing/Haedong-SimGuided-Kv-Modeling.Rmd
Writing - Introduction

---
title: Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes
author: "Haedong Kim, Hui Yang, Andrew R. Ednie, and Eric S. Bennet"
date: "10/18/2021"
output: pdf_document
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Abstract

Electrical signaling initiates and controls the contraction of cardiac muscles. Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes, so they have been important subjects of research in cardiac electrophysiology. Among major cardiac VGICs, K\textsuperscript{+} channel (K\textsubscript{v}) has a distinctive feature in that it has multiple isoforms. Each isoform produces a different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}). Despite the importance of studying each K\textsubscript{v} isoform rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments. Hence, we need a data assimilation method that combines mathematical models with experimental data, I\textsubscript{Ksum}, to estimate individual K\textsuperscript{+} currents. Exponential fitting is a traditional method to estimate the current functions of K\textsubscript{v} isoforms from I\textsubscript{Ksum} data based on the observation that K\textsuperscript{+} currents have the shape of an exponential decay function. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents. However, this classical method does not consider the kinetics of ion channels at the cellular level but only the shape of each I\textsubscript{Ksum} recording. On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities in the heart. This study proposes a data assimilation method based on simulation of K\textsubscript{v} kinetics that combines in-vitro experimental results I\textsubscript{Ksum} with in-silico models of K\textsubscript{v} isoforms by calibrating the computer models. The proposed method differs from the exponential fitting in a way that it estimates the characteristics of current functions and underlying kinetics. In addition, the suggested approach assimilates I\textsubscript{Ksum} recordings at cellular level rather than analyzing each recording independently. An accompanying graphical user interface (GUI) is developed to provide a user-friendly environment to test the proposed method for practitioners. We conducted a case study with our in-vitro experimental data that investigated the effect of reduced glycosylation on K\textsubscript{v}.

**Keywords:** Cardiac electrophysiology models, data assimilation, model calibration, K\textsubscript{v} isoforms, mouse cardiomyocytes

## Introduction

Electrical signaling plays a vital role in the heart conduction system that controls electrical excitation propagation through the heart, which triggers and synchronizes the contraction of a myriad of cardiac muscle cells [@veeraraghavan2014mechanisms]. In turn, this electrical stimulation is responsible for the heart to pump blood throughout the body properly. Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane [@bezanilla2005vgic]. They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure \ref{fig:variety_currents} [@nerbonne2003molecular; @grant2009cardiac]. This electrical simulations is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels. Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase).
```{r variety_currents, fig.align='center', fig.lp='fig:', fig.cap='A schematic illustration of action potentials in mouse ventricular cardiomyocytes and underlying ionic currents of Na\textsubscript{v}, Ca\textsubscript{v}, and K\textsubscript{v} isoforms'}
knitr::include_graphics('figs/variety_currents.pdf')
```

In contrast to Na\textsubscript{v} and Ca\textsubscript{v}, K\textsubscript{v} have a distinctive feature that they have multiple isoforms. Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization. To be specific, as shown in Figure \ref{fig:variety_currents}, there are various K\textsubscript{v} isoforms and currents they produce: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) [@brouillette2004functional; @liu2011dissection]. Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos}, only one of them is presented and referred to as I\textsubscript{Kto} because I\textsubscript{Ktof} is only found in ventricular apex myocytes, while I\textsubscript{Ktos} is reported to appear only in the septum [@xu1999four]. There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research. Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.

Hence, we need a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}. Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations. To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents. Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation~\ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters. I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents and served as a standard method to analyze the sum of K\textsuperscript{+} currents [@brunet2004heterogeneous; liu2011dissection]. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level.
\begin{equation}
    \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} 
\end{equation}

On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes [@winslow2011integrative]. Therefore, this paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level. K\textsubscript{v} models are developed on biophysical principles of VGICs that are encoded in terms of mathematical equations. These equations have parameters that act like ``knobs'', allowing to control the behavior of the models. Tunning computer models to couple with experimental data requires trial-and-error investigations and domain knowledge in both the kinetics of cardiomyocytes and mathematics of computer models. Since modeling of cardiac electrophysiology is an interdisciplinary field that incorporates biology and engineering, calibration of computer models and interpretation of modeling results can be challenging for researchers in biology or medicine communities without engineering training. This hurdle hinders researchers in physiology from utilizing computer models to analyze their experimental data. A graphical user interface (GUI) is provided in which researchers can easily calibrate K\textsubscript{v} models to their data and visualize the modeling results. A new model calibration method will allow researchers to utilize computer models to analyze complex potassium dynamics in cardiomyocytes without engineering background. A GUI will provide an interactive environment to use the calibration method and analyze the results easily with visualizations.

## Material and Methods

## Results

## Conclusions

## References

### 2021-11-06 writing/draft.tex
Draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex

% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.

% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

%\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
\documentclass[utf8]{frontiersHLTH} % for Health articles
%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% Leave a blank line between paragraphs instead of using \\

%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png
%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%

%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
%%% Frontiers will add the figures at the end of the provisional pdf automatically
%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

%% Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet). It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{H. Kim {et~al.}} %use et al only if is more than 1 author
\def\Authors{Haedong Kim\,$^{1}$, Hui Yang\,$^{1,*}$, Andrew R. Ednie\,$^{2}$ and Eric S. Bennett\,$^{2}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Complex Systems Monitoring, Modeling, and Control Laboratory, The Pennsylvania State University, University Park, PA, USA \\
$^{2}$Department of Neuroscience, Cell Biology and Physiology, Wright State University, Dayton, OH, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{310 Leonhard Building, University Park, PA 16802-4400, USA}
\def\corrEmail{huiyang@psu.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[Simulation-guided Data Assimilation of K\textsubscript{v}]{Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}
Electrical signaling plays vital roles in the heart that initiates and controls the synchronous contraction of cardiac muscles. Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes. Among major cardiac VGICs, K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have multiple isoforms. Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments. Hence, there is an urgent need to develop a data assimilation method that combines mathematical models with experimental data, I\textsubscript{Ksum} recordings, to estimate individual K\textsuperscript{+} currents. Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents have the shape of an exponential decay function. It has been proven that exponential fitting describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks the cellular-level analysis. This paper presents a data assimilation method based on the cellular-level simulation of the underlying gating kinetics of K\textsubscript{v} isoforms. Computer models are designed with parameters that allow us to control the behavior of the ion channel models. A fractional factorial design is utilized to identify the parameters having significant effects on the current shape. Then, various non-linear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell. Further, we develop a graphical user interface (GUI) to provide a user-friendly environment to test the proposed method for practitioners unfamiliar with coding. We conduct a case study with our in-vitro experimental data investigating the effect of reduced glycosylation on K\textsubscript{v}. Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms that gives a better understanding of disease-related perturbations.

% You may provide up to 8 keywords; at least 5 are mandatory.
\tiny
\keyFont{\section{Keywords:} Cardiac electrophysiology models, data assimilation, model calibration, K\textsubscript{v} isoforms, mouse cardiomyocytes} 
\end{abstract}

\section{Introduction}
Electrical signaling plays a vital role in the heart conduction system that controls electrical excitation propagation through the heart, which triggers and synchronizes the contraction of a myriad of cardiac muscle cells \cite{veeraraghavan2014mechanisms}. In turn, this electrical stimulation is responsible for the heart to pump blood throughout the body properly. Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane \cite{bezanilla2005vgic}. They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure~\ref{fig:variety_ionc} \cite{grant2009cardiac}. This electrical simulations is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels. Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase). 
\begin{figure}[!ht]
    \centering
    \includegraphics[scale=1]{figures/variety_currents.pdf}
    \caption{A schematic illustration of action potentials in mouse ventricular cardiomyocytes and underlying ionic currents of Na\textsubscript{v}, Ca\textsubscript{v}, and K\textsubscript{v} isoforms \cite{nerbonne2003molecular}.}
    \label{fig:variety_ionc}
\end{figure}

In contrast to Na\textsubscript{v} and Ca\textsubscript{v}, K\textsubscript{v} have a distinctive feature that they have multiple isoforms. Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization. To be specific, as shown in Figure~\ref{fig:variety_ionc}, there are various K\textsubscript{v} isoforms and currents they produce: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) \cite{brouillette2004functional, liu2011dissection}. Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos}, only one of them is presented and referred to as I\textsubscript{Kto} because I\textsubscript{Ktof} is only found in ventricular apex myocytes, while I\textsubscript{Ktos} is reported to appear only in the septum \cite{xu1999four}. There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research. Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.

Hence, we need a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}. Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations. To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents. Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation~\ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters. I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$. It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents and has served as a standard method to analyze the sum of K\textsuperscript{+} currents \cite{brunet2004heterogeneous, liu2011dissection}. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level. 
\begin{equation}
    \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} \label{eq:sumexp_model}
\end{equation}

On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes \cite{winslow2011integrative}. Therefore, this paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level. K\textsubscript{v} models are developed on biophysical principles of VGICs that are encoded in terms of mathematical equations. These equations have parameters that act like ``knobs'', allowing to control the behavior of the models. Tunning computer models to couple with experimental data requires trial-and-error investigations and domain knowledge in both the kinetics of cardiomyocytes and mathematics of computer models. Since modeling of cardiac electrophysiology is an interdisciplinary field that incorporates biology and engineering, calibration of computer models and interpretation of modeling results can be challenging for researchers in biology or medicine communities without engineering training. This hurdle hinders researchers in physiology from utilizing computer models to analyze their experimental data. A graphical user interface (GUI) is provided in which researchers can easily calibrate K\textsubscript{v} models to their data and visualize the modeling results. A new model calibration method will allow researchers to utilize computer models to analyze complex potassium dynamics in cardiomyocytes without engineering background. A GUI will provide an interactive environment to use the calibration method and analyze the results easily with visualizations.

\textbf{add a paragraph for case study with glycosylation and heart failure}

Cardiac electrophysiology has provided insights into the relationships between the complex dynamics of electrical signaling in the heart at the molecular level and the behavior of cardiomyocytes. Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes. Among major cardiac VGICs, potassium channels have a distinctive feature in that they have multiple isoforms. Although each potassium isoform produces a unique current having different functionality, only the sum of these individual potassium currents (IKsum) can be recorded through in-vitro experiments. Hence there is an urgent need to develop a data analysis tool to estimate individual potassium currents from IKsum. Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations. Exponential fitting is a traditional method based on the assumption that prominent potassium currents have the shape of an exponential decay function. However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding potassium channel isoforms. In addition, exponential fitting handles IKsum recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level. Therefore, this paper proposes a data assimilation method based on the computer models of potassium channel isoforms that facilitate kinetics modeling at the cellular level. A new data assimilation method is expected to allow researchers to estimate the kinetics of potassium isoforms from experimental IKsum data.

% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\section{Material and Methods}
\subsection{Procedure}
Least squares method \cite{reali2017optimization}.

\subsection{Data}

\section{Results}

\section{Conclusions}

\section{Formatting}

\subsection{Tables}
Tables should be inserted at the end of the manuscript. Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted. Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space. These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance. The author will be notified during the typesetting of the final article if this is the case.

\section*{Conflict of Interest Statement}
%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to confirm the following statement:

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

The Author Contributions section is mandatory for all articles, including articles by sole authors. If an appropriate statement is not provided on submission, a standard one will be inserted during the production process. The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work. Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.

\section*{Funding}
Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.

\section*{Acknowledgments}
This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.

\section*{Supplemental Data}
 \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.

\section*{Data Availability Statement}
The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{frontiers}

%%% Make sure to upload the bib file along with the tex file and PDF
\end{document}

### 2021-11-10 scratch.m
Voltage-clamp protocol for Overall Diagram

@@ -1,3 +1,29 @@
+%% voltage-clamp protocol
+clc
+clearvars
+close all
+
+v = -50:10:50;
+holdv = -70;
+
+holdt = 100;
+endt = 4.6*1000;
+t = 1:endt;
+
+figure('Color','w')
+cv = ones(length(t),1);
+cv(1:holdt) = holdv;
+cv(holdt+1:end) = v(1);
+plot(t,cv, 'LineWidth',1.5)
+hold on
+for i=2:length(v)
+    cv(holdt+1:end) = v(i);
+    plot(t,cv, 'LineWidth',1.5)
+end
+xlabel('Time (ms)')
+ylabel('Clamp Voltage (mV)')
+set(gca, 'XLimSpec','tight', 'LineWidth',1.5)
+set(gca, 'FontName','Arial', 'FontWeight','bold', 'FontSize',10)
 %% new function form of taua in Bondarenko 2020
 clc
 close all

### 2021-11-11 scratch.m
Current curves for Overall Diagram

@@ -1,3 +1,137 @@
+%% exponential fitting example
+clc
+clearvars
+close all
+
+v = -50:10:50;
+holdv = -70;
+
+holdt = 100;
+endt = 4.6*1000;
+t = 1:endt;
+tt = 0:(endt-holdt);
+
+amp = [22.8, 13.1, 7.3, 3.7];
+tau = [105, 1119, 7271];
+
+f1 = @(t) amp(1)*exp(-t./tau(1));
+f2 = @(t) amp(2)*exp(-t./tau(2));
+f3 = @(t) amp(3)*exp(-t./tau(3));
+
+figure('Name','ikto', 'Color','w')
+y1 = zeros(length(t),1);
+y1(holdt:end) = f1(tt);
+plot(t, y1, 'Color','black', 'LineWidth',1.5)
+ylim([0, sum(amp)])
+set(gca, 'XLimSpec','Tight', 'LineWidth',1.5)
+xlabel('Time (ms)')
+ylabel('I_{Kto} (pA/pF)')
+set(gca, 'FontName','Arial', 'FontSize',10, 'FontWeight','bold')
+
+figure('Name','ikslow1', 'Color','w')
+y2 = zeros(length(t),1);
+y2(holdt:end) = f2(tt);
+plot(t, y2, 'Color','black', 'LineWidth',1.5)
+ylim([0, sum(amp)])
+set(gca, 'XLimSpec','Tight', 'LineWidth',1.5)
+xlabel('Time (ms)')
+ylabel('I_{Kslow1} (pA/pF)')
+set(gca, 'FontName','Arial', 'FontSize',10, 'FontWeight','bold')
+
+figure('Name','ikslow2', 'Color','w')
+y3 = zeros(length(t),1);
+y3(holdt:end) = f3(tt);
+plot(t, y3, 'Color','black', 'LineWidth',1.5)
+ylim([0, sum(amp)])
+set(gca, 'XLimSpec','Tight', 'LineWidth',1.5)
+xlabel('Time (ms)')
+ylabel('I_{Kslow2} (pA/pF)')
+set(gca, 'FontName','Arial', 'FontSize',10, 'FontWeight','bold')
+
+figure('Name','ikss', 'Color','w')
+y4 = zeros(length(t),1);
+y4(holdt:end) = amp(4);
+plot(t, y4, 'Color','black', 'LineWidth',1.5)
+ylim([0, sum(amp)])
[…]

### 2021-11-11 writing/Haedong-SimGuided-Kv-Modeling.Rmd
Revised abstract

@@
 
-Electrical signaling initiates and controls the contraction of cardiac muscles.
-Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes, so they have been important subjects of research in cardiac electrophysiology.
-Among major cardiac VGICs, K\textsuperscript{+} channel (K\textsubscript{v}) has a distinctive feature in that it has multiple isoforms.
-Each isoform produces a different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}).
-Despite the importance of studying each K\textsubscript{v} isoform rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.
-Hence, we need a data assimilation method that combines mathematical models with experimental data, I\textsubscript{Ksum}, to estimate individual K\textsuperscript{+} currents.
-Exponential fitting is a traditional method to estimate the current functions of K\textsubscript{v} isoforms from I\textsubscript{Ksum} data based on the observation that K\textsuperscript{+} currents have the shape of an exponential decay function.
-It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents.
-However, this classical method does not consider the kinetics of ion channels at the cellular level but only the shape of each I\textsubscript{Ksum} recording.
-On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities in the heart.
-This study proposes a data assimilation method based on simulation of K\textsubscript{v} kinetics that combines in-vitro experimental results I\textsubscript{Ksum} with in-silico models of K\textsubscript{v} isoforms by calibrating the computer models.
-The proposed method differs from the exponential fitting in a way that it estimates the characteristics of current functions and underlying kinetics.
-In addition, the suggested approach assimilates I\textsubscript{Ksum} recordings at cellular level rather than analyzing each recording independently.
-An accompanying graphical user interface (GUI) is developed to provide a user-friendly environment to test the proposed method for practitioners.
-We conducted a case study with our in-vitro experimental data that investigated the effect of reduced glycosylation on K\textsubscript{v}.
+* **Motivation** - Among major cardiac voltage-gated ion channels (VGICs), K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have various isoforms.
+Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
+A data analysis tool is used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
+
+* **Gaps & Needs** - A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents.
+This method directly fits individual K\textsuperscript{+} currents by searching the shape parameters of the current functions that have the best fit to experimental I\textsubscript{Ksum} recordings.
+It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents.
+However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
+
+* **Objective** - As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms.
+
+* **Approach**
+  1. haedong
+  2. kim
+
+* **Experiment & Performance** - Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms that gives a better understanding of disease-related perturbations.
+
+* **Significance** - impacts on the research community
 
@@
 ## References
+
+<div id="refs"></div>
+
+## Appendix
+
+equations of potassium models

### 2021-11-11 writing/draft.tex
Writing backup

@@
 A fractional factorial design is utilized to identify the parameters having significant effects on the current shape.
-Then, various non-linear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell.
+Then, various nonlinear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell.
 Further, we develop a graphical user interface (GUI) to provide a user-friendly environment to test the proposed method for practitioners unfamiliar with coding.
@@
 \section{Introduction}
-Electrical signaling plays a vital role in the heart conduction system that controls electrical excitation propagation through the heart, which triggers and synchronizes the contraction of a myriad of cardiac muscle cells \cite{veeraraghavan2014mechanisms}.
-In turn, this electrical stimulation is responsible for the heart to pump blood throughout the body properly.
+Cardiac muscle cells have a conduction system that controls the propagation of electrical excitation, which triggers and synchronizes the mechanical contraction of the network of a large number of cardiomyocytes \cite{veeraraghavan2014mechanisms}.
+In turn, this electrical signaling system is responsible for the heart to pump blood throughout the body properly.
 Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane \cite{bezanilla2005vgic}.
 They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure~\ref{fig:variety_ionc} \cite{grant2009cardiac}.
-This electrical simulations is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels.
+This electrical simulation is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels.
 Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase).
@@
 Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization.
-To be specific, as shown in Figure~\ref{fig:variety_ionc}, there are various K\textsubscript{v} isoforms and currents they produce: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) \cite{brouillette2004functional, liu2011dissection}.
-Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos}, only one of them is presented and referred to as I\textsubscript{Kto} because I\textsubscript{Ktof} is only found in ventricular apex myocytes, while I\textsubscript{Ktos} is reported to appear only in the septum \cite{xu1999four}.
+To be specific, as shown in Figure~\ref{fig:variety_ionc}, there are various K\textsubscript{v} isoforms and currents conducted through them: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) \cite{brouillette2004functional, liu2011dissection}.
+Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos},
+they are referred to as one, I\textsubscript{Kto}, because I\textsubscript{Ktos} is reported to appear only in the ventricular septum myocytes \cite{xu1999four}.
 There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research.
-Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded during in-vitro experiments.
+Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded through in-vitro experiments.
 
-Hence, we need a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}.
+Hence, there is an urgent need to develop a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}.
 Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations.
@@
 On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes \cite{winslow2011integrative}.
-Therefore, this paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level.
-K\textsubscript{v} models are developed on biophysical principles of VGICs that are encoded in terms of mathematical equations.
-These equations have parameters that act like ``knobs'', allowing to control the behavior of the models.
-Tunning computer models to couple with experimental data requires trial-and-error investigations and domain knowledge in both the kinetics of cardiomyocytes and mathematics of computer models.
-Since modeling of cardiac electrophysiology is an interdisciplinary field that incorporates biology and engineering, calibration of computer models and interpretation of modeling results can be challenging for researchers in biology or medicine communities without engineering training.
-This hurdle hinders researchers in physiology from utilizing computer models to analyze their experimental data.
-A graphical user interface (GUI) is provided in which researchers can easily calibrate K\textsubscript{v} models to their data and visualize the modeling results.
-A new model calibration method will allow researchers to utilize computer models to analyze complex potassium dynamics in cardiomyocytes without engineering background.
-A GUI will provide an interactive environment to use the calibration method and analyze the results easily with visualizations.
+This paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level.
+Hodgkin-Huxley-type computer models of potassium isoforms are developed with parameters that act like ``knobs'', allowing to control the behavior of the models.
+The more parameters the models have, the more flexible they are, but surplus parameters increase the risk of overfitting, make interpretation of calibration results harder, and have a chance to acerbate calibration accuracy.
+We utilize a fractional factorial design to identify parameters that significantly affect the current shape, which is a standard sensitivity analysis tool \cite{du2015statistical, du2017silico}.
+A calibration process consists of an objective function and an optimization method to minimize that objective.
+We derive the objective function measuring discrepancies between experimental I\textsubscript{Ksum} recordings and simulated ones for each cell.
+VGIC kinetics is a nonlinear process \cite{rudy2006computational}, so nonlinear optimization methods are utilized.
+A graphical user interface (GUI) is provided in which researchers can easily apply the proposed method to their data and visualize the modeling results.
 
-\textbf{add a paragraph for case study with glycosylation and heart failure}
+We conduct a case study with
 
-Cardiac electrophysiology has provided insights into the relationships between the complex dynamics of electrical signaling in the heart at the molecular level and the behavior of cardiomyocytes.
-Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes.
-Among major cardiac VGICs, potassium channels have a distinctive feature in that they have multiple isoforms.
-Although each potassium isoform produces a unique current having different functionality, only the sum of these individual potassium currents (IKsum) can be recorded through in-vitro experiments.
-Hence there is an urgent need to develop a data analysis tool to estimate individual potassium currents from IKsum.
-Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations.
-Exponential fitting is a traditional method based on the assumption that prominent potassium currents have the shape of an exponential decay function.
-However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding potassium channel isoforms.
-In addition, exponential fitting handles IKsum recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level.
-Therefore, this paper proposes a data assimilation method based on the computer models of potassium channel isoforms that facilitate kinetics modeling at the cellular level.
-A new data assimilation method is expected to allow researchers to estimate the kinetics of potassium isoforms from experimental IKsum data.
+\textbf{impacts}
+\begin{enumerate}
+    \item contribution 1
+    \item contribution 2
+\end{enumerate}
 
@@
 \section{Material and Methods}
-\subsection{Procedure}
-Least squares method \cite{reali2017optimization}.
+\subsection{Computer Models}
+Hodgkin-Huxley type modeling; current equations; full model equations in appendix
 
-\subsection{Data}
+\subsection{Sensitivity Analysis and Calibration Parameters}
+Half-normal plots; calibration parameters; selected calibration parameters
+
+\subsection{Model Calibration}
+Least squares method with multiple starting points \cite{reali2017optimization}.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figures/overall_diagram.pdf}
+    \caption{Caption}
+    \label{fig:overall}
+\end{figure}
+
+\subsection{GUI}
+screenshots of GUI
 
@@
 \section{Conclusions}
-
-\section{Formatting}
-
-\subsection{Tables}
-Tables should be inserted at the end of the manuscript.
-Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
-Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
-These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
-The author will be notified during the typesetting of the final article if this is the case.
 

### 2021-11-17 writing/Haedong-SimGuided-Kv-Modeling.Rmd
Abstract revised

@@
 author: "Haedong Kim, Hui Yang, Andrew R. Ednie, and Eric S. Bennet"
-date: "10/18/2021"
+date: "`r Sys.Date()`"
 output: pdf_document
@@
 Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
-A data analysis tool is used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
+Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
 
 * **Gaps & Needs** - A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents.
-This method directly fits individual K\textsuperscript{+} currents by searching the shape parameters of the current functions that have the best fit to experimental I\textsubscript{Ksum} recordings.
+This method directly fits individual K\textsuperscript{+} currents to experimental I\textsubscript{Ksum} recordings by searching the shape parameters of the current functions that have the best fit.
 It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents.
-However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
+However, the mathematical form of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
+In addition, the curve fitting method handles experimental I\textsubscript{Ksum} recordings from the same cell independently, so it lacks the cellular-level analysis.
 
-* **Objective** - As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms.
+* **Objective** - As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms based on computer models of K\textsuperscript{+} currents that can simulate underlying gating kinetics of the currents.
 
 * **Approach**
-  1. haedong
-  2. kim
+  1. Computer models are designed with calibration parameters that allow us to control the behavior of the ion channel models.
+  2. A fractional factorial design is utilized to identify the parameters having significant effects on the current shape.
+  3.
+Then, various non-linear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell.
+  4. Further, we develop a graphical user interface (GUI) to provide a user-friendly environment to test the proposed method for practitioners unfamiliar with coding. 
+  5. We conduct a case study with our in-vitro experimental data investigating the effect of reduced glycosylation on K\textsubscript{v}.
 
-* **Experiment & Performance** - Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms that gives a better understanding of disease-related perturbations.
+* **Experiment & Performance** - The proposed framework outperforms the curve fitting method for minimizing discrepancies between experimental and estimated total K\textsuperscript{+} currents.
+We also provide the kinetics modeling results that are not possible in the traditional method.
 
-* **Significance** - impacts on the research community
+* **Significance** - Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms.
 
@@
 To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents.
-Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation~\ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters.
+Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation \ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters.
 I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$.
@@
 \begin{equation}
+\label{eq:sumexp_model}
     \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} 
@@
 
+### Computer Models
+
+As shown in Figure \ref{fig:variety_currents}, there are four primary currents yield the most portion of the total K\textsuperscript{+} current: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}.
+Computer models of the four currents are developed based on the Hodgkin-Huxley type modeling principle.
+The Hodgkin-Huxley model treats ion channels as electrical components in a circuit and represents the permeability changes of ion channels as the electrical conductance changes [@fang2021memristive].
+In the typical Hodgkin-Huxley modeling, K\textsuperscript{+} current is defined by Equation \ref{eq:hh_current}
+\begin{equation}
+\label{eq:hh_current}
+  I=g_Ka^{n}i^{m}(V-E_K),
+\end{equation}
+where $g_K$ is the maximum conductance, $a$ and $i$ are state variables of gating kinetics, $n$ and $m$ are non-negative integers, $V$ is the membrane potential, and the $E_K$ is the Nernst potential.
+
+### Sensitivity Analysis and Calibration Parameters
+
+figures of half-normal probability plots and selected parameters and their principles (it was not purely data-driven but I selected a few subjectively)
+
+### Model Calibration
+
+objective function and optimization methods
+
+### Graphical User Interface
+
+we develope gui
+
 ## Results
@@
 
-equations of potassium models
+full equations of potassium models

### 2021-12-01 calibration/opt_ko1.m
Parameterization from DoE

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'ko';
-save_dir = strcat('calib_exp22_extra_', group_name);
+save_dir = strcat('calib_exp24', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -16,20 +16,20 @@ num_currents = length(current_names);
 volt_range = 3:11;
 
 % tunning index in individual current models
-tune_idx1_kto = [1, 2, 6, 10, 13, 14, 15, 16, 17];
-tune_idx1_kslow1 = [1, 2, 3, 4, 5, 8, 9, 11, 12, 13];
+tune_idx1_kto = [1, 2, 6, 7, 9, 13, 15, 16, 17];
+tune_idx1_kslow1 = [1, 2, 4, 5, 8, 9, 11, 12, 13];
 tune_idx1_kslow2 = [1, 3];
-tune_idx1_kss = [3, 4];
+tune_idx1_kss = [1, 2, 3, 4];
 tune_idx1_kur = [1, 3];
 tune_idx1_k1 = [1, 3, 5, 7];
 
 % optimization options
 max_evals = 1e+6;
-num_iters = 10;
+num_iters = 30;
 options = optimoptions(@fmincon, ...
-    'Algorithm','active-set', 'Display','off', ...
+    'Algorithm','interior-point', 'Display','off', ...
     'MaxFunctionEvaluations',max_evals, ...
-    'SpecifyObjectiveGradient',true);
+    'SpecifyObjectiveGradient',false);
 
 % protocol
 hold_volt = -70;
@@ -91,7 +91,7 @@ for i = 1:num_files
     end
     loop_idx = [loop_idx, i];
 end
-loop_idx = loop_idx([12, 13, 14, 22, 24]);
+% loop_idx = loop_idx([12, 13, 14, 22, 24]);
 len_loop_idx = length(loop_idx);
 
 % default values
@@ -223,8 +223,9 @@ for l = 1:len_loop_idx
     time_space{5} = ideal_end_idx;
     
     % objective function
-%    obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
-    opt_fun = @(p) obj_rmse_grad(p, model_struct, volt_space, time_space, yksum);
+%     obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
[…]

### 2021-12-01 calibration/opt_wt1.m

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'wt';
-save_dir = strcat('calib_exp_22', group_name);
+save_dir = strcat('calib_exp_24', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -16,20 +16,20 @@ num_currents = length(current_names);
 volt_range = 3:11;
 
 % tunning index in individual current models
-tune_idx1_kto = [1, 2, 6, 10, 13, 14, 15, 16, 17];
-tune_idx1_kslow1 = [1, 2, 3, 4, 5, 8, 9, 11, 12, 13];
+tune_idx1_kto = [1, 2, 6, 7, 9, 13, 15, 16, 17];
+tune_idx1_kslow1 = [1, 2, 4, 5, 8, 9, 11, 12, 13];
 tune_idx1_kslow2 = [1, 3];
-tune_idx1_kss = [3, 4];
+tune_idx1_kss = [1, 2, 3, 4];
 tune_idx1_kur = [1, 3];
 tune_idx1_k1 = [1, 3, 5, 7];
 
 % optimization options
 max_evals = 1e+6;
-num_iters = 100;
+num_iters = 30;
 options = optimoptions(@fmincon, ...
-    'Algorithm','sqp', 'Display','off', ...
+    'Algorithm','interior-point', 'Display','off', ...
     'MaxFunctionEvaluations',max_evals, ...
-    'SpecifyObjectiveGradient',true);
+    'SpecifyObjectiveGradient',false);
 
 % protocol
 hold_volt = -70;
@@ -91,7 +91,7 @@ for i = 1:num_files
     end
     loop_idx = [loop_idx, i];
 end
-loop_idx = loop_idx(3);
+% loop_idx = loop_idx(3);
 len_loop_idx = length(loop_idx);
 
 % default values
@@ -223,8 +223,9 @@ for l = 1:len_loop_idx
     time_space{5} = ideal_end_idx;
     
     % objective function
-%    obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
-    opt_fun = @(p) obj_rmse_grad(p, model_struct, volt_space, time_space, yksum);
+%     obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
+%     obj_rmse_grad(p0, model_struct, volt_space, time_space, yksum)
[…]

### 2021-12-01 writing/draft.tex
Writing draft

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex
+\documentclass[11pt]{article}
 
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\pdfminorversion=4
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
 
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%% IISE Transactions margin specifications %%%%%%%%%%%%%%%%%%%
+% DON'T change margins - should be 1 inch all around.
+\addtolength{\oddsidemargin}{-.5in}%
+\addtolength{\evensidemargin}{-.5in}%
+\addtolength{\textwidth}{1in}%
+\addtolength{\textheight}{1.3in}%
+\addtolength{\topmargin}{-.8in}%
+\makeatletter
+\renewcommand\section{\@startsection {section}{1}{\z@}%
+                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
+                                   {2.3ex \@plus.2ex}%
+                                   {\normalfont\fontfamily{phv}\fontsize{16}{19}\bfseries}}
+\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
+                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
+                                     {1.5ex \@plus .2ex}%
+                                     {\normalfont\fontfamily{phv}\fontsize{14}{17}\bfseries}}
+\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
+                                    {-3.25ex\@plus -1ex \@minus -.2ex}%
+                                     {1.5ex \@plus .2ex}%
+                                     {\normalfont\normalsize\fontfamily{phv}\fontsize{14}{17}\selectfont}}
+\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%% Version 3.4 Generated 2018/06/15 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%
+%%%%% IISE Transactions package list %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{enumerate}
+\usepackage{natbib} %comment out if you do not have the package
+\usepackage{url} % not crucial - just used below for the URL
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-\documentclass[utf8]{frontiersHLTH} % for Health articles
-%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
-
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-
-\linenumbers
-
-% Leave a blank line between paragraphs instead of using \\
-
-%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
-%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
-%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
-%%% If using *.tif files convert them to .jpg or .png
-%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
-
-%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
-%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
-%%% Frontiers will add the figures at the end of the provisional pdf automatically
-%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
-
-%% Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
-It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
-
-\def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{H. Kim {et~al.}} %use et al only if is more than 1 author
-\def\Authors{Haedong Kim\,$^{1}$, Hui Yang\,$^{1,*}$, Andrew R. Ednie\,$^{2}$ and Eric S. Bennett\,$^{2}$}
-% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Complex Systems Monitoring, Modeling, and Control Laboratory, The Pennsylvania State University, University Park, PA, USA \\
-$^{2}$Department of Neuroscience, Cell Biology and Physiology, Wright State University, Dayton, OH, USA}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{310 Leonhard Building, University Park, PA 16802-4400, USA}
-\def\corrEmail{huiyang@psu.edu}
+%%%%% Author package list and commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%% Here are some examples %%%%%%%%%%%%%%
+%	\usepackage{amsfonts, amsthm, latexsym, amssymb}
+%	\usepackage{lineno}
+%	\newcommand{\mb}{\mathbf}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\onecolumn
-\firstpage{1}
 
-\title[Simulation-guided Data Assimilation of K\textsubscript{v}]{Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes}
+		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\def\spacingset#1{\renewcommand{\baselinestretch}%
+		{#1}\small\normalsize} \spacingset{1}
+	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
+	\if0\blind
+	{
+		\title{\bf Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes}
+		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
+		$^a$ Department of Industrial and Manufacturing Engineering, \\
+		The Pennsylvania State University, State College, PA USA \\
+        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
+        Wright State University, City, OH USA}
+		\date{}
+		\maketitle
+	} \fi
 
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
+	\if1\blind
+	{
+        \title{\bf \emph{IISE Transactions} \LaTeX \ Template}
+		\author{Author information is purposely removed for double-blind review}
 
-\maketitle
+\bigskip
+		\bigskip
+		\bigskip
+		\begin{center}
+			{\LARGE\bf \emph{IISE Transactions} \LaTeX \ Template}
+		\end{center}
+		\medskip
+	} \fi
+	\bigskip
 
 \begin{abstract}
-Electrical signaling plays vital roles in the heart that initiates and controls the synchronous contraction of cardiac muscles.
-Voltage-gated ion channels (VGICs) are the functional unit of electrical activities in cardiomyocytes.
-Among major cardiac VGICs, K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have multiple isoforms.
-Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
-Hence, there is an urgent need to develop a data assimilation method that combines mathematical models with experimental data, I\textsubscript{Ksum} recordings, to estimate individual K\textsuperscript{+} currents.
-Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents have the shape of an exponential decay function.
-It has been proven that exponential fitting describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents.
-However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
-In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks the cellular-level analysis.
-This paper presents a data assimilation method based on the cellular-level simulation of the underlying gating kinetics of K\textsubscript{v} isoforms.
-Computer models are designed with parameters that allow us to control the behavior of the ion channel models.
-A fractional factorial design is utilized to identify the parameters having significant effects on the current shape.
-Then, various nonlinear optimization methods were applied and compared to calibrate the selected parameters to minimize the discrepancies between the simulated currents and I\textsubscript{Ksum} recordings from the same cell.
-Further, we develop a graphical user interface (GUI) to provide a user-friendly environment to test the proposed method for practitioners unfamiliar with coding.
-We conduct a case study with our in-vitro experimental data investigating the effect of reduced glycosylation on K\textsubscript{v}.
-Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms that gives a better understanding of disease-related perturbations.
-
-% You may provide up to 8 keywords; at least 5 are mandatory.
-\tiny
-\keyFont{\section{Keywords:} Cardiac electrophysiology models, data assimilation, model calibration, K\textsubscript{v} isoforms, mouse cardiomyocytes} 
+This document provides a \LaTeX \ template for \emph{IISE Transactions}.
+Your paper should be compiled in the following order: title; abstract; keywords; main text, including an introduction and a conclusion or summary; acknowledgments; declaration of interest statement; references; appendices (as appropriate).
+Figures and tables should be inserted into the text as close to first mention as possible (NOT appended to the end of the manuscript).
+In-text citations and the reference list must follow \emph{IISE Transactions} guidelines.
+Use 11 point font, 1 inch margins, and double-spacing for the manuscript.
+A typical paper for this journal should be no more than 30 pages in manuscript format, counting from the title page to references.
+Appendices should be included as supplemental online materials.
+Do not use footnotes.
+\emph{IISE Transactions} uses a double-blind review process.
+Please make sure that you submit the \textbf{blind version} of your manuscript, which does not contain any information identifying the authors.
+This includes removing the authors information on the title page as well as the information that may be identifying in the Acknowledgment section.
+We thank you for your attention to these details.
 \end{abstract}
 
-\section{Introduction}
-Cardiac muscle cells have a conduction system that controls the propagation of electrical excitation, which triggers and synchronizes the mechanical contraction of the network of a large number of cardiomyocytes \cite{veeraraghavan2014mechanisms}.
-In turn, this electrical signaling system is responsible for the heart to pump blood throughout the body properly.
-Cardiac voltage-gated ion channels (VGICs) are membrane proteins in a cardiomyocyte that control specific ions to flow through the membrane \cite{bezanilla2005vgic}.
-They are fundamental functional units of electrical activities in cardiomyocytes as the electrical excitation results from the orchestrated dynamics of different types of ionic currents controlled by the gating kinetics of VGICs as shown in Figure~\ref{fig:variety_ionc} \cite{grant2009cardiac}.
-This electrical simulation is called the action potentials (AP), which is the net change of membrane potential, and there are three major cardiac VIGCs contributing to the shape and length of the AP: Na\textsuperscript{+}  (Na\textsubscript{v}), Ca\textsuperscript{2+}  (Ca\textsubscript{v}), and K\textsuperscript{+}  (K\textsubscript{v}) channels.
-Na\textsubscript{v} are responsible for AP initiation (depolarization phase), Ca\textsubscript{v} are responsible for the prolonged depolarization phase (``plateau'') particularly in larger species like human, while various isoforms of K\textsubscript{v}s are collectively responsible for AP deactivation (repolarization phase).
-\begin{figure}[!ht]
+\noindent%
+{\it Keywords:} \emph{IISE Transactions}; \LaTeX; Manuscript format; Taylor \& Francis.
+
+%\newpage
+\spacingset{1.5} % DON'T change the spacing!
+
+\section{Introduction} \label{s:intro}
+\begin{itemize}
+    \item \textbf{Motivation} Among major cardiac voltage-gated ion channels (VGICs), K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have various isoforms.
+Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
+Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
+    \item \textbf{Gaps \& Needs} A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents.
+This method directly fits individual K\textsuperscript{+} currents to experimental I\textsubscript{Ksum} recordings by searching the shape parameters of the current functions that have the best fit.
+It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents.
+However, the mathematical form of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
+In addition, the curve fitting method handles experimental I\textsubscript{Ksum} recordings from the same cell independently, so it lacks the cellular-level analysis.
+    \item \textbf{Objective} As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms based on computer models of K\textsuperscript{+} currents that can simulate underlying gating kinetics of the currents.
+    \item \textbf{Approach} 
+    \item \textbf{Experiment \& Performance} The proposed framework outperforms the curve fitting method for minimizing discrepancies between experimental and estimated total K\textsuperscript{+} currents.
+We also provide the kinetics modeling results that are not possible in the traditional method.
+    \item \textbf{Significance} Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms.
+\end{itemize}
+
+The review of literature can be a separate section but it can be merged with the introduction section. One can find a number of reference examples at the end of this template, like,
+\begin{itemize}
+\item A book: \cite{Hastie2009}.
+\item A journal article: \cite{Byon2013}.
+\item A conference paper: \cite{Breunig2000}.
+\item A book chapter: \cite{Yang2010}.
+\item A technical report: \cite{Kelley2007}.
+\end{itemize}
+
+\section{Methods} \label{s:methods}
+
+\subsection{\emph{Computer Models of Potassium Channel Isoforms}} \label{s:methods.1}
+modeling (mathematical equations)
+\begin{figure}
     \centering
-    \includegraphics[scale=1]{figures/variety_currents.pdf}
-    \caption{A schematic illustration of action potentials in mouse ventricular cardiomyocytes and underlying ionic currents of Na\textsubscript{v}, Ca\textsubscript{v}, and K\textsubscript{v} isoforms \cite{nerbonne2003molecular}.}
-    \label{fig:variety_ionc}
+    \includegraphics{figs/ap_underlying_currents.pdf}
+    \caption{Caption}
+    \label{fig:ap_underlying_currents}
 \end{figure}
 
-In contrast to Na\textsubscript{v} and Ca\textsubscript{v}, K\textsubscript{v} have a distinctive feature that they have multiple isoforms.
-Each isoform produces different but slightly overlapping component of the total K\textsuperscript{+} current (I\textsubscript{Ksum}) and contributes to different portions of the AP repolarization.
-To be specific, as shown in Figure~\ref{fig:variety_ionc}, there are various K\textsubscript{v} isoforms and currents conducted through them: the transient outward K\textsuperscript{+} current I\textsubscript{Kto} (K\textsubscript{v}4.2/K\textsubscript{v}4.3) several components of delayed rectifier K\textsuperscript{+} currents including I\textsubscript{Ks} (K\textsubscript{v}7.1 and minK), I\textsubscript{Kr} (mERG), I\textsubscript{Kslow1} (K\textsubscript{v}1.5), and I\textsubscript{Kslow2} (K\textsubscript{v}2.1), non-inactivating steady-state K\textsuperscript{+} current I\textsubscript{Kss} (K\textsubscript{2P}), and time independent K\textsuperscript{+} current I\textsubscript{K1} (Kir2 family) \cite{brouillette2004functional, liu2011dissection}.
-Note that although there are two types of transient outward currents, I\textsubscript{Ktof} and I\textsubscript{Ktos},
-they are referred to as one, I\textsubscript{Kto}, because I\textsubscript{Ktos} is reported to appear only in the ventricular septum myocytes \cite{xu1999four}.
-There are several K\textsubscript{v} isoforms whose exact roles in shaping the cardiac AP are still subject of intensive research.
-Although it is highly desirable to study each unique K\textsuperscript{+} current rigorously, only I\textsubscript{Ksum} can be recorded through in-vitro experiments.
-
-Hence, there is an urgent need to develop a data analysis tool to estimate individual K\textsuperscript{+} currents from I\textsubscript{Ksum}.
-Data assimilation is a scientific method that combines mathematical models and experimental data to calibrate the models; thereby, the in-silico models are compatible with the in-vitro observations.
-To be specific, for our case, a data assimilation method is required that couples mathematical models of K\textsuperscript{+} models with experimental data of I\textsubscript{Ksum} using whole-cell patch-clamp recordings to estimate various K\textsuperscript{+} currents.
-Exponential fitting is a traditional method based on the assumption that prominent K\textsuperscript{+} currents from patch-clamp recordings such as I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} have the shape of an exponential decay function $A_{i}e^{-t/\tau_{i}}$ as in Equation~\ref{eq:sumexp_model}, where $t$ is time variable, and $A_{i}$ and $\tau_{i}$ are the shape parameters.
-I\textsubscript{kss} is simplified as constant $A\textsubscript{Kss}$.
-It has been proven that exponential fitting accurately describes I\textsubscript{Ksum} with estimated K\textsuperscript{+} currents and has served as a standard method to analyze the sum of K\textsuperscript{+} currents \cite{brunet2004heterogeneous, liu2011dissection}.
-However, the mathematical model of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
-In addition, exponential fitting handles I\textsubscript{Ksum} recordings from the same cell with different conditions independently, so it lacks systematic analyses at the cellular level.
-\begin{equation}
-    \text{I}\textsubscript{Ksum} = \sum_{i} A_{i} e^{-t/\tau_{i}} + A\textsubscript{Kss}, \; i \in I \subseteq \{\text{Kto, } \text{Kslow1, } \text{Kslow2}\} \label{eq:sumexp_model}
-\end{equation}
-
-On the other hand, computer models of cardiac electrophysiology can simulate complex dynamics of electrical activities at the cellular level that enable the quantitative understanding of biophysical functions such as gating kinetics of VGICs in the behavior of cardiomyocytes \cite{winslow2011integrative}.
-This paper proposes a data assimilation method based on the computer models of K\textsubscript{v} isoforms that facilitates kinetics modeling at the cellular level.
-Hodgkin-Huxley-type computer models of potassium isoforms are developed with parameters that act like ``knobs'', allowing to control the behavior of the models.
-The more parameters the models have, the more flexible they are, but surplus parameters increase the risk of overfitting, make interpretation of calibration results harder, and have a chance to acerbate calibration accuracy.
-We utilize a fractional factorial design to identify parameters that significantly affect the current shape, which is a standard sensitivity analysis tool \cite{du2015statistical, du2017silico}.
-A calibration process consists of an objective function and an optimization method to minimize that objective.
-We derive the objective function measuring discrepancies between experimental I\textsubscript{Ksum} recordings and simulated ones for each cell.
-VGIC kinetics is a nonlinear process \cite{rudy2006computational}, so nonlinear optimization methods are utilized.
-A graphical user interface (GUI) is provided in which researchers can easily apply the proposed method to their data and visualize the modeling results.
-
-We conduct a case study with
-
-\textbf{impacts}
-\begin{enumerate}
-    \item contribution 1
-    \item contribution 2
-\end{enumerate}
-
-% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
-
-\section{Material and Methods}
-\subsection{Computer Models}
-Hodgkin-Huxley type modeling; current equations; full model equations in appendix
-
-\subsection{Sensitivity Analysis and Calibration Parameters}
-Half-normal plots; calibration parameters; selected calibration parameters
-
-\subsection{Model Calibration}
-Least squares method with multiple starting points \cite{reali2017optimization}.
-\begin{figure}[!ht]
+\subsection{\emph{Sensitivity Analysis of Calibration Parameters}} \label{s:methods.2}
+this section shows the results of the variable screening.
+\begin{figure}
     \centering
-    \includegraphics{figures/overall_diagram.pdf}
-    \caption{Caption}
-    \label{fig:overall}
+    \includegraphics{figs/doe_kto.pdf}
+    \caption{Half-normal plots I\textsubscript{Kto}}
+    \label{fig:doe_kto}
 \end{figure}
 
-\subsection{GUI}
-screenshots of GUI
+\begin{figure}
+    \centering
+    \includegraphics{figs/doe_kslow1.pdf}
+    \caption{Half-normal plots I\textsubscript{Kslow1}}
+    \label{fig:doe_kslow1}
+\end{figure}
 
-\section{Results}
+\begin{figure}
+    \centering
+    \includegraphics{figs/doe_kslow2.pdf}
+    \caption{Half-normal plots I\textsubscript{Kslow2}}
+    \label{fig:doe_kslow2}
+\end{figure}
 
-\section{Conclusions}
+\begin{figure}
+    \centering
+    \includegraphics{figs/doe_kss.pdf}
+    \caption{Half-normal plots I\textsubscript{Kss}}
+    \label{fig:doe_kss}
+\end{figure}
 
-\section*{Conflict of Interest Statement}
-%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
-If no such relationship exists, authors will be asked to confirm the following statement:
+\subsection{\emph{Model Calibration}} \label{s:methods.3}
+calibration procedure; non-linear optimization methods
 
-The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+\subsection{\emph{Graphical User Interface}} \label{s:methods.4}
+show how to use it
 
-\section*{Author Contributions}
+\subsection{\emph{Case Study Glycosylations}} \label{s:methods.5}
+case study
 
-The Author Contributions section is mandatory for all articles, including articles by sole authors.
-If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
-The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work.
-Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.
+\section{Results} \label{s:results}
+Perhaps numerical analysis of performance, comparison, and demonstration of applicability and impact using real data.
 
-\section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+\section{Conclusion}\label{s:conclusion}
+A paper ends with a conclusion or summary section.
 
-\section*{Acknowledgments}
-This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
+\if0\blind{
+\section*{Acknowledgements}
+The authors acknowledge the generous support from the funding agency of XYZ.	} \fi
 
-\section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+\bibliographystyle{chicago}
+\spacingset{1}
+\bibliography{IISE-Trans}
 
-\section*{Data Availability Statement}
-The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
-% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
-
-\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
-\bibliography{frontiers}
-
-%%% Make sure to upload the bib file along with the tex file and PDF
 \end{document}

### 2021-12-01 calibration/scratch.m
Current model with default parameters

@@ -1,3 +1,119 @@
+%% run kcurrent_model with default parameter values
+clc
+close all
+clear variables
+
+current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
+num_currents = length(current_names);
+
+% tunning index in individual current models
+tune_idx1_kto = [1, 2, 6, 10, 13, 14, 15, 16, 17];
+tune_idx1_kslow1 = [1, 2, 3, 4, 5, 8, 9, 11, 12, 13];
+tune_idx1_kslow2 = [1, 3];
+tune_idx1_kss = [3, 4];
+tune_idx1_kur = [1, 3];
+tune_idx1_k1 = [1, 3, 5, 7];
+
+% protocol
+hold_volt = -70;
+min_volt = -50;
+volt = 30;
+ek = -91.1;
+ideal_hold_time = 120;
+ideal_end_time = 4.6*1000;
+
+idx_info1 = cell(1, num_currents);
+for i = 1:num_currents
+    switch current_names{i}
+    case 'ikto'
+        idx_info1{i} = tune_idx1_kto;
+    case 'ikslow1'
+        idx_info1{i} = tune_idx1_kslow1;
+    case 'ikslow2'
+        idx_info1{i} = tune_idx1_kslow2;
+    case 'ikss'
+        idx_info1{i} = tune_idx1_kss;
+    case 'ikur'
+        idx_info1{i} = tune_idx1_kur;
+    case 'ik1'
+        idx_info1{i} = tune_idx1_k1;
+    end
+end
+
+% tunning index in decision variable, p
+idx_info2 = cell(1, num_currents);
+cul_idx_len = 0;
+for i = 1: num_currents
+    cul_idx_len_old = cul_idx_len;
+    cul_idx_len = cul_idx_len + length(idx_info1{i});
+    idx_info2{i} = (1+cul_idx_len_old):(cul_idx_len);
+end
+
+% create model structure
+model_info = [current_names; idx_info1; idx_info2];
+field_names = {'name', 'idx1', 'idx2'};
+model_struct = cell2struct(model_info, field_names, 1);
+
+% default values
+kto_default = [33, 15.5, 20, 16, 8, 7, 0.03577, 0.06237, 0.18064, 0.3956, 0.000152, 0.067083, 0.00095, 0.051335, 0.2087704319, 0.14067, 0.387];
[…]

### 2021-12-03 calibration/opt_ko1.m
Change parameterization

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'ko';
-save_dir = strcat('calib_exp24', group_name);
+save_dir = strcat('calib_exp25_', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -16,7 +16,7 @@ num_currents = length(current_names);
 volt_range = 3:11;
 
 % tunning index in individual current models
-tune_idx1_kto = [1, 2, 6, 7, 9, 13, 15, 16, 17];
+tune_idx1_kto = [1, 2, 6, 7, 9, 13, 14, 15, 16, 17];
 tune_idx1_kslow1 = [1, 2, 4, 5, 8, 9, 11, 12, 13];
 tune_idx1_kslow2 = [1, 3];
 tune_idx1_kss = [1, 2, 3, 4];

### 2021-12-03 calibration/opt_wt1.m

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'wt';
-save_dir = strcat('calib_exp_24', group_name);
+save_dir = strcat('calib_exp25', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -16,7 +16,7 @@ num_currents = length(current_names);
 volt_range = 3:11;
 
 % tunning index in individual current models
-tune_idx1_kto = [1, 2, 6, 7, 9, 13, 15, 16, 17];
+tune_idx1_kto = [1, 2, 6, 7, 9, 13, 14, 15, 16, 17];
 tune_idx1_kslow1 = [1, 2, 4, 5, 8, 9, 11, 12, 13];
 tune_idx1_kslow2 = [1, 3];
 tune_idx1_kss = [1, 2, 3, 4];

### 2021-12-03 calibration/opt_wt1.m
Minor edits

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'wt';
-save_dir = strcat('calib_exp25', group_name);
+save_dir = strcat('calib_exp25_', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};

### 2021-12-03 calibration/scratch.m
Figure for biomarkers

@@ -17,7 +17,7 @@ tune_idx1_k1 = [1, 3, 5, 7];
 % protocol
 hold_volt = -70;
 min_volt = -50;
-volt = 30;
+volt = 0;
 ek = -91.1;
 ideal_hold_time = 120;
 ideal_end_time = 4.6*1000;
@@ -101,18 +101,34 @@ protocol{4} = volt_space{3};
 
 [yksum_hat, comp_currents] = kcurrent_model1(p0, model_struct, protocol);
 
-figure('Color','w', 'Position',[100,100,700,250])
-subplot(1,2,1)
+figure('Color','w', 'Position',[100,100,550,500])
+subplot(2,2,1)
 plot(t,comp_currents{1},'LineWidth',2,'Color','red')
 xlabel('Time (ms)')
 ylabel('Current (pA/pF)')
 set(gca,'FontName','Arial','FontSize',10','FontWeight','bold')
+box off
 
-subplot(1,2,2)
+subplot(2,2,2)
 plot(t,comp_currents{2}, 'LineWidth',2, 'Color','red')
 xlabel('Time (ms)')
 ylabel('Current (pA/pF)')
 set(gca,'FontName','Arial','FontSize',10','FontWeight','bold')
+box off
+
+subplot(2,2,3)
+plot(t,comp_currents{3},'LineWidth',2,'Color','red')
+xlabel('Time (ms)')
+ylabel('Current (pA/pF)')
+set(gca,'FontName','Arial','FontSize',10','FontWeight','bold')
+box off
+
+subplot(2,2,4)
+plot(t,comp_currents{4}, 'LineWidth',2, 'Color','red')
+xlabel('Time (ms)')
+ylabel('Current (pA/pF)')
+set(gca,'FontName','Arial','FontSize',10','FontWeight','bold')
+box off
 
 %% test trust-region-reflective
 % options = optimoptions('fmincon','Algorithm','trust-region-reflective','SpecifyObjectiveGradient',true);
@@ -348,13 +364,13 @@ for l = 1:len_loop_idx
     obj_rmse_grad(p0, model_struct, volt_space, time_space, yksum);
 end
 
-%% calculate RMSEs from calibration results (current model 2)
+%% calculate RMSEs from calibration results
 clc
 close all
 clear variables
 
 group_name = 'ko';
[…]

### 2022-01-07 calibration/log_analysis.R
Log process

@@ -46,29 +46,31 @@ find_min_rmse <- function(log_list) {
 log_dir <- "./calibration/log/"
 
 ## WT -----
-num_files <- 34
-log1_name <- "exp25"
-log2_name <- "exp21"
-log3_name <- "exp22"
-log4_name <- "exp23"
+num_files <- 33
+group_name <- "ko"
+log1_name <- "exp16"
+log2_name <- "exp26"
+# log3_name <- "exp26"
+# log4_name <- "exp23"
 
-file_names <- seq(1, num_files) %>% as.character()
-extra_idx <- c(18, 19, 29, 31, 32)
-
-log1 <- read_log(str_c(log_dir, log1_name, "_wt.txt"), num_files, 30)
-log2 <- read_log(str_c(log_dir, log2_name, "_wt.txt"), num_files, 1)
-log3 <- read_log(str_c(log_dir, log3_name, "_wt.txt"), num_files, 1)
-log4 <- read_log(str_c(log_dir, log4_name, "_wt.txt"), num_files, 1)
+log1 <- read_log(str_c(log_dir, log1_name, "_", group_name, ".txt"), num_files, 30)
+log2 <- read_log(str_c(log_dir, log2_name, "_", group_name, ".txt"), num_files, 1)
+# log3 <- read_log(str_c(log_dir, log3_name, "_", group_name, ".txt"), num_files, 1)
+# log4 <- read_log(str_c(log_dir, log4_name, "_", group_name, ".txt"), num_files, 1)
 
 rmse_df1 <- find_min_rmse(log1)
-write_csv(rmse_df1, str_c(log_dir, log1_name, "_wt.csv"))
-rmse_val2 <- find_min_rmse(log2)
-rmse_val3 <- find_min_rmse(log3)
-rmse_val4 <- find_min_rmse(log4)
+rmse_df2 <- find_min_rmse(log2)
+# rmse_val3 <- find_min_rmse(log3)
+# rmse_val4 <- find_min_rmse(log4)
 
-rmse_val1[extra_idx]
-rmse_val2
+write_csv(rmse_df1, str_c(log_dir, log1_name, "_", group_name, ".csv"))
+write_csv(rmse_df2, str_c(log_dir, log2_name, "_", group_name, ".csv"))
 
+# file_names <- seq(1, num_files) %>% as.character()
+# extra_idx <- c(18, 19, 29, 31, 32)
+# rmse_val1[extra_idx]
+
[…]

### 2022-01-09 calibration/opt_ko1.m
Exp 24

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'ko';
-save_dir = strcat('calib_exp26_', group_name);
+save_dir = strcat('calib_exp24_', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -29,7 +29,7 @@ num_iters = 30;
 options = optimoptions(@fmincon, ...
     'Algorithm','interior-point', 'Display','off', ...
     'MaxFunctionEvaluations',max_evals, ...
-    'SpecifyObjectiveGradient',true);
+    'SpecifyObjectiveGradient',false);
 
 % protocol
 hold_volt = -70;
@@ -225,8 +225,9 @@ for l = 1:len_loop_idx
     % objective function
 %     obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
 %     obj_rmse_grad1(p0, model_struct, volt_space, time_space, yksum)
-    opt_fun = @(p) obj_rmse_grad1(p, model_struct, volt_space, time_space, yksum);
-
+%     opt_fun = @(p) obj_rmse_grad1(p, model_struct, volt_space, time_space, yksum);
+    opt_fun = @(p) obj_rmse(p, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
+    
     % run optimization
     rmse_list = zeros(num_iters, 1);
     sol_list = cell(num_iters, 1);

### 2022-01-09 calibration/opt_wt1.m

@@ -6,7 +6,7 @@ warning('off', 'all')
 
 % code arguments for calibration
 group_name = 'wt';
-save_dir = strcat('calib_exp26_', group_name);
+save_dir = strcat('calib_exp24_', group_name);
 
 % selection of currents
 current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
@@ -29,7 +29,7 @@ num_iters = 30;
 options = optimoptions(@fmincon, ...
     'Algorithm','interior-point', 'Display','off', ...
     'MaxFunctionEvaluations',max_evals, ...
-    'SpecifyObjectiveGradient',true);
+    'SpecifyObjectiveGradient',false);
 
 % protocol
 hold_volt = -70;
@@ -225,8 +225,9 @@ for l = 1:len_loop_idx
     % objective function
 %     obj_rmse(p0, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
 %     obj_rmse_grad1(p0, model_struct, volt_space, time_space, yksum)
-    opt_fun = @(p) obj_rmse_grad1(p, model_struct, volt_space, time_space, yksum);
-
+%     opt_fun = @(p) obj_rmse_grad1(p, model_struct, volt_space, time_space, yksum);
+    opt_fun = @(p) obj_rmse(p, @kcurrent_model1, model_struct, volt_space, time_space, yksum);
+    
     % run optimization
     rmse_list = zeros(num_iters, 1);
     sol_list = cell(num_iters, 1);

### 2022-01-09 writing/draft.tex
Abstract draft

@@
 \begin{abstract}
-This document provides a \LaTeX \ template for \emph{IISE Transactions}.
-Your paper should be compiled in the following order: title; abstract; keywords; main text, including an introduction and a conclusion or summary; acknowledgments; declaration of interest statement; references; appendices (as appropriate).
-Figures and tables should be inserted into the text as close to first mention as possible (NOT appended to the end of the manuscript).
-In-text citations and the reference list must follow \emph{IISE Transactions} guidelines.
-Use 11 point font, 1 inch margins, and double-spacing for the manuscript.
-A typical paper for this journal should be no more than 30 pages in manuscript format, counting from the title page to references.
-Appendices should be included as supplemental online materials.
-Do not use footnotes.
-\emph{IISE Transactions} uses a double-blind review process.
-Please make sure that you submit the \textbf{blind version} of your manuscript, which does not contain any information identifying the authors.
-This includes removing the authors information on the title page as well as the information that may be identifying in the Acknowledgment section.
-We thank you for your attention to these details.
+Abstract goes here.
 \end{abstract}
@@
 \begin{itemize}
-    \item \textbf{Motivation} Among major cardiac voltage-gated ion channels (VGICs), K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature that they have various isoforms.
-Although each K\textsubscript{v} isoform produces a unique but slightly overlapping current, having different functionality, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
+    \item \textbf{Motivation} Voltage-gated ion channels (VGICs) are the functional unit of the electrical conduction system of the heart that controls the synchronous contraction of numerous cardiomyocytes.
+Among major cardiac VGICs, K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature: they have various isoforms that contribute to K\textsuperscript{+} current, while the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
+Although each K\textsubscript{+} generated through K\textsubscript{v} isoform has a different role in the repolarization phase, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
 Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
@@
 
+action potential and various ioninc currents
+\begin{figure}
+    \centering
+    \includegraphics{figs/ap_underlying_currents.pdf}
+    \caption{The action potentials and underlying ionic currents in \textbf{(A)} human ventricular cardiomyocytes and \textbf{(B)} mouse ventricular cardiomyocytes \cite{nerbonne2003molecular}.}
+    \label{fig:ap_underlying_currents}
+\end{figure}
+
 \section{Methods} \label{s:methods}
@@
 \subsection{\emph{Computer Models of Potassium Channel Isoforms}} \label{s:methods.1}
-modeling (mathematical equations)
+Various K\textsuperscript{+} currents play different roles in cardiac repolarization. 
 \begin{figure}
     \centering
-    \includegraphics{figs/ap_underlying_currents.pdf}
-    \caption{Caption}
-    \label{fig:ap_underlying_currents}
+    \includegraphics{figs/biomarkers.pdf}
+    \caption{Six biomarkers of K\textsuperscript{+} currents that are subjects of sensitivity analysis.
+(A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, (D) I\textsubscript{Kss}.
+All currents are simulated for the purpose of illustration and biomarker labels refer to a: at 10 ms, b: 25\%, c: 50\%, d: 75\%, e: peak, and f: time constant}
+    \label{fig:biomarkers}
 \end{figure}
@@
 calibration procedure; non-linear optimization methods
+\begin{figure}
+    \centering
+    \includegraphics{figs/pkto.pdf}
+    \caption{parameter distribution}
+    \label{fig:pkto}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figs/pkslow1.pdf}
+    \caption{parameter distribution}
+    \label{fig:pkslow1}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figs/pkslow2.pdf}
+    \caption{parameter distribution}
+    \label{fig:pkslow2}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{figs/pkss.pdf}
+    \caption{parameter distribution}
+    \label{fig:pkss}
+\end{figure}
 

### 2022-03-21 KvApp/run_calib.m
Code structure

@@ -1,23 +1,23 @@
-function [sol,fval] = run_calib(y, model_info, protocol_info, init_pts, lb, ub)            
+function [sol,fval] = run_calib(y, model_info, protocol_info, init_pts, pdefault, lb, ub)            
     % optimization options
     options = optimoptions(@fmincon, ...
-        'Algorithm','interior-point', 'Display','off', ...
-        'MaxFunctionEvaluations',1e+6, ...
+        'Algorithm','sqp', 'Display','off', ...
+        'MaxFunctionEvaluations',1e+7, ...
         'SpecifyObjectiveGradient',false);
     A = [];
     b = [];
     Aeq = [];
     beq = [];
     nonlcon = [];
-
-    opt_fun = @(p) obj_rmse(p, y, model_info, protocol_info);
+    
+    obj_rmse(init_pts, y, model_info, protocol_info, pdefault)
+    opt_fun = @(p) obj_rmse(p, y, model_info, protocol_info, pdefault);
     [sol,fval] = fmincon(opt_fun, init_pts, A, b, Aeq, beq, lb, ub, nonlcon, options);
 end
 
-function z = obj_rmse(p, y, model_info, protocol_info)
+function z = obj_rmse(p, y, model_info, protocol_info, pdefault)
     time_space = protocol_info{4};
     hold_idx = length(time_space{2});
-    
     volts = protocol_info{3};
     num_volts = length(volts);
 
@@ -26,29 +26,24 @@ function z = obj_rmse(p, y, model_info, protocol_info)
         y_i = y(:, i);
         protocol_info{3} = volts(i);
 
-        [y_hat, ~] = kcurrent_model(p, model_info, protocol_info);
-        rmse_list(i) = sqrt(mean((y_i((hold_idx + 1):end) - y_hat((hold_idx + 1):end)).^2));
+        [y_hat, ~] = kcurrent_model(p, model_info, protocol_info, pdefault);
+        rmse_list(i) = sqrt(mean((y_i((hold_idx+1):end) - y_hat((hold_idx+1):end)).^2));
     end
     z = sum(rmse_list);
 end
 
-function [yksum, comp_currents] = kcurrent_model(p, model_info, protocol_info)
-    global param_kslow1
-    
+function [yksum, comp_currents] = kcurrent_model(p, model_info, protocol_info, pdefault)
     num_currents = length(model_info);
     current_names = cell(num_currents, 1);
-    
     for i = 1:num_currents
[…]

### 2022-03-23 KvApp/run_calib.m
Minor edits

@@ -1,8 +1,8 @@
 function [sol,fval] = run_calib(y, model_info, protocol_info, init_pts, pdefault, lb, ub)            
     % optimization options
     options = optimoptions(@fmincon, ...
-        'Algorithm','sqp', 'Display','off', ...
-        'MaxFunctionEvaluations',1e+7, ...
+        'Algorithm','interior-point', 'Display','off', ...
+        'MaxFunctionEvaluations',1e+6, ...
         'SpecifyObjectiveGradient',false);
     A = [];
     b = [];
@@ -10,7 +10,6 @@ function [sol,fval] = run_calib(y, model_info, protocol_info, init_pts, pdefault
     beq = [];
     nonlcon = [];
     
-    obj_rmse(init_pts, y, model_info, protocol_info, pdefault)
     opt_fun = @(p) obj_rmse(p, y, model_info, protocol_info, pdefault);
     [sol,fval] = fmincon(opt_fun, init_pts, A, b, Aeq, beq, lb, ub, nonlcon, options);
 end

### 2022-03-23 KvAppUserManual.Rmd
User Guide

---
title: "User Manual for K~v~App"
author: "Haedong Kim"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Started

Although K~v~App is developed in MATLAB, it does not require MATLAB installed on the machine. Instead, running `MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without MATLAB. A machine needs to be connected to the Internet to complete the installation.

## Modules and Functions

KvApp

## Example

example with 4.5-sec dat and 25-sec data

### 2022-03-24 KvApp_guide.Rmd
Rename

--- 2022-03-23 KvAppUserManual.Rmd
@@
 ---
-title: "User Manual for K~v~App"
+title: "User Guide for K~v~App"
 author: "Haedong Kim"
@@
 
+K~v~App
+
 Although K~v~App is developed in MATLAB, it does not require MATLAB installed on the machine.
 Instead, running `MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without MATLAB.
-A machine needs to be connected to the Internet to complete the installation.
+Your computer needs to be connected to the Internet to complete the installation.
+You will get the error message below if the right version of MATLAB Runtime.
+![]()
 

### 2022-03-27 KvApp_guide.Rmd
User guide

@@
 
-K~v~App
+K~v~App consists of five files as shown in Figure 1 below: KvAppBeta, MyAppInstaller_mcr, MyAppInstaller_web, readme, and splash.
+
+![Consisting files of the applications](KvApp_guide_img/folder_struct.png){width=50%}
+
+KvAppBeta is the file that runs the application, but you will get the error message in Figure 2 if the correct version of MATLAB is not on the machine.
+
+![Consisting files of the applications](KvApp_guide_img/folder_struct.png){width=50%}
 
@@
 Your computer needs to be connected to the Internet to complete the installation.
+If
 You will get the error message below if the right version of MATLAB Runtime.
-![]()
 
@@
 
-KvApp
+K~v~App
 

### 2022-03-27 KvApp_guide.Rmd
Getting started

--- 2022-03-24 KvApp_guide.Rmd
@@
 ---
-title: "User Guide for K~v~App"
+title: "User Guide for KvApp"
 author: "Haedong Kim"
@@
 
-K~v~App
+There are four associated files as shown in Figure 1 below: `KvAppBeta`, `MyAppInstaller_mcr`, `MyAppInstaller_web`, and `splash`.
 
-Although K~v~App is developed in MATLAB, it does not require MATLAB installed on the machine.
-Instead, running `MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without MATLAB.
+![Consisting files of the applications.](KvApp_guide_img/folder_struct.png){width=50%}
+
+`KvAppBeta` is the file that runs the application, but you will get the error message in Figure 2 if the correct version of MATLAB is not installed on the machine.
+
+![Error message when the right version of MATLAB is not installed on the target machine.](KvApp_guide_img/runtime_error.png){width=50%}
+
+`MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without having MATLAB installed.
 Your computer needs to be connected to the Internet to complete the installation.
-You will get the error message below if the right version of MATLAB Runtime.
-![]()
+If the Internet connection is not available, `MyAppInstaller_mcr` will install the necessary programs locally.
+A few screenshots below show the installation process.
+
+![](KvApp_guide_img/install1.png){width=50%} ![](KvApp_guide_img/install4.png){width=50%}
+![](KvApp_guide_img/install5.png){width=50%} ![](KvApp_guide_img/install6.png){width=50%}
+
+The other file `splash` is just an image file for the application icon.
 

### 2022-03-28 calibration/viz_calib_results.m
Typo

@@ -120,11 +120,11 @@ clc
 close all 
 clearvars
 
-file_group = 'ko';
-exp_num = 'exp27';
+file_group = "wt";
+exp_num = "exp45";
 
 % specify result files and voltages to check
-file_name = '15o29024.xlsx';
+file_name = '15o29015.xlsx';
 % file_name = file_names(1);
 save_dir = strcat('calib_', exp_num);
 
@@ -141,8 +141,7 @@ calib1 = table2array(readtable(fullfile(pwd, strcat(save_dir, '_', file_group), 
 
 % specify model structure
 % field 1: selection of currents
-current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
-num_currents = length(current_names);
+current_names = {'iktof', 'ikslow1', 'ikslow2', 'ikss'};
 
 % field 2: tunning index in individual current models
 tune_idx1{1} = [1, 2, 4, 5, 7, 11, 13];
@@ -153,7 +152,6 @@ tune_idx1{5} = [1, 3];
 tune_idx1{6} = [1, 3, 5, 7];
 
 [model_struct1, psize1] = gen_mdl_struct(current_names, tune_idx1);
-
 sol1 = gen_sol_vec(calib1, model_struct1, psize1);
 
 % experimental data to be compared

### 2022-03-28 KvApp_guide.Rmd
User guide

@@
 If the Internet connection is not available, `MyAppInstaller_mcr` will install the necessary programs locally.
-A few screenshots below show the installation process.
+A few screenshots below in Figure 3 show the installation process.
+The other file `splash` is just an image file for the application icon.
 
-![](KvApp_guide_img/install1.png){width=50%} ![](KvApp_guide_img/install4.png){width=50%}
-![](KvApp_guide_img/install5.png){width=50%} ![](KvApp_guide_img/install6.png){width=50%}
-
-The other file `splash` is just an image file for the application icon.
+![Screenshots of the installation process.](KvApp_guide_img/install.png)
 
@@
 
-KvApp
+KvApp consists of three modules: Experimental Data, Computer Model, and Calibration.
+Steps follows the same order.
+There are two panels in this module: Experimental Protocol and Experimental Data.
+Need to specify protocol first that matches experimental data to be investigated.
+If the defined protocol does not compatible with the imported experiemntal data, it will give error or does not show the data properly.
+Below the example - example 1: voltages not matching, exmaple 2: time not matching.
+This module has Update and Data upload functions.
+Use it appropriately.
+
+The Computer Model module shows equations and kinetic parameters in it that controls the behavior of the model.
+Computer models for different K+ isoforms are separated and grouped by tabs.
+Kv4.2 tab contanes IKto,f, the sencond tab (Kv1.5, Kv2.1, K2P) has three models as they are closed connected.
+They share some parameters and variables.
+The thrid tab (Kv1.4) has a model of IKto,s it is not included in default.
+You can select the currents you want to include in IKsum.
+You can manually change the kinetic parameters if you want.
+The kinetic parameters in red color have more significant impacts on the model kineics and can be calibrated via optimization methods in the thrid module.
+
+The previous discussion naturally brings us to the third module.
+It shows the model output of current, kinetics variables such as steady states, time constants, or transition rates.
+The first tab shows the sum current, while the other tabs illustrate kinetics modeling of each Kv isoforms.
+In the Model Output tab, tou can select a voltage of interest using the drop-down button that has voltage steps spcified in the Experimental Protocol.
+There should be some discrepancies between the model out of IKsum and the experimental recordings uploaded on the app.
+Rather than manually explore appropriate kinetic parameters, Run Calibration button provides an automatic calibration function.
+Once you finish the calibration and want to save the result, Model Parameters Save function gets the job done.
 
@@
 
-example with 4.5-sec dat and 25-sec data
+example with 9 voltage steps rather than the full data.
+
+example with 25-sec data

### 2022-04-01 KvApp_guide.Rmd
Minor edtis
User guide

--- 2022-03-27 KvApp_guide.Rmd
@@
 
-KvApp
+KvApp consists of three modules, as shown in Figure 4: Experimental Protocol & Data, Computer Model, and Model Outputs & Kinetics Modeling Results.
 
-## Example
+![Startup window of KvApp.](KvApp_guide_img/window.png){width=50%}
 
-example with 4.5-sec dat and 25-sec data
+The data to be investigated *in silico* should be stored in an Excel file, with the first column corresponding to time and the other columns normalized currents for voltage steps in increasing order.
+This is a default format when I~Ksum~ traces are transferred to a tabular format in ClampFit.
+Figure 5 is an example of I~Ksum~ recordings.
+To import a dataset, we need to specify the experimental protocol of the voltage-clamp method used to generate that dataset.
+If the defined protocol does not match the data, it will give an error or does not display the data correctly.
+
+![Example of raw I~Ksum~ data in ClampFit.](KvApp_guide_img/pClamp.png)
+
+The Computer Model module shows equations and kinetics parameters of computers models of K~v~ isoforms separated by tabs.
+The Computer Model module shows equations and kinetics parameters of computers models of potassium isoforms, separated and grouped by tabs.
+The Kv4.2 model in the first tab is the only stand-alone model in which every kinetic variables and parameter are independent of other models.
+The second tab contains three closely connected models.
+Each model can be chosen to be or not be included in calculating the total potassium current by checking or unchecking the Include boxes.
+The kinetic parameters can be adjusted manually by entering the number in the corresponding box and pressing the enter/return key.
+Parameters in red are identified as having significant impacts on the model behavior.
+
+These parameters in red can be calibrated automatically using the Turn Calibration button in the Model Output tab of the third module.
+This function calibrates the models by minimizing discrepancies between model prediction and experimental I~Ksum~ traces.
+The fitness of calibration results is visualized, and the drop-down button allows to select different voltage steps.
+The calibrated parameter values can be saved using the Model Parameters Save button in an Excel file.
+It is desirable to perform model calibration separately, utilizing more computing power and time for better results.
+Other tabs named after the K~v~ isoforms show kinetic variables such as steady states, time constants, and transition rates as shown on Figure 6.
+
+![Kinetics modeling results.](KvApp_guide_img/kvoutput.png)
+
+## Examples
+
+### Example 1: 4.5-sec Data
+
+example with 9 voltage steps rather than the full data.
+
+### Example 2: 25-sec Data
+
+example with 25-sec data

### 2022-04-01 calibration/log_analysis.R
Log process

@@ -49,24 +49,24 @@ log_dir <- "./calibration/log/"
 num_files1 <- 34
 num_files2 <- 33
 num_iters <- 30
-log1_name <- "exp35"
+log1_name <- "exp42"
 log2_name <- "exp36"
 log3_name <- "exp37"
 # log4_name <- "exp23"
 
 log1 <- read_log(str_c(log_dir,log1_name,"_wt",".txt"), num_files1, num_iters)
-log2 <- read_log(str_c(log_dir,log2_name,"_wt",".txt"), num_files1, num_iters)
-log3 <- read_log(str_c(log_dir,log3_name,"_wt",".txt"), num_files1, num_iters)
+log2 <- read_log(str_c(log_dir,log1_name,"_ko",".txt"), num_files2, num_iters)
+# log3 <- read_log(str_c(log_dir,log3_name,"_wt",".txt"), num_files1, num_iters)
 # log4 <- read_log(str_c(log_dir, log4_name, "_", group_name, ".txt"), num_files, 1)
 
 rmse_df1 <- find_min_rmse(log1)
 rmse_df2 <- find_min_rmse(log2)
-rmse_df3 <- find_min_rmse(log3)
+# rmse_df3 <- find_min_rmse(log3)
 # rmse_val4 <- find_min_rmse(log4)
 
 write_csv(rmse_df1, str_c(log_dir,log1_name,"_wt.csv"))
-write_csv(rmse_df2, str_c(log_dir,log2_name,"_wt.csv"))
-write_csv(rmse_df3, str_c(log_dir,log3_name,"_wt.csv"))
+write_csv(rmse_df2, str_c(log_dir,log1_name,"_ko.csv"))
+# write_csv(rmse_df3, str_c(log_dir,log3_name,"_wt.csv"))
 
 # file_names <- seq(1, num_files) %>% as.character()
 # extra_idx <- c(18, 19, 29, 31, 32)

### 2022-04-01 KvApp_guide.Rmd
User manual

@@
 date: "`r Sys.Date()`"
-output: pdf_document
+output: word_document
 ---
@@
 
-## Examples
+## Example
 
-### Example 1: 4.5-sec Data
+This section provides an example of using the application.
+A data set will be used having voltage steps from -30 to 50 mV lasting 4.5 s from holding voltage -70 mV.
+The data is saved in the provided file, `15o27010.xlsx`.
+First update the protocol accordingly and upload the data as in Figure 7.
 
-example with 9 voltage steps rather than the full data.
+![Data upload.](KvApp_guide_img/exp1-1.png)
 
-### Example 2: 25-sec Data
+Discrepancies between the model prediction and exerrmental I~Ksum~ traces can be observed in the bottom right figure.
+To calibrate the models according to the data, click the Run Calibration button.
+It will change to a progress bar as in Figure 8.
 
-example with 25-sec data
+![Run calibration.](KvApp_guide_img/ex1-2.png)
+
+Once the calibration is completed, the model outputs and kinetics are updated as in Figure 9.
+
+![Calibration results.](KvApp_guide_img/ex1-3.png)

### 2022-05-30 calibration/param_dist_viz/param_preproc.py
Figures
For meeting in May 2022

@@ -40,7 +40,7 @@ def compare_param_dist(pdf_wt, pdf_ko, idx):
     p = sns.displot(pdf, x=str(idx), hue='Group', kind='kde', fill=True)
     p.set_xlabels('Parameter '+str(idx), fontsize=12)
     
-exp_num =  '45'
+exp_num =  '41'
 currents = ['iktof', 'ikslow1', 'ikslow2', 'ikss']
 idx_dict = {'iktof': [1, 2, 4, 5, 7, 11, 13],
     'iktos': [2, 3],

### 2022-05-30 calibration/viz_calib_results.m

@@ -44,28 +44,42 @@ clearvars
 
 group = "ko";
 exp_num = "exp41";
-file_name = "15o27006.xlsx";
+file_name = "15n03018.xlsx";
 
-data_path = fullfile(pwd,"data",strcat(group,"-preprocessed-25s"),file_name);
+data_path = fullfile(pwd,"mgat1ko_data",strcat(group,"-preprocessed-25s"),file_name);
 calib_path = fullfile(pwd,strcat("calib_",exp_num,"_",group),file_name);
 
-current_names = {'iktof', 'ikslow1', 'ikslow2', 'ikss'};
-tune_idx1{1} = [1, 2, 4, 5, 7, 11, 13];
-tune_idx1{2} = [1, 2, 4, 5, 9, 10, 11];
-tune_idx1{3} = [2, 3];
-tune_idx1{4} = [1, 2, 3, 4];
+calib = table2array(readtable(calib_path));
 
-% import the experimental data and calibration results
+%----- Model structure -----%
+current_names = {'iktof', 'ikslow1', 'ikslow2', 'ikss'};
+tune_idx = cell(5,1); % iktof, iktos, ikslow1, ikslow2, ikss
+tune_idx{1} = [1, 2, 4, 5, 7, 11, 13];
+tune_idx{2} = [2, 3];
+tune_idx{3} = [1, 2, 4, 5, 9, 10, 11];
+tune_idx{4} = [2, 3];
+tune_idx{5} = [1, 2, 3, 4];
+
+[mdl_struct, psize] = gen_mdl_struct(current_names, tune_idx);
+sol = gen_sol_vec(calib, mdl_struct, psize);
+
+pdefault = cell(5,1);
+pdefault{1} = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, ...
+    0.000152, 0.067083, 0.00095, 0.051335, 0.3846];
+pdefault{2} = [-1050, 270, 0.0629];
+pdefault{3} = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18, 2.058, 803.0, 0.16];
+pdefault{4} = [4912, 5334, 0.16];
+pdefault{5} = [0.0862, 1235.5, 13.17, 0.0611];
+
+%----- Protocol & Experimental data -----%
 exp_mx = table2array(readtable(data_path));
-sol_mx = table2array(readtable(calib_path));
[…]

### 2022-05-30 calibration/opt_wt1.m
Uploaded to ROAR

@@ -1,353 +1,0 @@
-clc
-close all
-clear variables
-
-warning('off', 'all')
-
-% code arguments for calibration
-group_name = 'wt';
-save_dir = strcat('calib_exp24_', group_name);
-
-% selection of currents
-current_names = {'ikto', 'ikslow1', 'ikslow2', 'ikss'};
-num_currents = length(current_names);
-
-% voltages
-volt_range = 3:11;
-
-% tunning index in individual current models
-tune_idx1_kto = [1, 2, 6, 7, 9, 13, 14, 15, 16, 17];
-tune_idx1_kslow1 = [1, 2, 4, 5, 8, 9, 11, 12, 13];
-tune_idx1_kslow2 = [1, 3];
-tune_idx1_kss = [1, 2, 3, 4];
-tune_idx1_kur = [1, 3];
-tune_idx1_k1 = [1, 3, 5, 7];
-
-% optimization options
-max_evals = 1e+6;
-num_iters = 30;
-options = optimoptions(@fmincon, ...
-    'Algorithm','interior-point', 'Display','off', ...
-    'MaxFunctionEvaluations',max_evals, ...
-    'SpecifyObjectiveGradient',false);
-
-% protocol
-hold_volt = -70;
-min_volt = -50;
-ideal_hold_time = 120;
-ideal_end_time = 4.6*1000;
-ek = -91.1;
-
-volts = NaN(length(volt_range), 1);
-for i = 1:length(volt_range)
-    volts(i) = min_volt + (volt_range(i)-1)*10;
-end
-
-idx_info1 = cell(1, num_currents);
-for i = 1:num_currents
-    switch current_names{i}
-    case 'ikto'
-        idx_info1{i} = tune_idx1_kto;
-    case 'ikslow1'
-        idx_info1{i} = tune_idx1_kslow1;
-    case 'ikslow2'
-        idx_info1{i} = tune_idx1_kslow2;
-    case 'ikss'
-        idx_info1{i} = tune_idx1_kss;
-    case 'ikur'
-        idx_info1{i} = tune_idx1_kur;
-    case 'ik1'
-        idx_info1{i} = tune_idx1_k1;
-    end
-end
-
-% tunning index in decision variable, p
-idx_info2 = cell(1, num_currents);
-cul_idx_len = 0;
-for i = 1: num_currents
-    cul_idx_len_old = cul_idx_len;
-    cul_idx_len = cul_idx_len + length(idx_info1{i});
[…]

### 2022-05-30 calibration/opt_wt_25.m

@@ -1,0 +1,282 @@
+clc
+close all
+clearvars
+warning('off', 'all')
+
+%----- Code arguments & Model information %-----
+group = "wt";
+exp_num = "exp41";
+save_dir = strcat("calib_",exp_num,"_",group);
+mkdir(fullfile(pwd,save_dir))
+
+current_names = {'iktof','ikslow1','ikslow2','ikss'};
+tune_idx = cell(5,1); % iktof, iktos, ikslow1, ikslow2, ikss
+tune_idx{1} = [1, 2, 4, 5, 7, 11, 13];
+% tune_idx{2} = [2, 3];
+tune_idx{3} = [1, 2, 4, 5, 9, 10, 11];
+tune_idx{4} = [2, 3];
+tune_idx{5} = [1, 2, 3, 4];
+
+pdefault = cell(5,1);
+pdefault{1} = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, ...
+    0.000152, 0.067083, 0.00095, 0.051335, 0.3846];
+pdefault{2} = [-1050, 270, 0.0629];
+pdefault{3} = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18, 2.058, 803.0, 0.16];
+pdefault{4} = [4912, 5334, 0.16];
+pdefault{5} = [0.0862, 1235.5, 13.17, 0.0611];
+
+[mdl_struct, psize] = gen_mdl_struct(current_names, tune_idx);
+[p0, lb, ub] = gen_param_bounds(mdl_struct, psize, pdefault);
+
+%----- Protocol -----%
+holdv = -70;
+minv = 50;
+vstep = 10;
+num_steps = 1;
+ek = -91.1;
+
+ideal_hold_time = 470;
+ideal_end_time = 25*1000;
+
+% save time information later when experimental data imlported
+protocol = cell(6,1);
+protocol{1} = holdv;
+protocol{2} = ek;
+protocol{3} = minv:vstep:(minv+vstep*(num_steps-1));
+
+% meta data of experimental datasets
+matching_table = readtable(fullfile(pwd,"mgat1ko_data",strcat("matching-table-",group,".xlsx")));
+file_names = matching_table.trace_file_name_25;
+data_dir = fullfile("mgat1ko_data",strcat(group,"-preprocessed-25s"));
+
+% exclude null rows
+file_names(cellfun(@isempty,file_names)) = [];
+file_names = string(file_names);
[…]

### 2022-05-30 calibration/pop_idv_25s.m

@@ -1,0 +1,359 @@
+clc
+close all
+clearvars
+warning('off', 'all')
+
+%----- Code arguments & Model information %-----
+group = "wt";
+exp_num = "exp_50";
+save_dir = strcat("calib_",exp_num,"_",group);
+mkdir(fullfile(pwd,save_dir))
+
+% currents to be included
+current_names = {'iktof', 'ikslow1', 'ikslow2', 'ikss'};
+
+tune_idx = cell(5,1); % iktof, iktos, ikslow1, ikslow2, ikss
+tune_idx{1} = [1, 2, 4, 5, 7, 11, 13];
+% tune_idx{2} = [2, 3];
+tune_idx{3} = [1, 2, 4, 5, 9, 10, 11];
+tune_idx{4} = [2, 3];
+tune_idx{5} = [1, 2, 3, 4];
+
+pdefault = cell(5,1);
+pdefault{1} = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, ...
+    0.000152, 0.067083, 0.00095, 0.051335, 0.3846];
+pdefault{2} = [-1050, 270, 0.0629];
+pdefault{3} = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18, 2.058, 803.0, 0.16];
+pdefault{4} = [4912, 5334, 0.16];
+pdefault{5} = [0.0862, 1235.5, 13.17, 0.0611];
+
+[mdl_struct, psize] = gen_mdl_struct(current_names, tune_idx);
+[p0, lb, ub] = gen_param_bounds(mdl_struct, psize, pdefault);
+
+%----- Protocol & Experimental data -----%
+holdv = -70;
+minv = 50;
+vstep = 10;
+num_steps = 1;
+ek = -91.1;
+
+protocol = cell(8,1);
+protocol{7} = 470; % ideal holding time
+protocol{8} = 25*1000; % ideal end time
+
+% save time information later when experimental data imlported
+protocol{1} = holdv;
+protocol{2} = ek;
+protocol{3} = minv:vstep:(minv+vstep*(num_steps-1));
+
+% meta data of experimental datasets
+matching_table = readtable(fullfile(pwd,"mgat1ko_data",strcat("matching-table-",group,".xlsx")));
+file_names = matching_table.trace_file_name_4half;
+data_dir = fullfile("mgat1ko_data",strcat(group,"-preprocessed-25s"));
+
+% exclude null rows
[…]

### 2022-05-30 writing/draft.tex
Draft

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{mathtools}
 \usepackage{graphicx}
@@
 \begin{abstract}
-Abstract goes here.
+Aberrant activities of voltage-gated potassium (K\textsubscript{v}) channels can cause fatal heart disease such as long QT syndrome.
+In our recent investigation on a disease-related perturbation, reduced glycosylation in cardiomyocytes, significant reduction in K\textsuperscript{+} currents was observed.
+Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
+On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
+Because only the sum of these K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded in experiments, data analytical methods are necessary to estimate the properties of individual K\textsuperscript{+} currents and their effects on the behaviors of the cardiomyocyte.
+However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
+These approaches assume the functional form of the shape of K\textsuperscript{+} currents and search the shape parameters in the current functions that the sum of estimated currents has the best fit with a single I\textsubscript{Ksum} recording.
+It does not provide information about underlying kinetics of the currents at the cellular level.
+First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
+Third, we design model calibration as nonlinear optimization that couple the \textit{in-silico} model and \textit{in-vitro} I\textsubscript{Ksum} recordings by minimizing discrepancies between them.
+Further, we develop a graphical user interface (GUI) application to make the proposed framework accessible for researchers and patricians without engineering backgrounds.
+We apply the proposed framework to our data, and the experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
+This method has strong potential to pave a new way to analyze K\textsuperscript{+} currents.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} \emph{IISE Transactions}; \LaTeX; Manuscript format; Taylor \& Francis.
+{\it Keywords:} Potassium channels, kinetics modeling, data assimilation, model calibration, reduced glycosylation
 
@@
 \begin{itemize}
-    \item \textbf{Motivation} Voltage-gated ion channels (VGICs) are the functional unit of the electrical conduction system of the heart that controls the synchronous contraction of numerous cardiomyocytes.
-Among major cardiac VGICs, K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature: they have various isoforms that contribute to K\textsuperscript{+} current, while the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
-Although each K\textsubscript{+} generated through K\textsubscript{v} isoform has a different role in the repolarization phase, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments.
-Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
-    \item \textbf{Gaps \& Needs} A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents.
-This method directly fits individual K\textsuperscript{+} currents to experimental I\textsubscript{Ksum} recordings by searching the shape parameters of the current functions that have the best fit.
-It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents.
-However, the mathematical form of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms.
-In addition, the curve fitting method handles experimental I\textsubscript{Ksum} recordings from the same cell independently, so it lacks the cellular-level analysis.
-    \item \textbf{Objective} As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms based on computer models of K\textsuperscript{+} currents that can simulate underlying gating kinetics of the currents.
-    \item \textbf{Approach} 
-    \item \textbf{Experiment \& Performance} The proposed framework outperforms the curve fitting method for minimizing discrepancies between experimental and estimated total K\textsuperscript{+} currents.
-We also provide the kinetics modeling results that are not possible in the traditional method.
-    \item \textbf{Significance} Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms.
+    \item \textbf{Motivation:} Aberrant activities of voltage-gated potassium (K\textsubscript{v}) channels can cause fatal heart disease such as long QT syndrome.
+In our recent investigation on a disease-related perturbation, reduced glycosylation in cardiomyocytes, significant reduction in K\textsuperscript{+} currents was observed.
+Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
+On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
+Because only the sum of these K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded in experiments, data analytical methods are necessary to estimate the properties of individual K\textsuperscript{+} currents and their effects on the behaviors of the cardiomyocyte.
+    \item \textbf{Gaps \& Needs:} However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
+These approaches assume the functional form of the shape of K\textsuperscript{+} currents and search the shape parameters in the current functions that the sum of estimated currents has the best fit with a single I\textsubscript{Ksum} recording.
+It does not provide information about underlying kinetics of the currents at the cellular level.
+    \item \textbf{Objective:} In this paper, we present a novel framework of data assimilation method that delineates kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings based on computer models of K\textsuperscript{+} currents that simulate underlying biophysics of the currents.
+As opposed to the traditional data-driven curve fitting method, the proposed biophysics-driven method estimates the cellular-level kinetics of K\textsubscript{v} isoforms.
+    \item \textbf{Approach:} First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
+Third, we design model calibration as nonlinear optimization that couple the in-silico model and in-vivo I\textsubscript{Ksum} recordings by minimizing discrepancies between them.
+Further, we develop a graphical user interface (GUI) application to make the proposed framework accessible for researchers and patricians without engineering backgrounds.
+    \item \textbf{Experiment \& Performance} We apply the proposed framework to our data, and the experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
+    \item \textbf{Significance/Contributions} 1) We model the underlying kinetics of observed currents based on biophysical principles rather than curve-fitting methods.
+2) The calibration process includes multiple protocols rather than a single protocol, resulting in comprehensive cellular-level modeling.
+3) Our method provides differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
 \end{itemize}
 
-The review of literature can be a separate section but it can be merged with the introduction section. One can find a number of reference examples at the end of this template, like,
-\begin{itemize}
-\item A book: \cite{Hastie2009}.
-\item A journal article: \cite{Byon2013}.
-\item A conference paper: \cite{Breunig2000}.
-\item A book chapter: \cite{Yang2010}.
-\item A technical report: \cite{Kelley2007}.
-\end{itemize}
-
-action potential and various ioninc currents
 \begin{figure}
     \centering
-    \includegraphics{figs/ap_underlying_currents.pdf}
-    \caption{The action potentials and underlying ionic currents in \textbf{(A)} human ventricular cardiomyocytes and \textbf{(B)} mouse ventricular cardiomyocytes \cite{nerbonne2003molecular}.}
-    \label{fig:ap_underlying_currents}
+    \includegraphics{figs/overall_diagram2.pdf}
+    \caption{Caption}
+    \label{fig:framework_diagram}
 \end{figure}
@@
 \subsection{\emph{Computer Models of Potassium Channel Isoforms}} \label{s:methods.1}
-Various K\textsuperscript{+} currents play different roles in cardiac repolarization. 
-\begin{figure}
+Computer models mouse ventricular cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP.
+In this paper, the AP is modeled as a differential equation of transmembrane currents and stimulus current I\textsubscript{stim} as below in Equation~\ref{eq:ap} where $C_{m}$ is the membrane capacitance, $t$ is time.
+\begin{equation}
+    \label{eq:ap}
+    \begin{split}
+    -C_{m}\frac{dV}{dt} &= \mathrm{I}_{\mathrm{CaL}}+\mathrm{I}_{\mathrm{p}(Ca)}+\mathrm{I}_{\mathrm{NaCa}}+\mathrm{I}_{\mathrm{Cab}}+\mathrm{I}_{\mathrm{Na}}+\mathrm{I}_{\mathrm{Nab}}+\mathrm{I}_{\mathrm{NaK}}+\mathrm{I}_{\mathrm{Cl},Ca} \\
+    &+\mathrm{I}_{\mathrm{Kto}}+\mathrm{I}_{\mathrm{Kslow1}}+\mathrm{I}_{\mathrm{Kslow2}}+\mathrm{I}_{\mathrm{Kss}}+\mathrm{I}_{\mathrm{Ks}}+\mathrm{I}_{\mathrm{Kr}}+\mathrm{I}_{\mathrm{K1}}+\mathrm{I}_{\mathrm{stim}}
+    \end{split}
+\end{equation}
+There are 15 transmembrane currents in this AP model: the L-type calcium current (I\textsubscript{CaL}), the calcium pump current (I\textsubscript{p(Ca)}), the Na\textsuperscript{+}/Ca\textsuperscript{2+} exchange current (I\textsubscript{NaCa}), the calcium background current (I\textsubscript{Cab}), the fast Na\textsuperscript{+} current (I\textsubscript{Na}), the background Na\textsuperscript{+} current (I\textsubscript{Nab}), the Na\textsuperscript{+}/K\textsuperscript{+} pump current (I\textsubscript{NaK}), the Ca\textsuperscript{2+}-activated Cl\textsuperscript{-} current (I\textsubscript{Cl,Ca}), the rapidly inactivating transient outward K\textsuperscript{+} current (I\textsubscript{Kto}), the two ultra-rapidly activating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}), the slow delayed rectifier K\textsuperscript{+} current (I\textsubscript{Ks}), the rapid delayed rectifier K\textsuperscript{+} current (I\textsubscript{Kr}), and the time independent K\textsuperscript{+} current (I\textsubscript{K1}).
+
+The three types of K\textsuperscript{+} currents have larger magnitude compared to the other K\textsuperscript{+} currents: the rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), the delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}).
+The rapidly inactivating transient outward currents have two components I\textsubscript{Kto,f} and I\textsubscript{Kto,s}, but the later is essentially absent in apex cardiomyocytes.
+We only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because we focus on apex myocytes in this investigation.
+However, the transient outward current can be easily modified according to the region of ventricular cardiomyocytes.
+Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}).
+\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{Ksum} given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
+\textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of protocols and consequent change in I\textsubscript{Ksum}.
+Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach.
+In this approach, K\textsuperscript{+} current can be defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
+\begin{figure}[!ht]
     \centering
-    \includegraphics{figs/biomarkers.pdf}
-    \caption{Six biomarkers of K\textsuperscript{+} currents that are subjects of sensitivity analysis.
-(A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, (D) I\textsubscript{Kss}.
-All currents are simulated for the purpose of illustration and biomarker labels refer to a: at 10 ms, b: 25\%, c: 50\%, d: 75\%, e: peak, and f: time constant}
-    \label{fig:biomarkers}
+    \includegraphics{figs/exemplary_kv_currents.pdf}
+    \caption{Example of voltage-clamp K\textsuperscript{+} currents.
+(A) Voltage step of 0 mV from holding potential -70 mV.
+(B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
+(C) Range of protocols from -50 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
+    \label{fig:kcurrent_example}
+\end{figure}
+\begin{equation}
+    \label{eq:expl_fitting}
+    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}
+\end{equation}
+
+Unlike the simple current functions, computer models of K\textsubscript{v} isoforms can simulate detailed gating kinetics.
+We developed them with Hodgkin-Huxley modeling that has two gating variables controlling conductance.
+For example, K\textsubscript{v} isoforms that conducts current I\textsubscript{K} can be defined by Equation~\ref{eq:hh_ex}
+\begin{equation}
+    \label{eq:hh_ex}
+    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}})
+\end{equation}
+where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential.
+$V-E_{\mathrm{K}}$ is the driving force of the ion movement.
+Important components in this equation are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$.
+These processes are governed by first-order kinetics and its voltage-dependent transition rates $\alpha$ and $\beta$.
+$\alpha$ is the rate at which a gate in closed state opens, whereas $\beta$ is the rates at which a gate in open state closes.
+Equation~\ref{eq:gv} shows a schematic relationship of this gating kinetics.
+\begin{align}
+    \label{eq:gv}
+    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i
+\end{align}
+
+The two biophysical processes can be modeled using differential equations in two ways as follows:
+\begin{align}
+    \frac{da}{dt} &=\alpha_{a}(1-a)-\beta_{a}a   &\frac{di}{dt} &=\alpha_{i}(1-i)-\beta_{i}i \\
+    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}}
+\end{align}
+where $a_{\infty}$ and $i_{\infty}$ are the steady-state values to which $a$ and $b$ converge; $\tau_{a}$ and $\tau_{i}$ are time constants determine the convergence speed defined by
+\begin{align}
+    a_{\infty} &= \frac{\alpha_{a}}{\alpha_{a}+\beta_{a}} & i_{\infty} &=  \frac{\alpha_{i}}{\alpha_{i}+\beta_{i}} \\
+    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}
+\end{align}
+The steady-state values and time constants can be defined directly by functions of voltage without transition rates in some cases.
+These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters that control the behavior of the kinetics of an ion channel.
+
+\subsubsection{In-silico Modeling of K\textsubscript{v}4.2 (I\textsubscript{Kto})}
+The rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation.
+It mainly contributes to the peak at the very beginning of activation in I\textsubscript{Ksum}.
+I\textsubscript{Kto} is defined by
+\begin{align}
+    &\mathrm{I}_{\mathrm{Kto}} = G_{\mathrm{Kto}}a_{\mathrm{Kto}}^{3}i_{\mathrm{Kto}}(V-E_{\mathrm{K}}) \\
+    &\frac{da_{\mathrm{Kto}}}{dt} = \alpha_{a}(1-a_{\mathrm{Kto}}) - \beta_{a}a_{\mathrm{Kto}} \\
+    &\frac{di_{\mathrm{Kto}}}{dt} = \alpha_{i}(1-i_{\mathrm{Kto}}) - \beta_{i}i_{\mathrm{Kto}} \\
+    &\alpha_{a} = p_{7}e^{p_{5}(V+p_{1})} \label{eq:ikto_alpha1} \\
+    &\beta_{a}= p_{8}e^{-p_6(V+p_{1})} \\
+    &\alpha_{i} = \frac{p_{9}e^{-(V+p_{2})/p_{4}}}{1+p_{10}e^{-(V+p_{2}+p_{3})/p_{4}}} \\
+    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}} \label{eq:ikto_beta2}
+\end{align}
+There are two gating variables $a_{\mathrm{Kto}}$ and $i_{\mathrm{Kto}}$ responsible for activation and inactivation, respectively.
+Their kinetics are governed by transition-rate functions from Equation~\ref{eq:ikto_alpha1} to Equation~\ref{eq:ikto_beta2}.
+Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like ``knobs'', allowing to control the behavior of the I\textsubscript{Kto} model.
+\textbf{Figure~\ref{fig:kinetic_iktof}} shows a representative example of the relationship between the transition-rate functions and the channel kinetics, given the specific value of $p_{i}$.
+In general, the transition rates for inactivation gates change more rapidly and have smaller ranges than for activation, which results in a drastic slope in the steady-state curve and a large time constant.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/kinetic_iktof.pdf}
+    \caption{Example of the behavior of the transition-rate functions with changes in voltage, given specific values of parameters that govern kinetics of I\textsubscript{Kto}.}
+    \label{fig:kinetic_iktof}
 \end{figure}
 
-\subsection{\emph{Sensitivity Analysis of Calibration Parameters}} \label{s:methods.2}
-this section shows the results of the variable screening.
-\begin{figure}
+\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})}
+There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are very rapidly activating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
+As shown in \textbf{Figure~\ref{fig:kcurrent_example}}, both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, but I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1}.
+The non-inactivating steady-state current, which conducted through K\textsubscript{2P} family, remains constant during the voltage-clamp recording.
+These three currents collectively address the most part of the decaying phase of I\textsubscript{Ksum}.
+We assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and I\textsubscript{Kss} has similar activation behavior with a slightly different rate to keep the model as simple as possible to reduce the structural risk of overfitting.
+
+I\textsubscript{Kslow1} is modeled without transition-rate functions as opposed to I\textsubscript{Kto}.
+Its gating variables, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$ are defined directly by steady-state ($a_{ss}$ and $i_{ss}$) and time-constant functions ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) as follows.
+\begin{align}
+    &\mathrm{I}_{\mathrm{Kslow1}} = G_{\mathrm{Kslow1}}a_{\mathrm{Kslow1}}i_{\mathrm{Kslow1}}(V-E_{\mathrm{K}}) \\
+    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(1)}} \\
+    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(1)}} \\
+    &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \\
+    &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \\
+    &\tau_{a}^{(1)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
+    &\tau_{i}^{(1)} = p_{10} - p_{8}i_{ss}
+\end{align}
+
+I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function for the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}.
+As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows.
+\begin{align}
+    &\mathrm{I}_{\mathrm{Kslow2}} = G_{\mathrm{Kslow2}}a_{\mathrm{Kslow2}}i_{\mathrm{Kslow2}}(V-E_{\mathrm{K}}) \\
+    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \\
+    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(2)}} \\
+    &\tau_{i}^{(2)} = p_{2} - p_{1}i_{ss}
+\end{align}
+
+I\textsubscript{Kss} does not have an inactivation variable as it is non-inactivating current.
+It shares the same steady-state function for activation $a_{ss}$ with the rectifier currents but have a separate time-constant function to address the different activation rate.
+I\textsubscript{Kss} is modeled as
+\begin{align}
+    &\mathrm{I}_{\mathrm{Kss}} = G_{\mathrm{Kss}}a_{\mathrm{Kss}}(V-E_{\mathrm{K}}) \\
+    &\frac{da_{\mathrm{Kss}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kss}}}{\tau_{a}^{(3)}} \\
+    &\tau_{a}^{(3)}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}
+\end{align}
+Note that $p_{3}^\prime$ is equal to $p_{3}$ in I\textsubscript{Kslow1}.
+\textbf{Figure~\ref{fig:kinetic_rectifiers}} illustrates general trends of the relationship of kinetic variables and voltage of the rectifier currents.
+\begin{figure}[!ht]
     \centering
-    \includegraphics{figs/doe_kto.pdf}
-    \caption{Half-normal plots I\textsubscript{Kto}}
-    \label{fig:doe_kto}
+    \includegraphics{figs/kinetic_rectifiers.pdf}
+    \caption{Example of the voltage dependence of kinetic variables of the rectifier currents.}
+    \label{fig:kinetic_rectifiers}
 \end{figure}
 
-\begin{figure}
+\subsection{\emph{Data Assimilation and Model Calibration}} \label{s:methods.2}
+\textit{Data assimilation} is a method to find the optimal configuration and state of computational models by coupling them with experimental data.
+Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation.
+The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment.
+Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term.
+\begin{equation}
+    \label{eq:rd_relation}
+    y^{\mathcal{R}}(x) = y^{\mathcal{D}}(x) + \epsilon
+\end{equation}
+Let $y^{\mathcal{M}}(x|\theta)$ denote the output from a computer model $\mathcal{M}$, given parameters $\theta$.
+Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in Equation~\ref{eq:dm_relation}.
+\begin{align}
+    \label{eq:dm_relation}
+    y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
+    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
+\end{align}
+Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints.
+By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.
+
+In this study, $\delta$ is defined by the sum of root-mean-square errors (RMSEs) as in Equation~\ref{eq:rmse}, which measures deviations between experimental data and model predictions from the end of a holding potential $t_i^{(h)}$ to the end of voltage step $t_i^{(e)}$ across a range of protocols, $i=1,2,\cdots,n$.
+\begin{equation}
+    \label{eq:rmse}
+    \delta = \sum_{i=1}^{n} \sqrt{\int_{t_i^{(h)}}^{t_i^{(e)}}\frac{(y_i^{\mathcal{D}}(t) - y_i^{\mathcal{M}}(t|\theta))^2}{t_i^{(e)}-t_i^{(h)}}dt}
+\end{equation}
+This approach differs from previous studies in two-fold.
+First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method includes multiple protocols simultaneously for comprehensive cellular-level modeling.
+Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}.
+Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.
+
+We developed the box-constrained nonlinear optimization routine using the BFGS algorithm with multi-random initial points to minimize $\delta$.
+Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube.
+In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling.
+Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting.
+Global optimum is not guaranteed due to the complexity of our models.
+We developed the multi-random starting scheme to avoid local optima and find the solution as close to the optimum as possible.
+Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
+This work is implemented in MATLAB R2022a, and the code is available at xyz.
+
+\subsection{\emph{Sensitivity Analysis of Calibration Parameters}} \label{s:methods.3}
+The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$.
+We performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output, i.e., current traces.
+We defined six \textit{anchor points} that capture representative characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments as illustrated in \textbf{Figure~\ref{fig:current_feature}}.
+Each point represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
+Feature f will be equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_feature}C} and \textbf{~\ref{fig:current_feature}D}.
+
+We developed factorial designs in which parameters varied at two levels, contrasting their effect.
+The anchor points were evaluated at each design to calculate factorial effects.
+Then, half-normal plots were drawn based on these contrasts, providing visual inspections of critical parameters that exert the most impact on each shape feature.
+All the half-normal plots can be found in Supplement A.
+\textbf{Table~\ref{tab:calib_param}} summarizes the selected parameters that will be calibrated to couple the \textit{in-silico} models with \textit{in-vitro} experimental data.
+\begin{figure}[!ht]
     \centering
-    \includegraphics{figs/doe_kslow1.pdf}
-    \caption{Half-normal plots I\textsubscript{Kslow1}}
-    \label{fig:doe_kslow1}
+    \includegraphics{figs/anchor_ponts.pdf}
+    \caption{Examples of the six feature of K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
+All currents are simulated for illustration, and the feature labels refer to a: the current magnitude 10 ms after voltage is applied, b: 25\% of the total recording time has elapsed, c: 50\%, d: 75\%, e: the peak magnitude, and f: the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
+    \label{fig:current_feature}
+\end{figure}
+\begin{table}
+    \centering
+    \caption{Calibration parameters that will be optimized to minimize the discrepancy between model predictions and experimental data.}
+    \begin{tabular}{l|l}
+        \hline
+        Model & Calibration parameters \\
+        \hline
+        K\textsubscript{v}4.2 (I\textsubscript{Kto}) & $p_{1},\quad p_{2},\quad p_{4},\quad p_{5},\quad p_{7},\quad p_{11},\quad G_{\mathrm{Kto}}$ \\
+        K\textsubscript{v}1.5 (I\textsubscript{Kslow1}) & $p_{1},\quad p_{2},\quad p_{4},\quad p_{5},\quad p_{9},\quad p_{10},\quad G_{\mathrm{Kslow1}}$ \\
+        K\textsubscript{v}2.1 (I\textsubscript{Kslow1}) & $p_{2},\quad G_{\mathrm{Kslow2}}$ \\
+        K\textsubscript{2P} (I\textsubscript{Kss}) & $p_{1},\quad p_{2},\quad p_{3},\quad G_{\mathrm{Kss}}$ \\
+        \hline
+    \end{tabular}
+    \label{tab:calib_param}
+\end{table}
+
+We categorized the calibration parameters into four classes according to their functional roles in kinetic functions, which are highlighted in different colors in \textbf{Figure~\ref{fig:calib_param}}.
+red: voltage threshold, green: voltage slope, blue: kinetic-function factor, and time-constant shift. 
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/calib_param.pdf}
+    \caption{Calibration parameters with highlights according to the class}
+    \label{fig:calib_param}
 \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics{figs/doe_kslow2.pdf}
-    \caption{Half-normal plots I\textsubscript{Kslow2}}
-    \label{fig:doe_kslow2}
-\end{figure}
+\subsection{\emph{Glycosylation and Dilated Cardiomyopathy (DCM)}} \label{s:methods.4}
+Case study with our own in-vitro experimental data.
 
-\begin{figure}
-    \centering
-    \includegraphics{figs/doe_kss.pdf}
-    \caption{Half-normal plots I\textsubscript{Kss}}
-    \label{fig:doe_kss}
-\end{figure}
+\section{Results} \label{s:results}
+\subsection{Experimental Design: Reduced Glycosylation}
+case study with the in-vitro data of investigating the casuality between reduced glycosylation and dilated cardiomyopathy.
 
-\subsection{\emph{Model Calibration}} \label{s:methods.3}
-calibration procedure; non-linear optimization methods
+\subsection{Experimental Results}
+case study results with the proposed data assimilation method
 \begin{figure}
@@
 
-\subsection{\emph{Graphical User Interface}} \label{s:methods.4}
-show how to use it
-
-\subsection{\emph{Case Study Glycosylations}} \label{s:methods.5}
-case study
-
-\section{Results} \label{s:results}
 Perhaps numerical analysis of performance, comparison, and demonstration of applicability and impact using real data.
@@
 \section*{Acknowledgements}
-The authors acknowledge the generous support from the funding agency of XYZ.	} \fi
+The authors acknowledge the generous support from the funding agency of XYZ and funding list} \fi
 

### 2022-06-02 writing/draft.tex
Figures removed
Kinetic functions & result figures

@@
 Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like ``knobs'', allowing to control the behavior of the I\textsubscript{Kto} model.
-\textbf{Figure~\ref{fig:kinetic_iktof}} shows a representative example of the relationship between the transition-rate functions and the channel kinetics, given the specific value of $p_{i}$.
-In general, the transition rates for inactivation gates change more rapidly and have smaller ranges than for activation, which results in a drastic slope in the steady-state curve and a large time constant.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/kinetic_iktof.pdf}
-    \caption{Example of the behavior of the transition-rate functions with changes in voltage, given specific values of parameters that govern kinetics of I\textsubscript{Kto}.}
-    \label{fig:kinetic_iktof}
-\end{figure}
 
@@
 Note that $p_{3}^\prime$ is equal to $p_{3}$ in I\textsubscript{Kslow1}.
-\textbf{Figure~\ref{fig:kinetic_rectifiers}} illustrates general trends of the relationship of kinetic variables and voltage of the rectifier currents.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/kinetic_rectifiers.pdf}
-    \caption{Example of the voltage dependence of kinetic variables of the rectifier currents.}
-    \label{fig:kinetic_rectifiers}
-\end{figure}
 
@@
 case study results with the proposed data assimilation method
-\begin{figure}
-    \centering
-    \includegraphics{figs/pkto.pdf}
-    \caption{parameter distribution}
-    \label{fig:pkto}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/pkslow1.pdf}
-    \caption{parameter distribution}
-    \label{fig:pkslow1}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/pkslow2.pdf}
-    \caption{parameter distribution}
-    \label{fig:pkslow2}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/pkss.pdf}
-    \caption{parameter distribution}
-    \label{fig:pkss}
-\end{figure}
 

### 2022-06-05 writing/draft.tex
Method section completed & New intro

@@
 \noindent%
-{\it Keywords:} Potassium channels, kinetics modeling, data assimilation, model calibration, reduced glycosylation
+{\it Keywords:} Data assimilation, potassium channels, kinetics modeling, model calibration, glycosylation, cardiomyopathy,
 
@@
 \section{Introduction} \label{s:intro}
+Cardiac voltage-gated potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential (AP) in cardiomyocytes.
+They are indispensable in the electrical conduction system of the heart.
+Even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
+Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, has shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics}.
+However, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
+Data from our \textit{in-vitro} experiments show that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
+We also conducted in-silico experiments to verify the data and investigate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the unique powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico}.
+As illustrated in \textbf{Figure~\ref{fig:framework_diagram}},
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/overall_diagram2.pdf}
+    \caption{Caption}
+    \label{fig:framework_diagram}
+\end{figure}
+
+However, unique feature of Kv raise challenges.
 \begin{itemize}
-    \item \textbf{Motivation:} Aberrant activities of voltage-gated potassium (K\textsubscript{v}) channels can cause fatal heart disease such as long QT syndrome.
-In our recent investigation on a disease-related perturbation, reduced glycosylation in cardiomyocytes, significant reduction in K\textsuperscript{+} currents was observed.
-Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
+    \item \textbf{Gaps \& Needs:} Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
 On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
 Because only the sum of these K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded in experiments, data analytical methods are necessary to estimate the properties of individual K\textsuperscript{+} currents and their effects on the behaviors of the cardiomyocyte.
-    \item \textbf{Gaps \& Needs:} However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
+However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
 These approaches assume the functional form of the shape of K\textsuperscript{+} currents and search the shape parameters in the current functions that the sum of estimated currents has the best fit with a single I\textsubscript{Ksum} recording.
@@
 As opposed to the traditional data-driven curve fitting method, the proposed biophysics-driven method estimates the cellular-level kinetics of K\textsubscript{v} isoforms.
-    \item \textbf{Approach:} First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+    \item \textbf{Approach: (need a flowchart} First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
 Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
@@
 \end{itemize}
-
-\begin{figure}
-    \centering
-    \includegraphics{figs/overall_diagram2.pdf}
-    \caption{Caption}
-    \label{fig:framework_diagram}
-\end{figure}
 
@@
 
-\subsection{\emph{Sensitivity Analysis of Calibration Parameters}} \label{s:methods.3}
+\subsection{\emph{Sensitivity Analysis and Model Regularization}} \label{s:methods.3}
 The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$.
-We performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output, i.e., current traces.
-We defined six \textit{anchor points} that capture representative characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments as illustrated in \textbf{Figure~\ref{fig:current_feature}}.
+We performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
+As illustrated in \textbf{Figure~\ref{fig:anchor_points}}, six \textit{anchor points} were defined that capture representative characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
 Each point represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
-Feature f will be equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_feature}C} and \textbf{~\ref{fig:current_feature}D}.
+Feature f will be equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:anchor_points}C} and \textbf{~\ref{fig:anchor_points}D}.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/anchor_ponts.pdf}
+    \caption{Illustration of the six anchor points of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
+All currents are simulated for illustration, and the labels refer to a) the current magnitude 10 ms after voltage is applied, b) 25\% of the total recording time has elapsed, c) 50\%, d) 75\%, e) the peak magnitude, and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
+    \label{fig:anchor_points}
+\end{figure}
 
@@
 All the half-normal plots can be found in Supplement A.
-\textbf{Table~\ref{tab:calib_param}} summarizes the selected parameters that will be calibrated to couple the \textit{in-silico} models with \textit{in-vitro} experimental data.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/anchor_ponts.pdf}
-    \caption{Examples of the six feature of K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
-All currents are simulated for illustration, and the feature labels refer to a: the current magnitude 10 ms after voltage is applied, b: 25\% of the total recording time has elapsed, c: 50\%, d: 75\%, e: the peak magnitude, and f: the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
-    \label{fig:current_feature}
-\end{figure}
-\begin{table}
-    \centering
-    \caption{Calibration parameters that will be optimized to minimize the discrepancy between model predictions and experimental data.}
-    \begin{tabular}{l|l}
-        \hline
-        Model & Calibration parameters \\
-        \hline
-        K\textsubscript{v}4.2 (I\textsubscript{Kto}) & $p_{1},\quad p_{2},\quad p_{4},\quad p_{5},\quad p_{7},\quad p_{11},\quad G_{\mathrm{Kto}}$ \\
-        K\textsubscript{v}1.5 (I\textsubscript{Kslow1}) & $p_{1},\quad p_{2},\quad p_{4},\quad p_{5},\quad p_{9},\quad p_{10},\quad G_{\mathrm{Kslow1}}$ \\
-        K\textsubscript{v}2.1 (I\textsubscript{Kslow1}) & $p_{2},\quad G_{\mathrm{Kslow2}}$ \\
-        K\textsubscript{2P} (I\textsubscript{Kss}) & $p_{1},\quad p_{2},\quad p_{3},\quad G_{\mathrm{Kss}}$ \\
-        \hline
-    \end{tabular}
-    \label{tab:calib_param}
-\end{table}
-
-We categorized the calibration parameters into four classes according to their functional roles in kinetic functions, which are highlighted in different colors in \textbf{Figure~\ref{fig:calib_param}}.
-red: voltage threshold, green: voltage slope, blue: kinetic-function factor, and time-constant shift. 
+\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
+We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters.
+Note that the voltage-dependence parameters in red and green appear multiple times across different equations.
+This parsimonious model design is intended to maximize the structural regularization to minimize overfitting.
 \begin{figure}[!ht]
@@
     \includegraphics{figs/calib_param.pdf}
-    \caption{Calibration parameters with highlights according to the class}
+    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
     \label{fig:calib_param}
@@
 \subsection{\emph{Glycosylation and Dilated Cardiomyopathy (DCM)}} \label{s:methods.4}
-Case study with our own in-vitro experimental data.
+We applied the proposed framework to our studies on the pathophysiology of electrical signaling by altered glycosylation.
+Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}.
+It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}.
+We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology.
+Electrical signaling is orchestrated activities of a variety of ion channels and transporters.
+Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
+
+Glycosylation is a multi-step process and usually ends with sialic acid added.
+We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
+More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
+MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
+To further investigate the role of altered glycosylation and pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
+However, it was difficult to determine channel kinetics rigorously from the \textit{in-vitro} experiments alone, especially for K\textsuperscript{+} current recordings, due to vast diversity of K\textsubscript{v} isoforms and small current magnitude of the MGAT1KO model.
+Therefore, we validate the suggested framework by modeling K\textsubscript{v} isoforms with our \textit{in-vitro} experimental data.
+Therefore, we leverage the suggested framework to delineate the channel kinetics of K\textsubscript{v} to provide new descriptive/predictive insights into our \textit{in-vitro} experiments.
 
 \section{Results} \label{s:results}
-\subsection{Experimental Design: Reduced Glycosylation}
-case study with the in-vitro data of investigating the casuality between reduced glycosylation and dilated cardiomyopathy.
 
-\subsection{Experimental Results}
-case study results with the proposed data assimilation method
+\subsection{Calibration Results}
 
+\subsection{Distributions Estimations}
+
+\subsection{Clustering}
 Perhaps numerical analysis of performance, comparison, and demonstration of applicability and impact using real data.
+
+\section{Discuss}
+Discuss of the results goes here.
 

### 2022-06-08 writing/draft.tex
New intro

@@
 Even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
-Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, has shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics}.
-However, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
-Data from our \textit{in-vitro} experiments show that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
-We also conducted in-silico experiments to verify the data and investigate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the unique powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico}.
-As illustrated in \textbf{Figure~\ref{fig:framework_diagram}},
+Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics}.
+Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
+Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
+In addition, we verified our experimental results via \textit{in-silico} studies and estimate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
+\textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases, which will be referred to as the data assimilation framework that combines experimental data with mathematical/computational models.
 \begin{figure}[!ht]
@@
     \includegraphics{figs/overall_diagram2.pdf}
-    \caption{Caption}
+    \caption{Conceptual diagram to illustrate a data assimilation framework.}
     \label{fig:framework_diagram}
@@
 
-However, unique feature of Kv raise challenges.
+However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies of them: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
+In contrast, there are predominant isoforms for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
+Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
+Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments.
+Typically, curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of current are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
+However, this traditional approach is limited in its ability to estimate information about kinetics and gating of the channels in two aspects: 1) curve fitting provides parameters pertinent to the current shape only, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
+These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
+Therefore, there is an urgent need to develop new data assimilation methods and tools that delineate the kinetics of K\textsubscript{v} isoforms for better understanding their roles in heart diseases.
+
+This paper presents a novel approach to data assimilation for analyzing whole-cell I\textsubscript{K} recordings and modeling the underlying kinetics of K\textsubscript{v} isoforms.
+\textbf{(need a flowchart)} First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting.
+Computer models of ion channels are a system of differential equations consisting of gating variables and kinetic functions that continuously simulate ion channel activities over time.
+Parameters in kinetic functions control the gating of ion channels and, in turn, determine current outputs.
+Second, a sensitivity analysis was performed to screen out parameters that marginally influence the model outputs.
+Third, we developed a model calibration that minimizes discrepancy between model outputs and experimental data generated from multiple protocols.
+Further, we packaged the suggested framework into a graphical user interface (GUI) application to make the proposed framework more accessible.
+Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
+
+The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases.
+Our contributions are summarized as follows:
 \begin{itemize}
-    \item \textbf{Gaps \& Needs:} Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
-On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
-Because only the sum of these K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded in experiments, data analytical methods are necessary to estimate the properties of individual K\textsuperscript{+} currents and their effects on the behaviors of the cardiomyocyte.
-However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
-These approaches assume the functional form of the shape of K\textsuperscript{+} currents and search the shape parameters in the current functions that the sum of estimated currents has the best fit with a single I\textsubscript{Ksum} recording.
-It does not provide information about underlying kinetics of the currents at the cellular level.
-    \item \textbf{Objective:} In this paper, we present a novel framework of data assimilation method that delineates kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings based on computer models of K\textsuperscript{+} currents that simulate underlying biophysics of the currents.
-As opposed to the traditional data-driven curve fitting method, the proposed biophysics-driven method estimates the cellular-level kinetics of K\textsubscript{v} isoforms.
-    \item \textbf{Approach: (need a flowchart} First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
-Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
-Third, we design model calibration as nonlinear optimization that couple the in-silico model and in-vivo I\textsubscript{Ksum} recordings by minimizing discrepancies between them.
-Further, we develop a graphical user interface (GUI) application to make the proposed framework accessible for researchers and patricians without engineering backgrounds.
-    \item \textbf{Experiment \& Performance} We apply the proposed framework to our data, and the experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
-    \item \textbf{Significance/Contributions} 1) We model the underlying kinetics of observed currents based on biophysical principles rather than curve-fitting methods.
-2) The calibration process includes multiple protocols rather than a single protocol, resulting in comprehensive cellular-level modeling.
-3) Our method provides differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
+    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
+    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
+    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
 \end{itemize}
+The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.
 

### 2022-06-08 calibration/param_dist_viz/param_preproc.py
Parameter distribution data

@@ -40,8 +40,8 @@ def compare_param_dist(pdf_wt, pdf_ko, idx):
     p = sns.displot(pdf, x=str(idx), hue='Group', kind='kde', fill=True)
     p.set_xlabels('Parameter '+str(idx), fontsize=12)
     
-exp_num =  '41'
-currents = ['iktof', 'ikslow1', 'ikslow2', 'ikss']
+exp_num =  '45'
+currents = ['iktof','ikslow1','ikslow2','ikss']
 idx_dict = {'iktof': [1, 2, 4, 5, 7, 11, 13],
     'iktos': [2, 3],
     'ikslow1': [1, 2, 4, 5, 9, 10, 11],

### 2022-06-13 writing/draft.tex
Flowchart added

@@
     \centering
-    \includegraphics{figs/overall_diagram2.pdf}
+    \includegraphics{figs/overall_diagram.pdf}
     \caption{Conceptual diagram to illustrate a data assimilation framework.}
@@
 
-However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies of them: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
+However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
 In contrast, there are predominant isoforms for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
@@
 Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments.
-Typically, curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of current are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
-However, this traditional approach is limited in its ability to estimate information about kinetics and gating of the channels in two aspects: 1) curve fitting provides parameters pertinent to the current shape only, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
+Typically, curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
+Then the shape parameters are used to calibrate computer models \citep{kim2022simulation}, or kinetics information is added to the data assimilation process with additional curve fitting if applicable \citep{du2017silico}.
+However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of the K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
 These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
@@
 This paper presents a novel approach to data assimilation for analyzing whole-cell I\textsubscript{K} recordings and modeling the underlying kinetics of K\textsubscript{v} isoforms.
-\textbf{(need a flowchart)} First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting.
+A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}.
+First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting.
 Computer models of ion channels are a system of differential equations consisting of gating variables and kinetic functions that continuously simulate ion channel activities over time.
@@
 Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
+\begin{figure}
+    \centering
+    \includegraphics{figs/flow_chart.pdf}
+    \caption{Flowcharts of (A) the current approach and (B) the proposed method.}
+    \label{fig:flow_chart}
+\end{figure}
 
@@
 \subsection{Calibration Results}
+first we compared fitting results.
 

### 2022-06-13 calibration/log_analysis.R
Log processed

@@ -46,50 +46,21 @@ find_min_rmse <- function(log_list) {
 log_dir <- "./calibration/log/"
 
 ## create data frame for bar plot -----
-num_files1 <- 34
-num_files2 <- 33
+num_files <- 33
 num_iters <- 30
+group <- "ko";
 log1_name <- "exp42"
-log2_name <- "exp36"
-log3_name <- "exp37"
-# log4_name <- "exp23"
+log2_name <- "exp43"
+log3_name <- "exp44"
 
-log1 <- read_log(str_c(log_dir,log1_name,"_wt",".txt"), num_files1, num_iters)
-log2 <- read_log(str_c(log_dir,log1_name,"_ko",".txt"), num_files2, num_iters)
-# log3 <- read_log(str_c(log_dir,log3_name,"_wt",".txt"), num_files1, num_iters)
-# log4 <- read_log(str_c(log_dir, log4_name, "_", group_name, ".txt"), num_files, 1)
+log1 <- read_log(str_c(log_dir,log1_name,"_",group,".txt"), num_files, num_iters)
+log2 <- read_log(str_c(log_dir,log1_name,"_",group,".txt"), num_files, num_iters)
+log3 <- read_log(str_c(log_dir,log3_name,"_",group,".txt"), num_files, num_iters)
 
 rmse_df1 <- find_min_rmse(log1)
 rmse_df2 <- find_min_rmse(log2)
-# rmse_df3 <- find_min_rmse(log3)
-# rmse_val4 <- find_min_rmse(log4)
+rmse_df3 <- find_min_rmse(log3)
 
-write_csv(rmse_df1, str_c(log_dir,log1_name,"_wt.csv"))
-write_csv(rmse_df2, str_c(log_dir,log1_name,"_ko.csv"))
-# write_csv(rmse_df3, str_c(log_dir,log3_name,"_wt.csv"))
-
-# file_names <- seq(1, num_files) %>% as.character()
-# extra_idx <- c(18, 19, 29, 31, 32)
-# rmse_val1[extra_idx]
-
-## create data frame for multiple bar plot -----
-bar_df <- tibble(
-  name = file_names,
-  mtd1 = rmse_val1,
-  mtd2 = rmse_val2,
-  mtd3 = rmse_val3,
-  mtd4 = rmse_val4)
-
-# export for MATLAB
-write_csv(bar_df, './calibration/bar_graph_norm.csv')
-
-bar_df <- bar_df %>% 
[…]

### 2022-06-15 calibration/kinetic_predict.m
Results analysis

@@ -1,0 +1,423 @@
+clc
+clearvars
+close all
+
+load('file_names-25s.mat')
+exp_num1 = "exp41";
+sol_dir = fullfile(pwd,strcat("calib_",exp_num1));
+
+current_names = {'iktof','ikslow1','ikslow2','ikss'};
+tune_idx = cell(5,1); % iktof, iktos, ikslow1, ikslow2, ikss
+tune_idx{1} = [1, 2, 4, 5, 7, 11, 13];
+tune_idx{2} = [2, 3];
+tune_idx{3} = [1, 2, 4, 5, 9, 10, 11];
+tune_idx{4} = [2, 3];
+tune_idx{5} = [1, 2, 3, 4];
+[mdl_struct,psize] = gen_mdl_struct(current_names,tune_idx);
+
+pdefault = cell(5,1);
+pdefault{1} = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, ...
+    0.000152, 0.067083, 0.00095, 0.051335, 0.3846];
+pdefault{2} = [-1050, 270, 0.0629];
+pdefault{3} = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18, 2.058, 803.0, 0.16];
+pdefault{4} = [4912, 5334, 0.16];
+pdefault{5} = [0.0862, 1235.5, 13.17, 0.0611];
+
+volts = -50:10:50;
+ideal_hold_time = 470;
+ideal_end_time = 25*1000;
+t = 0:ideal_end_time;
+pulset = t(ideal_hold_time+1:end);
+
+protocol = cell(6,1);
+protocol{1} = -70; % hold_volt
+protocol{2} = -91.1; % ek
+protocol{4} = t;
+protocol{5} = t(1:ideal_hold_time);
+protocol{6} = pulset - pulset(1);
+
+atof_wt = NaN(length(files_wt),length(volts));
+itof_wt = NaN(length(files_wt),length(volts));
+taua_tof_wt = NaN(length(files_wt),length(volts));
+taui_tof_wt = NaN(length(files_wt),length(volts));
+akslow1_wt = NaN(length(files_wt),length(volts));
+ikslow1_wt = NaN(length(files_wt),length(volts));
+taua_kslow1_wt = NaN(length(files_wt),length(volts));
+taui_kslow1_wt = NaN(length(files_wt),length(volts));
+akslow2_wt = NaN(length(files_wt),length(volts));
[…]

### 2022-06-15 calibration/log_analysis.R
Log processed

@@ -46,10 +46,10 @@ find_min_rmse <- function(log_list) {
 log_dir <- "./calibration/log/"
 
 ## create data frame for bar plot -----
-num_files <- 33
+num_files <- 35
 num_iters <- 30
-group <- "ko";
-log1_name <- "exp42"
+group <- "wt";
+log1_name <- "exp41"
 log2_name <- "exp43"
 log3_name <- "exp44"
 

### 2022-06-15 calibration/log_analysis.R
Log processed

@@ -46,12 +46,12 @@ find_min_rmse <- function(log_list) {
 log_dir <- "./calibration/log/"
 
 ## create data frame for bar plot -----
-num_files <- 35
+num_files <- 34
 num_iters <- 30
 group <- "wt";
-log1_name <- "exp41"
-log2_name <- "exp43"
-log3_name <- "exp44"
+log1_name <- "exp45"
+log2_name <- "exp46"
+log3_name <- "exp47"
 
 log1 <- read_log(str_c(log_dir,log1_name,"_",group,".txt"), num_files, num_iters)
 log2 <- read_log(str_c(log_dir,log1_name,"_",group,".txt"), num_files, num_iters)

### 2022-06-15 calibration/log_analysis.R
Log processed

@@ -46,9 +46,9 @@ find_min_rmse <- function(log_list) {
 log_dir <- "./calibration/log/"
 
 ## create data frame for bar plot -----
-num_files <- 34
+num_files <- 33
 num_iters <- 30
-group <- "wt";
+group <- "ko";
 log1_name <- "exp45"
 log2_name <- "exp46"
 log3_name <- "exp47"

### 2022-06-16 calibration/viz_calib_results.m
Result analysis

@@ -1,67 +1,16 @@
-%% multiple bar graph
-clc
-clearvars
-close all
-
-exp_num1 = "exp35";
-exp_num2 = "exp36";
-exp_num3 = "exp37";
-group = "wt";
-
-rmse_df1 = readtable(fullfile(pwd,"log",strcat(exp_num1,"_",group,".csv")));
-rmse_df2 = readtable(fullfile(pwd,"log",strcat(exp_num2,"_",group,".csv")));
-rmse_df3 = readtable(fullfile(pwd,"log",strcat(exp_num3,"_",group,".csv")));
-rmse_df_comb = join(rmse_df1,rmse_df2, 'Keys','File');
-rmse_df_comb = join(rmse_df_comb,rmse_df3, 'Keys','File');
-
-file_names = categorical(rmse_df_comb.File);
-rmse = table2array(rmse_df_comb(:,2:end));
-
-figure('Color','w', 'Position',[50,50,900,770])
-barh(file_names,rmse)
-legend("Interior point","SQP","Active set")
-set(gca, 'FontName','Arial', 'FontSize',11, 'FontWeight','bold')
-
-%% bar graph of RMSE
+%% RMSE bar graph and current fitting
 clc
 close all
 
-file_group = "wt";
-exp_num = "exp35";
-
-rmse_df = readtable(fullfile(pwd,"log",strcat(exp_num,"_",file_group,".csv")));
-file_names = categorical(rmse_df.File);
-rmse = rmse_df.RMSE;
-
-figure('Color','w', 'Position',[100,100,670,550])
-barh(file_names,rmse, 'blue')
-set(gca, 'FontName','Arial', 'FontSize',11, 'FontWeight','bold')
-
-%% compare with experimental data (25-sec data)
-clc
-close all
-clearvars
-
-group = "ko";
-exp_num = "exp41";
-file_name = "15n03018.xlsx";
-
-data_path = fullfile(pwd,"mgat1ko_data",strcat(group,"-preprocessed-25s"),file_name);
-calib_path = fullfile(pwd,strcat("calib_",exp_num,"_",group),file_name);
-
-calib = table2array(readtable(calib_path));
-
-%----- Model structure -----%
-current_names = {'iktof', 'ikslow1', 'ikslow2', 'ikss'};
[…]

### 2022-06-16 calibration/viz_calib_results.m
Kinetics modeling results

@@ -190,11 +190,7 @@ akslow1_wt = NaN(length(files_wt),length(volts));
 ikslow1_wt = NaN(length(files_wt),length(volts));
 taua_kslow1_wt = NaN(length(files_wt),length(volts));
 taui_kslow1_wt = NaN(length(files_wt),length(volts));
-akslow2_wt = NaN(length(files_wt),length(volts));
-ikslow2_wt = NaN(length(files_wt),length(volts));
-taua_kslow2_wt = NaN(length(files_wt),length(volts));
 taui_kslow2_wt = NaN(length(files_wt),length(volts));
-akss_wt = NaN(length(files_wt),length(volts));
 taua_kss_wt = NaN(length(files_wt),length(volts));
 
 atof_ko = NaN(length(files_ko),length(volts));
@@ -205,11 +201,7 @@ akslow1_ko = NaN(length(files_ko),length(volts));
 ikslow1_ko = NaN(length(files_ko),length(volts));
 taua_kslow1_ko = NaN(length(files_ko),length(volts));
 taui_kslow1_ko = NaN(length(files_ko),length(volts));
-akslow2_ko = NaN(length(files_ko),length(volts));
-ikslow2_ko = NaN(length(files_ko),length(volts));
-taua_kslow2_ko = NaN(length(files_ko),length(volts));
 taui_kslow2_ko = NaN(length(files_ko),length(volts));
-akss_ko = NaN(length(files_ko),length(volts));
 taua_kss_ko = NaN(length(files_ko),length(volts));
 for i=1:length(files_wt)
     sol_mx = table2array(readtable(fullfile(sol_dir,"wt",strcat(files_wt(i),".xlsx"))));
@@ -226,11 +218,7 @@ for i=1:length(files_wt)
         ikslow1_wt(i,j) = kv{2}(2);
         taua_kslow1_wt(i,j) = kv{2}(3);
         taui_kslow1_wt(i,j) = kv{2}(4);
-        akslow2_wt(i,j) = kv{3}(1);
-        ikslow2_wt(i,j) = kv{3}(2);
-        taua_kslow2_wt(i,j) = kv{3}(3);
         taui_kslow2_wt(i,j) = kv{3}(4);
-        akss_wt(i,j) = kv{4}(1);
[…]

### 2022-06-16 calibration/param_preproc.py
Preprocessing
Calibration parameters and their distribution estimations

@@ -48,11 +48,13 @@ idx_dict = {'iktof': [1, 2, 4, 5, 7, 11, 13],
     'ikslow2': [2, 3],
     'ikss': [1, 2, 3, 4]}
 
-base_dir = Path.cwd().parent
-wt_dir = base_dir / ('calib_exp'+exp_num+'_wt')
-ko_dir = base_dir / ('calib_exp'+exp_num+'_ko')
-fpaths_wt = wt_dir.glob('*.xlsx')
-fpaths_ko = ko_dir.glob('*.xlsx')
+file_names = pd.read_excel('file_names_4half4.xlsx', index_col=None)
+
+wt_dir = Path.cwd() / ('calib_exp'+exp_num) / 'wt'
+ko_dir = Path.cwd() / ('calib_exp'+exp_num) / 'mgat1ko'
+
+fpaths_wt = [Path(wt_dir/(p+'.xlsx')) for p in file_names['WT'].tolist()]
+fpaths_ko = [Path(ko_dir/(p+'.xlsx')) for p in file_names['MGAT1KO'].tolist()]
 
 plist_wt = read_param_file(fpaths_wt, currents, idx_dict)
 plist_ko = read_param_file(fpaths_ko, currents, idx_dict)

### 2022-06-16 calibration/param_preproc.py
Code streamlined

@@ -1,7 +1,6 @@
 # -*- coding: utf-8 -*-
 from pathlib import Path
 import pandas as pd
-import seaborn as sns
 
 def read_param_file(paths, currents, idx_dict):
     plist = []
@@ -31,14 +30,6 @@ def melt_param(pdf_wt, pdf_ko):
     pdf_ko['Group'] = 'Mgat1KO'
     pdf = pd.concat([pdf_wt, pdf_ko], ignore_index=True)
     return pdf.melt(id_vars=['Group','file'], var_name='param')
-
-def compare_param_dist(pdf_wt, pdf_ko, idx):
-    pdf_wt['Group'] = 'WT'
-    pdf_ko['Group'] = 'Mgat1KO'
-    pdf = pd.concat([pdf_wt, pdf_ko], ignore_index=True)
-
-    p = sns.displot(pdf, x=str(idx), hue='Group', kind='kde', fill=True)
-    p.set_xlabels('Parameter '+str(idx), fontsize=12)
     
 exp_num =  '45'
 currents = ['iktof','ikslow1','ikslow2','ikss']
@@ -50,11 +41,11 @@ idx_dict = {'iktof': [1, 2, 4, 5, 7, 11, 13],
 
 file_names = pd.read_excel('file_names_4half4.xlsx', index_col=None)
 
-wt_dir = Path.cwd() / ('calib_exp'+exp_num) / 'wt'
-ko_dir = Path.cwd() / ('calib_exp'+exp_num) / 'mgat1ko'
+wt_dir = Path.cwd()/('calib_exp'+exp_num)/'wt'
+ko_dir = Path.cwd()/('calib_exp'+exp_num)/'mgat1ko'
 
-fpaths_wt = [Path(wt_dir/(p+'.xlsx')) for p in file_names['WT'].tolist()]
-fpaths_ko = [Path(ko_dir/(p+'.xlsx')) for p in file_names['MGAT1KO'].tolist()]
+fpaths_wt = [Path(wt_dir/(p+'.xlsx')) for p in file_names['WT'].tolist() if pd.isnull(p) == False]
+fpaths_ko = [Path(ko_dir/(p+'.xlsx')) for p in file_names['MGAT1KO'].tolist() if pd.isnull(p) == False]
 
 plist_wt = read_param_file(fpaths_wt, currents, idx_dict)
[…]

### 2022-06-16 calibration/result_analysis.m
Result analysis
Parameter distributions and clustering

@@ -397,6 +397,193 @@ ylabel("\tau_{a}^{(4)}")
 set(gca,'FontWeight','bold','LineWidth',1.5)
 
 %% parameter distribution
+clc
+clearvars
+close all
+load('file_names_4half4.mat')
+
+exp_num = "exp45";
+base_dir = fullfile(pwd,strcat("calib_",exp_num));
+
+% pktof
+pktof = readtable(fullfile(base_dir,"pktof.csv"));
+pidx = unique(pktof.param);
+
+fig = figure('Color','w','Position',[50,50,830,320]);
+orient(fig,'landscape')
+for i=1:length(pidx)
+    psub = pktof(pktof.param==pidx(i),:);
+    psub_wt = psub(string(psub.Group)=="WT",:);
+    psub_ko = psub(string(psub.Group)=="Mgat1KO",:);
+
+    [f1,xi1] = ksdensity(psub_wt.value);
+    [f2,xi2] = ksdensity(psub_ko.value);
+    
+    subplot(2,4,i)
+    plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+    hold on
+    plot(xi2,f2, 'Color','red', 'LineWidth',2)
+    hold off
+    axis tight
+    grid on
+    xlabel(strcat('Parameter', num2str(pidx(i))))
+    ylabel('Density')
+    set(gca, 'FontName','Arial','FontWeight','bold')
+
+    if i==length(pidx)
+        legend(["WT","Mgat1KO"])
+        legend box off
+        xlabel("G_{Kto}")
+    end
+end
+nrow = size(pktof,1)/length(pidx);
+pktof2 = array2table(NaN(nrow,length(pidx)+2));
+
+pktof2.Properties.VariableNames(1:2) = {'Group','File'};
+pktof2.Group = pktof(pktof.param==pidx(1),:).Group;
+pktof2.File = pktof(pktof.param==pidx(1),:).file;
+
+for i=1:length(pidx)
+    pktof2.Properties.VariableNames(i+2) = {strcat('P',num2str(pidx(i)))};
+    pktof2(:,i+2)= pktof(pktof.param==pidx(i),'value');
+end
+
+% pkslow1
+pkslow1 = readtable(fullfile(base_dir,"pkslow1.csv"));
+pidx = unique(pkslow1.param);
+
+fig = figure('Color','w','Position',[100,100,830,640]);
[…]

### 2022-07-20 calibration/result_analysis.m
Parameter distribution
Use histogram

@@ -426,7 +426,7 @@ legend(["WT","MGAT1KO"],'Location','northeast')
 %% parameter distribution
 clc
 clearvars
-% close all
+close all
 load('file_names_4half4.mat')
 
 exp_num = "exp45";
@@ -443,13 +443,15 @@ for i=1:length(pidx)
     psub_wt = psub(string(psub.Group)=="WT",:);
     psub_ko = psub(string(psub.Group)=="Mgat1KO",:);
 
-    [f1,xi1] = ksdensity(psub_wt.value);
-    [f2,xi2] = ksdensity(psub_ko.value);
+%     [f1,xi1] = ksdensity(psub_wt.value);
+%     [f2,xi2] = ksdensity(psub_ko.value);
     
     subplot(2,4,i)
-    plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+%     plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+    histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5)
     hold on
-    plot(xi2,f2, 'Color','red', 'LineWidth',2)
+%     plot(xi2,f2, 'Color','red', 'LineWidth',2)
+    histogram(psub_ko.value,15,'FaceColor','r','FaceAlpha',0.5)
     hold off
     axis tight
     grid on
@@ -486,18 +488,21 @@ for i=1:length(pidx)
     psub_wt = psub(string(psub.Group)=="WT",:);
     psub_ko = psub(string(psub.Group)=="Mgat1KO",:);
 
-    [f1,xi1] = ksdensity(psub_wt.value);
-    [f2,xi2] = ksdensity(psub_ko.value);
+%     [f1,xi1] = ksdensity(psub_wt.value);
+%     [f2,xi2] = ksdensity(psub_ko.value);
     
     subplot(4,4,i)
-    plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+%     plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+    histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5)
     hold on
-    plot(xi2,f2, 'Color','red', 'LineWidth',2)
+%     plot(xi2,f2, 'Color','red', 'LineWidth',2)
+    histogram(psub_ko.value,15,'FaceColor','r','FaceAlpha',0.5)
     hold off
     axis tight
     grid on
     xlabel(strcat('Parameter', num2str(pidx(i))))
[…]

### 2022-07-21 KvApp_guide.Rmd
KvApp Guide

--- 2022-04-01 KvApp_guide.Rmd
@@
 ---
-title: "User Guide for KvApp"
-author: "Haedong Kim"
+title: "A Novel Tool for Analyzing Whole-cell Potassium Current Recordings in Mouse Cardiomyocytes"
+author: "Haedong Kim and Hui Yang"
 date: "`r Sys.Date()`"
@@
 
+## Introduction
+
+This manual describes the installation and use of a novel tool for analyzing whole-cell K^+^ current traces (I~K~).
+Cardiac voltage-gated K^+^ channels (K~v~) are responsible for repolarizing the action potential propagating thorugh cardiomyocytes.
+K~v~ are indispensable in the electrical conduction system of the heart, and even modest changes lead to fatal heart failures and diseases, such as long QT syndrome.
+Whole-cell patch-clamp recordings of K^+^ currents have been critical methods for studying the effects of disease-related alteration in activities of K~v~.
+These *in-vitro* experiments alone are not enough to determine the component currents of I~K~ recordings generated from various K~v~ isofroms and underlying gating mechanism.
+Typically, curve fitting is used to analyze I~K~ traces.
+However, this traditional approach is limited in its ability to estimate channel kinetics rigorously in two aspects.
+1) It applies multiple curve fittings to the channel-gating processes that work together.
+2) This approach is only applicable to a single I~K~ recording at each time.
+Hence, methods based on curve fitting do not consider complex relationships of channel-gating processes and cellular-level dynamics of multiple responses (i.e., I~K~ recordings) measured in a cell.
+This software implements a novel concurrent assimilation method of multi-response data for modeling and calibrating K~v~ kinetics in mouse cardiomyocytes.
+
 ## Getting Started
 
-There are four associated files as shown in Figure 1 below: `KvAppBeta`, `MyAppInstaller_mcr`, `MyAppInstaller_web`, and `splash`.
-
-![Consisting files of the applications.](KvApp_guide_img/folder_struct.png){width=50%}
-
-`KvAppBeta` is the file that runs the application, but you will get the error message in Figure 2 if the correct version of MATLAB is not installed on the machine.
-
-![Error message when the right version of MATLAB is not installed on the target machine.](KvApp_guide_img/runtime_error.png){width=50%}
-
-`MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, which enables the application to run without having MATLAB installed.
-Your computer needs to be connected to the Internet to complete the installation.
+`MyAppInstaller_web` will install the packaged application and MATLAB Runtime from the Internet, *which enables the application to run without having MATLAB installed.* Your computer needs to be connected to the Internet to complete the installation.
 If the Internet connection is not available, `MyAppInstaller_mcr` will install the necessary programs locally.
-A few screenshots below in Figure 3 show the installation process.
-The other file `splash` is just an image file for the application icon.
+A few screenshots below in Figure 1 show the installation process.
+`KvAppBeta` is the execution file that runs the application.
+Another file `splash` is just an image file for the application icon.
 
@@
 
-## Modules and Functions
+## Importing Experimental Data
 
-KvApp consists of three modules, as shown in Figure 4: Experimental Protocol & Data, Computer Model, and Model Outputs & Kinetics Modeling Results.
-
-![Startup window of KvApp.](KvApp_guide_img/window.png){width=50%}
-
+The application consists of three modules, as shown in Figure 2: Experimental Protocol & Data, Computer Model, and Model Outputs & Kinetics Modeling Results.
 The data to be investigated *in silico* should be stored in an Excel file, with the first column corresponding to time and the other columns normalized currents for voltage steps in increasing order.
 This is a default format when I~Ksum~ traces are transferred to a tabular format in ClampFit.
-Figure 5 is an example of I~Ksum~ recordings.
+Figure 3 is an example of I~Ksum~ recordings.
 To import a dataset, we need to specify the experimental protocol of the voltage-clamp method used to generate that dataset.
@@
 
-![Example of raw I~Ksum~ data in ClampFit.](KvApp_guide_img/pClamp.png)
+![Startup window of the application](KvApp_guide_img/window.png)
+
+![Example of raw I~K~ data in ClampFit.](KvApp_guide_img/pClamp.png)
+
+## Model Calibration and Results
+
+![Kinetics modeling results.](KvApp_guide_img/kvoutput.png)
 
@@
 The Computer Model module shows equations and kinetics parameters of computers models of potassium isoforms, separated and grouped by tabs.
-The Kv4.2 model in the first tab is the only stand-alone model in which every kinetic variables and parameter are independent of other models.
+The K~v~4.2 model in the first tab is the only stand-alone model in which every kinetic variables and parameter are independent of other models.
 The second tab contains three closely connected models.
@@
 These parameters in red can be calibrated automatically using the Turn Calibration button in the Model Output tab of the third module.
-This function calibrates the models by minimizing discrepancies between model prediction and experimental I~Ksum~ traces.
+This function calibrates the models by minimizing discrepancies between model prediction and experimental I~K~ traces.
 The fitness of calibration results is visualized, and the drop-down button allows to select different voltage steps.
@@
 It is desirable to perform model calibration separately, utilizing more computing power and time for better results.
-Other tabs named after the K~v~ isoforms show kinetic variables such as steady states, time constants, and transition rates as shown on Figure 6.
+Other tabs named after the K~v~ isoforms show kinetic variables such as steady states, time constants, and transition rates as shown on Figure 4.
 
-![Kinetics modeling results.](KvApp_guide_img/kvoutput.png)
+\newpage
 
-## Examples
+## Example
 
-### Example 1: 4.5-sec Data
+This section provides an example of using the application.
+A data set will be used having voltage steps from -30 to 50 mV lasting 4.5 s from holding voltage -70 mV.
+The data is saved in the provided file, `15o27010.xlsx`.
+First update the protocol accordingly and upload the data as in Figure 5.
+Change the protocol of the smallest clamp voltage from -50 to -30 mV, and exclude I~Kslow2~.
+Discrepancies between the model prediction and experimental I~K~ traces can be observed in the bottom right figure.
+To calibrate the models according to the data, click the Run Calibration button.
+It will change to a progress bar as in Figure 6.
+Once the calibration is completed, the model outputs and kinetics are updated as in Figure 7.
 
-example with 9 voltage steps rather than the full data.
+![Importing the experimental data saved in `15o27010.xlsx`.](KvApp_guide_img/exp1-1.png)
 
-### Example 2: 25-sec Data
+![Progress bar showing the calibration progress.](KvApp_guide_img/ex1-2.png)
 
-example with 25-sec data
+![Updated model outputs with the calibrated parameters.](KvApp_guide_img/ex1-3.png)

### 2022-07-22 sensitivity_analysis/sensitivity_analysis.m
Consolidated half-normal plots

@@ -3,39 +3,27 @@ clearvars
 close all
 
 % import design matrices
-dgn_kto = table2array(readtable('ikto_dgn2.csv'));
+dgn_kto = table2array(readtable('ikto_dgn.csv'));
 dgn_kslow1 = table2array(readtable('ikslow1_dgn.csv'));
 dgn_kslow2 = table2array(readtable('ikslow2_dgn.csv'));
 dgn_kss = table2array(readtable('ikss_dgn.csv'));
 
 % bounds
 kto_default = [33, 15.5, 20, 16, 8, 7, 0.03577, 0.06237, 0.18064, 0.3956, 0.000152, 0.067083, 0.00095, 0.051335, 0.2087704319, 0.14067, 0.387];
-kto_default2 = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, 0.000152, 0.067083, 0.00095, 0.051335, 0.4067];
 kslow1_default = [22.5, 45.2, 40.0, 7.7, 5.7, 6.1, 0.0629, 2.058, 803.0, 18.0, 0.9214774521, 0.05766, 0.07496];
-kslow1_default2 = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18.0, 2.058, 803.0, 0.16];
 kslow2_default = [5334, 4912, 0.05766];
 kss_default = [0.0862, 1235.5, 13.17, 0.0428];
 
-% lower bounds
-lb_kto = [-70, -70, eps, 1, eps, eps, eps, eps, eps, eps, eps, eps];
-lb_kto(7:14) = kto_default(5:12)*0.05;
-lb_kto2 = [-70, -70, eps, 1, eps, eps, eps, eps, eps, eps, eps, eps];
-lb_kto2(5:12) = kto_default2(5:12)*0.05;
-
+lb_kto = [-70, eps, eps, eps, eps, 1, eps, eps, eps, eps, eps, eps, eps, eps];
+lb_kto(7:14) = kto_default(7:14)*0.05;
 lb_kslow1 = [-70, -70, -70, 1, 1, eps, eps, eps, 50, eps];
-lb_kslow12 = [-70, -70, -70, 1, 1, eps, eps, eps, eps, 50+eps];
+
 lb_kslow2 = [5000, eps];
 lb_kss = [eps, eps, eps];
 
-% upper bounds
[…]

### 2022-07-22 calibration/result_analysis.m
Bin edges

@@ -448,21 +448,23 @@ for i=1:length(pidx)
     
     subplot(2,4,i)
 %     plot(xi1,f1, 'Color','blue', 'LineWidth',2)
-    histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5)
+    h1 = histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5);
+    bin_edges1 = h1.BinEdges;
     hold on
 %     plot(xi2,f2, 'Color','red', 'LineWidth',2)
-    histogram(psub_ko.value,15,'FaceColor','r','FaceAlpha',0.5)
+    h2 = histogram(psub_ko.value,'BinEdges',bin_edges1,'FaceColor','r','FaceAlpha',0.5);
     hold off
     axis tight
     grid on
-    xlabel(strcat('Parameter', num2str(pidx(i))))
-    ylabel('Density')
-    set(gca, 'FontName','Arial','FontWeight','bold')
+    xlabel(strcat("p_{", num2str(pidx(i)), "} Range"))
+%     ylabel('Density')
+    ylabel("Frequency")
+    set(gca, 'FontName','Arial','FontWeight','bold','LineWidth',1.5)
 
     if i==length(pidx)
         legend(["WT","Mgat1KO"])
         legend box off
-        xlabel("G_{Kto}")
+        xlabel("G_{Kto} Range")
     end
 end
 nrow = size(pktof,1)/length(pidx);
@@ -493,20 +495,21 @@ for i=1:length(pidx)
     
     subplot(4,4,i)
 %     plot(xi1,f1, 'Color','blue', 'LineWidth',2)
-    histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5)
+    h1 = histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5);
+    bin_edges1 = h1.BinEdges;
     hold on
 %     plot(xi2,f2, 'Color','red', 'LineWidth',2)
-    histogram(psub_ko.value,15,'FaceColor','r','FaceAlpha',0.5)
+    histogram(psub_ko.value,'BinEdges',bin_edges1,'FaceColor','r','FaceAlpha',0.5)
     hold off
     axis tight
     grid on
-    xlabel(strcat('Parameter', num2str(pidx(i))))
+    xlabel(strcat("p_{", num2str(pidx(i)), "} Range"))
 %     ylabel('Density')
     ylabel("Frequency")
[…]

### 2022-07-23 sensitivity_analysis/sensitivity_analysis.m
Parameter histograms

@@ -773,13 +773,14 @@ for i=1:length(theta)
 end
 
 % kss
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,1),'range'),dgn_kslow2);
+[theta,idx,phiinv] = doe_plot(normalize(response_kss(:,1),'range'),dgn_kss);
 plot(phiinv,theta,'+','MarkerSize',10,'Color',c(4,:),'LineWidth',1.5)
 for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','bottom');
 end
 hold off
-legend(["I_{Kto}","I_{Kslow1}","I_{Kslow2}","I_{Kss}"])
+legend(["I_{Kto}","I_{Kslow1}","I_{Kslow2}","I_{Kss}"],'Location','northwest')
+title("Anchor Point 1")
 xlabel('Half-normal Quantiles')
 ylabel('Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -809,13 +810,13 @@ for i=1:length(theta)
 end
 
 % kss
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,2),'range'),dgn_kslow2);
+[theta,idx,phiinv] = doe_plot(normalize(response_kss(:,2),'range'),dgn_kss);
 plot(phiinv,theta,'+','MarkerSize',10,'Color',c(4,:),'LineWidth',1.5)
 for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','bottom');
 end
 hold off
-legend(["I_{Kto}","I_{Kslow1}","I_{Kslow2}","I_{Kss}"])
+title("Anchor Point 2")
 xlabel('Half-normal Quantiles')
 ylabel('Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -845,13 +846,13 @@ for i=1:length(theta)
 end
 
 % kss
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,3),'range'),dgn_kslow2);
+[theta,idx,phiinv] = doe_plot(normalize(response_kss(:,3),'range'),dgn_kss);
 plot(phiinv,theta,'+','MarkerSize',10,'Color',c(4,:),'LineWidth',1.5)
 for i=1:length(theta)
[…]

### 2022-07-23 sensitivity_analysis/sensitivity_analysis.m
Plot specifications

@@ -745,224 +745,225 @@ dgn_kss = table2array(readtable('ikss_dgn.csv'));
 
 c = parula(5);
 
-fig = figure('Color','w','Position',[100,100,900,1200]);
+fig = figure('Color','w','Position',[100,100,500,680]);
 orient(fig,'landscape')
 
 % biomarker 1
 % kto
 subplot(3,2,1)
 [theta,idx,phiinv] = doe_plot(normalize(response_kto(:,1),'range'),dgn_kto);
-plot(phiinv,theta,'+','MarkerSize',10,'Color',c(1,:),'LineWidth',1.5)
+plot(phiinv,theta,'+','MarkerSize',8,'Color',c(1,:),'LineWidth',1.5)
 for i=1:length(theta)
-    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','bottom');
+    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 
 hold on
 % kslow1
 [theta,idx,phiinv] = doe_plot(normalize(response_kslow1(:,1),'range'),dgn_kslow1);
-plot(phiinv,theta,'+','MarkerSize',10,'Color',c(2,:),'LineWidth',1.5)
+plot(phiinv,theta,'x','MarkerSize',8,'Color',c(2,:),'LineWidth',1.5)
 for i=1:length(theta)
-    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','bottom');
+    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 
 % kslow2
 [theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,1),'range'),dgn_kslow2);
-plot(phiinv,theta,'+','MarkerSize',10,'Color',c(3,:),'LineWidth',1.5)
+plot(phiinv,theta,'*','MarkerSize',8,'Color',c(3,:),'LineWidth',1.5)
 for i=1:length(theta)
-    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','bottom');
+    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 
 % kss
[…]

### 2022-08-08 sensitivity_analysis/sensitivity_analysis.m
Appearance changes

@@ -805,7 +805,7 @@ for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 hold off
-title("Anchor Point 1")
+title("Marker a")
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -841,7 +841,7 @@ for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 hold off
-title("Anchor Point 2")
+title("Marker b")
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -877,7 +877,9 @@ for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 hold off
-title("Anchor Point 3")
+legend(["I_{Kto}","I_{Kslow1}","I_{Kslow2}","I_{Kss}"],...
+    'Location','northeast')
+title("Marker c")
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -913,7 +915,7 @@ for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 hold off
-title("Anchor Point 4")
+title("Marker d")
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -949,7 +951,7 @@ for i=1:length(theta)
     text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 hold off
-title("Anchor Point 5")
+title("Marker e")
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -985,9 +987,7 @@ for i=1:length(theta)
[…]

### 2022-08-09 sensitivity_analysis/sensitivity_analysis.m
Debug

@@ -792,10 +792,11 @@ for i=1:length(theta)
 end
 
 % kslow2
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,1),'range'),dgn_kslow2);
+idx_kslow2 = [2,1];
+[theta,~,phiinv] = doe_plot(normalize(response_kslow2(:,1),'range'),dgn_kslow2);
 plot(phiinv,theta,'*','MarkerSize',8,'Color',c(3,:),'LineWidth',1.5)
 for i=1:length(theta)
-    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
+    text(phiinv(i),theta(i),int2str(i),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 
 % kss
@@ -806,6 +807,8 @@ for i=1:length(theta)
 end
 hold off
 title("Marker a")
+legend(["I_{Kto}","I_{Kslow1}","I_{Kslow2}","I_{Kss}"],...
+    'Location','northeast')
 xlabel('Half-normal Quantiles')
 ylabel('Normalized Absolute Effects')
 set(gca,'FontWeight','bold','LineWidth',1.5)
@@ -828,10 +831,10 @@ for i=1:length(theta)
 end
 
 % kslow2
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,2),'range'),dgn_kslow2);
+[theta,~,phiinv] = doe_plot(normalize(response_kslow2(:,2),'range'),dgn_kslow2);
 plot(phiinv,theta,'*','MarkerSize',8,'Color',c(3,:),'LineWidth',1.5)
 for i=1:length(theta)
-    text(phiinv(i),theta(i),int2str(idx(i)),'HorizontalAlignment','right','VerticalAlignment','top');
+    text(phiinv(i),theta(i),int2str(i),'HorizontalAlignment','right','VerticalAlignment','top');
 end
 
 % kss
@@ -864,10 +867,10 @@ for i=1:length(theta)
 end
 
 % kslow2
-[theta,idx,phiinv] = doe_plot(normalize(response_kslow2(:,3),'range'),dgn_kslow2);
+[theta,~,phiinv] = doe_plot(normalize(response_kslow2(:,3),'range'),dgn_kslow2);
 plot(phiinv,theta,'*','MarkerSize',8,'Color',c(3,:),'LineWidth',1.5)
 for i=1:length(theta)
[…]

### 2022-08-10 writing/draft.tex
Draft update

@@
 	{
-		\title{\bf Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes}
+		\title{\bf Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
 		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
@@
         $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
-        Wright State University, City, OH USA}
+        Wright State University, Dayton, OH USA}
 		\date{}
@@
 		\bigskip
-		\bigskip
+		\bigs   kip
 		\begin{center}
@@
 \begin{abstract}
-Aberrant activities of voltage-gated potassium (K\textsubscript{v}) channels can cause fatal heart disease such as long QT syndrome.
-In our recent investigation on a disease-related perturbation, reduced glycosylation in cardiomyocytes, significant reduction in K\textsuperscript{+} currents was observed.
+Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart, and even slight changes in function can cause fatal heart disease such as long QT syndrome.
+Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death.
+Activities of K\textsubscript{v} channels were also affected, and K\textsuperscript{+} currents were significantly reduced.
 Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
 On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
-Because only the sum of these K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded in experiments, data analytical methods are necessary to estimate the properties of individual K\textsuperscript{+} currents and their effects on the behaviors of the cardiomyocyte.
-However, traditional approaches estimate only one property, the shape of individual K\textsuperscript{+} currents, from a single I\textsubscript{Ksum} recording rather than analyzing I\textsubscript{Ksum} recordings from the same cell with different conditions.
-These approaches assume the functional form of the shape of K\textsuperscript{+} currents and search the shape parameters in the current functions that the sum of estimated currents has the best fit with a single I\textsubscript{Ksum} recording.
-It does not provide information about underlying kinetics of the currents at the cellular level.
+This characteristic raises challenges for studying K\textsubscript{v} channels.
+Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
+In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured.
+Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}.
+Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents.
+However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously.
+Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols.
+Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces.
 First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
 Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
-Third, we design model calibration as nonlinear optimization that couple the \textit{in-silico} model and \textit{in-vitro} I\textsubscript{Ksum} recordings by minimizing discrepancies between them.
-Further, we develop a graphical user interface (GUI) application to make the proposed framework accessible for researchers and patricians without engineering backgrounds.
-We apply the proposed framework to our data, and the experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
-This method has strong potential to pave a new way to analyze K\textsuperscript{+} currents.
+Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings.
+We apply the proposed framework to our  glycosylation data.
+Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
+This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} Data assimilation, potassium channels, kinetics modeling, model calibration, glycosylation, cardiomyopathy,
+{\it Keywords:} Data assimilation, potassium channels, kinetics modeling, model calibration, glycosylation
 
@@
 
-\section{Introduction} \label{s:intro}
+\section{Introduction}
 Cardiac voltage-gated potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential (AP) in cardiomyocytes.
@@
 Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics}.
+A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics,nagai2016aberrant}.
 Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
@@
 However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
-In contrast, there are predominant isoforms for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
-Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
+In contrast, there is a predominant isoform for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
+Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}.
 Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments.
-Typically, curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
+Typically, a form of curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
 Then the shape parameters are used to calibrate computer models \citep{kim2022simulation}, or kinetics information is added to the data assimilation process with additional curve fitting if applicable \citep{du2017silico}.
-However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of the K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
+However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
 These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
@@
 
-This paper presents a novel approach to data assimilation for analyzing whole-cell I\textsubscript{K} recordings and modeling the underlying kinetics of K\textsubscript{v} isoforms.
+This paper presents a novel approach to data assimilation for analyzing whole-cell K\textsubscript{v} current traces (I\textsubscript{K}) and modeling the underlying kinetics of K\textsubscript{v} isoforms.
 A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}.
@@
 Third, we developed a model calibration that minimizes discrepancy between model outputs and experimental data generated from multiple protocols.
-Further, we packaged the suggested framework into a graphical user interface (GUI) application to make the proposed framework more accessible.
+Last, we made estimations and inferences of impacts of disease-related perturbations on potassium channel kinetics and used low-dimensional embedding to show cellular variability.
 Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
@@
     \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
+    \item cell variability.
     \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
@@
 
-\section{Methods} \label{s:methods}
+\section{Research Background}
+\subsection{Computational Modeling of Cardiomyocytes}
+computational electrophysiology (pain points in analyziyng or modeling the data with focus on Kv). add citation \citep{chen2019heterogeneous}
 
-\subsection{\emph{Computer Models of Potassium Channel Isoforms}} \label{s:methods.1}
-Computer models mouse ventricular cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP.
+Computational approaches for studying heart (cite papers of Bing or Rui).
+Narrow down the topics into molecular-level studies: action potential and iom channels.
+Computational electrophyisiology has been a successful story of system biology.
+It has provided new insights how excitable cells work (neurons, cardiomyocytes, etc.).
+It can be also used to test new drugs.
+Other many application are also available.
+It is one of the vital part of disease-related investigation.
+
+Example: studies of the effects of altered glycosylation in the heart and pain points (elaborate the pain points mentioned in the earlier paragraph)
+
+calibration is necessary to apply this computational approaches (or models). How the calibration has been done in the literature.
+
+Point out problems of curve fitting
+
+\subsection{Altered Glycosylation and Heart Diseases}
+DCM and altered glycosylation
+
+\section{Methods}
+\subsection{Computer Models of Potassium Channel Isoforms}
+Mouse models have been very useful methods for studying human cardiac electric signaling \citep{nerbonne2004studying,milani2014small,kim2022irx5}.
+Computer models of mouse cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP.
 In this paper, the AP is modeled as a differential equation of transmembrane currents and stimulus current I\textsubscript{stim} as below in Equation~\ref{eq:ap} where $C_{m}$ is the membrane capacitance, $t$ is time.
@@
 There are 15 transmembrane currents in this AP model: the L-type calcium current (I\textsubscript{CaL}), the calcium pump current (I\textsubscript{p(Ca)}), the Na\textsuperscript{+}/Ca\textsuperscript{2+} exchange current (I\textsubscript{NaCa}), the calcium background current (I\textsubscript{Cab}), the fast Na\textsuperscript{+} current (I\textsubscript{Na}), the background Na\textsuperscript{+} current (I\textsubscript{Nab}), the Na\textsuperscript{+}/K\textsuperscript{+} pump current (I\textsubscript{NaK}), the Ca\textsuperscript{2+}-activated Cl\textsuperscript{-} current (I\textsubscript{Cl,Ca}), the rapidly inactivating transient outward K\textsuperscript{+} current (I\textsubscript{Kto}), the two ultra-rapidly activating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}), the slow delayed rectifier K\textsuperscript{+} current (I\textsubscript{Ks}), the rapid delayed rectifier K\textsuperscript{+} current (I\textsubscript{Kr}), and the time independent K\textsuperscript{+} current (I\textsubscript{K1}).
+Detailed implementation of mouse AP and VGIC models can be found in \cite{bondarenko2014compartmentalized,asfaw2020compartmentalized}.
 
@@
 Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}).
-\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{Ksum} given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
+\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{K} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
 \textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of protocols and consequent change in I\textsubscript{Ksum}.
@@
 (B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
-(C) Range of protocols from -50 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
+(C) Range of protocols from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
     \label{fig:kcurrent_example}
@@
 Unlike the simple current functions, computer models of K\textsubscript{v} isoforms can simulate detailed gating kinetics.
-We developed them with Hodgkin-Huxley modeling that has two gating variables controlling conductance.
+We developed them with Hodgkin-Huxley modeling, which has two gating variables controlling conductance, a popular modeling scheme in various species \citep{ten2004model,mahajan2008rabbit}.
 For example, K\textsubscript{v} isoforms that conducts current I\textsubscript{K} can be defined by Equation~\ref{eq:hh_ex}
@@
 Their kinetics are governed by transition-rate functions from Equation~\ref{eq:ikto_alpha1} to Equation~\ref{eq:ikto_beta2}.
-Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like ``knobs'', allowing to control the behavior of the I\textsubscript{Kto} model.
+Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
 
-\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})}
+\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})} \label{s:method.model2}
 There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are very rapidly activating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
@@
     &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(1)}} \\
-    &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \\
-    &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \\
-    &\tau_{a}^{(1)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
-    &\tau_{i}^{(1)} = p_{10} - p_{8}i_{ss}
+    &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \label{eq:ass} \\
+    &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \label{eq:iss} \\
+    &\tau_{a}^{(2)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
+    &\tau_{i}^{(2)} = p_{10} - p_{8}i_{ss}
 \end{align}
@@
     &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(2)}} \\
-    &\tau_{i}^{(2)} = p_{2} - p_{1}i_{ss}
+    &\tau_{i}^{(3)} = p_{2} - p_{1}i_{ss}
 \end{align}
@@
     &\frac{da_{\mathrm{Kss}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kss}}}{\tau_{a}^{(3)}} \\
-    &\tau_{a}^{(3)}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}
+    &\tau_{a}^{(4)}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}
 \end{align}
@@
 
-\subsection{\emph{Data Assimilation and Model Calibration}} \label{s:methods.2}
+\subsection{Data Assimilation and Model Calibration}
 \textit{Data assimilation} is a method to find the optimal configuration and state of computational models by coupling them with experimental data.
@@
 
-We developed the box-constrained nonlinear optimization routine using the BFGS algorithm with multi-random initial points to minimize $\delta$.
+We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$.
 Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube.
@@
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
+Different nonlinear optimization algorithms are tested and compared.
 This work is implemented in MATLAB R2022a, and the code is available at xyz.
 
-\subsection{\emph{Sensitivity Analysis and Model Regularization}} \label{s:methods.3}
+\subsection{Model Regularization and Sensitivity Analysis}
 The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$.
-We performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
-As illustrated in \textbf{Figure~\ref{fig:anchor_points}}, six \textit{anchor points} were defined that capture representative characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
-Each point represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
-Feature f will be equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:anchor_points}C} and \textbf{~\ref{fig:anchor_points}D}.
+Excessive flexibility of a model has a risk of causing overfitting that occurs when the model fits data too closely, even including noise and random effects in the data.
+Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results of parameters.
+It is worth noting that the models presented in Section 3.1 are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simply the model structure.
+Specifically, the same parameters that control voltage thresholds, such as $p_{1}$, $p_{2}$, $p_{3}$, and $p_{4}$, are used in equations having a similar function.
+The steady-state functions (Equation~\ref{eq:ass} and \ref{eq:iss}) are reused in the three rectifier-type currents.
+
+We also performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
+In order to do that, factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output.
+As illustrated in \textbf{Figure~\ref{fig:current_markers}}, six markers were defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
+Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
+Marker f will equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_markers}C} and \textbf{~\ref{fig:current_markers}D}.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/anchor_ponts.pdf}
-    \caption{Illustration of the six anchor points of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
+    \includegraphics{figs/current_markers.pdf}
+    \caption{Illustration of the six markers of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
 All currents are simulated for illustration, and the labels refer to a) the current magnitude 10 ms after voltage is applied, b) 25\% of the total recording time has elapsed, c) 50\%, d) 75\%, e) the peak magnitude, and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
-    \label{fig:anchor_points}
+    \label{fig:current_markers}
 \end{figure}
 
-We developed factorial designs in which parameters varied at two levels, contrasting their effect.
-The anchor points were evaluated at each design to calculate factorial effects.
-Then, half-normal plots were drawn based on these contrasts, providing visual inspections of critical parameters that exert the most impact on each shape feature.
-All the half-normal plots can be found in Supplement A.
-\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
-We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters.
-Note that the voltage-dependence parameters in red and green appear multiple times across different equations.
-This parsimonious model design is intended to maximize the structural regularization to minimize overfitting.
+We adopted a fractional factorial design of 1024 runs for I\textsubscript{Kto}, which results in resolution VIII.
+The resolution ensures that the main effects and 2-/3-factor interactions are strongly clear.
+Full factorial designs are used for the other three currents.
+The marker points are evaluated at each design point and factorial effects are calculated using the linear model in Equation~\ref{eq:fact_effect} for each current.
+\begin{equation}
+    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\},
+    \label{eq:fact_effect}
+\end{equation}
+The least squares method is used to estimate $\beta_{i}^{k}$. Then, half-normal plots are drawn to test significance of the estimated effects.
+
+\subsection{Variability Visualization and Low-dimensional Embedding}
+Most cardiac electrophysiology studies utilize statistical tests for the mean to support hypotheses; in turn, the majority of ion channel models are based on population-averaged data.
+However, fine-grained analyses that consider the variability of cells can provide new insights into data.
+For example, it is possible there are differences between control and disease-altered groups but also within the groups according to characteristics of cells such as size or age.
+Hence, there is an urgent need to develop a tool to investigate cellular variability.
+Because the proposed approach calibrates models for the individual dataset from each cell, it allows us to quantify and visualize cellular variability in the tuned parameters.
+Histograms are used in this study to visualize distributions of calibration parameters and the effects of disease-related perturbations on the parameter distributions.
+Although a histogram is a simple yet powerful tool, one of the limitations of a histogram is that it cannot visualize multiple variables more than three on the same plot.
+It is desirable to inspect how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy control and disease groups.
+Low-dimensional embedding methods transform high-dimensional data into a plane or 3D space while preserving relative locations of data points as much as possible so that the structure of data can be visualized.
+\textbf{Figure~\ref{fig:embd}} illustrates how a low-dimensional embedding method is used to map the high-dimensional parameter space of data points from wild-type and disease groups into 3D space.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/calib_param.pdf}
-    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
-    \label{fig:calib_param}
+    \includegraphics{figs/embd.pdf}
+    \caption{Diagram illustrating the procedure of low-dimensional embedding to reveal cellular-level structure of the parameter space.}
+    \label{fig:embd}
 \end{figure}
 
-\subsection{\emph{Glycosylation and Dilated Cardiomyopathy (DCM)}} \label{s:methods.4}
-We applied the proposed framework to our studies on the pathophysiology of electrical signaling by altered glycosylation.
+For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \cite{van2008visualizing}.
+It has proven to be an effective method for visualizing high-dimensional data.
+t-SNE is a statistical method that constructs two sets of probability distributions $p_{ij}$ and $q_{ij}$ over pairs of data points $i$ and $j$ in a high-/low-dimensional space, respectively.
+These are probabilities of similarities of data points such that neighboring points have a higher probability while dissimilar points have a lower probability.
+We first define the conditional probability of $j$ given $i$:
+\begin{equation}
+    p_{j|i}=
+    \begin{cases}\frac{\mathrm{exp}(-d(x_{i},x_{j})^{2}/(2\sigma_{i}^{2}))}{\sum\limits_{k \neq i}\mathrm{exp}(-d(x_{i},x_{k})^{2}/(2\sigma_{i}^{2}))}, \ j \neq i, \\
+    0, \ j=i,
+    \end{cases}
+\end{equation}
+where $d(\cdot,\cdot)$ is a distance function such as Euclidean distance, and $\sum\limits_{j} p_{j|i}=1$ for all $i$.
+Then $p_{ij}$ can be defined by the symmetric property of the joint probabilities:
+\begin{equation}
+    p_{ij} = \frac{p_{j|i} + p_{i|j}}{2N},
+\end{equation}
+where $N$ is the number of total data points, and $\sum\limits_{i,j}p_{ij}=1$.
+It can be calculated from data once the standard deviation $\sigma_{i}$ is given.
+$\sigma_{i}$ is set in a way that the perplexity of the conditional probability distribution over other data points given $x_{i}$ equals a prefixed value that is a hyperparameter of t-SNE.
+Let $P_{i}$ denote the conditional probability distribution, then the perplexity of the distribution is
+\begin{equation}
+    \mathrm{perplexity}(P_{i}) = 2^{H(P_{i})}
+\end{equation}
+where $H(P_{i})$ is the Shannon entropy of $P_{i}$ defined by
+\begin{equation}
+    H(P_{i}) = -\sum\limits_{j}p_{j|i}\log_{2}(p_{j|i}).
+\end{equation}
+
+Then the probability distribution $q_{ij}$ is defined by the similarity of data points $y_i$ and $y_j$ in a low-dimensional space:
+\begin{equation}
+    q_{ij} = 
+    \begin{cases}\frac{(1+||y_{i}-y_{j}||^{2})^{-1}}{\sum\limits_{k}\sum\limits_{l \neq k}(1+||y_{k}-y_{l}||^{2})^{-1}}, \ j \neq i, \\
+    0, \ j=i,
+    \end{cases}
+\end{equation}
+where $\sum\limits_{i,j}q_{ij}=1$.
+Note that $q_{ij}$ is modeled by a heavy-tailed Student's t-distribution with one degree of freedom, from which the name ``t-SNE'' originates.
+The objective of t-SNE is to learn $y$ that minimizes discrepancies between $P$ and $Q$, so that the low-dimensional distribution preserves the structure of $p_{ij}$ constructed from the original high-dimensional data.
+To learn $y$, t-SNE maps $y$ by minimizing the Kullback-Leibler (KL) divergence, which measures the similarity between two probability distributions:
+\begin{equation}
+    \mathrm{KL}(P||Q) = \sum\limits_{j}\sum\limits_{i \neq j}p{ij}\log \frac{p_{ij}}{q_{ij}}.
+\end{equation}
+Typically, a gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.
+
+\section{Experimental Design}
+We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation.
 Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
@@
 To further investigate the role of altered glycosylation and pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
-However, it was difficult to determine channel kinetics rigorously from the \textit{in-vitro} experiments alone, especially for K\textsuperscript{+} current recordings, due to vast diversity of K\textsubscript{v} isoforms and small current magnitude of the MGAT1KO model.
-Therefore, we validate the suggested framework by modeling K\textsubscript{v} isoforms with our \textit{in-vitro} experimental data.
-Therefore, we leverage the suggested framework to delineate the channel kinetics of K\textsubscript{v} to provide new descriptive/predictive insights into our \textit{in-vitro} experiments.
+To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
+There were 31 and 30 observations for each WT and MGAT1KO.
 
-\section{Results} \label{s:results}
+It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes.
+A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents.
+All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably.
+Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously.
+For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model.
+Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{K} recordings.
+This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics with considering the cellular-level dynamics of multiple protocols.
 
-\subsection{Calibration Results}
-first we compared fitting results.
+\section{Results}
+\subsection{Parameter Screening}
+\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six marker points.
+The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points.
+The farther parameters fall above the straight lines, the more significant impacts they have on the marker.
+We picked the parameters to be calibrated as a union of parameter sets above the straight lines on the six half-normal plots.
+For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
+\begin{figure}
+    \centering
+    \includegraphics{figs/half-norm.pdf}
+    \caption{Half-normal plots of factorial effects of parameters on the six maker points.}
+    \label{fig:half-norm}
+\end{figure}
 
-\subsection{Distributions Estimations}
+\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
+We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters.
+Note that the voltage-dependence parameters in red and green appear multiple times across different equations.
+This parsimonious model design is intended to maximize the structural regularization to minimize overfitting.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/calib_param.pdf}
+    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
+    \label{fig:calib_param}
+\end{figure}
 
-\subsection{Clustering}
-Perhaps numerical analysis of performance, comparison, and demonstration of applicability and impact using real data.
+\subsection{Model Fitness}
+We compared three nonlinear optimization algorithms satisfying box constraints: BFGS, sequential quadratic programming (SQP), and active set methods.
+\textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} shows values of the objective function (sum of RMSEs) of the three optimization algorithms for WT and MGAT1KO, respectively.
+The three algorithms barely showed differences in the objective values.
+We compared the actual model predictions and experimental potassium I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
+Representative data were selected randomly for each group and plotted at -30 mV and 50 mV voltage steps to illustrate discrepancies at low and high voltages.
+Sum of RMSEs values are lower in MGAT1KO than WT in general, because magnitudes are smaller for the disease group.
+Calibration results of the BFGS algorithm will be used in analysis of kinetics modeling in the rest the sections.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/model_fitness_4half4.pdf}
+    \caption{Bar graphs of the sum of RMSEs of (A) WT and (C) MGAT1KO, and representative actual fitness between model outputs and experimental I\textsubscript{K} recordings (in red) at -30 mV and 50 mV voltage steps of (B) observation 26 in WT and (D) 4 in MGAT1KO.}
+    \label{fig:model_fitness}
+\end{figure}
 
-\section{Discuss}
-Discuss of the results goes here.
+\subsection{Kinetics Modeling} \label{s:results.kinetics}
+The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior.
+\textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results of showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
+There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.2 at voltages lower 0 mV (\textbf{Figure~\ref{fig:kinetics_ikto}B}).
+Activation and inactivation time constants $\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$ to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
+$\tau_{a}^{(1)}$ for MGAT1KO is higher in range of $>$ -10 mV than WT.
+$\tau_{i}^{(1)}$ for MGAT1KO is slightly shifted toward left.
+The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increases time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/kinetics_ikto.pdf}
+    \caption{Kinetics functions in I\textsubscript{Kto} and their relationships to voltage.
+(A) steady-state activation, (B) steady-state inactivation, (C) time-constant activation, and (D) time-constant inactivation.}
+    \label{fig:kinetics_ikto}
+\end{figure}
 
-\section{Conclusion}\label{s:conclusion}
-A paper ends with a conclusion or summary section.
+\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}.
+The most prominent result is that the time constants of I\textsubscript{Kslow1} are estimated to be significantly slow in MGAT1KO $\sim$ 3.4 and 235.8 ms (\textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}).
+Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO.
+SSI is also shifted $\sim$ 10 mV to the right.
+Slow time constants and increased SSI are likely to address the flattened shape of MGAT1KO currents as illustrated in \textbf{Figure~\ref{fig:model_fitness}D}).
+Time constant inactivation of I\textsubscript{Kslow2} (\textbf{Figure~\ref{fig:kinetics_rec}C}) and activation of I\textsubscript{Kss} (\textbf{Figure~\ref{fig:kinetics_rec}F}) do not show significant differences between the two groups.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/kinetics_rec.pdf}
+    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (D) time-constant activation, (E) time-constant inactivation, (C) I\textsubscript{Kslow2} time-constant inactivation, and (F) I\textsubscript{Kss} time-constant activation.}
+    \label{fig:kinetics_rec}
+\end{figure}
+
+\subsection{Parameter Distributions Between WT and MGAT1KO} \label{s:results.distribution}
+We estimated distributions of the calibrated kinetic parameters to analyze the impacts of reduced N-glycosylation further.
+\textbf{Figure~\ref{fig:hist_pkto}} shows the distributional differences of the kinetic parameters in the K\textsubscript{4.2} model between the two group.
+A notable fact about the figure is that MGAT1KO shows small variances for Parameters 1, 5, and 7 in transition $\alpha_{a}$ and $\beta_{a}$, which control kinetics activation of I\textsubscript{Kto}.
+Reduced glycosylation affects activation gating that plays a vital role in I\textsubscript{Kto}, as it is characterized by a high peak at the early depolarization and rapid inactivation.
+In addition, MGAT1KO causes a leftward shift in the distribution of the maximum conductance $G_{\mathrm{Kto}}$, resulting in smaller conduction than WT in general.
+It is likely due to the observed current reductions in MGAT1KO.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/hist_pkto.pdf}
+    \caption{Distribution estimations of calibration parameters in the K\textsubscript{Kto} model.}
+    \label{fig:hist_pkto}
+\end{figure}
+
+\textbf{Figure~\ref{fig:hist_prectifiers}} shows distribution estimations of the calibration parameters in the rectifier and steady-state currents models.
+\textbf{Figure~\ref{fig:hist_prectifiers}A} shows significant distributional differences in Parameters 9 and 10, which control time constants of I\textsubscript{Kslow1}.
+It is compatible with the previous observations of significant differences in time constants to voltage relationships of I\textsubscript{Kslow1} between WT and MGAT1KO.
+There are slight shifts in Parameters 1 and 4 of I\textsubscript{Kslow1} that affect steady-state activation.
+The voltage-threshold parameter for MGAT1KO in steady-state inactivation has higher densities at low values around 30 than WT, which has the peak mode around 48.
+It accounts for the shift at low voltages less than 0 mV in SSI.
+The maximum conductance $G_{\mathrm{Kslow1}}$ of MGAT1KO is centered around relatively small values compared to WT.
+This leftward shift in the maximum conductance is also observed in I\textsubscript{Kslow2} (\textbf{Figure~\ref{fig:kinetics_rec}B}).
+Although there is a difference in variance of $G_{\mathrm{Kss}}$, the two groups have similar means around 0.03 (i.e.
+WT: $0.0375\pm0.0027$ vs MGAT1KO: $0.0325\pm0.015$).
+It is compatible with the observation that I\textsubscript{Kss} does not show significant reductions.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/hist_prectifiers.pdf}
+    \caption{Distribution estimations of calibration parameters in (A) the I\textsubscript{Kslow1} model, (B) the I\textsubscript{Kslow2} model, and (C) the I\textsubscript{Kss} model.}
+    \label{fig:hist_prectifiers}
+\end{figure}
+
+\subsection{Low-dimensional Embedding and Visualization}
+We aggregated calibration results into one to analyze the parameters in a collective way.
+It resulted in a $61\times20$ matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters).
+Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point.
+\textbf{Figure~\ref{fig:cluster}} presents the visualization of the t-SNE embedding of the calibration results.
+It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/cluster.pdf}
+    \caption{t-SNE embedding of calibration parameters.}
+    \label{fig:cluster}
+\end{figure}
+
+\section{Conclusions}\label{s:conclusions}
+We propose a novel data assimilation framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
+This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights.
+Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in protocols.
+The proposed method focuses on more comprehensive modeling of underlying channel kinetics of K\textsubscript{v} isoforms than traditional approaches.
+Experimental results show the strong potential of the proposed framework to discover channel kinetics from various angles.
+Predictions of kinetics variables to voltage relationships provide information about changes in the voltage dependence of channels.
+In addition, calibrated parameters of the computer models present a micro-level analysis of channel gating.
+The embedding method allows visualizing the variability of each cell.
+In summary, the proposed framework paves a new way for investigating the electrical signaling of potassium channels and heart diseases.
 

### 2022-08-23 calibration/result_analysis.m
Fitness figure

@@ -1,3 +1,186 @@
+% Bar graphs
+clc
+close all
+clearvars
+
+exp_num = "exp51";
+calib_dir = strcat("calib_",exp_num);
+
+pa_wt = readtable(fullfile(calib_dir,"r2l2_measures_wt.csv"));
+pa_ko = readtable(fullfile(calib_dir,"r2l2_measures_mgat1ko.csv"));
+
+files_wt = strings(length(pa_wt.file_name),1);
+files_ko = strings(length(pa_ko.file_name),1);
+
+for i=1:length(pa_wt.file_name)
+    [~,files_wt(i),~] = fileparts(pa_wt.file_name{i});
+end
+
+for i=1:length(pa_ko.file_name)
+    [~,files_ko(i),~] = fileparts(pa_ko.file_name{i});
+end
+
+r2_wt = pa_wt.mean_R2;
+l2_wt = pa_wt.mean_L2;
+r2_ko = pa_ko.mean_R2;
+l2_ko = pa_ko.mean_L2;
+
+% Without the worst fitting point in MGAT1KO
+[~,rmv_idx1] = mink(r2_wt,2);
+[~,rmv_idx2] = min(r2_ko);
+
+files_wt(rmv_idx1) = [];
+files_wt = categorical(files_wt);
+r2_wt(rmv_idx1) = [];
+files_ko(rmv_idx2) = [];
+files_ko = categorical(files_ko);
+r2_ko(rmv_idx2) = [];
+
+figure('Color','w','Position',[0,0,500,650])
+subplot(6,4,[1,5,9])
+barh(files_wt, r2_wt,'b');
+xlim([0,1])
+grid on
+title("WT")
+% xlabel("Nonlinear R^{2}")
+set(gca,'FontWeight','bold')
+
+subplot(6,4,[13,17,21])
+barh(files_ko,r2_ko,'r')
+xlim([0,1])
+grid on
+title("MGAT1KO")
+xlabel("Nonlinear R^{2}")
+set(gca,'FontWeight','bold')
+
+% Representative example of showing actual fitness
+volts = -30:10:50;
+[~,max_wt] = max(r2_wt);
+exp_data = table2array(...
+    readtable(fullfile("mgat1ko_data/wt-preprocessed",pa_wt.file_name{max_wt})));
+yksum = table2array(...
+    readtable(fullfile(calib_dir,"wt_yhat",pa_wt.file_name{max_wt})));
[…]

### 2022-08-23 calibration/result_analysis.m
Fitness figure

@@ -1,4 +1,4 @@
-% Bar graphs
+%----- Bar graphs
 clc
 close all
 clearvars
@@ -25,35 +25,36 @@ l2_wt = pa_wt.mean_L2;
 r2_ko = pa_ko.mean_R2;
 l2_ko = pa_ko.mean_L2;
 
-% Without the worst fitting point in MGAT1KO
+%----- Without the worst fitting point in MGAT1KO
 [~,rmv_idx1] = mink(r2_wt,2);
 [~,rmv_idx2] = min(r2_ko);
 
 files_wt(rmv_idx1) = [];
-files_wt = categorical(files_wt);
+% files_wt = categorical(files_wt);
 r2_wt(rmv_idx1) = [];
 files_ko(rmv_idx2) = [];
-files_ko = categorical(files_ko);
+% files_ko = categorical(files_ko);
 r2_ko(rmv_idx2) = [];
 
-figure('Color','w','Position',[0,0,500,650])
+fig = figure('Color','w','Position',[0,0,500,610]);
+orient(fig,'landscape')
 subplot(6,4,[1,5,9])
-barh(files_wt, r2_wt,'b');
+barh(categorical(1:length(files_wt)), r2_wt,'b');
 xlim([0,1])
 grid on
-title("WT")
+% title("WT")
 % xlabel("Nonlinear R^{2}")
 set(gca,'FontWeight','bold')
 
 subplot(6,4,[13,17,21])
-barh(files_ko,r2_ko,'r')
+barh(categorical(1:length(files_ko)),r2_ko,'r')
 xlim([0,1])
 grid on
-title("MGAT1KO")
-xlabel("Nonlinear R^{2}")
+% title("MGAT1KO")
+% xlabel("Nonlinear R^{2}")
 set(gca,'FontWeight','bold')
 
-% Representative example of showing actual fitness
+%----- Representative example of showing actual fitness
 volts = -30:10:50;
 [~,max_wt] = max(r2_wt);
 exp_data = table2array(...
@@ -69,7 +70,7 @@ for i=1:size(yksum,2)
     r2 = pa_row{1,2+(i-1)*2};
     l2 = pa_row{1,3+(i-1)*2};
 
-    if ismember(i,1:3)
+    if i==1
         subplot(6,4,1+i)
         plot(t,yi,'Color','magenta')
         hold on
@@ -78,10 +79,26 @@ for i=1:size(yksum,2)
         axis tight
         grid on
         title(strcat(...
[…]

### 2022-08-23 calibration/param_preproc.py
Minor edits

@@ -31,9 +31,9 @@ def melt_param(pdf_wt, pdf_ko):
     pdf = pd.concat([pdf_wt, pdf_ko], ignore_index=True)
     return pdf.melt(id_vars=['Group','file'], var_name='param')
     
-exp_num =  '45'
+exp_num =  '51'
 currents = ['iktof','ikslow1','ikslow2','ikss']
-idx_dict = {'iktof': [1, 2, 4, 5, 7, 11, 13],
+idx_dict = {'iktof': [1, 2, 3, 4, 5, 7, 11, 13],
     'iktos': [2, 3],
     'ikslow1': [1, 2, 4, 5, 9, 10, 11],
     'ikslow2': [2, 3],

### 2022-08-23 calibration/result_output.m
Renamed

@@ -627,9 +627,9 @@ legend(["WT","MGAT1KO"],'Location','northeast')
 clc
 clearvars
 close all
-load('file_names_4half4.mat')
+load('r2_file_names.mat')
 
-exp_num = "exp45";
+exp_num = "exp51";
 base_dir = fullfile(pwd,strcat("calib_",exp_num));
 
 % pktof
@@ -638,6 +638,8 @@ pidx = unique(pktof.param);
 
 fig = figure('Color','w','Position',[50,50,830,320]);
 orient(fig,'landscape')
+tl = tiledlayout(2,4);
+ylabel(tl,"Frequency",'FontWeight','bold')
 for i=1:length(pidx)
     psub = pktof(pktof.param==pidx(i),:);
     psub_wt = psub(string(psub.Group)=="WT",:);
@@ -646,8 +648,9 @@ for i=1:length(pidx)
 %     [f1,xi1] = ksdensity(psub_wt.value);
 %     [f2,xi2] = ksdensity(psub_ko.value);
     
-    subplot(2,4,i)
+%     subplot(2,4,i)
 %     plot(xi1,f1, 'Color','blue', 'LineWidth',2)
+    nexttile
     h1 = histogram(psub_wt.value,15,'FaceColor','b','FaceAlpha',0.5);
     bin_edges1 = h1.BinEdges;
     hold on
@@ -658,7 +661,7 @@ for i=1:length(pidx)
     grid on
     xlabel(strcat("p_{", num2str(pidx(i)), "} Range"))
 %     ylabel('Density')
-    ylabel("Frequency")
+%     ylabel("Frequency")
     set(gca, 'FontName','Arial','FontWeight','bold','LineWidth',1.5)
 
     if i==length(pidx)
@@ -705,7 +708,7 @@ for i=1:length(pidx)
     grid on
     xlabel(strcat("p_{", num2str(pidx(i)), "} Range"))
 %     ylabel('Density')
-    ylabel("Frequency")
+%     ylabel("Frequency")
     set(gca, 'FontName','Arial','FontWeight','bold','LineWidth',1.5)
     
     if i==length(pidx)
@@ -750,7 +753,7 @@ for i=1:length(pidx)
     grid on
     xlabel(strcat("p_{", num2str(pidx(i)), "} Range"))
 %     ylabel('Density')
-    ylabel("Frequency")
+%     ylabel("Frequency")
[…]

### 2022-08-24 writing/draft.tex
Draft revised

@@
 \begin{abstract}
-Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart, and even slight changes in function can cause fatal heart disease such as long QT syndrome.
+Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart.
+Even slight changes in K\textsubscript{v} functions can cause fatal heart disease, such as long QT syndrome.
 Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death.
-Activities of K\textsubscript{v} channels were also affected, and K\textsuperscript{+} currents were significantly reduced.
-Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of cardiac action potential (AP).
+Activities of K\textsubscript{v} channels were also affected by altered glycosylation, and K\textsuperscript{+} currents were significantly reduced.
+Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of the cardiac action potential (AP).
 On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
@@
 Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols.
-Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces.
+Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces and apply the proposed method to our glycosylation data.
 First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
@@
 Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings.
-We apply the proposed framework to our  glycosylation data.
+Last, low-dimensional embedding is used to visualize distributional difference of kinetic parameters of cells, showing the cellular-level variability.
 Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
@@
 \noindent%
-{\it Keywords:} Data assimilation, potassium channels, kinetics modeling, model calibration, glycosylation
+{\it Keywords:} Data assimilation, model calibration, potassium channels, kinetics modeling, glycosylation
 
@@
     \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
-    \item cell variability.
+    \item The proposed approach builds separate K\textsubscript{v} models and calibrates them for each cell, which allows analyzing the cellular variability. 
     \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
@@
 \section{Research Background}
-\subsection{Computational Modeling of Cardiomyocytes}
-computational electrophysiology (pain points in analyziyng or modeling the data with focus on Kv). add citation \citep{chen2019heterogeneous}
+\subsection{Computational Modeling of the Heart}
+Computer models of the heart have been extensively studied and used in cardiology to understand the mechanism of the organ, investigate fatal diseases, and develop drugs.
+Models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
+Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
+For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}.
+With the advancement of sensing technology and data analysis, computational cardiology is expanding its field.
+Personalized heart models pave a new way for monitoring and managing the cardiac health of individuals by characterizing individual heart signals from the large-scale IoT network and calculating dissimilarities to detect potential cardiac risk \citep{yang2020network}.
+Hence, cardiac computer models provide a great opportunity to mitigate heart disease, the number one cause of death in the U.S.
+\citep{murphy2021mortality}.
 
-Computational approaches for studying heart (cite papers of Bing or Rui).
-Narrow down the topics into molecular-level studies: action potential and iom channels.
-Computational electrophyisiology has been a successful story of system biology.
-It has provided new insights how excitable cells work (neurons, cardiomyocytes, etc.).
-It can be also used to test new drugs.
-Other many application are also available.
-It is one of the vital part of disease-related investigation.
+The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself.
+Hence, the electrical activities and waves spreading through the whole heart have been studied considerably.
+For examle, \citet{yang2016whole} a new reaction-diffusion and pattern recognition method to model spatiotemporal dynamics of electrical waves propagating throughout the heart surface.
+Sensing methods, such as vectorcardiogram (VCG) and electrocardiogram (ECG), measure the electrical activities of the heart.
+Various analytics methods have been proposed to cope with their complex nonlinear dynamics in different heart regions over time.
+\citet{chen2019heterogeneous} developed a heterogeneous recurrence method that characterizes spatiotemporal patterns in disease-altered multi-channel cardiac signals.
+Nonlinear recurrence analysis captures and quantifies topological relationships of recurring patterns in signals, which has been proven to be an effective tool for cardiac signals that are naturally recurring over time \citep{yang2010multiscale,meyers2020cross}.
 
-Example: studies of the effects of altered glycosylation in the heart and pain points (elaborate the pain points mentioned in the earlier paragraph)
+However, to perform a fine-grained analysis, it is necessary to study the cellular-level activities of the heart as opposed to the organ-level signals.
+Voltage-gated ion channels (VGICs) are membrane proteins in cardiomyocytes that are the functional unit of the electrical activities of cardiac muscle cells.
+The whole-cell patch clamp recording is a powerful experimental technique for investigating the electrical properties of excitable cells such as cardiomyocytes and neurons.
+It allows measuring ionic currents conducted through designated VGICs and the action potential (AP).
+Although the laboratory technique reveals essential properties of VGICs and cardiomyocytes, the detailed gating mechanism remains ambiguous.
+For example, it is hard to determine rigorously how different ion channels collectively lead to changes in the AP.
+Computer models, coupled with experimental observations, provide integrative insights into the data, and they have become an essential part of studying the heart, as illustrated in Figure~\ref{fig:framework_diagram} \citep{winslow2011integrative}.
 
-calibration is necessary to apply this computational approaches (or models). How the calibration has been done in the literature.
-
-Point out problems of curve fitting
+Hence, modeling VGICs involves a data assimilation procedure that combines experimental data with computer models.
+Traditionally, computer models are calibrated to characteristic curves estimated from ionic-current recordings \citep{colman2018description}.
+These estimation processes include curve fitting, which assumes a functional form of the characteristic curve and finds shape parameters best fit data.
+However, curve fitting only considers a single current recording at each time independently, so it cannot capture cellular-level dynamics from the current recordings by changes in input stimuli, i.e., voltage-clamp protocols.
+It raises an additional challenge for analyzing I\textsubscript{K} recordings, because one-more curve fitting is used to separate individual current traces generated by various K\textsubscript{v} isoforms.
+I\textsubscript{Ksum} traces are decomposed only based on the assumption of the shapes of the component currents but not the underlying biophysical gating kinetics \citep{kim2022simulation}.
+Therefore, there is an urgent need to develop a novel data assimilation method for modeling and calibrating K\textsubscript{v} kinetics with considering cellular dynamics.
 
 \subsection{Altered Glycosylation and Heart Diseases}
-DCM and altered glycosylation
+Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}.
+It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}.
+However, the underlying pathological mechanism still remains elusive.
+
+We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology.
+Electrical signaling is orchestrated activities of a variety of ion channels and transporters.
+Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
+Glycosylation is a multi-step process and usually ends with sialic acid added.
+We also reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
+Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.
+However, these studies also rely on characteristic curves and statistics estimated from curve fitting.
 
@@
 Global optimum is not guaranteed due to the complexity of our models.
-We developed the multi-random starting scheme to avoid local optima and find the solution as close to the optimum as possible.
+We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible.
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
-Different nonlinear optimization algorithms are tested and compared.
 This work is implemented in MATLAB R2022a, and the code is available at xyz.
@@
 \section{Experimental Design}
-We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation.
-Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}.
-It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}.
-We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology.
-Electrical signaling is orchestrated activities of a variety of ion channels and transporters.
-Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
-
-Glycosylation is a multi-step process and usually ends with sialic acid added.
-We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
 More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
 MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
-To further investigate the role of altered glycosylation and pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
+To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
 To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
@@
 
+We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation.
 It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes.
@@
 \subsection{Parameter Screening}
-\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six marker points.
+\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six markers.
 The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points.
 The farther parameters fall above the straight lines, the more significant impacts they have on the marker.
-We picked the parameters to be calibrated as a union of parameter sets above the straight lines on the six half-normal plots.
+We picked the parameters to be calibrated as a union of the sets of parameters falling above the straight lines on the six half-normal plots.
 For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
@@
 \end{figure}
-
 \textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
@@
 This parsimonious model design is intended to maximize the structural regularization to minimize overfitting.
+In general, voltage-threshold and time-constant-shifting parameters impact the current traces more than others.
 \begin{figure}[!ht]
@@
 \subsection{Model Fitness}
+We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints.
+
+box-constrained BFGS / goodness-of-fit measure / problem of rmse (not normalized) / problem of classical r2 / so, we chose a new r2 measure developed by UCLA guys.
+
 We compared three nonlinear optimization algorithms satisfying box constraints: BFGS, sequential quadratic programming (SQP), and active set methods.
@@
     \centering
-    \includegraphics{figs/model_fitness_4half4.pdf}
+    \includegraphics{figs/model_fitness_4half.pdf}
     \caption{Bar graphs of the sum of RMSEs of (A) WT and (C) MGAT1KO, and representative actual fitness between model outputs and experimental I\textsubscript{K} recordings (in red) at -30 mV and 50 mV voltage steps of (B) observation 26 in WT and (D) 4 in MGAT1KO.}

### 2022-08-28 calibration/result_output.m
Figures revised

@@ -1,4 +1,4 @@
-%----- Bar graphs
+%% Model fitness - nonlinear R-squared measure
 clc
 close all
 clearvars
@@ -25,7 +25,7 @@ l2_wt = pa_wt.mean_L2;
 r2_ko = pa_ko.mean_R2;
 l2_ko = pa_ko.mean_L2;
 
-%----- Without the worst fitting point in MGAT1KO
+%----- Without the worst fitting point in MGAT1KO -----%
 [~,rmv_idx1] = mink(r2_wt,2);
 [~,rmv_idx2] = min(r2_ko);
 
@@ -36,7 +36,7 @@ files_ko(rmv_idx2) = [];
 % files_ko = categorical(files_ko);
 r2_ko(rmv_idx2) = [];
 
-fig = figure('Color','w','Position',[0,0,500,610]);
+fig = figure('Color','w','Position',[0,0,750,600]);
 orient(fig,'landscape')
 subplot(6,4,[1,5,9])
 barh(categorical(1:length(files_wt)), r2_wt,'b');
@@ -54,7 +54,7 @@ grid on
 % xlabel("Nonlinear R^{2}")
 set(gca,'FontWeight','bold')
 
-%----- Representative example of showing actual fitness
+%----- Representative example of showing actual fitness -----%
 volts = -30:10:50;
 [~,max_wt] = max(r2_wt);
 exp_data = table2array(...
@@ -132,6 +132,11 @@ for i=1:size(yksum,2)
 %         ylabel("Current (pA/pF)")
         set(gca,'LineWidth',1.5,'FontWeight','bold');
         set(gca,'GridLineStyle','--')        
+
+        if i==9
+            legend(["Experimental I_{K}", "Model Prediction"])
+            legend Box off
+        end
     end
 end
 
@@ -198,7 +203,8 @@ for i=1:size(yksum,2)
     end
 end
 
-%% protocol
+%% Model fitness - RMSE
+%----- protocol -----%
 clc
 clearvars
 close all
@@ -230,7 +236,7 @@ protocol = cell(6,1);
 protocol{1} = -70; % hold_volt
 protocol{2} = -91.1; % ek
 
-%% RMSE bar graph and current fitting
+%----- RMSE bar graph and current fitting -----%
 % plot 1
[…]

### 2022-08-29 calibration/result_output.m
Protocol

@@ -390,8 +390,37 @@ ylabel("Current (pA/pF)")
 set(gca,'FontWeight','bold')
 
 %% Kinetics modeling - Estimation
+clc
+close all
+
+%----- protocol -----%
+load("file_names_4half4.mat")
+
 exp_num = "exp51";
+current_names = {'iktof','ikslow1','ikslow2','ikss'};
+tune_idx = cell(5,1);
+tune_idx{1} = [1, 2, 3, 4, 5, 7, 11, 13];
+tune_idx{2} = [2, 3];
+tune_idx{3} = [1, 2, 4, 5, 9, 10 11];
+tune_idx{4} = [2, 3];
+tune_idx{5} = [1, 2, 3, 4];
+[mdl_struct,psize] = gen_mdl_struct(current_names,tune_idx);
+
+pdefault = cell(5,1);
+pdefault{1} = [33, 15.5, 20, 7, 0.03577, 0.06237, 0.18064, 0.3956, ...
+    0.000152, 0.067083, 0.00095, 0.051335, 0.3846];
+pdefault{2} = [-1050, 270, 0.0629];
+pdefault{3} = [22.5, 45.2, 40.0, 7.7, 5.7, 0.0629, 6.1, 18, 2.058, 803.0, 0.16];
+pdefault{4} = [4912, 5334, 0.16];
+pdefault{5} = [0.0862, 1235.5, 13.17, 0.0611];
+
 volts = -100:10:70;
+ideal_hold_time = 120;
+ideal_end_time = 4.6*1000;
+
+protocol = cell(6,1);
+protocol{1} = -70; % hold_volt
+protocol{2} = -91.1; % ek
 sol_dir = fullfile(pwd,strcat("calib_",exp_num));
 
 atof_wt = NaN(length(files_wt),length(volts));
@@ -845,7 +874,7 @@ zlabel("Embedding Dim3")
 axis tight
 grid off
 legend(["WT","MGAT1KO"],'location','best')
-set(gca,'FontWeight','bold','LineWidth',1.5)
+set(gca,'FontWeight','bold','FontSize',14,'LineWidth',1.5)
 
 
 %% custom functions

### 2022-08-29 writing/draft.tex
Draft revised

@@
 
-I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function for the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}.
+I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}.
 As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows.
@@
 To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
-There were 31 and 30 observations for each WT and MGAT1KO.
+There were 32 observations for each WT and MGAT1KO.
 
@@
 We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints.
+A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than utilizing the final objective function value of the calibration routine to report the fitness.
+The objective function, the sum of RMSEs over voltage steps, does not provide a straightforward interpretation that makes it difficult to measure the performance of the models.
+$R^{2}$, defined in Equation~\ref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model.
+However, this interpretation is not applicable, and its value can be negative for nonlinear models because the variance decomposition in Equation~\ref{eq:var_decomp} does not hold anymore.
+\begin{align}
+    R^{2} &= 1 - \frac{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2}{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2} \label{eq:r2} \\
+    \sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2 &= \sum_{k=1}^{n}(y^{\mathcal{M}}(k|\theta))-\bar{y})^2+\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2 \label{eq:var_decomp}
+\end{align}
 
-box-constrained BFGS / goodness-of-fit measure / problem of rmse (not normalized) / problem of classical r2 / so, we chose a new r2 measure developed by UCLA guys.
-
-We compared three nonlinear optimization algorithms satisfying box constraints: BFGS, sequential quadratic programming (SQP), and active set methods.
-\textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} shows values of the objective function (sum of RMSEs) of the three optimization algorithms for WT and MGAT1KO, respectively.
-The three algorithms barely showed differences in the objective values.
-We compared the actual model predictions and experimental potassium I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
-Representative data were selected randomly for each group and plotted at -30 mV and 50 mV voltage steps to illustrate discrepancies at low and high voltages.
-Sum of RMSEs values are lower in MGAT1KO than WT in general, because magnitudes are smaller for the disease group.
-Calibration results of the BFGS algorithm will be used in analysis of kinetics modeling in the rest the sections.
+Therefore, a pseudo-R-squared measure, proposed in \citet{li2019prediction} was used to evaluate the model fitness in this study.
+This nonlinear $R^{2}$ measure is based on the concept of linear correction of prediction function, allowing the variance decomposition for nonlinear functions, which guarantees straightforward interpretation and normalized scores between 0 and 1 as in the classical $R^{2}$.
+\textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} present averages of $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively.
+In most cases, more than 80 \% of variances on average in multiple current recordings are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
+The model fitness measures in MGAT1KO are less than WT in general.
+It is expected that the small current magnitude of MGAT1KO challenges the calibration process.
+We compared the actual model predictions and experimental I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
+As illustrated in the comparison plot of MGAT1KO at -30 mV clamp voltage, small and noisy current trace results in decreased $R^{2}$, 0.8335.
 \begin{figure}[!ht]
@@
     \includegraphics{figs/model_fitness_4half.pdf}
-    \caption{Bar graphs of the sum of RMSEs of (A) WT and (C) MGAT1KO, and representative actual fitness between model outputs and experimental I\textsubscript{K} recordings (in red) at -30 mV and 50 mV voltage steps of (B) observation 26 in WT and (D) 4 in MGAT1KO.}
+    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (A) WT and (C) MGAT1KO.
+Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{K} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
     \label{fig:model_fitness}
@@
 The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior.
-\textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results of showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
-There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.2 at voltages lower 0 mV (\textbf{Figure~\ref{fig:kinetics_ikto}B}).
-Activation and inactivation time constants $\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$ to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
-$\tau_{a}^{(1)}$ for MGAT1KO is higher in range of $>$ -10 mV than WT.
-$\tau_{i}^{(1)}$ for MGAT1KO is slightly shifted toward left.
-The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increases time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
+\textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
+There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}).
+Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
+$\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV.
+$\tau_{i}^{(1)}$ for MGAT1KO shifted toward left.
+The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
 \begin{figure}[!ht]
@@
     \includegraphics{figs/kinetics_ikto.pdf}
-    \caption{Kinetics functions in I\textsubscript{Kto} and their relationships to voltage.
-(A) steady-state activation, (B) steady-state inactivation, (C) time-constant activation, and (D) time-constant inactivation.}
+    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
     \label{fig:kinetics_ikto}
@@
 \textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}.
-The most prominent result is that the time constants of I\textsubscript{Kslow1} are estimated to be significantly slow in MGAT1KO $\sim$ 3.4 and 235.8 ms (\textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}).
+The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}).
 Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO.
 SSI is also shifted $\sim$ 10 mV to the right.
-Slow time constants and increased SSI are likely to address the flattened shape of MGAT1KO currents as illustrated in \textbf{Figure~\ref{fig:model_fitness}D}).
-Time constant inactivation of I\textsubscript{Kslow2} (\textbf{Figure~\ref{fig:kinetics_rec}C}) and activation of I\textsubscript{Kss} (\textbf{Figure~\ref{fig:kinetics_rec}F}) do not show significant differences between the two groups.
+The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO.
+Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}C}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
 \begin{figure}[!ht]
@@
     \includegraphics{figs/kinetics_rec.pdf}
-    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (D) time-constant activation, (E) time-constant inactivation, (C) I\textsubscript{Kslow2} time-constant inactivation, and (F) I\textsubscript{Kss} time-constant activation.}
+    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
     \label{fig:kinetics_rec}
@@
 \subsection{Parameter Distributions Between WT and MGAT1KO} \label{s:results.distribution}
-We estimated distributions of the calibrated kinetic parameters to analyze the impacts of reduced N-glycosylation further.
+We estimated distributions of the calibrated kinetic parameters using histograms to analyze the impacts of reduced N-glycosylation further.
 \textbf{Figure~\ref{fig:hist_pkto}} shows the distributional differences of the kinetic parameters in the K\textsubscript{4.2} model between the two group.
@@
 Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point.
-\textbf{Figure~\ref{fig:cluster}} presents the visualization of the t-SNE embedding of the calibration results.
+\textbf{Figure~\ref{fig:embd_result}} presents the visualization of the t-SNE embedding of the calibration results.
 It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
@@
     \centering
-    \includegraphics{figs/cluster.pdf}
+    \includegraphics{figs/embd_result.pdf}
     \caption{t-SNE embedding of calibration parameters.}
-    \label{fig:cluster}
+    \label{fig:embd_result}
 \end{figure}

### 2022-09-27 writing/draft.tex
Abstract revised

@@
 \begin{abstract}
-Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart.
-Even slight changes in K\textsubscript{v} functions can cause fatal heart disease, such as long QT syndrome.
-Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death.
-Activities of K\textsubscript{v} channels were also affected by altered glycosylation, and K\textsuperscript{+} currents were significantly reduced.
-Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of the cardiac action potential (AP).
-On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
-This characteristic raises challenges for studying K\textsubscript{v} channels.
-Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
-In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured.
-Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}.
-Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents.
-However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously.
-Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols.
-Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces and apply the proposed method to our glycosylation data.
+Potassium channels (K\textsubscript{v}) are critical to the electrical conduction system of the heart, and aberrant activities in K\textsubscript{v} functions can cause fatal heart disease.
+K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte.
+Repolarization is a complex process that consists of multiple phases to which different K\textsubscript{v} isoforms contribute.
+For example, in the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively.
+It is a unique characteristic of K\textsubscript{v} compared to other cardiac ion channels where dominant isoforms exist.
+Although it is highly desirable to study individual K\textsubscript{v} isoforms to understand pathological mechanisms in the cardiac conduction system, only the sum of their activities as a whole can be measured in experiments that record whole-cell voltage clamp K\textsubscript{v} currents (I\textsubscript{K}).
+A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms.
+Most existing studies separate individual K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure, assuming a curve of I\textsubscript{K} consists of the summation of exponential decay functions and a constant term; each represents the shape of K\textsuperscript{+} currents.
+However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting I\textsubscript{K} decomposition only relies on the shape of K\textsuperscript{+} current without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell with different conditions one-by-one individually.
+There is an urgent need to develop new data assimilation methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of the ion channels to provide insights into clinical applications, such as drug design.
+Therefore, we propose a novel \textit{concurrent data assimilation} framework for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{K} traces measured in the same cell.
 First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
@@
 Last, low-dimensional embedding is used to visualize distributional difference of kinetic parameters of cells, showing the cellular-level variability.
-Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
-This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
+We apply the proposed framework to data of glyco-altered I\textsubscript{K} recordings from our previous studies.
+Experimental results show that the proposed method can estimate differences in kinetics between the control and experiment groups.
+This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
 \end{abstract}

### 2022-09-28 writing/draft.tex
Abstract shortened

--- 2022-08-29 writing/draft.tex
@@
 \begin{abstract}
-Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart.
-Even slight changes in K\textsubscript{v} functions can cause fatal heart disease, such as long QT syndrome.
-Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death.
-Activities of K\textsubscript{v} channels were also affected by altered glycosylation, and K\textsuperscript{+} currents were significantly reduced.
-Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of the cardiac action potential (AP).
-On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
-This characteristic raises challenges for studying K\textsubscript{v} channels.
-Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
-In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured.
-Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}.
-Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents.
-However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously.
-Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols.
-Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces and apply the proposed method to our glycosylation data.
-First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
-Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
-Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings.
-Last, low-dimensional embedding is used to visualize distributional difference of kinetic parameters of cells, showing the cellular-level variability.
-Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
-This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
+Potassium channels (K\textsubscript{v}) are responsible for repolarizing the cardiac action potential.
+Different K\textsubscript{v} isoforms contribute to various phases of repolarization.
+In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
+A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities can be measured in the form of currents (I\textsubscript{K}).
+Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure.
+However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually.
+There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms.
+We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics.
+We apply the proposed framework to glyco-altered I\textsubscript{K} recordings from our previous studies.
+Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups.
+This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
 \end{abstract}

### 2022-10-10 writing/draft.tex
Manuscript revised

@@
 	{
-		\title{\bf Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
+		\title{\bf Concurrent Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
 		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
 		$^a$ Department of Industrial and Manufacturing Engineering, \\
-		The Pennsylvania State University, State College, PA USA \\
+		The Pennsylvania State University, University Park, PA USA \\
         $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
@@
 \begin{abstract}
-Potassium channels (K\textsubscript{v}) are responsible for repolarizing the cardiac action potential.
-Different K\textsubscript{v} isoforms contribute to various phases of repolarization.
-In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
-A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities can be measured in the form of currents (I\textsubscript{K}).
-Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure.
+Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes.
+Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization.
+In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
+Data assimilation is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities is measured in the form of currents (I\textsubscript{K}).
+Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} recordings and make inference of kinetics by a curve-fitting decomposition procedure.
 However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually.
@@
 We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics.
-We apply the proposed framework to glyco-altered I\textsubscript{K} recordings from our previous studies.
+We apply the proposed method to glyco-altered I\textsubscript{K} recordings from our previous studies.
 Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups.
@@
 \noindent%
-{\it Keywords:} Data assimilation, model calibration, potassium channels, kinetics modeling, glycosylation
+{\it Keywords:} Data assimilation, computational electrophysiology, potassium channels, kinetics modeling, glycosylation
 
@@
 \section{Introduction}
-Cardiac voltage-gated potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential (AP) in cardiomyocytes.
-They are indispensable in the electrical conduction system of the heart.
-Even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
+Voltage-gated potassium channels (K\textsubscript{v}) are critical to the electrical conduction system of the heart.
+K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time, representing the net electrical activity in a myocyte.
+The AP is produced by the orchestrated activities of ion channels.
+Voltage-gated sodium channels (Na\textsubscript{v}) initiates fast depolarization that drives the sharp upstroke of the AP.
+K\textsubscript{v} play antagonistic roles that repolarize the AP with the rapid inactivation of Na\textsubscript{v}.
+Repolarization consists of multiple phases, to which different K\textsubscript{v} isoforms contribute with fluxes of Ca\textsuperscript{2+} and Cl\textsuperscript{-} involved through designated ion channels.
+This complex process can be affected by even modest changes in K\textsubscript{v} activities.
+Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
 Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
@@
 
-However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
-In contrast, there is a predominant isoform for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
-Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}.
-Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments.
-Typically, a form of curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
-Then the shape parameters are used to calibrate computer models \citep{kim2022simulation}, or kinetics information is added to the data assimilation process with additional curve fitting if applicable \citep{du2017silico}.
-However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
+However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have various isoforms playing different roles in AP repolarization \citep{nerbonne2005molecular}, whereas dominant isoforms exist for other major VGICs \citep{abriel2010cardiac,benitah2010type}.
+In the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively.
+Although it is highly desirable to study individual K\textsubscript{v} isoforms to understand pathological mechanisms in the cardiac conduction system, only the collective activities as a whole can be measured in the form of currents using whole-cell voltage-clamp K\textsubscript{v} currents (I\textsubscript{K}) because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}.
+A data assimilation method, combining experimental data with mathematical/computational models, can used to estimate information on K\textsubscript{v} isoforms.
+Most existing studies separate individual K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure, assuming a curve of I\textsubscript{K} consists of the summation of exponential decay functions and a constant term; each represents the shape of K\textsuperscript{+} currents \citep{brunet2004heterogeneous}.
+However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting I\textsubscript{K} decomposition only relies on the shape of K\textsuperscript{+} current without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell with different conditions (i.e., clamp voltages) one-by-one individually.
 These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
-Therefore, there is an urgent need to develop new data assimilation methods and tools that delineate the kinetics of K\textsubscript{v} isoforms for better understanding their roles in heart diseases.
+There is an urgent need to develop new data assimilation methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of the ion channels to provide insights into clinical applications, such as drug design.
 
-This paper presents a novel approach to data assimilation for analyzing whole-cell K\textsubscript{v} current traces (I\textsubscript{K}) and modeling the underlying kinetics of K\textsubscript{v} isoforms.
+Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{K} traces measured in the same cell.
 A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}.
-First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting.
-Computer models of ion channels are a system of differential equations consisting of gating variables and kinetic functions that continuously simulate ion channel activities over time.
-Parameters in kinetic functions control the gating of ion channels and, in turn, determine current outputs.
-Second, a sensitivity analysis was performed to screen out parameters that marginally influence the model outputs.
-Third, we developed a model calibration that minimizes discrepancy between model outputs and experimental data generated from multiple protocols.
-Last, we made estimations and inferences of impacts of disease-related perturbations on potassium channel kinetics and used low-dimensional embedding to show cellular variability.
-Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
+First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents.
+Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation.
+Several markers are defined that best represent the characteristics of the currents in various perspectives.
+Third, \textit{in-vitro} data of I\textsubscript{K} recordings are coupled with the proposed models by calibrating the parameters selected by the sensitivity analysis.
+We apply the proposed framework to glyco-altered IK recordings from our previous studies.
+Our data showed that reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
+Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells.
+Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform kinetics of both the wild type and disease-altered groups and successfully captures cellular variability.
 \begin{figure}
@@
 
-The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases.
+The main advantage of this approach is that it analyzes multiple I\textsubscript{K} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases.
 Our contributions are summarized as follows:
 \begin{itemize}
-    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
-    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
-    \item The proposed approach builds separate K\textsubscript{v} models and calibrates them for each cell, which allows analyzing the cellular variability. 
+    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting procedures.
+    \item We developed a concurrent data assimilation routine that harnesses multiple protocols simultaneously rather than a single protocol independently one-by-one, resulting in comprehensive cellular-level modeling.
+    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
     \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
 \end{itemize}
-The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.
+The proposed data assimilation framework has strong potential to serve as a novel complementary data analytical tool to \textit{in-vitro} experiments.
 
@@
 
-\subsection{Data Assimilation and Model Calibration}
-\textit{Data assimilation} is a method to find the optimal configuration and state of computational models by coupling them with experimental data.
+\subsection{Concurrent Assimilation of Functional Data}
+Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling it with experimental data.
 Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation.
@@
 
-In this study, $\delta$ is defined by the sum of root-mean-square errors (RMSEs) as in Equation~\ref{eq:rmse}, which measures deviations between experimental data and model predictions from the end of a holding potential $t_i^{(h)}$ to the end of voltage step $t_i^{(e)}$ across a range of protocols, $i=1,2,\cdots,n$.
+Traditionally, bi-/tri-exponential functions as described in Equation~\ref{eq:expl_fitting} have served as $y^{\mathcal{M}}(x|\theta)$, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$.
+Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{K} recordings.
+In general, multiple input protocols are applied to a myocyte producing a set of I\textsubscript{K} recordings to study the voltage-dependent characteristics of VGICs.
+However, these exponential functions cannot analyze simultaneously, leading to potential information loss in the cellular-level dynamics of ion channels.
+In contrast, computer models of VGICs can generate currents given one set of parameters because they are designed by biophysical principles encoding voltage dependence.
+In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in Equation~\ref{eq:rmse}.
+RMSE evaluates the goodness-of-fit between a I\textsubscript{K} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$.
+RMSEs from the same cell are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltage $v=1,2,\cdots,n$.
 \begin{equation}
     \label{eq:rmse}
-    \delta = \sum_{i=1}^{n} \sqrt{\int_{t_i^{(h)}}^{t_i^{(e)}}\frac{(y_i^{\mathcal{D}}(t) - y_i^{\mathcal{M}}(t|\theta))^2}{t_i^{(e)}-t_i^{(h)}}dt}
+    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y_v^{\mathcal{M}}(t|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
 \end{equation}
+
 This approach differs from previous studies in two-fold.
@@
 Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.
-
 We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$.
@@
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
-This work is implemented in MATLAB R2022a, and the code is available at xyz.
+This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.
 
@@
 \textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
-There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}).
+There is a very slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}).
 Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
@@
 \textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}.
-The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}).
-Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO.
-SSI is also shifted $\sim$ 10 mV to the right.
+The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}C} and \textbf{D}).
+Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO (see \textbf{Figure~\ref{fig:kinetics_rec}A}).
+SSI is also shifted $\sim$ 10 mV to the right (see \textbf{Figure~\ref{fig:kinetics_rec}B}).
 The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO.
-Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}C}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
+Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}E}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
+It is worth mentioning that the proposed method extrapolates $\tau_{i}^{(2)}$ that is greater than 5 s from the 4.5-s-pulse protocol.
 \begin{figure}[!ht]
@@
 
-\subsection{Parameter Distributions Between WT and MGAT1KO} \label{s:results.distribution}
-We estimated distributions of the calibrated kinetic parameters using histograms to analyze the impacts of reduced N-glycosylation further.
-\textbf{Figure~\ref{fig:hist_pkto}} shows the distributional differences of the kinetic parameters in the K\textsubscript{4.2} model between the two group.
-A notable fact about the figure is that MGAT1KO shows small variances for Parameters 1, 5, and 7 in transition $\alpha_{a}$ and $\beta_{a}$, which control kinetics activation of I\textsubscript{Kto}.
-Reduced glycosylation affects activation gating that plays a vital role in I\textsubscript{Kto}, as it is characterized by a high peak at the early depolarization and rapid inactivation.
-In addition, MGAT1KO causes a leftward shift in the distribution of the maximum conductance $G_{\mathrm{Kto}}$, resulting in smaller conduction than WT in general.
-It is likely due to the observed current reductions in MGAT1KO.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/hist_pkto.pdf}
-    \caption{Distribution estimations of calibration parameters in the K\textsubscript{Kto} model.}
-    \label{fig:hist_pkto}
-\end{figure}
-
-\textbf{Figure~\ref{fig:hist_prectifiers}} shows distribution estimations of the calibration parameters in the rectifier and steady-state currents models.
-\textbf{Figure~\ref{fig:hist_prectifiers}A} shows significant distributional differences in Parameters 9 and 10, which control time constants of I\textsubscript{Kslow1}.
-It is compatible with the previous observations of significant differences in time constants to voltage relationships of I\textsubscript{Kslow1} between WT and MGAT1KO.
-There are slight shifts in Parameters 1 and 4 of I\textsubscript{Kslow1} that affect steady-state activation.
-The voltage-threshold parameter for MGAT1KO in steady-state inactivation has higher densities at low values around 30 than WT, which has the peak mode around 48.
-It accounts for the shift at low voltages less than 0 mV in SSI.
-The maximum conductance $G_{\mathrm{Kslow1}}$ of MGAT1KO is centered around relatively small values compared to WT.
-This leftward shift in the maximum conductance is also observed in I\textsubscript{Kslow2} (\textbf{Figure~\ref{fig:kinetics_rec}B}).
-Although there is a difference in variance of $G_{\mathrm{Kss}}$, the two groups have similar means around 0.03 (i.e.
-WT: $0.0375\pm0.0027$ vs MGAT1KO: $0.0325\pm0.015$).
-It is compatible with the observation that I\textsubscript{Kss} does not show significant reductions.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/hist_prectifiers.pdf}
-    \caption{Distribution estimations of calibration parameters in (A) the I\textsubscript{Kslow1} model, (B) the I\textsubscript{Kslow2} model, and (C) the I\textsubscript{Kss} model.}
-    \label{fig:hist_prectifiers}
-\end{figure}
-
-\subsection{Low-dimensional Embedding and Visualization}
 We aggregated calibration results into one to analyze the parameters in a collective way.
@@
 \section{Conclusions}\label{s:conclusions}
-We propose a novel data assimilation framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
+We propose a novel \textit{concurrent data assimilation} framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
+In this study, we propose a novel \textit{concurrent data assimilation} method for modeling K\textsubscript{v} kinetics from multiple I\textsubscript{v} recordings measured from the same cell with different conditions for investigating disease-altered functionalities in the ion channels.
 This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights.
@@
 \section*{Acknowledgements}
-The authors acknowledge the generous support from the funding agency of XYZ and funding list} \fi
+This work was supported in part by the National Science Foundation (MCB-1856132 to HY) and (MCB-1856199, IOS-1146882, and IOS-1660926 to EB), an American Heart Association, Greater Southeast Affiliate Grant-In-Aid (14GRNT20450148 to EB), and Postdoctoral Fellowship (15POST25710010 to AE).
+HK and HY would also like to thank the NSF I/UCRC Center for Health Organization Transformation (CHOT) award NSF IIP-1624727 for the support of their research work.
+} \fi
 

### 2022-10-10 writing/draft_old.tex
Draft v3

--- 2022-09-28 writing/draft.tex
@@
 \begin{abstract}
-Potassium channels (K\textsubscript{v}) are responsible for repolarizing the cardiac action potential.
-Different K\textsubscript{v} isoforms contribute to various phases of repolarization.
-In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
-A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities can be measured in the form of currents (I\textsubscript{K}).
-Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure.
-However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually.
-There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms.
-We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics.
-We apply the proposed framework to glyco-altered I\textsubscript{K} recordings from our previous studies.
-Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups.
-This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
+Activities of cardiac voltage-gated potassium channels (K\textsubscript{v}) are essential in the electrical conduction system of the heart.
+Even slight changes in K\textsubscript{v} functions can cause fatal heart disease, such as long QT syndrome.
+Our recent studies found that altered glycosylation causes aberrant electrical signaling in mouse cardiomyocytes and leads to heart failure and early death.
+Activities of K\textsubscript{v} channels were also affected by altered glycosylation, and K\textsuperscript{+} currents were significantly reduced.
+Among major cardiac voltage-gated ion channels (VGIC), K\textsubscript{v} channels have a distinctive feature: various isoforms generate their own unique currents contributing to different phases of repolarization of the cardiac action potential (AP).
+On the other hand, the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents.
+This characteristic raises challenges for studying K\textsubscript{v} channels.
+Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating.
+In whole-cell recordings, the sum of K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be measured.
+Traditional methods apply curve fitting to estimate components of K\textsubscript{v} isoforms from I\textsubscript{Ksum}.
+Additional curve fitting is used to find kinetics information for the individual K\textsuperscript{+} currents.
+However, this traditional approach is limited in its ability to determine gating kinetics of K\textsubscript{v} isoforms rigorously.
+Curve fittings do not account for interactions of multiple gating processes of various isoforms and dynamics according to changes in protocols.
+Therefore, we propose a novel data assimilation framework for investigating whole-cell K\textsubscript{v} current traces and apply the proposed method to our glycosylation data.
+First, computer models of K\textsuperscript{+} currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+Second, we develop fractional factorial designs to identify the parameters that have significant impacts on the current shapes.
+Third, we develop model calibration as nonlinear optimization that couple the \textit{in-silico} K\textsubscript{v} models and \textit{in-vitro} I\textsubscript{K} recordings.
+Last, low-dimensional embedding is used to visualize distributional difference of kinetic parameters of cells, showing the cellular-level variability.
+Experimental results show that the data assimilation method can estimate differences in kinetic rate between the control and experiment groups.
+This method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
 \end{abstract}

### 2022-10-10 writing/draft_diff.tex
Test

\documentclass[11pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

%%%%%%% IISE Transactions margin specifications %%%%%%%%%%%%%%%%%%%
% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%
\makeatletter
\renewcommand\section{\@startsection {section}{1}{\z@}%
                                   {-3.5ex \@plus -1ex \@minus -.2ex}%
                                   {2.3ex \@plus.2ex}%
                                   {\normalfont\fontfamily{phv}\fontsize{16}{19}\bfseries}}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                                     {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\fontfamily{phv}\fontsize{14}{17}\bfseries}}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                                    {-3.25ex\@plus -1ex \@minus -.2ex}%
                                     {1.5ex \@plus .2ex}%
                                     {\normalfont\normalsize\fontfamily{phv}\fontsize{14}{17}\selectfont}}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% IISE Transactions package list %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib} %comment out if you do not have the package
\usepackage{url} % not crucial - just used below for the URL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% Author package list and commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% Here are some examples %%%%%%%%%%%%%%
%	\usepackage{amsfonts, amsthm, latexsym, amssymb}
%	\usepackage{lineno}
%	\newcommand{\mb}{\mathbf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\def\spacingset#1{\renewcommand{\baselinestretch}%
		{#1}\small\normalsize} \spacingset{1}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\if0\blind
	{
		\title{\bf Simulation-guided Data Assimilation for Kinetics Modeling of Potassium Channel Isoforms in Mouse Cardiomyocytes}
		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
		$^a$ Department of Industrial and Manufacturing Engineering, \\
		The Pennsylvania State University, State College, PA USA \\
        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
        Wright State University, City, OH USA}
		\date{}
		\maketitle
	} \fi

	\if1\blind
	{
        \title{\bf \emph{IISE Transactions} \LaTeX \ Template}
		\author{Author information is purposely removed for double-blind review}

\bigskip
		\bigskip
		\bigskip
		\begin{center}
			{\LARGE\bf \emph{IISE Transactions} \LaTeX \ Template}
		\end{center}
		\medskip
	} \fi
	\bigskip

\begin{abstract}
This document provides a \LaTeX \ template for \emph{IISE Transactions}. Your paper should be compiled in the following order: title; abstract; keywords; main text, including an introduction and a conclusion or summary; acknowledgments; declaration of interest statement; references; appendices (as appropriate). Figures and tables should be inserted into the text as close to first mention as possible (NOT appended to the end of the manuscript). In-text citations and the reference list must follow \emph{IISE Transactions} guidelines. Use 11 point font, 1 inch margins, and double-spacing for the manuscript. A typical paper for this journal should be no more than 30 pages in manuscript format, counting from the title page to references. Appendices should be included as supplemental online materials. Do not use footnotes. \emph{IISE Transactions} uses a double-blind review process. Please make sure that you submit the \textbf{blind version} of your manuscript, which does not contain any information identifying the authors.  This includes removing the authors information on the title page as well as the information that may be identifying in the Acknowledgment section. We thank you for your attention to these details.
\end{abstract}

\noindent%
{\it Keywords:} \emph{IISE Transactions}; \LaTeX; Manuscript format; Taylor \& Francis.

%\newpage
\spacingset{1.5} % DON'T change the spacing!

\section{Introduction} \label{s:intro}
\begin{itemize}
    \item \textbf{Motivation} Voltage-gated ion channels (VGICs) are the functional unit of the electrical conduction system of the heart that controls the synchronous contraction of numerous cardiomyocytes. Among major cardiac VGICs, K\textsuperscript{+} channels (K\textsubscript{v}) have a distinctive feature: they have various isoforms that contribute to K\textsuperscript{+} current, while the dominant isoform is responsible for Na\textsuperscript{+} and Ca\textsuperscript{2+} currents. Although each K\textsubscript{+} generated through K\textsubscript{v} isoform has a different role in the repolarization phase, only the sum of these individual K\textsuperscript{+} currents (I\textsubscript{Ksum}) can be recorded through in-vitro experiments. Data analysis tools are used to estimate information about K\textsubscript{v} isoforms from experimental I\textsubscript{Ksum} recordings.
    \item \textbf{Gaps \& Needs} A traditional approach is based on the data-driven curve fitting method that assumes the functional forms for the shape of individual K\textsuperscript{+} currents. This method directly fits individual K\textsuperscript{+} currents to experimental I\textsubscript{Ksum} recordings by searching the shape parameters of the current functions that have the best fit. It has been proven that the curve fitting approach describes I\textsubscript{Ksum} well with estimated K\textsuperscript{+} currents. However, the mathematical form of this classical method only characterizes the shape of the currents, not the kinetics of corresponding K\textsubscript{v} isoforms. In addition, the curve fitting method handles experimental I\textsubscript{Ksum} recordings from the same cell independently, so it lacks the cellular-level analysis.
    \item \textbf{Objective} As opposed to the traditional data-driven curve fitting method, we propose a novel framework of biophysics-driven data assimilation method for kinetics modeling of K\textsubscript{v} isoforms based on computer models of K\textsuperscript{+} currents that can simulate underlying gating kinetics of the currents.
    \item \textbf{Approach} 
    \item \textbf{Experiment \& Performance} The proposed framework outperforms the curve fitting method for minimizing discrepancies between experimental and estimated total K\textsuperscript{+} currents. We also provide the kinetics modeling results that are not possible in the traditional method.
    \item \textbf{Significance} Experimental results show that the proposed method has strong potential to be a novel analysis tool to decompose I\textsubscript{Ksum} recordings to estimate not only the current characteristics but also kinetics of K\textsubscript{v} isoforms.
\end{itemize}

The review of literature can be a separate section but it can be merged with the introduction section. One can find a number of reference examples at the end of this template, like,
\begin{itemize}
\item A book: \cite{Hastie2009}.
\item A journal article: \cite{Byon2013}.
\item A conference paper: \cite{Breunig2000}.
\item A book chapter: \cite{Yang2010}.
\item A technical report: \cite{Kelley2007}.
\end{itemize}

\section{Methods} \label{s:methods}

\subsection{\emph{Computer Models of Potassium Channel Isoforms}} \label{s:methods.1}
modeling (mathematical equations)
\begin{figure}
    \centering
    \includegraphics{figs/ap_underlying_currents.pdf}
    \caption{Caption}
    \label{fig:ap_underlying_currents}
\end{figure}

\subsection{\emph{Sensitivity Analysis of Calibration Parameters}} \label{s:methods.2}
this section shows the results of the variable screening.
\begin{figure}
    \centering
    \includegraphics{figs/doe_kto.pdf}
    \caption{Half-normal plots I\textsubscript{Kto}}
    \label{fig:doe_kto}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figs/doe_kslow1.pdf}
    \caption{Half-normal plots I\textsubscript{Kslow1}}
    \label{fig:doe_kslow1}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figs/doe_kslow2.pdf}
    \caption{Half-normal plots I\textsubscript{Kslow2}}
    \label{fig:doe_kslow2}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figs/doe_kss.pdf}
    \caption{Half-normal plots I\textsubscript{Kss}}
    \label{fig:doe_kss}
\end{figure}

\subsection{\emph{Model Calibration}} \label{s:methods.3}
calibration procedure; non-linear optimization methods

\subsection{\emph{Graphical User Interface}} \label{s:methods.4}
show how to use it

\subsection{\emph{Case Study Glycosylations}} \label{s:methods.5}
case study

\section{Results} \label{s:results}
Perhaps numerical analysis of performance, comparison, and demonstration of applicability and impact using real data.

\section{Conclusion}\label{s:conclusion}
A paper ends with a conclusion or summary section.

\if0\blind{
\section*{Acknowledgements}
The authors acknowledge the generous support from the funding agency of XYZ.	} \fi

\bibliographystyle{chicago}
\spacingset{1}
\bibliography{IISE-Trans}

\end{document}

### 2022-11-02 writing/draft.tex
Draft revision

@@
 	{
-		\title{\bf Concurrent Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
+		\title{\bf Model-guided Concurrent Data Assimilation for Calibrating Cardiac Ion-channel Kinetics}
 		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
@@
 Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes.
-Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization.
-In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
-Data assimilation is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities is measured in the form of currents (I\textsubscript{K}).
-Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} recordings and make inference of kinetics by a curve-fitting decomposition procedure.
-However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually.
-There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms.
-We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics.
-We apply the proposed method to glyco-altered I\textsubscript{K} recordings from our previous studies.
-Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups.
-This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
+Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization, e.g., the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
+Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate different currents of K\textsubscript{v} isoforms.
+Most existing studies separate and make inference of K\textsubscript{v} kinetics via a curve-fitting procedure (e.g., tri-exponential fitting), but are limited in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be decomposed and fitted for one clamp voltage at a time.
+In this paper, we develop a novel \textit{concurrent data assimilation} framework that calibrates biophysics-based computer models to delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
+The proposed method is evaluated and validated with voltage-clamp I\textsubscript{Ksum} recordings from our \textit{in-vitro} experiments on both wild-type and glyco-altered myocytes.
+Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes disease-altered perturbations to various K\textsubscript{v} isoforms.
+In addition, we develop a graphical-user-interface (GUI) software that provides an enabling tool to medical scientists for model calibration with voltage-clamp data.
+The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in variety of heart diseases.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} Data assimilation, computational electrophysiology, potassium channels, kinetics modeling, glycosylation
+{\it Keywords:} Data assimilation, simulation model, ion channels, congenital glycosylation disorder, model calibration
 
@@
 \section{Introduction}
-Voltage-gated potassium channels (K\textsubscript{v}) are critical to the electrical conduction system of the heart.
-K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time, representing the net electrical activity in a myocyte.
-The AP is produced by the orchestrated activities of ion channels.
-Voltage-gated sodium channels (Na\textsubscript{v}) initiates fast depolarization that drives the sharp upstroke of the AP.
-K\textsubscript{v} play antagonistic roles that repolarize the AP with the rapid inactivation of Na\textsubscript{v}.
-Repolarization consists of multiple phases, to which different K\textsubscript{v} isoforms contribute with fluxes of Ca\textsuperscript{2+} and Cl\textsuperscript{-} involved through designated ion channels.
-This complex process can be affected by even modest changes in K\textsubscript{v} activities.
-Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
+Cardiac voltage-gated potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart.
+K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time representing the net electrical activity in a cardiomyocyte \citep{ravens2008role}.
+The shape and duration of the AP are primarily determined by various K\textsubscript{v} isoforms, contributing to different phases of repolarization with unique biophysical properties as illustrated in \textbf{Figure~\ref{fig:kv_ap}} \citep{nerbonne2005molecular}.
+Diverse distribution is a unique feature of K\textsubscript{v}.
+Dominant isoforms exist for other major voltage-gated ion channels (VGICs) such as Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
+Repolarization begins with the rapid inactivation of Na\textsubscript{v}, reducing inward Na\textsuperscript{+} flux, and fast open and close of K\textsubscript{v} isoforms, generating rapidly inactivating K\textsuperscript{+} current.
+There are other K\textsubscript{v} that remain open producing slowly inactivating currents, but in large species such as human, this phase 2 is characterized by prolonged repolarization due to intracellular Ca\textsuperscript{2+} flux (``plateau'').
+During phase 3, Ca\textsubscript{v} close and the cell begins to repolarize rapidly (``rapid repolarization'').
+In small species such as mice, phase 2 and 3 are not very distinctive.
+Hence, in the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating delayed rectifier-type currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively \citep{brunet2004heterogeneous}.
+Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{montpetit2009regulated}.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/kv_ap.pdf}
+    \caption{Ventricular action potential in mouse cardiomyocytes and the diverse K\textsubscript{v} isoforms and other ion channels contributing to shaping the AP.}
+    \label{fig:kv_ap}
+\end{figure}
+
+However, only the collective activities of K\textsubscript{v} isoforms can be measured in the form of currents using whole-cell voltage-clamp recording (I\textsubscript{Ksum}) due to their slightly overlapping voltage dependence of gating \citep{brouillette2004functional}.
+Hence, there is a need to decompose I\textsubscript{Ksum} into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological mechanisms in disease-altered cardiomyocytes.
+Data assimilation is a method of data analysis that combines experimental data with mathematical/computational models to make inferences on information of interest, such as individual K\textsuperscript{+} currents.
+Most existing studies separate individual K\textsuperscript{+} currents mathematically via a curve-fitting procedure, assuming a curve of I\textsubscript{Ksum} as the summation of exponential decay functions and a constant term; each represents the shape of individual K\textsuperscript{+} currents \citep{brunet2004heterogeneous}.
+However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition is sorely determined by the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) although a set of I\textsubscript{Ksum} traces are recorded with varying clamp voltages for a cell to investigate the voltage dependence of gating kinetics, only a single I\textsubscript{Ksum} trace is decomposed and fitted to individual K\textsuperscript{+} currents for one clamp voltage.
+ion methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{Ksum} traces for a better understanding of the ion channels to provide insights into mechanism of cardiomyocytes and clinical applications, such as drug design.
+
+Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{Ksum} traces measured in the same cell.
+First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents.
+\textbf{Figure~\ref{fig:exp-vs-com}} shows two different approaches to assimilate the experimental data to extract information of K\textsubscript{v} isoforms.
+Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation.
+Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in various perspectives.
+Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings concurrently.
+Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells.
+We apply the proposed framework to glyco-altered I\textsubscript{Ksum} recordings from our previous studies.
 Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics,nagai2016aberrant}.
-Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
 Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
-In addition, we verified our experimental results via \textit{in-silico} studies and estimate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
-\textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases, which will be referred to as the data assimilation framework that combines experimental data with mathematical/computational models.
+To be specific, reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/exp-vs-com.pdf}
+    \caption{Comparison between the traditional curve-fitting approach and the proposed model-guided data assimilation of \textit{in-vitro} experimental data of I\textsubscript{Ksum} recordings.}
+    \label{fig:exp-vs-com}
+\end{figure}
+
+Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform  currents and underlying kinetics of both the wild-type and disease-altered groups and successfully captures cellular variability.
+\textit{In-silico} modeling provides rich information that \textit{in-vitro} data cannot alone.
+For example, in our previous studies and this paper, computer models are used to verify experimental results, estimate channel kinetics, and show how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
+\textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases.
+Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to medical scientists for investigating K\textsubscript{v} channelopathy, which is available as online supplemental material.
+The main advantage of the proposed approach is that it analyzes multiple I\textsubscript{Ksum} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases.
+Our contributions are summarized as follows:
+\begin{itemize}
+    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{Ksum} recordings based on biophysical principles rather than curve-fitting procedures.
+    \item We developed a concurrent data assimilation routine that harnesses a set of I\textsubscript{Ksum} traces simultaneously rather than a single current trace independently one-by-one, resulting in comprehensive cellular-level modeling.
+    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
+    \item Our method captures differences in channel kinetics of the wild-type and the disease-altered groups.
+\end{itemize}
 \begin{figure}[!ht]
@@
 
-However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have various isoforms playing different roles in AP repolarization \citep{nerbonne2005molecular}, whereas dominant isoforms exist for other major VGICs \citep{abriel2010cardiac,benitah2010type}.
-In the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively.
-Although it is highly desirable to study individual K\textsubscript{v} isoforms to understand pathological mechanisms in the cardiac conduction system, only the collective activities as a whole can be measured in the form of currents using whole-cell voltage-clamp K\textsubscript{v} currents (I\textsubscript{K}) because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}.
-A data assimilation method, combining experimental data with mathematical/computational models, can used to estimate information on K\textsubscript{v} isoforms.
-Most existing studies separate individual K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure, assuming a curve of I\textsubscript{K} consists of the summation of exponential decay functions and a constant term; each represents the shape of K\textsuperscript{+} currents \citep{brunet2004heterogeneous}.
-However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting I\textsubscript{K} decomposition only relies on the shape of K\textsuperscript{+} current without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell with different conditions (i.e., clamp voltages) one-by-one individually.
-These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
-There is an urgent need to develop new data assimilation methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of the ion channels to provide insights into clinical applications, such as drug design.
+\section{Research Background}
+\subsection{Data Assimilation and Calibration of Cardiac Models}
+The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself.
+Cardiac electrophysiology has contributed significantly to our understanding of heart and disease-related modifications, such as congenital disorders of glycosylation (CDG).
+Advances in laboratory techniques allow studying molecular-level activities in cardiomyocytes via measuring currents of ion channels, which are functional units of the electrical conduction system of the heart.
+For example, whole-cell current recordings are used to show the roles of a certain gene, encoding $\alpha$ subunits that form Na\textsubscript{v} in human cardiomyocytes, which causes a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}.
+Analysis of ionic currents showed pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities \citep{ednie2013sialicNav1, ednie2015sialicKv}; O-glycosylation modulations cause aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
+However, \textit{in-vitro} experiments alone are limited to investigate channel activities.
+There is a need to assimilate experimental data to estimate information that cannot be observed directly.
+Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.
 
-Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{K} traces measured in the same cell.
-A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}.
-First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
-It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents.
-Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation.
-Several markers are defined that best represent the characteristics of the currents in various perspectives.
-Third, \textit{in-vitro} data of I\textsubscript{K} recordings are coupled with the proposed models by calibrating the parameters selected by the sensitivity analysis.
-We apply the proposed framework to glyco-altered IK recordings from our previous studies.
-Our data showed that reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
-Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells.
-Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform kinetics of both the wild type and disease-altered groups and successfully captures cellular variability.
-\begin{figure}
-    \centering
-    \includegraphics{figs/flow_chart.pdf}
-    \caption{Flowcharts of (A) the current approach and (B) the proposed method.}
-    \label{fig:flow_chart}
-\end{figure}
-
-The main advantage of this approach is that it analyzes multiple I\textsubscript{K} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases.
-Our contributions are summarized as follows:
-\begin{itemize}
-    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting procedures.
-    \item We developed a concurrent data assimilation routine that harnesses multiple protocols simultaneously rather than a single protocol independently one-by-one, resulting in comprehensive cellular-level modeling.
-    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
-    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
-\end{itemize}
-The proposed data assimilation framework has strong potential to serve as a novel complementary data analytical tool to \textit{in-vitro} experiments.
-
-\section{Research Background}
-\subsection{Computational Modeling of the Heart}
-Computer models of the heart have been extensively studied and used in cardiology to understand the mechanism of the organ, investigate fatal diseases, and develop drugs.
-Models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
+However, the curve-fitting approach has limitations in its ability to assimilate the experimental data, because it only considers the shape of information to infer rather than reflecting underlying kinetics and interactions of various gating processes.
+Further, as curve fitting is only based on the shape of a single curve to fit, the traditional approach is not able to analyze a set of current recordings that are supposed to be analyzed together.
+In contrast, computer models of the heart consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
 Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
 For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}.
-With the advancement of sensing technology and data analysis, computational cardiology is expanding its field.
-Personalized heart models pave a new way for monitoring and managing the cardiac health of individuals by characterizing individual heart signals from the large-scale IoT network and calculating dissimilarities to detect potential cardiac risk \citep{yang2020network}.
-Hence, cardiac computer models provide a great opportunity to mitigate heart disease, the number one cause of death in the U.S.
-\citep{murphy2021mortality}.
+Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}.
+In the computational modeling literature, discrete-event simulation has been a primary focus, which only models events that occur apart over time.
+However, computer models of ion channels simulate gating kinetics that changes continuously, expressed mathematically in differential equations.
+The complex nonlinear structure of cardiac computer models also challenges calibration.
+However, little work has been done to calibrate continuous simulation-guided model.
 
-The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself.
-Hence, the electrical activities and waves spreading through the whole heart have been studied considerably.
-For examle, \citet{yang2016whole} a new reaction-diffusion and pattern recognition method to model spatiotemporal dynamics of electrical waves propagating throughout the heart surface.
-Sensing methods, such as vectorcardiogram (VCG) and electrocardiogram (ECG), measure the electrical activities of the heart.
-Various analytics methods have been proposed to cope with their complex nonlinear dynamics in different heart regions over time.
-\citet{chen2019heterogeneous} developed a heterogeneous recurrence method that characterizes spatiotemporal patterns in disease-altered multi-channel cardiac signals.
-Nonlinear recurrence analysis captures and quantifies topological relationships of recurring patterns in signals, which has been proven to be an effective tool for cardiac signals that are naturally recurring over time \citep{yang2010multiscale,meyers2020cross}.
-
-However, to perform a fine-grained analysis, it is necessary to study the cellular-level activities of the heart as opposed to the organ-level signals.
-Voltage-gated ion channels (VGICs) are membrane proteins in cardiomyocytes that are the functional unit of the electrical activities of cardiac muscle cells.
-The whole-cell patch clamp recording is a powerful experimental technique for investigating the electrical properties of excitable cells such as cardiomyocytes and neurons.
-It allows measuring ionic currents conducted through designated VGICs and the action potential (AP).
-Although the laboratory technique reveals essential properties of VGICs and cardiomyocytes, the detailed gating mechanism remains ambiguous.
-For example, it is hard to determine rigorously how different ion channels collectively lead to changes in the AP.
-Computer models, coupled with experimental observations, provide integrative insights into the data, and they have become an essential part of studying the heart, as illustrated in Figure~\ref{fig:framework_diagram} \citep{winslow2011integrative}.
-
-Hence, modeling VGICs involves a data assimilation procedure that combines experimental data with computer models.
-Traditionally, computer models are calibrated to characteristic curves estimated from ionic-current recordings \citep{colman2018description}.
-These estimation processes include curve fitting, which assumes a functional form of the characteristic curve and finds shape parameters best fit data.
-However, curve fitting only considers a single current recording at each time independently, so it cannot capture cellular-level dynamics from the current recordings by changes in input stimuli, i.e., voltage-clamp protocols.
-It raises an additional challenge for analyzing I\textsubscript{K} recordings, because one-more curve fitting is used to separate individual current traces generated by various K\textsubscript{v} isoforms.
-I\textsubscript{Ksum} traces are decomposed only based on the assumption of the shapes of the component currents but not the underlying biophysical gating kinetics \citep{kim2022simulation}.
-Therefore, there is an urgent need to develop a novel data assimilation method for modeling and calibrating K\textsubscript{v} kinetics with considering cellular dynamics.
-
-\subsection{Altered Glycosylation and Heart Diseases}
+\subsection{Congenital Disorders of Glycosylation}
 Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
@@
 Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.
-However, these studies also rely on characteristic curves and statistics estimated from curve fitting.
 
-\section{Methods}
+However, most of existing works utilized the curve-fitting method, e.g., bi-exponential or tri-exponential fitting, to assimilate the \textit{in-vitro} data from glycosylayion experiments.
+Curve-fitting provides estimates of characteristic curves (e.g., steady-state activation or inactivation) and statistics (e.g., time constants) for data-driven insights into CDG diseases.
+Few, if any, previous approaches have leveraged computer models to calibrate and delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
+
+\section{Research Methodology}
 \subsection{Computer Models of Potassium Channel Isoforms}
@@
 Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}).
-\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{K} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
-\textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of protocols and consequent change in I\textsubscript{Ksum}.
+\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{Ksum} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
+\textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of voltage-clamp protocol and consequent change in I\textsubscript{Ksum}.
 Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach.
-In this approach, K\textsuperscript{+} current can be defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
+In this approach, K\textsuperscript{+} current is defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
 \begin{figure}[!ht]
@@
 (B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
-(C) Range of protocols from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
+(C) Range of voltage-clamp protocol from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
     \label{fig:kcurrent_example}
@@
 Traditionally, bi-/tri-exponential functions as described in Equation~\ref{eq:expl_fitting} have served as $y^{\mathcal{M}}(x|\theta)$, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$.
-Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{K} recordings.
-In general, multiple input protocols are applied to a myocyte producing a set of I\textsubscript{K} recordings to study the voltage-dependent characteristics of VGICs.
+Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{Ksum} recordings.
+In general, multiple input voltage steps are applied to a myocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics of VGICs.
 However, these exponential functions cannot analyze simultaneously, leading to potential information loss in the cellular-level dynamics of ion channels.
@@
 In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in Equation~\ref{eq:rmse}.
-RMSE evaluates the goodness-of-fit between a I\textsubscript{K} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$.
+RMSE evaluates the goodness-of-fit between a I\textsubscript{Ksum} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$.
 RMSEs from the same cell are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltage $v=1,2,\cdots,n$.
@@
 This approach differs from previous studies in two-fold.
-First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method includes multiple protocols simultaneously for comprehensive cellular-level modeling.
+First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method uses computer models of K\textsubscript{v} and incorporate a set of I\textsubscript{Ksum} traces simultaneously for comprehensive cellular-level modeling.
+Cardiac models enable this concurrent data assimilation, because they generate multiple current traces with different input voltages by simulating underlying gating kinetics.
 Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}.
+The proposed model-guided concurrent data assimilation method is summarized and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}}.
 Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.
@@
 This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics{figs/flow_chart.pdf}
+    \caption{Schematic diagram of (A) the current curve-fitting approach that searches for the best fit of K\textsuperscript{+}-currents curves to I\textsubscript{Ksum} recordings one-by-one and estimates feature kinetics using additional curve fitting if available; (B) the proposed concurrent model-guided data assimilation that calibrates model parameters to multiple I\textsubscript{Ksum} traces and simulates K\textsubscript{+} currents and underlying kinetics from the model.}
+    \label{fig:flow_chart}
+\end{figure}
 
@@
 To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
-To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
+To be specific, whole-cell I\textsubscript{Ksum} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
 There were 32 observations for each WT and MGAT1KO.
@@
 For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model.
-Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{K} recordings.
-This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics with considering the cellular-level dynamics of multiple protocols.
+Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{Ksum} recordings.
+This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics concurrently.
 
@@
 It is expected that the small current magnitude of MGAT1KO challenges the calibration process.
-We compared the actual model predictions and experimental I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
+We compared the actual model predictions and experimental I\textsubscript{Ksum} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
 As illustrated in the comparison plot of MGAT1KO at -30 mV clamp voltage, small and noisy current trace results in decreased $R^{2}$, 0.8335.
@@
     \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (A) WT and (C) MGAT1KO.
-Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{K} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
+Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{Ksum} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
     \label{fig:model_fitness}
@@
 This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights.
-Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in protocols.
+Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in
+clamp voltages.
 The proposed method focuses on more comprehensive modeling of underlying channel kinetics of K\textsubscript{v} isoforms than traditional approaches.
@@
 The embedding method allows visualizing the variability of each cell.
+The proposed method focuses on the mouse model, but the same framework is applicable to other species easily with modifications to model structures.
 In summary, the proposed framework paves a new way for investigating the electrical signaling of potassium channels and heart diseases.

### 2023-05-16 writing/draft.tex
Initial submission

@@
 	{
-		\title{\bf Model-guided Concurrent Data Assimilation for Calibrating Cardiac Ion-channel Kinetics}
-		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
-		$^a$ Department of Industrial and Manufacturing Engineering, \\
-		The Pennsylvania State University, University Park, PA USA \\
-        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
-        Wright State University, Dayton, OH USA}
+		\title{\bf Concurrent Data Assimilation for Model-guided Learning of Cardiac Potassium Channel Activities}
+		\author{Author information is purposely removed for double-blind review}
 		\date{}
@@
 Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes.
-Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization, e.g., the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
-Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate different currents of K\textsubscript{v} isoforms.
-Most existing studies separate and make inference of K\textsubscript{v} kinetics via a curve-fitting procedure (e.g., tri-exponential fitting), but are limited in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be decomposed and fitted for one clamp voltage at a time.
-In this paper, we develop a novel \textit{concurrent data assimilation} framework that calibrates biophysics-based computer models to delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
-The proposed method is evaluated and validated with voltage-clamp I\textsubscript{Ksum} recordings from our \textit{in-vitro} experiments on both wild-type and glyco-altered myocytes.
-Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes disease-altered perturbations to various K\textsubscript{v} isoforms.
-In addition, we develop a graphical-user-interface (GUI) software that provides an enabling tool to medical scientists for model calibration with voltage-clamp data.
-The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in variety of heart diseases.
+There is a variety of K\textsubscript{v} isoforms and corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}) that contribute to different phases of repolarization.
+Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate K\textsubscript{v} currents and kinetics.
+Most existing studies separate and make inference of K\textsubscript{v} kinetics via curve-fitting procedures (e.g., tri-exponential fitting), but are limited because: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later.
+Here, we develop a novel concurrent data assimilation method that calibrates biophysics-based subject-specific computer models to decompose and delineate kinetics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
+The proposed method is evaluated and validated with whole-cell I\textsubscript{Ksum} recordings from wild-type and chronically glycosylation-deficient cardiomyocytes.
+Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes the glycosylation-conferred perturbations observed experimentally to various K\textsubscript{v} isoforms.
+In addition, we develop a graphical-user-interface (GUI) application that provides an enabling tool to biomedical scientists.
+The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in various heart diseases.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} Data assimilation, simulation model, ion channels, congenital glycosylation disorder, model calibration
+{\it Keywords:} Data assimilation, in-silico modeling, model calibration, cardiac potassium channels, congenital disorders of glycosylation
 
@@
 \section{Introduction}
-Cardiac voltage-gated potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart.
-K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time representing the net electrical activity in a cardiomyocyte \citep{ravens2008role}.
-The shape and duration of the AP are primarily determined by various K\textsubscript{v} isoforms, contributing to different phases of repolarization with unique biophysical properties as illustrated in \textbf{Figure~\ref{fig:kv_ap}} \citep{nerbonne2005molecular}.
-Diverse distribution is a unique feature of K\textsubscript{v}.
-Dominant isoforms exist for other major voltage-gated ion channels (VGICs) such as Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
-Repolarization begins with the rapid inactivation of Na\textsubscript{v}, reducing inward Na\textsuperscript{+} flux, and fast open and close of K\textsubscript{v} isoforms, generating rapidly inactivating K\textsuperscript{+} current.
-There are other K\textsubscript{v} that remain open producing slowly inactivating currents, but in large species such as human, this phase 2 is characterized by prolonged repolarization due to intracellular Ca\textsuperscript{2+} flux (``plateau'').
-During phase 3, Ca\textsubscript{v} close and the cell begins to repolarize rapidly (``rapid repolarization'').
-In small species such as mice, phase 2 and 3 are not very distinctive.
-Hence, in the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating delayed rectifier-type currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively \citep{brunet2004heterogeneous}.
-Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{montpetit2009regulated}.
+Potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart, particularly in the repolarization phase of the action potential (AP).
+The AP is a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte, and the AP shape and duration are primarily determined by K\textsubscript{v} isoforms.
+As shown in Figure~\ref{fig:ap_currents}, the different phases of AP repolarization in mouse cardiomyocytes are the result of the coordinated activity of a variety of K\textsubscript{v} isoforms (e.g., K\textsubscript{v}4.2, K\textsubscript{v}1.5, and K\textsubscript{v}2.1) and their corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2}).
+There are other major voltage-gated ion channels (VGICs), such as for Na\textsuperscript{+} (Na\textsubscript{v}) and Ca\textsuperscript{2+} (Ca\textsubscript{v}), that also contribute to the AP.
+The Na\textsubscript{v} are primarily responsible for the AP upstroke in the depolarization phase, while the Ca\textsubscript{v} control the cellular contraction.
+Aberrant activities of ion channels can significantly impact the AP and lead to fatal arrhythmias.
+Using whole-cell voltage-clamp recording methods, only the collective activities of K\textsubscript{v} isoforms can be measured reliably as the sum of all K\textsuperscript{+} currents (I\textsubscript{Ksum}), and these K\textsubscript{v} isoforms have overlapping biophysical properties (i.e., voltage dependence of gating and kinetics), which complicate the ability to separate one type of I\textsubscript{K} from another \citep{brouillette2004functional}.
+K\textsubscript{v} activities can be altered in various cardiomyopathies \citep{tristani2001molecular,giudicessi2012potassium}, thus there is an urgent need to decompose I\textsubscript{Ksum} more rigorously into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological roles by comparing their activities in healthy versus diseased hearts and cardiomyocytes.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/kv_ap.pdf}
-    \caption{Ventricular action potential in mouse cardiomyocytes and the diverse K\textsubscript{v} isoforms and other ion channels contributing to shaping the AP.}
-    \label{fig:kv_ap}
+    \includegraphics{figs/ap_currents.pdf}
+    \caption{(a) Ventricular action potential in mouse cardiomyocytes and (b) diverse K\textsubscript{v} isoforms and their currents contributing to the AP shape.}
+    \label{fig:ap_currents}
 \end{figure}
 
-However, only the collective activities of K\textsubscript{v} isoforms can be measured in the form of currents using whole-cell voltage-clamp recording (I\textsubscript{Ksum}) due to their slightly overlapping voltage dependence of gating \citep{brouillette2004functional}.
-Hence, there is a need to decompose I\textsubscript{Ksum} into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological mechanisms in disease-altered cardiomyocytes.
-Data assimilation is a method of data analysis that combines experimental data with mathematical/computational models to make inferences on information of interest, such as individual K\textsuperscript{+} currents.
-Most existing studies separate individual K\textsuperscript{+} currents mathematically via a curve-fitting procedure, assuming a curve of I\textsubscript{Ksum} as the summation of exponential decay functions and a constant term; each represents the shape of individual K\textsuperscript{+} currents \citep{brunet2004heterogeneous}.
-However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition is sorely determined by the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) although a set of I\textsubscript{Ksum} traces are recorded with varying clamp voltages for a cell to investigate the voltage dependence of gating kinetics, only a single I\textsubscript{Ksum} trace is decomposed and fitted to individual K\textsuperscript{+} currents for one clamp voltage.
-ion methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{Ksum} traces for a better understanding of the ion channels to provide insights into mechanism of cardiomyocytes and clinical applications, such as drug design.
+Transgenic mouse models are the most commonly used animal models that have provided insights into cardiac research despite the differences between rodents and humans; the isoforms expressed in both small and large mammals share many similarities \citep{milani2014small}.
+Most existing studies using mouse models separate individual K\textsuperscript{+} currents mathematically via a curve-fitting procedure \citep{costantini2005homeodomain,ednie2019reduced2,teng2022tmem65}, assuming the I\textsubscript{Ksum} current trace represents the summation of exponential decay functions (two or three, usually) and a constant term; each represents the shape of individual K\textsuperscript{+} currents.
+However, the ability to scrutinize the kinetics and dynamics of individual K\textsubscript{v} isoforms is limited by this traditional method because: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces as discreet exponential functions which does not discern the underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be decomposed for one clamp voltage at a time, and estimated information from each fitting is analyzed in a population-averaged way later.
+New methods are urgently needed to simultaneously delineate the underlying kinetics of multiple I\textsubscript{Ksum} traces from the same cell concurrently for a better understanding of K\textsubscript{v} channels and their roles as well as cellular variability in diseased cardiomyocytes.
 
-Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{Ksum} traces measured in the same cell.
-First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+This paper presents a new approach of subject-specific concurrent data assimilation for model-guided learning of K\textsubscript{v} isoforms.
+First, computer models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
 It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents.
-\textbf{Figure~\ref{fig:exp-vs-com}} shows two different approaches to assimilate the experimental data to extract information of K\textsubscript{v} isoforms.
-Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation.
-Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in various perspectives.
-Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings concurrently.
-Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells.
-We apply the proposed framework to glyco-altered I\textsubscript{Ksum} recordings from our previous studies.
-Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
-To be specific, reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/exp-vs-com.pdf}
-    \caption{Comparison between the traditional curve-fitting approach and the proposed model-guided data assimilation of \textit{in-vitro} experimental data of I\textsubscript{Ksum} recordings.}
-    \label{fig:exp-vs-com}
-\end{figure}
+Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant effects on the current generation.
+Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in different perspectives.
+Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple \textit{in-silico} models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings from the same cell concurrently.
+Last, low-dimensional embedding is utilized to visualize calibrated parameters in a collective way across the cells.
+The proposed methodology is evaluated and validated by comparing I\textsubscript{Ksum} recordings on healthy control versus chronically glycosylation-deficient cardiomyocyte.
+Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other VGICs \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+Our \textit{in-vitro} experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced,ednie2019reduced2}.
+In fact, these studies also showed that reduced cardiomyocyte complex N-glycosylation is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
 
-Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform  currents and underlying kinetics of both the wild-type and disease-altered groups and successfully captures cellular variability.
-\textit{In-silico} modeling provides rich information that \textit{in-vitro} data cannot alone.
-For example, in our previous studies and this paper, computer models are used to verify experimental results, estimate channel kinetics, and show how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
-\textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases.
-Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to medical scientists for investigating K\textsubscript{v} channelopathy, which is available as online supplemental material.
-The main advantage of the proposed approach is that it analyzes multiple I\textsubscript{Ksum} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases.
+Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} currents with underlying kinetics and successfully captures cellular variability.
+For example, it verifies the experimental data that showed the significant reduction in the current magnitude and elongated decaying for the glycosylation-deficient cardiomyocyte group.
+Additionally, computational modeling estimates channel kinetics, which was not possible in the electrophysiological experiments alone.
+Cellular variability is visualized by low-dimensional embedding of calibration parameters into 3D space, which shows one of the powers of systematic computer models that allow one to predict how changes in VGICs impact cellular functions.
+Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to biomedical scientists for investigating K\textsubscript{v}-related channelopathy that is available as online supplemental material.
+The proposed method shows strong potential for modeling the kinetics and gating mechanism of K\textsubscript{v} to study heart diseases.
 Our contributions are summarized as follows:
 \begin{itemize}
-    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{Ksum} recordings based on biophysical principles rather than curve-fitting procedures.
-    \item We developed a concurrent data assimilation routine that harnesses a set of I\textsubscript{Ksum} traces simultaneously rather than a single current trace independently one-by-one, resulting in comprehensive cellular-level modeling.
-    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
-    \item Our method captures differences in channel kinetics of the wild-type and the disease-altered groups.
+    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles,
+    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics caused by voltage dependence,
+    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding,
+    \item A GUI application is provided to make the proposed method an enabling tool for biomedical scientists.
 \end{itemize}
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/overall_diagram.pdf}
-    \caption{Conceptual diagram to illustrate a data assimilation framework.}
-    \label{fig:framework_diagram}
-\end{figure}
 
@@
 \subsection{Data Assimilation and Calibration of Cardiac Models}
-The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself.
-Cardiac electrophysiology has contributed significantly to our understanding of heart and disease-related modifications, such as congenital disorders of glycosylation (CDG).
-Advances in laboratory techniques allow studying molecular-level activities in cardiomyocytes via measuring currents of ion channels, which are functional units of the electrical conduction system of the heart.
-For example, whole-cell current recordings are used to show the roles of a certain gene, encoding $\alpha$ subunits that form Na\textsubscript{v} in human cardiomyocytes, which causes a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}.
-Analysis of ionic currents showed pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities \citep{ednie2013sialicNav1, ednie2015sialicKv}; O-glycosylation modulations cause aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
-However, \textit{in-vitro} experiments alone are limited to investigate channel activities.
-There is a need to assimilate experimental data to estimate information that cannot be observed directly.
+The functionality of the heart to pump blood through the body is controlled by electrical signals generated by the organ itself.
+Cardiac electrophysiology has contributed significantly to our understanding of the heart and disease-related modifications, such as congenital disorders of glycosylation (CDG).
+Advancements in laboratory techniques allow researchers to study molecular-level activities in cardiomyocytes via measuring ionic currents conducted through VGICs.
+For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2, ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
+However, \textit{in-vitro} experiments alone are limited to investigate detailed channel activities.
+There is a need to assimilate experimental data to estimate information that are not able to be observed directly.
 Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.
 
-However, the curve-fitting approach has limitations in its ability to assimilate the experimental data, because it only considers the shape of information to infer rather than reflecting underlying kinetics and interactions of various gating processes.
-Further, as curve fitting is only based on the shape of a single curve to fit, the traditional approach is not able to analyze a set of current recordings that are supposed to be analyzed together.
-In contrast, computer models of the heart consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
-Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
+However, curve-fitting approaches have limitations in that they only consider the shape of informational functions to infer rather than incorporating underlying kinetics and interactions of various gating processes.
+Further, as curve fitting works for a single curve independently, the traditional approach is less concerned about the dynamics of a set of multiple responses measured from the same cell (e.g., the coincidental activity of multiple K\textsubscript{v} isoforms).
+In contrast, computer models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
+Mathematical and computational modeling allows for studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
+Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}.
 For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}.
-Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}.
-In the computational modeling literature, discrete-event simulation has been a primary focus, which only models events that occur apart over time.
-However, computer models of ion channels simulate gating kinetics that changes continuously, expressed mathematically in differential equations.
-The complex nonlinear structure of cardiac computer models also challenges calibration.
-However, little work has been done to calibrate continuous simulation-guided model.
+From an engineering perspective, computer models of ion channels are dynamic system simulations that represent continuously changing gating kinetics.
+The dynamic nature of the models complicates the calibration process.
+
+Complex nonlinear structures of cardiac computer models also make calibration difficult.
+We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous \textit{in-silico} studies to cope with the complexity of cardiac models.
+However, all these studies are based on current-shape parameters via curve fitting, or characteristic curves/statistics from additional curve fitting applied to estimated currents.
+In addition, only population-average models are derived and calibrated via learning the entire dataset for each healthy and diseased group.
+Little work has been done to leverage computer models to calibrate and delineate kinetic dynamics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
 
@@
 Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}.
-It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}.
-However, the underlying pathological mechanism still remains elusive.
-
+A growing number of studies have shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ednie2019reduced2,ohtsubo2006glycosylation}.
+It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly show modest reductions in protein glycosylation, present with cardiac deficits, including idiopathic DCM \citep{marques2017cardiac}.
+However, uncovering the underlying pathological mechanisms still remains elusive.
 We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology.
 Electrical signaling is orchestrated activities of a variety of ion channels and transporters.
-Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
+VGICs are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
 Glycosylation is a multi-step process and usually ends with sialic acid added.
-We also reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
+We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
 Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.
-
-However, most of existing works utilized the curve-fitting method, e.g., bi-exponential or tri-exponential fitting, to assimilate the \textit{in-vitro} data from glycosylayion experiments.
-Curve-fitting provides estimates of characteristic curves (e.g., steady-state activation or inactivation) and statistics (e.g., time constants) for data-driven insights into CDG diseases.
-Few, if any, previous approaches have leveraged computer models to calibrate and delineate kinetic dynamics of various K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
 
@@
 \subsection{Computer Models of Potassium Channel Isoforms}
-Mouse models have been very useful methods for studying human cardiac electric signaling \citep{nerbonne2004studying,milani2014small,kim2022irx5}.
-Computer models of mouse cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP.
-In this paper, the AP is modeled as a differential equation of transmembrane currents and stimulus current I\textsubscript{stim} as below in Equation~\ref{eq:ap} where $C_{m}$ is the membrane capacitance, $t$ is time.
-\begin{equation}
-    \label{eq:ap}
-    \begin{split}
-    -C_{m}\frac{dV}{dt} &= \mathrm{I}_{\mathrm{CaL}}+\mathrm{I}_{\mathrm{p}(Ca)}+\mathrm{I}_{\mathrm{NaCa}}+\mathrm{I}_{\mathrm{Cab}}+\mathrm{I}_{\mathrm{Na}}+\mathrm{I}_{\mathrm{Nab}}+\mathrm{I}_{\mathrm{NaK}}+\mathrm{I}_{\mathrm{Cl},Ca} \\
-    &+\mathrm{I}_{\mathrm{Kto}}+\mathrm{I}_{\mathrm{Kslow1}}+\mathrm{I}_{\mathrm{Kslow2}}+\mathrm{I}_{\mathrm{Kss}}+\mathrm{I}_{\mathrm{Ks}}+\mathrm{I}_{\mathrm{Kr}}+\mathrm{I}_{\mathrm{K1}}+\mathrm{I}_{\mathrm{stim}}
-    \end{split}
-\end{equation}
-There are 15 transmembrane currents in this AP model: the L-type calcium current (I\textsubscript{CaL}), the calcium pump current (I\textsubscript{p(Ca)}), the Na\textsuperscript{+}/Ca\textsuperscript{2+} exchange current (I\textsubscript{NaCa}), the calcium background current (I\textsubscript{Cab}), the fast Na\textsuperscript{+} current (I\textsubscript{Na}), the background Na\textsuperscript{+} current (I\textsubscript{Nab}), the Na\textsuperscript{+}/K\textsuperscript{+} pump current (I\textsubscript{NaK}), the Ca\textsuperscript{2+}-activated Cl\textsuperscript{-} current (I\textsubscript{Cl,Ca}), the rapidly inactivating transient outward K\textsuperscript{+} current (I\textsubscript{Kto}), the two ultra-rapidly activating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}), the slow delayed rectifier K\textsuperscript{+} current (I\textsubscript{Ks}), the rapid delayed rectifier K\textsuperscript{+} current (I\textsubscript{Kr}), and the time independent K\textsuperscript{+} current (I\textsubscript{K1}).
-Detailed implementation of mouse AP and VGIC models can be found in \cite{bondarenko2014compartmentalized,asfaw2020compartmentalized}.
-
-The three types of K\textsuperscript{+} currents have larger magnitude compared to the other K\textsuperscript{+} currents: the rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), the delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}).
-The rapidly inactivating transient outward currents have two components I\textsubscript{Kto,f} and I\textsubscript{Kto,s}, but the later is essentially absent in apex cardiomyocytes.
-We only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because we focus on apex myocytes in this investigation.
-However, the transient outward current can be easily modified according to the region of ventricular cardiomyocytes.
-Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}).
-\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{Ksum} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
-\textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of voltage-clamp protocol and consequent change in I\textsubscript{Ksum}.
+Because of the relative ease at which their genetics can be altered, mice have been useful models for studying cardiac electric signaling \citep{milani2014small,nerbonne2004studying}.
+In ventricular and atrial mouse cardiomyocytes, there are three major components of I\textsubscript{Ksum}: rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), which are conducted through K\textsubscript{v}4.2 and K\textsubscript{v}1.4 respectively, slowly inactivating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), which are conducted through K\textsubscript{v}1.5/K\textsubscript{v}2.1, and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}) \citep{xu1999four}.
+I\textsubscript{Kto,s} (K\textsubscript{v}1.4) is primarily found in septal cardiomyocytes \citep{bondarenko2004computer}.
+In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our \textit{in-vitro} studies focused on the ventricular apex myocytes.
+However, I\textsubscript{Kto} can be easily modified according to the region of ventricular cardiomyocytes.
+We model these four dominant K\textsuperscript{+} currents: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}.
+Figure~\ref{fig:kcurrent_example}(b) illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the whole-cell I\textsubscript{K} trace (I\textsubscript{Ksum}) given the protocol in Figure~\ref{fig:kcurrent_example}(a) that applies 0 mV voltage pulse from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
+Figure~\ref{fig:kcurrent_example}(c) and (d) show a range of voltage steps from -30 mV to 50 mV in 10 mV increments and consequent changes in I\textsubscript{Ksum}.
 Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach.
-In this approach, K\textsuperscript{+} current is defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
+In this approach, if all four current models are included , I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
+In some cases, it is reduced to a bi-exponential function combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2}.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/exemplary_kv_currents.pdf}
-    \caption{Example of voltage-clamp K\textsuperscript{+} currents.
-(A) Voltage step of 0 mV from holding potential -70 mV.
-(B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
-(C) Range of voltage-clamp protocol from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
+    \includegraphics{figs/kcurrent_example.pdf}
+    \caption{Example of voltage-clamp protocol and K\textsuperscript{+} current traces.
+(a) Clamp-voltage pulse of 0 mV from the holding potential -70 mV.
+(b) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
+(c) Series of clamp-voltage pulses (-30-50 mV in 10 mV increments) from the holding potential -70 mV and (d) consequent I\textsubscript{Ksum} traces.}
     \label{fig:kcurrent_example}
@@
     \label{eq:expl_fitting}
-    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}
+    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}.
 \end{equation}
 
-Unlike the simple current functions, computer models of K\textsubscript{v} isoforms can simulate detailed gating kinetics.
-We developed them with Hodgkin-Huxley modeling, which has two gating variables controlling conductance, a popular modeling scheme in various species \citep{ten2004model,mahajan2008rabbit}.
-For example, K\textsubscript{v} isoforms that conducts current I\textsubscript{K} can be defined by Equation~\ref{eq:hh_ex}
+Unlike the simple current functions, K\textsubscript{v} computer models simulate detailed gating kinetics underlying the currents.
+We develop mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}.
+This type of model consists of two gating variables controlling the channel conductance, and its canonical form is defined by \eqref{eq:hh_ex}
 \begin{equation}
     \label{eq:hh_ex}
-    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}})
+    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}}),
 \end{equation}
 where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential.
-$V-E_{\mathrm{K}}$ is the driving force of the ion movement.
-Important components in this equation are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$.
-These processes are governed by first-order kinetics and its voltage-dependent transition rates $\alpha$ and $\beta$.
-$\alpha$ is the rate at which a gate in closed state opens, whereas $\beta$ is the rates at which a gate in open state closes.
-Equation~\ref{eq:gv} shows a schematic relationship of this gating kinetics.
+$V-E_{\mathrm{K}}$ implies the driving force of the ion movement.
+Important components in \eqref{eq:hh_ex} are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$.
+These processes are governed by first-order kinetics and voltage-dependent transition rates $\alpha$ and $\beta$.
+$\alpha$ is the rate at which a gate in a closed state opens, whereas $\beta$ is the rate at which a gate in an open state closes.
+Equation~\eqref{eq:gv} shows a schematic relationship of this gating kinetics.
 \begin{align}
     \label{eq:gv}
-    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i
+    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i.
 \end{align}
 
-The two biophysical processes can be modeled using differential equations in two ways as follows:
+These two biophysical processes $a$ and $i$ can be expressed using differential equations in two ways \eqref{eq:1stkinetic1} and \eqref{eq:1stkenetic2}
 \begin{align}
+    \label{eq:1stkinetic1}
     \frac{da}{dt} &=\alpha_{a}(1-a)-\beta_{a}a   &\frac{di}{dt} &=\alpha_{i}(1-i)-\beta_{i}i \\
-    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}}
+    \label{eq:1stkenetic2}
+    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}},
 \end{align}
@@
     a_{\infty} &= \frac{\alpha_{a}}{\alpha_{a}+\beta_{a}} & i_{\infty} &=  \frac{\alpha_{i}}{\alpha_{i}+\beta_{i}} \\
-    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}
+    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}.
 \end{align}
-The steady-state values and time constants can be defined directly by functions of voltage without transition rates in some cases.
-These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters that control the behavior of the kinetics of an ion channel.
+The steady-state values and time constants can be defined directly by functions of voltage $V$ without transition rates in some cases.
+These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters $p_{i}$ that control the behavior of the kinetics of an ion channel.
 
-\subsubsection{In-silico Modeling of K\textsubscript{v}4.2 (I\textsubscript{Kto})}
-The rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation.
+\subsubsection{In-silico Modeling of I\textsubscript{Kto}}
+The rapidly inactivating transient outward current I\textsubscript{Kto}, which is conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation.
 It mainly contributes to the peak at the very beginning of activation in I\textsubscript{Ksum}.
@@
     &\alpha_{i} = \frac{p_{9}e^{-(V+p_{2})/p_{4}}}{1+p_{10}e^{-(V+p_{2}+p_{3})/p_{4}}} \\
-    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}} \label{eq:ikto_beta2}
+    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}}.
+    \label{eq:ikto_beta2}
 \end{align}
 There are two gating variables $a_{\mathrm{Kto}}$ and $i_{\mathrm{Kto}}$ responsible for activation and inactivation, respectively.
-Their kinetics are governed by transition-rate functions from Equation~\ref{eq:ikto_alpha1} to Equation~\ref{eq:ikto_beta2}.
+Their kinetics are governed by transition-rate functions from \eqref{eq:ikto_alpha1} to \eqref{eq:ikto_beta2}.
 Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
 
-\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})} \label{s:method.model2}
-There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are very rapidly activating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
-As shown in \textbf{Figure~\ref{fig:kcurrent_example}}, both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, but I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1}.
-The non-inactivating steady-state current, which conducted through K\textsubscript{2P} family, remains constant during the voltage-clamp recording.
-These three currents collectively address the most part of the decaying phase of I\textsubscript{Ksum}.
-We assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and I\textsubscript{Kss} has similar activation behavior with a slightly different rate to keep the model as simple as possible to reduce the structural risk of overfitting.
+\subsubsection{In-silico Modeling of I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}}
+There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are rapidly activating and slowly inactivating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
+As shown in Figure~\ref{fig:kcurrent_example}(b), both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, and I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1} \citep{liu2011dissection}.
+The non-inactivating steady-state current I\textsubscript{Kss}, which is likely conducted through K\textsubscript{v} isoforms K\textsubscript{2P} family \citep{feliciangeli2015family}, remains constant during the voltage-clamp recording.
+These three currents contribute to most part of the decaying portion of I\textsubscript{Ksum}.
+To keep the models as simple as possible to reduce the structural risk of overfitting, we assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and they have a similar inactivation pattern; I\textsubscript{Kss} has similar activation behavior with the two delayed rectifier currents but a slightly different rate.
 
-I\textsubscript{Kslow1} is modeled without transition-rate functions as opposed to I\textsubscript{Kto}.
-Its gating variables, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$ are defined directly by steady-state ($a_{ss}$ and $i_{ss}$) and time-constant functions ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) as follows.
+I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss} are modeled without transition-rate functions as opposed to I\textsubscript{Kto}, and their steady-state and time-constant functions are directly defined.
+First, the gating variables of I\textsubscript{Kslow1}, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$, are defined by \eqref{eq:akslow1} and \eqref{eq:ikslow1}.
+The steady-state functions ($a_{ss}$ and $i_{ss}$) in \eqref{eq:ass} and \eqref{eq:iss} will be shared with the other two current models.
+A full description of I\textsubscript{Kslow1} is given as follows:
 \begin{align}
     &\mathrm{I}_{\mathrm{Kslow1}} = G_{\mathrm{Kslow1}}a_{\mathrm{Kslow1}}i_{\mathrm{Kslow1}}(V-E_{\mathrm{K}}) \\
-    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(1)}} \\
-    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(1)}} \\
+    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(\mathrm{Kslow1})}} \label{eq:akslow1} \\
+    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(\mathrm{Kslow1})}} \label{eq:ikslow1} \\
     &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \label{eq:ass} \\
     &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \label{eq:iss} \\
-    &\tau_{a}^{(2)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
-    &\tau_{i}^{(2)} = p_{10} - p_{8}i_{ss}
+    &\tau_{a}^{(\mathrm{Kslow1})} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
+    &\tau_{i}^{(\mathrm{Kslow1})} = p_{10} - p_{8}i_{ss}.
 \end{align}
 
-I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}.
-As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows.
+As a structural regularization, I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1} as in \eqref{eq:act_var}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ shares the same steady-state function $i_{ss}$ \eqref{eq:iss} with I\textsubscript{Kslow1}.
+As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows:
 \begin{align}
     &\mathrm{I}_{\mathrm{Kslow2}} = G_{\mathrm{Kslow2}}a_{\mathrm{Kslow2}}i_{\mathrm{Kslow2}}(V-E_{\mathrm{K}}) \\
-    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \\
-    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(2)}} \\
-    &\tau_{i}^{(3)} = p_{2} - p_{1}i_{ss}
+    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \label{eq:act_var} \\
+    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(\mathrm{Kslow2})}} \\
+    &\tau_{i}^{(\mathrm{Kslow2})} = p_{2} - p_{1}i_{ss}.
 \end{align}
 
-I\textsubscript{Kss} does not have an inactivation variable as it is non-inactivating current.
-It shares the same steady-state function for activation $a_{ss}$ with the rectifier currents but have a separate time-constant function to address the different activation rate.
+I\textsubscript{Kss} does not have an inactivation variable because it is non-inactivating current.
+It shares the same steady-state function for activation $a_{ss}$ \eqref{eq:ass} with the other two delayed rectifier currents but have a separate time-constant function \eqref{eq:time_ikss} to address the different activation rate.
 I\textsubscript{Kss} is modeled as
@@
     &\mathrm{I}_{\mathrm{Kss}} = G_{\mathrm{Kss}}a_{\mathrm{Kss}}(V-E_{\mathrm{K}}) \\
-    &\frac{da_{\mathrm{Kss}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kss}}}{\tau_{a}^{(3)}} \\
-    &\tau_{a}^{(4)}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}
+    &\frac{da_{\mathrm{Kss}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kss}}}{\tau_{a}^{(\mathrm{Kss})}} \\
+    &\tau_{a}^{(\mathrm{Kss})}= \frac{p_{2}}{e^{p_{1}(V+p_{3}^\prime)}+e^{-p_{1}(V+p_{3}^\prime)}} + p_{3}. \label{eq:time_ikss}
 \end{align}
@@
 \subsection{Concurrent Assimilation of Functional Data}
-Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling it with experimental data.
+Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling them with experimental data.
 Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation.
 The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment.
-Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term.
+Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in \eqref{eq:rd_relation}, where $\epsilon$ is the error term.
 \begin{equation}
@@
 Let $y^{\mathcal{M}}(x|\theta)$ denote the output from a computer model $\mathcal{M}$, given parameters $\theta$.
-Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in Equation~\ref{eq:dm_relation}.
+Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in \eqref{eq:dm_relation}:
 \begin{align}
@@
     y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
-    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
+    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon.
 \end{align}
@@
 
-Traditionally, bi-/tri-exponential functions as described in Equation~\ref{eq:expl_fitting} have served as $y^{\mathcal{M}}(x|\theta)$, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$.
-Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{Ksum} recordings.
-In general, multiple input voltage steps are applied to a myocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics of VGICs.
-However, these exponential functions cannot analyze simultaneously, leading to potential information loss in the cellular-level dynamics of ion channels.
-In contrast, computer models of VGICs can generate currents given one set of parameters because they are designed by biophysical principles encoding voltage dependence.
-In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in Equation~\ref{eq:rmse}.
-RMSE evaluates the goodness-of-fit between a I\textsubscript{Ksum} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$.
-RMSEs from the same cell are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltage $v=1,2,\cdots,n$.
+From this perspective, bi-/tri-exponential functions in \eqref{eq:expl_fitting} serve as $y^{\mathcal{M}}(x|\theta)$ in the curve-fitting approach, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$.
+Note that $\mathrm{A}_{i}(x)$ and $\tau_{i}(x)$ are dependent on input data $x$.
+In general, multiple input voltage steps are applied to a cardiomyocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics.
+In contrast, computer models of VGICs generate currents for different input voltages given one set of parameters, because they are designed by biophysical principles encoding voltage dependence.
+In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in \eqref{eq:rmse}.
+RMSE evaluates the goodness-of-fit between experimental I\textsubscript{Ksum} and model prediction.
+RMSEs for a set of I\textsubscript{Ksum} traces are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltages $v=1,2,\cdots,n$.
 \begin{equation}
     \label{eq:rmse}
-    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y_v^{\mathcal{M}}(t|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
+    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y^{\mathcal{M}}(t,v|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
 \end{equation}
 
-This approach differs from previous studies in two-fold.
-First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method uses computer models of K\textsubscript{v} and incorporate a set of I\textsubscript{Ksum} traces simultaneously for comprehensive cellular-level modeling.
 Cardiac models enable this concurrent data assimilation, because they generate multiple current traces with different input voltages by simulating underlying gating kinetics.
-Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}.
-The proposed model-guided concurrent data assimilation method is summarized and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}}.
-Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.
-We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$.
+Note that the proposed method calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data via curve-fitting \citep{du2013silico,du2015statistical, du2017silico, kim2022simulation}.
+We develop the box-constrained nonlinear optimization routine with multi-random initial points to minimize $\delta$.
 Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube.
 In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling.
-Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting.
-Global optimum is not guaranteed due to the complexity of our models.
-We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible.
+The multi-random-start scheme helps escape local optima and find the solution as close to the optimum as possible.
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
 This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/flow_chart.pdf}
-    \caption{Schematic diagram of (A) the current curve-fitting approach that searches for the best fit of K\textsuperscript{+}-currents curves to I\textsubscript{Ksum} recordings one-by-one and estimates feature kinetics using additional curve fitting if available; (B) the proposed concurrent model-guided data assimilation that calibrates model parameters to multiple I\textsubscript{Ksum} traces and simulates K\textsubscript{+} currents and underlying kinetics from the model.}
-    \label{fig:flow_chart}
-\end{figure}
 
-\subsection{Model Regularization and Sensitivity Analysis}
-The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$.
-Excessive flexibility of a model has a risk of causing overfitting that occurs when the model fits data too closely, even including noise and random effects in the data.
-Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results of parameters.
-It is worth noting that the models presented in Section 3.1 are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simply the model structure.
-Specifically, the same parameters that control voltage thresholds, such as $p_{1}$, $p_{2}$, $p_{3}$, and $p_{4}$, are used in equations having a similar function.
-The steady-state functions (Equation~\ref{eq:ass} and \ref{eq:iss}) are reused in the three rectifier-type currents.
+\subsection{Sensitivity Analysis and Model Regularization}
+The principle of parsimony is critical in model calibration to enhance fitting accuracy, prevent overfitting, and improve interpretability of $\theta$.
+Excessive flexibility has a risk of overfitting that occurs when the model fits data too closely, even including noise and random effects in data.
+Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results.
+It is worth emphasizing that the presented models are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simplify the model structure.
 
-We also performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
-In order to do that, factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output.
-As illustrated in \textbf{Figure~\ref{fig:current_markers}}, six markers were defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
-Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
-Marker f will equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_markers}C} and \textbf{~\ref{fig:current_markers}D}.
+We also perform a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
+Factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output.
+As illustrated in Figure~\ref{fig:current_markers}, six markers are defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
+Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate over time; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
+Marker f will be equal to the total recording time if current does not decline enough as in Figure~\ref{fig:current_markers}(c) and (d).
 \begin{figure}[!ht]
@@
     \includegraphics{figs/current_markers.pdf}
-    \caption{Illustration of the six markers of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (A) I\textsubscript{Kto}, (B) I\textsubscript{Kslow1}, (C) I\textsubscript{Kslow2}, and (D) I\textsubscript{Kss}.
+    \caption{Illustration of the six markers of voltage-clamp K\textsuperscript{+} currents that quantify characteristics of the current shape of (a) I\textsubscript{Kto}, (b) I\textsubscript{Kslow1}, (c) I\textsubscript{Kslow2}, and (d) I\textsubscript{Kss}.
 All currents are simulated for illustration, and the labels refer to a) the current magnitude 10 ms after voltage is applied, b) 25\% of the total recording time has elapsed, c) 50\%, d) 75\%, e) the peak magnitude, and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.}
@@
 
-We adopted a fractional factorial design of 1024 runs for I\textsubscript{Kto}, which results in resolution VIII.
+A fractional factorial design of 1024 runs is adopted for I\textsubscript{Kto}, which results in resolution VIII.
 The resolution ensures that the main effects and 2-/3-factor interactions are strongly clear.
 Full factorial designs are used for the other three currents.
-The marker points are evaluated at each design point and factorial effects are calculated using the linear model in Equation~\ref{eq:fact_effect} for each current.
+The marker points are evaluated at each design sample and factorial effects are calculated via the linear model in \eqref{eq:fact_effect} for each current.
+The least squares method is used to estimate $\beta_{i}^{k}$.
+Then, half-normal plots are drawn to test significance of the estimated factorial effects.
 \begin{equation}
-    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\},
     \label{eq:fact_effect}
+    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\}.
 \end{equation}
-The least squares method is used to estimate $\beta_{i}^{k}$. Then, half-normal plots are drawn to test significance of the estimated effects.
 
-\subsection{Variability Visualization and Low-dimensional Embedding}
+\subsection{Low-dimensional Embedding}
 Most cardiac electrophysiology studies utilize statistical tests for the mean to support hypotheses; in turn, the majority of ion channel models are based on population-averaged data.
-However, fine-grained analyses that consider the variability of cells can provide new insights into data.
-For example, it is possible there are differences between control and disease-altered groups but also within the groups according to characteristics of cells such as size or age.
-Hence, there is an urgent need to develop a tool to investigate cellular variability.
-Because the proposed approach calibrates models for the individual dataset from each cell, it allows us to quantify and visualize cellular variability in the tuned parameters.
-Histograms are used in this study to visualize distributions of calibration parameters and the effects of disease-related perturbations on the parameter distributions.
-Although a histogram is a simple yet powerful tool, one of the limitations of a histogram is that it cannot visualize multiple variables more than three on the same plot.
-It is desirable to inspect how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy control and disease groups.
-Low-dimensional embedding methods transform high-dimensional data into a plane or 3D space while preserving relative locations of data points as much as possible so that the structure of data can be visualized.
-\textbf{Figure~\ref{fig:embd}} illustrates how a low-dimensional embedding method is used to map the high-dimensional parameter space of data points from wild-type and disease groups into 3D space.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/embd.pdf}
-    \caption{Diagram illustrating the procedure of low-dimensional embedding to reveal cellular-level structure of the parameter space.}
-    \label{fig:embd}
-\end{figure}
+However, subject-specific analyses that consider the cellular-level variability can provide new insights into data.
+For example, it is possible there are differences between control and experimental groups but also within the groups.
+Hence, there is an urgent need to develop a tool to investigate cell-specfic characteristics.
+Because the proposed approach calibrates models using the dataset for each cell, it allows us to quantify cellular variability in the tuned parameters.
+We adopt low-dimensional embedding that transforms high-dimensional data into a plane or 3D space while preserving relative locations of data points to visualize how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy and diseased groups.
 
-For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \cite{van2008visualizing}.
+For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \citep{van2008visualizing}.
 It has proven to be an effective method for visualizing high-dimensional data.
 t-SNE is a statistical method that constructs two sets of probability distributions $p_{ij}$ and $q_{ij}$ over pairs of data points $i$ and $j$ in a high-/low-dimensional space, respectively.
-These are probabilities of similarities of data points such that neighboring points have a higher probability while dissimilar points have a lower probability.
+These are probabilities of similarities such that neighboring points have a higher probability while dissimilar points have a lower probability.
 We first define the conditional probability of $j$ given $i$:
@@
 \begin{equation}
-    \mathrm{perplexity}(P_{i}) = 2^{H(P_{i})}
+    \mathrm{perplexity}(P_{i}) = 2^{H(P_{i})},
 \end{equation}
@@
 \end{equation}
-Typically, a gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.
+A gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.
 
-\section{Experimental Design}
-More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
-MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
+\section{Experimental Design and Results}
+We apply the proposed framework to our data for investigating the pathophysiology of electrical signaling altered by reduced glycosylation.
+Recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM, achieved through genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced,ednie2019reduced2}.
+MGAT1KO mice developed DCM that deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
 To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
-To be specific, whole-cell I\textsubscript{Ksum} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
-There were 32 observations for each WT and MGAT1KO.
+To be specific, whole-cell I\textsubscript{Ksum} traces were measured in left ventricular apex cardiomyocytes of $\sim$14-week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 to +50 mV) from holding potential of -70 mV.
+There were 31 sets of whole-cell I\textsubscript{K} recordings from different WT cells and 30 from MGAT1KO cells.
+Animals were used and cared for as outlined by the NIH’s Guide for the Care and Use of Laboratory Animals.
+All animal protocols were reviewed and approved by the Wright State University Institutional Animal Use and Care Committee.
 
-We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation.
-It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes.
-A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents.
-All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably.
-Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously.
-For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model.
-Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{Ksum} recordings.
+It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1KO ventricular myocytes \citep{ednie2019reduced}.
+Because of the overlapping inactivation rates of the K\textsubscript{v} isoforms responsible for I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1), a curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in \eqref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents.
+All three component currents were reduced, but the rectifier current I\textsubscript{Kslow} was significantly reduced and slowed notably.
+Although this \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with chronic glycosylation deficiency, it was difficult to determine channel kinetics rigorously.
+For example, curve fitting was not able to provide reliable results due to small current magnitude of the MGAT1KO model, particularly at the lower voltage steps.
+Therefore, here we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{Ksum} recordings.
 This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics concurrently.
 
-\section{Results}
 \subsection{Parameter Screening}
-\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six markers.
+Figure~\ref{fig:half-norm} shows the half-normal plots of factorial effects of parameters on the six markers.
 The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points.
@@
 For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
+In addition, all the maximum conductance variables are included in the calibration parameter set.
 \begin{figure}
@@
     \includegraphics{figs/half-norm.pdf}
-    \caption{Half-normal plots of factorial effects of parameters on the six maker points.}
+    \caption{Half-normal plots of factorial effects for identifying significant parameters in the four current models on (a) Marker a, (b) Marker b, (c) Marker c, (d) Marker d, (e) Marker e, and (d) Marker d.}
     \label{fig:half-norm}
 \end{figure}
-\textbf{Figure~\ref{fig:calib_param}} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
+
+Figure~\ref{fig:calib_param} shows the selected parameters, highlighted in different colors according to their functional roles in channel kinetics.
 We categorized the calibration parameters into four classes: The red represents the voltage-threshold parameters and the green voltage slopes, controlling the voltage dependence, the blue scale factors of kinetic functions, and purple time-constant shifters.
@@
     \includegraphics{figs/calib_param.pdf}
-    \caption{Selected parameters from the sensitivity analysis, highlighted in different colors according to their functional roles in channel kinetics.}
+    \caption{Selected parameters by the sensitivity analysis of (a) I\textsubscript{Kto}, (b) I\textsubscript{Kslow1}, (c) I\textsubscript{Kslow2}, and I\textsubscript{Kss}.
+Parameters are highlighted in different colors according to their functional roles in channel kinetics.}
     \label{fig:calib_param}
@@
 \subsection{Model Fitness}
-We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints.
-A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than utilizing the final objective function value of the calibration routine to report the fitness.
-The objective function, the sum of RMSEs over voltage steps, does not provide a straightforward interpretation that makes it difficult to measure the performance of the models.
-$R^{2}$, defined in Equation~\ref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model.
-However, this interpretation is not applicable, and its value can be negative for nonlinear models because the variance decomposition in Equation~\ref{eq:var_decomp} does not hold anymore.
+We tested various nonlinear optimization algorithms in the proposed model calibration routine that support box constraints, and the BFGS algorithm provided the most reliable results.
+A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than reporting the final objective function value.
+The objective function (i.e., sum of RMSEs) does not provide a straightforward interpretation that makes it difficult to evaluate the calibration performance.
+$R^{2}$, defined in \eqref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model.
+However, this interpretation is not applicable to nonlinear models because the variance decomposition in \eqref{eq:var_decomp} does not hold anymore.
 \begin{align}
-    R^{2} &= 1 - \frac{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2}{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2} \label{eq:r2} \\
-    \sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2 &= \sum_{k=1}^{n}(y^{\mathcal{M}}(k|\theta))-\bar{y})^2+\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2 \label{eq:var_decomp}
+    \label{eq:r2}
+    &R^{2} = 1 - \frac{\sum_{i=1}^{n}(y_{i}-\hat{y_{i}})^2}{\sum_{i=1}^{n}(y_{i}-\bar{y})^2} \\
+    \label{eq:var_decomp}
+    &\sum_{i=1}^{n}(y_{i}-\bar{y})^2 = \sum_{i=1}^{n}(\hat{y_{i}}-\bar{y})^2+\sum_{i=1}^{n}(y_{i}-\hat{y_{i}})^2
 \end{align}
 
-Therefore, a pseudo-R-squared measure, proposed in \citet{li2019prediction} was used to evaluate the model fitness in this study.
+Therefore, a pseudo-R-squared measure, proposed in \cite{li2019prediction} is used to evaluate the model fitness in this study.
 This nonlinear $R^{2}$ measure is based on the concept of linear correction of prediction function, allowing the variance decomposition for nonlinear functions, which guarantees straightforward interpretation and normalized scores between 0 and 1 as in the classical $R^{2}$.
-\textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} present averages of $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively.
-In most cases, more than 80 \% of variances on average in multiple current recordings are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
-The model fitness measures in MGAT1KO are less than WT in general.
+Figure~\ref{fig:model_fitness}(a) and (c) present averages of nonlinear $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively.
+In most cases, more than 90 \% of variances on average are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
+The model fitness measures in MGAT1KO are lower than WT in general.
 It is expected that the small current magnitude of MGAT1KO challenges the calibration process.
-We compared the actual model predictions and experimental I\textsubscript{Ksum} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
-As illustrated in the comparison plot of MGAT1KO at -30 mV clamp voltage, small and noisy current trace results in decreased $R^{2}$, 0.8335.
+We randomly select a cell from each group and compare the actual model predictions and experimental I\textsubscript{Ksum} traces in Figure~\ref{fig:model_fitness}(b) and (d) to validate the model calibration results further.
+As shown in Figure~\ref{fig:model_fitness}(d), MGAT1KO at -30 mV clamp voltage, small and noisy current trace, results in decreased $R^{2}$, 0.8335.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/model_fitness_4half.pdf}
-    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (A) WT and (C) MGAT1KO.
-Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{Ksum} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
+    \includegraphics{figs/model_fitness.pdf}
+    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (a) WT and (c) MGAT1KO.
+Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{Ksum} recordings (in pink) at six voltage steps of (b) cell 6 in WT and (d) cell 12 in MGAT1KO.}
     \label{fig:model_fitness}
@@
 
-\subsection{Kinetics Modeling} \label{s:results.kinetics}
-The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior.
-\textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
-There is a very slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}).
-Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
-$\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV.
-$\tau_{i}^{(1)}$ for MGAT1KO shifted toward left.
-The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
+\subsection{Prediction of K\textsubscript{v} Activities}
+The \textit{in-silico} modeling predicts that chronic reduction in cariomyocyte N-glycosylations results in significant changes in channel steady states and kinetic behaviors.
+The most notable result in the \textit{in-vitro} experimental data was the significant reduction in the current density in response to a 50-mV test potential.
+To verify this observation, Figure~\ref{fig:current_density} presents the prediction results of the current-density relationship of the four K\textsuperscript{+} currents.
+The model predicts the compatible result showing the I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} densities are significantly reduced in MGAT1KO cardiomyocytes, while there are no remarkable differences in I\textsubscript{Kss}.
+Note that in the \textit{in-vitro} experiments, bi-exponential fitting was used, in which the two delayed rectifier currents are combined (I\textsubscript{Kslow}), because it is difficult to reliably fit a tri-exponential function with a relatively short voltage step (4.5 s here) \citep{liu2011dissection}.
+In contrast, although the proposed model-guided data assimilation distinguishes K\textsubscript{v}1.5 and K\textsubscript{v}2.1, it shows compatible prediction results with the experimental data as well as high fitness accuracy ($\sim$90 \% $R^{2}$) \citep{ednie2019reduced}.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/kinetics_ikto.pdf}
-    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
-    \label{fig:kinetics_ikto}
+    \includegraphics{figs/current_density.pdf}
+    \caption{Current-density relationships of the major K\textsubscript{+} currents: (a) I\textsubscript{Kto} (b) I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}.}
+    \label{fig:current_density}
 \end{figure}
 
-\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}.
-The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}C} and \textbf{D}).
-Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO (see \textbf{Figure~\ref{fig:kinetics_rec}A}).
-SSI is also shifted $\sim$ 10 mV to the right (see \textbf{Figure~\ref{fig:kinetics_rec}B}).
-The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO.
-Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}E}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
-It is worth mentioning that the proposed method extrapolates $\tau_{i}^{(2)}$ that is greater than 5 s from the 4.5-s-pulse protocol.
+Another aberrant activity of K\textsuperscript{+} currents was slower inactivation of the delayed rectifier currents in MGAT1KO cardiomyocytes.
+Figure~\ref{fig:time_constants} shows the prediction results of the inactivation time constants of I\textsubscript{Kslow1} and I\textsubscript{Kslow2}.
+The computer models successfully capture this trend, resulting in the time constant of inactivation in for I\textsubscript{Kslow1} ($\tau_{i}^{(\mathrm{Kslow1})}$) being estimated to be significantly slowed in MGAT1KO compared to WT, i.e., $\sim$348.9 ms, as shown in Figure~\ref{fig:time_constants} (a).
+I\textsubscript{Kslow2} is also slowed (see Figure~\ref{fig:time_constants} (b)).
+A polot is not provided here, but there is no significant effect on I\textsubscript{Kto} for membrane potential greater than 0 mV.
+It is worth mentioning that the proposed method was able to extrapolate $\tau_{i}^{(\mathrm{Kslow2})}$ that is greater than 5 s from the 4.5-s-pulse protocol, which shows the generalization power of the \textit{in-silico} modeling.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/kinetics_rec.pdf}
-    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
-    \label{fig:kinetics_rec}
+    \includegraphics{figs/time_constants.pdf}
+    \caption{Inactivation time constants of (a) I\textsubscript{Kslow1}, and (b) I\textsubscript{Kslow2}.}
+    \label{fig:time_constants}
 \end{figure}
 
-We aggregated calibration results into one to analyze the parameters in a collective way.
-It resulted in a $61\times20$ matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters).
-Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point.
-\textbf{Figure~\ref{fig:embd_result}} presents the visualization of the t-SNE embedding of the calibration results.
-It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
+\subsection{Cellular Variability}
+The calibrated parameters are in a 61×20 matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters), and we apply t-SNE to encapsulate this high-dimensional data into three-dimensional space to visualize each data point.
+Figure~\ref{fig:embd_result} (a) presents the visualization of the t-SNE embedding of the calibration results.
+It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups across the entire data, indicating that MGAT1KO cells have higher variance in the aggregated kinetic parameter space.
+Determinants of the covariance matrices of the 3D embedding data to quantify the variability of WT and MGAT1KO in the kinetic parameter space.
+It turns out that WT has 38 \% higher variability than MGAT1KO.
+
+To further investigate it, we generated the histograms of the maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant parameter determining the magnitude of the current (see Figure~\ref{fig:embd_result} (b)), and experimental I\textsubscript{K} density from the \textit{in-vitro} data.
+There is clear skewness in MGAT1KO I\textsubscript{Ksum} density (small variance), while WT I\textsubscript{K} density is more disperse (high variance).
+Calibrated parameters, for example \textit{G}\textsubscript{Kslow1}, likely show smaller variations in MGAT1KO to address this trend in the experimental data.
+It is likely due to significant I\textsubscript{K} reduction in the diseased group combined with the elimination of complex N-glycosylation, thereby minimizing potential variability of N-glycan structures
 \begin{figure}[!ht]
@@
     \includegraphics{figs/embd_result.pdf}
-    \caption{t-SNE embedding of calibration parameters.}
+    \caption{(a) Low-dimensional embedding of calibration parameters into 3D space.
+Blue circles - WT myocytes (n=31); Red triangles - MGAT1KO myocytes (n=30).
+(b) Histogram of the calibrated kinetic parameter for maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant factor determining the magnitude of I\textsubscript{Kslow1}.
+(c) Histogram of experimental I\textsubscript{Ksum} density.}
     \label{fig:embd_result}
@@
 
-\section{Conclusions}\label{s:conclusions}
+\section{Conclusions}
 We propose a novel \textit{concurrent data assimilation} framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
@@
 \section*{Acknowledgements}
-This work was supported in part by the National Science Foundation (MCB-1856132 to HY) and (MCB-1856199, IOS-1146882, and IOS-1660926 to EB), an American Heart Association, Greater Southeast Affiliate Grant-In-Aid (14GRNT20450148 to EB), and Postdoctoral Fellowship (15POST25710010 to AE).
-HK and HY would also like to thank the NSF I/UCRC Center for Health Organization Transformation (CHOT) award NSF IIP-1624727 for the support of their research work.
+Funding information is purposely removed for double-blind review
 } \fi
@@
 \spacingset{1}
-\bibliography{IISE-Trans}
+\bibliography{Bibliography}
 

### 2023-07-10 calibration/opt_4half.m
New plot for the second revision

@@ -5,7 +5,7 @@ warning('off', 'all')
 
 %----- Code arguments & Model information -----%
 group = "wt";
-exp_num = "exp45";
+exp_num = "exp51";
 save_dir = strcat("calib_",exp_num,"_",group);
 mkdir(fullfile(pwd,save_dir))
 

### 2023-07-10 calibration/result_output.m

@@ -775,52 +775,56 @@ set(gca,'FontWeight','bold','LineWidth',1.5)
 clc
 close all
 
-figure('Color','w','Position',[50,50,600,500])
-t = tiledlayout(2,2);
+figure('Color','w','Position',[50,50,600,250])
+t = tiledlayout(1,2);
 xlabel(t,"Voltage (mV)",'FontWeight','bold')
 
-nexttile
-r = 3:12;
-taui_tof_sem = std(taui_tof_wt(:,r),0,1)/sqrt(length(files_wt));
-errorbar(volts(r),mean(taui_tof_wt(:,r),1),taui_tof_sem,'-o','Color','blue','MarkerFaceColor','blue','LineWidth',1.5)
-hold on
-taui_tof_sem = std(taui_tof_ko(:,r),0,1)/sqrt(length(files_ko));
-errorbar(volts(r),mean(taui_tof_ko(:,r),1),taui_tof_sem,'--s','Color','red','MarkerFaceColor','red','LineWidth',1.5)
-hold off
-grid on
-axis tight
-xticks(volts(r))
-ylabel("\tau_{i}^{(Kto)}")
-set(gca,'FontWeight','bold','LineWidth',1.5)
+% nexttile
+% r = 3:12;
+% taui_tof_sem = std(taui_tof_wt(:,r),0,1)/sqrt(length(files_wt));
+% errorbar(volts(r),mean(taui_tof_wt(:,r),1),taui_tof_sem,'-o','Color','blue','MarkerFaceColor','blue','LineWidth',1.5)
+% hold on
+% taui_tof_sem = std(taui_tof_ko(:,r),0,1)/sqrt(length(files_ko));
+% errorbar(volts(r),mean(taui_tof_ko(:,r),1),taui_tof_sem,'--s','Color','red','MarkerFaceColor','red','LineWidth',1.5)
+% hold off
+% grid on
+% axis tight
+% xticks(volts(r))
+% ylabel("\tau_{i}^{(Kto)}")
+% set(gca,'FontWeight','bold','LineWidth',1.5)
 
 nexttile
 r = 3:16;
 taui_kslow1_sem = std(taui_kslow1_wt(:,r),0,1)/sqrt(length(files_wt));
+errorbar(volts(r),mean(taui_kslow1_wt(:,r),1),taui_kslow1_sem,'-o','Color','blue','MarkerFaceColor','blue','LineWidth',1.5)
[…]

### 2023-07-10 writing/draft_new.tex
2nd revision

--- 2023-05-16 writing/draft.tex
@@
 There is a variety of K\textsubscript{v} isoforms and corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}) that contribute to different phases of repolarization.
-Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose and estimate K\textsubscript{v} currents and kinetics.
-Most existing studies separate and make inference of K\textsubscript{v} kinetics via curve-fitting procedures (e.g., tri-exponential fitting), but are limited because: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later.
+Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose into individual K\textsuperscript{+} currents and their characteristics.
+Most existing studies separate and make inference of K\textsubscript{v} activities via curve-fitting procedures, but there are limitations such that: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later.
 Here, we develop a novel concurrent data assimilation method that calibrates biophysics-based subject-specific computer models to decompose and delineate kinetics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
@@
 The AP is a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte, and the AP shape and duration are primarily determined by K\textsubscript{v} isoforms.
+The ability of the heart to pump blood through the body in an appropriate rhythm is controlled by electrical signaling.
+Hence, even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
 As shown in Figure~\ref{fig:ap_currents}, the different phases of AP repolarization in mouse cardiomyocytes are the result of the coordinated activity of a variety of K\textsubscript{v} isoforms (e.g., K\textsubscript{v}4.2, K\textsubscript{v}1.5, and K\textsubscript{v}2.1) and their corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2}).
@@
 First, computer models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
-It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents.
+It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism.
 Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant effects on the current generation.
@@
 Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other VGICs \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-Our \textit{in-vitro} experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced,ednie2019reduced2}.
+Our in-vitro experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
 In fact, these studies also showed that reduced cardiomyocyte complex N-glycosylation is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
@@
 \begin{itemize}
-    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles,
-    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics caused by voltage dependence,
-    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding,
+    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles.
+    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics.
+    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding.
     \item A GUI application is provided to make the proposed method an enabling tool for biomedical scientists.
 \end{itemize}
+The software package of the GUI application, tutorial, and reproducible MATLAB codes for modeling and analysis results are available as supplementary material.
 
@@
 Advancements in laboratory techniques allow researchers to study molecular-level activities in cardiomyocytes via measuring ionic currents conducted through VGICs.
-For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2, ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
-However, \textit{in-vitro} experiments alone are limited to investigate detailed channel activities.
+For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2,ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
+However, in-vitro experiments alone are limited to investigate detailed channel activities.
 There is a need to assimilate experimental data to estimate information that are not able to be observed directly.
@@
 
-However, curve-fitting approaches have limitations in that they only consider the shape of informational functions to infer rather than incorporating underlying kinetics and interactions of various gating processes.
-Further, as curve fitting works for a single curve independently, the traditional approach is less concerned about the dynamics of a set of multiple responses measured from the same cell (e.g., the coincidental activity of multiple K\textsubscript{v} isoforms).
 In contrast, computer models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
@@
 The dynamic nature of the models complicates the calibration process.
-
 Complex nonlinear structures of cardiac computer models also make calibration difficult.
-We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous \textit{in-silico} studies to cope with the complexity of cardiac models.
+We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous in-silico studies to cope with the complexity of cardiac models.
 However, all these studies are based on current-shape parameters via curve fitting, or characteristic curves/statistics from additional curve fitting applied to estimated currents.
@@
 We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
-Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.
+Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical,du2017silico}.
 
@@
 I\textsubscript{Kto,s} (K\textsubscript{v}1.4) is primarily found in septal cardiomyocytes \citep{bondarenko2004computer}.
-In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our \textit{in-vitro} studies focused on the ventricular apex myocytes.
+In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our in-vitro studies focused on the ventricular apex myocytes.
 However, I\textsubscript{Kto} can be easily modified according to the region of ventricular cardiomyocytes.
@@
 Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach.
-In this approach, if all four current models are included , I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
+In this approach, if all four current models are included, I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
 In some cases, it is reduced to a bi-exponential function combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2}.
@@
 
-Unlike the simple current functions, K\textsubscript{v} computer models simulate detailed gating kinetics underlying the currents.
-We develop mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}.
+We developed mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}.
 This type of model consists of two gating variables controlling the channel conductance, and its canonical form is defined by \eqref{eq:hh_ex}
@@
 Equation~\eqref{eq:gv} shows a schematic relationship of this gating kinetics.
+$n$ and $m$ represent the numbers of activation and inactivation gates, which are dependent on a particular K\textsubscript{v} isoform and determined by specific kinetics and properties of that channel.
+In general, the number of gates depends on the complexity of the channel behavior and the level of detail required in the model.
 \begin{align}
@@
 The steady-state values and time constants can be defined directly by functions of voltage $V$ without transition rates in some cases.
-These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters $p_{i}$ that control the behavior of the kinetics of an ion channel.
+These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters $p_{s}$ that control the behavior of the kinetics of an ion channel.
 
@@
 Their kinetics are governed by transition-rate functions from \eqref{eq:ikto_alpha1} to \eqref{eq:ikto_beta2}.
-Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
+Parameters $p_{s}$ for $s=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
 
@@
 Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints.
-By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.
+By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{D}$, which provides two complementary angles to study the real process $\mathcal{R}$.
 
-From this perspective, bi-/tri-exponential functions in \eqref{eq:expl_fitting} serve as $y^{\mathcal{M}}(x|\theta)$ in the curve-fitting approach, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$.
-Note that $\mathrm{A}_{i}(x)$ and $\tau_{i}(x)$ are dependent on input data $x$.
+From this perspective, bi-/tri-exponential function in \eqref{eq:expl_fitting} serves as $y^{\mathcal{M}}(t,v|\theta)$ in the curve-fitting approach, where $\theta=\{\mathrm{A}_{i}(v),\tau_{i}(v)\}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
+$v$ represents voltage.
+Note that $\mathrm{A}_{i}(v)$ and $\tau_{i}(v)$ are dependent on input data $v$, so for each voltage, we need to perform a data assimilation procedure separately.
 In general, multiple input voltage steps are applied to a cardiomyocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics.
-In contrast, computer models of VGICs generate currents for different input voltages given one set of parameters, because they are designed by biophysical principles encoding voltage dependence.
-In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in \eqref{eq:rmse}.
+Suppose the sum of in-silico models of K\textsuperscript{+} currents $\mathrm{I}_\mathrm{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$ serve as a computer model for data assimilation as in \eqref{eq:comp_model_as_insilico}:
+\begin{equation}
+    \label{eq:comp_model_as_insilico}
+    y^{\mathcal{M}}(t,v|\theta) = \sum_{i} \mathrm{I}_{i},
+\end{equation}
+where $\theta$ is the union of kinetic parameters $p_{s}$ for each $\mathrm{I}_{i}$, which are constants.
+In contrast, this computer model generates I\textsubscript{Ksum} for different input voltages given one set of parameters, because the in-silico models are designed by biophysical principles encoding voltage dependence.
+Therefore, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in \eqref{eq:rmse}.
 RMSE evaluates the goodness-of-fit between experimental I\textsubscript{Ksum} and model prediction.
@@
 Cardiac models enable this concurrent data assimilation, because they generate multiple current traces with different input voltages by simulating underlying gating kinetics.
-Note that the proposed method calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data via curve-fitting \citep{du2013silico,du2015statistical, du2017silico, kim2022simulation}.
+Note that the proposed method calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data via curve-fitting \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
 We develop the box-constrained nonlinear optimization routine with multi-random initial points to minimize $\delta$.
@@
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
-This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.
+This work is implemented in MATLAB R2022a.
 
@@
 Full factorial designs are used for the other three currents.
-The marker points are evaluated at each design sample and factorial effects are calculated via the linear model in \eqref{eq:fact_effect} for each current.
-The least squares method is used to estimate $\beta_{i}^{k}$.
+The marker points $\lambda^{k}$ defined as in Figure~\ref{fig:current_markers} are evaluated at each design sample and factorial effects are calculated via the linear model in \eqref{eq:fact_effect} for each current.
+The least squares method is used to estimate $\beta_{s}^{k}$.
 Then, half-normal plots are drawn to test significance of the estimated factorial effects.
@@
     \label{eq:fact_effect}
-    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\}.
+    \lambda^{k} = \beta_{o}^{k} + \sum\limits_{s}\beta_{s}^{k}p_{s}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\}.
 \end{equation}
@@
 For example, it is possible there are differences between control and experimental groups but also within the groups.
-Hence, there is an urgent need to develop a tool to investigate cell-specfic characteristics.
+Hence, there is an urgent need to develop a tool to investigate cell-specific characteristics.
 Because the proposed approach calibrates models using the dataset for each cell, it allows us to quantify cellular variability in the tuned parameters.
@@
     q_{ij} = 
-    \begin{cases}\frac{(1+||y_{i}-y_{j}||^{2})^{-1}}{\sum\limits_{k}\sum\limits_{l \neq k}(1+||y_{k}-y_{l}||^{2})^{-1}}, \ j \neq i, \\
+    \begin{cases}\frac{(1+||z_{i}-z_{j}||^{2})^{-1}}{\sum\limits_{k}\sum\limits_{l \neq k}(1+||z_{k}-z_{l}||^{2})^{-1}}, \ j \neq i, \\
     0, \ j=i,
@@
 We apply the proposed framework to our data for investigating the pathophysiology of electrical signaling altered by reduced glycosylation.
-Recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM, achieved through genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced,ednie2019reduced2}.
+Recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM, achieved through genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
 MGAT1KO mice developed DCM that deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
@@
 All three component currents were reduced, but the rectifier current I\textsubscript{Kslow} was significantly reduced and slowed notably.
-Although this \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with chronic glycosylation deficiency, it was difficult to determine channel kinetics rigorously.
+Although this in-vitro investigation discovered aberrant reductions in K\textsuperscript{+} currents with chronic glycosylation deficiency, it was difficult to determine channel kinetics rigorously.
 For example, curve fitting was not able to provide reliable results due to small current magnitude of the MGAT1KO model, particularly at the lower voltage steps.
@@
 We picked the parameters to be calibrated as a union of the sets of parameters falling above the straight lines on the six half-normal plots.
-For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
+For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
 In addition, all the maximum conductance variables are included in the calibration parameter set.
@@
 
-Therefore, a pseudo-R-squared measure, proposed in \cite{li2019prediction} is used to evaluate the model fitness in this study.
+Therefore, a pseudo-R-squared measure, proposed in \citep{li2019prediction} is used to evaluate the model fitness in this study.
 This nonlinear $R^{2}$ measure is based on the concept of linear correction of prediction function, allowing the variance decomposition for nonlinear functions, which guarantees straightforward interpretation and normalized scores between 0 and 1 as in the classical $R^{2}$.
@@
 As shown in Figure~\ref{fig:model_fitness}(d), MGAT1KO at -30 mV clamp voltage, small and noisy current trace, results in decreased $R^{2}$, 0.8335.
+The fitting was performed on Intel 4-core Xeon E7-4830 processors.
+Without parallel computing on multiple cores, each cell took an average of $438.1 \pm 6.8$ s to complete the data assimilation procedure for 5 random initial points.
+It took an average of $333.8 \pm 6.0$ s with multicore processing, which is ~31 \% efficiexnt.
 \begin{figure}[!ht]
@@
 \subsection{Prediction of K\textsubscript{v} Activities}
-The \textit{in-silico} modeling predicts that chronic reduction in cariomyocyte N-glycosylations results in significant changes in channel steady states and kinetic behaviors.
-The most notable result in the \textit{in-vitro} experimental data was the significant reduction in the current density in response to a 50-mV test potential.
+The in-silico modeling predicts that chronic reduction in cardiomyocyte N-glycosylation results in significant changes in channel steady states and kinetic behaviors.
+The most notable result in the in-vitro experimental data was the significant reduction in the current density in response to a 50-mV test potential.
 To verify this observation, Figure~\ref{fig:current_density} presents the prediction results of the current-density relationship of the four K\textsuperscript{+} currents.
 The model predicts the compatible result showing the I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} densities are significantly reduced in MGAT1KO cardiomyocytes, while there are no remarkable differences in I\textsubscript{Kss}.
-Note that in the \textit{in-vitro} experiments, bi-exponential fitting was used, in which the two delayed rectifier currents are combined (I\textsubscript{Kslow}), because it is difficult to reliably fit a tri-exponential function with a relatively short voltage step (4.5 s here) \citep{liu2011dissection}.
+Note that in the in-vitro experiments, bi-exponential fitting was used, in which the two delayed rectifier currents are combined (I\textsubscript{Kslow}), because it is difficult to reliably fit a tri-exponential function with a relatively short voltage step (4.5 s here) \citep{liu2011dissection}.
 In contrast, although the proposed model-guided data assimilation distinguishes K\textsubscript{v}1.5 and K\textsubscript{v}2.1, it shows compatible prediction results with the experimental data as well as high fitness accuracy ($\sim$90 \% $R^{2}$) \citep{ednie2019reduced}.
@@
 I\textsubscript{Kslow2} is also slowed (see Figure~\ref{fig:time_constants} (b)).
-A polot is not provided here, but there is no significant effect on I\textsubscript{Kto} for membrane potential greater than 0 mV.
-It is worth mentioning that the proposed method was able to extrapolate $\tau_{i}^{(\mathrm{Kslow2})}$ that is greater than 5 s from the 4.5-s-pulse protocol, which shows the generalization power of the \textit{in-silico} modeling.
+The inactivation time constants are important because they determine how slowly I\textsubscript{Kslow1} and I\textsubscript{Kslow2} inactivate and, in turn, affect the repolarization of the AP significantly.
+$\tau_{i}^{(Kslow1)}$ show higher variability than other currents to address most of the decaying portion of I\textsubscript{Ksum}.
+This is likely due to the fact that while K\textsubscript{v}4.2 and K\textsubscript{v}2.1 are O-glycosylated, K\textsubscript{v}1.5 possess a single occupied N-linked glycosylation.
+Hence, I\textsubscript{Kslow1} is more affected by reduced N-glycosylation.
+A plot is not provided here, but there is no significant effect on I\textsubscript{Kto} for membrane potential greater than 0 mV.
+It is worth mentioning that the proposed method was able to extrapolate $\tau_{i}^{(\mathrm{Kslow2})}$ that is greater than 5 s from the 4.5-s-pulse protocol, which shows the generalization power of the in-silico modeling.
 \begin{figure}[!ht]
@@
 
-To further investigate it, we generated the histograms of the maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant parameter determining the magnitude of the current (see Figure~\ref{fig:embd_result} (b)), and experimental I\textsubscript{K} density from the \textit{in-vitro} data.
+To further investigate it, we generated the histograms of the maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant parameter determining the magnitude of the current (see Figure~\ref{fig:embd_result} (b)), and experimental I\textsubscript{K} density from the in-vitro data.
 There is clear skewness in MGAT1KO I\textsubscript{Ksum} density (small variance), while WT I\textsubscript{K} density is more disperse (high variance).
@@
     \caption{(a) Low-dimensional embedding of calibration parameters into 3D space.
-Blue circles - WT myocytes (n=31); Red triangles - MGAT1KO myocytes (n=30).
+Blue circles - WT myocytes ($n=31$); Red triangles - MGAT1KO myocytes ($n=30$).
 (b) Histogram of the calibrated kinetic parameter for maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant factor determining the magnitude of I\textsubscript{Kslow1}.
@@
 
-\section{Conclusions}
-We propose a novel \textit{concurrent data assimilation} framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
-In this study, we propose a novel \textit{concurrent data assimilation} method for modeling K\textsubscript{v} kinetics from multiple I\textsubscript{v} recordings measured from the same cell with different conditions for investigating disease-altered functionalities in the ion channels.
-This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights.
-Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in
-clamp voltages.
-The proposed method focuses on more comprehensive modeling of underlying channel kinetics of K\textsubscript{v} isoforms than traditional approaches.
-Experimental results show the strong potential of the proposed framework to discover channel kinetics from various angles.
-Predictions of kinetics variables to voltage relationships provide information about changes in the voltage dependence of channels.
-In addition, calibrated parameters of the computer models present a micro-level analysis of channel gating.
-The embedding method allows visualizing the variability of each cell.
-The proposed method focuses on the mouse model, but the same framework is applicable to other species easily with modifications to model structures.
-In summary, the proposed framework paves a new way for investigating the electrical signaling of potassium channels and heart diseases.
+\section{Discussion and Conclusions}
+Repolarization is a complex process that involves various K\textsubscript{v} isoforms.
+It is critical to understand the unique properties and functional roles of each K\textsubscript{v} to investigate pathological mechanisms of diseased cardiomyocytes that contribute to fatal heart diseases.
+However, current laboratory techniques, such as whole-cell patch-clamp recording, are not able to measure individual activities of K\textsubscript{v} isoforms reliably that activate, inactivate, and close at overlapping times during recordings, except through the attempt to remove K\textsubscript{v} isoform activity through less-than-fully-specific pharmacological intervention.
+Thus, only the sum of the different K\textsubscript{v} isoform activities can be measured as a single current trace, I\textsubscript{Ksum}.
+Hence, it is necessary to decompose I\textsubscript{Ksum} into individual K\textsuperscript{+} currents and estimate their channel activity via data assimilation.
+This paper presents a subject-specific concurrent data assimilation method for learning K\textsubscript{v} activities using multiple I\textsubscript{Ksum} recordings simultaneously for each cell.
+A case study is provided using our in-vitro experimental data of mouse cardiomyocyte I\textsubscript{K} in control conditions (WT) and under conditions of reduced complex N-glycosylation (MGAT1KO).
+We evaluate the calibration results using an adjusted $R^2$ measure for nonlinear models that preserves the interpretability of the classical $R^2$ based on the variance decomposition for linear models.
+Experimental results show the proposed method explains more than 90\% of variances by calibrated models in most cases (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
 
-\if0\blind{
-\section*{Acknowledgements}
-Funding information is purposely removed for double-blind review
-} \fi
+In addition to achieving a high degree of goodness-of-fit, it is important to determine the source of uncertainty and variability in model predictions to build trustworthy models.
+Until recently, the conventional approach to developing cardiac models and fitting model parameters has involved using single values.
+As a result, most mathematical models currently in use only provide point estimates without quantifying uncertainty \citep{johnstone2016uncertainty}.
+Although variances across cells are provided in the model predictions of current density and inactivation times in Figure~\ref{fig:current_density} and Figure~\ref{fig:time_constants}, it remains elusive whether the variability arises from model uncertainty or true cellular variability in the data.
+Statistical methods for uncertainty quantification (UQ) in cardiac models, such as Bayesian inferences or Gaussian process emulators, have been proposed \citep{johnstone2016uncertainty,coveney2020sensitivity}.
+Exploring the incorporation of UQ techniques into data assimilation is a topic for future research.
+A GUI application is provided as an enabling tool for biomedical scientists without full expertise in modeling and computational analysis.
+
+The estimation of kinetics provides novel insights into potential mechanisms by which specific K\textsubscript{v} isoforms contribute to the overall reduction in I\textsubscript{K} observed in in-vitro experiments following chronic reductions in N-glycosylation.
+Thanks to the cell-specific approach, prediction uncertainty is quantified, and error bars are provided in the prediction results.
+Further, calibrated parameters are visualized via low-dimensional embedding that allows for encapsulating calibrated parameters in 3D space and, in turn, visualizing the variability across cells.
+WT cells show higher variability than MGAT1KO myocytes, which is likely due to significant I\textsubscript{K} reduction in the diseased group, combined with the elimination of complex N-glycosylation, thereby minimizing potential variability of N-glycan structures.
+The proposed method and pertinent software show strong potential for studying K\textsubscript{v} kinetics in various heart diseases.
+However, there are still some limitations and challenges that need to be addresses in future work.
+One of the main limitations of this study is that it does not address the interactions of K\textsubscript{v} with other major VGICs, namely Na\textsubscript{v} and Ca\textsubscript{v}, and cellular-level processes.
+Thus, causal inference of the effects of chronic glycosylation deficiency is limited.
+We plan to develop an integrative modeling method for learning activities of major VGICs together.
+
+\section*{Funding}
+This work is supported by the National Science Foundation (MCB-1856132 to HY and HK), (MCB-1856199 to EB and AE), (IOS-1146882 and IOS-1660926 to EB), and the American Heart Association (Postdoctoral Fellowship 15POST25710010 to AE).
+HY and HK would also like to thank the NSF I/UCRC Center for Health Organization Transformation (CHOT) award NSF IIP-1624727 for the support of their research work.
+Any opinions, findings, or conclusions found in this research are those of the authors and do not necessarily reflect the views of the sponsors.
+
+\section*{Disclosure of Interest}
+The authors report there are no competing interests to declare.
+
+\section*{Consent and Approval}
+This research does not involve any human participants or data. Therefore, no consent or approval was required for this study.
 

### 2023-07-10 writing/draft_old.tex
2nd revision

--- 2022-09-28 writing/draft.tex
@@
 	{
-		\title{\bf Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
-		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
-		$^a$ Department of Industrial and Manufacturing Engineering, \\
-		The Pennsylvania State University, State College, PA USA \\
-        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
-        Wright State University, Dayton, OH USA}
+		\title{\bf Concurrent Data Assimilation for Model-guided Learning of Cardiac Potassium Channel Activities}
+		\author{Author information is purposely removed for double-blind review}
 		\date{}
@@
 \begin{abstract}
-Potassium channels (K\textsubscript{v}) are responsible for repolarizing the cardiac action potential.
-Different K\textsubscript{v} isoforms contribute to various phases of repolarization.
-In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1).
-A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities can be measured in the form of currents (I\textsubscript{K}).
-Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure.
-However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually.
-There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms.
-We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics.
-We apply the proposed framework to glyco-altered I\textsubscript{K} recordings from our previous studies.
-Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups.
-This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
+Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes.
+There is a variety of K\textsubscript{v} isoforms and corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}) that contribute to different phases of repolarization.
+Because only the sum of their activities can be measured in the form of currents (I\textsubscript{Ksum}), there is a need to decompose into individual K\textsuperscript{+} currents and their characteristics.
+Most existing studies separate and make inference of K\textsubscript{v} activities via curve-fitting procedures, but there are limitations such that: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces, which does not discern the underlying kinetics and interactions; 2) I\textsubscript{Ksum} traces can only be fitted for one clamp voltage at each time, and estimated information is analyzed in a population-averaged way later.
+Here, we develop a novel concurrent data assimilation method that calibrates biophysics-based subject-specific computer models to decompose and delineate kinetics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
+The proposed method is evaluated and validated with whole-cell I\textsubscript{Ksum} recordings from wild-type and chronically glycosylation-deficient cardiomyocytes.
+Experimental results show that the proposed method effectively handles multiple responses in the voltage-clamp protocol and describes the glycosylation-conferred perturbations observed experimentally to various K\textsubscript{v} isoforms.
+In addition, we develop a graphical-user-interface (GUI) application that provides an enabling tool to biomedical scientists.
+The proposed method and pertinent software are shown to have strong potential to study K\textsubscript{v} kinetics in various heart diseases.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} Data assimilation, model calibration, potassium channels, kinetics modeling, glycosylation
+{\it Keywords:} Data assimilation, in-silico modeling, model calibration, cardiac potassium channels, congenital disorders of glycosylation
 
@@
 \section{Introduction}
-Cardiac voltage-gated potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential (AP) in cardiomyocytes.
-They are indispensable in the electrical conduction system of the heart.
-Even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
-Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics,nagai2016aberrant}.
-Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression.
-Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
-In addition, we verified our experimental results via \textit{in-silico} studies and estimate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
-\textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases, which will be referred to as the data assimilation framework that combines experimental data with mathematical/computational models.
+Potassium channels (K\textsubscript{v}) play critical roles in the electrical conduction system of the heart, particularly in the repolarization phase of the action potential (AP).
+The AP is a change of membrane potential over time, representing the net electrical activity in a cardiomyocyte, and the AP shape and duration are primarily determined by K\textsubscript{v} isoforms.
+The ability of the heart to pump blood through the body in an appropriate rhythm is controlled by electrical signaling.
+Hence, even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}.
+As shown in Figure~\ref{fig:ap_currents}, the different phases of AP repolarization in mouse cardiomyocytes are the result of the coordinated activity of a variety of K\textsubscript{v} isoforms (e.g., K\textsubscript{v}4.2, K\textsubscript{v}1.5, and K\textsubscript{v}2.1) and their corresponding currents (e.g., I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2}).
+There are other major voltage-gated ion channels (VGICs), such as for Na\textsuperscript{+} (Na\textsubscript{v}) and Ca\textsuperscript{2+} (Ca\textsubscript{v}), that also contribute to the AP.
+The Na\textsubscript{v} are primarily responsible for the AP upstroke in the depolarization phase, while the Ca\textsubscript{v} control the cellular contraction.
+Aberrant activities of ion channels can significantly impact the AP and lead to fatal arrhythmias.
+Using whole-cell voltage-clamp recording methods, only the collective activities of K\textsubscript{v} isoforms can be measured reliably as the sum of all K\textsuperscript{+} currents (I\textsubscript{Ksum}), and these K\textsubscript{v} isoforms have overlapping biophysical properties (i.e., voltage dependence of gating and kinetics), which complicate the ability to separate one type of I\textsubscript{K} from another \citep{brouillette2004functional}.
+K\textsubscript{v} activities can be altered in various cardiomyopathies \citep{tristani2001molecular,giudicessi2012potassium}, thus there is an urgent need to decompose I\textsubscript{Ksum} more rigorously into individual K\textsuperscript{+} currents and estimate gating kinetics of K\textsubscript{v} isoforms to understand their pathological roles by comparing their activities in healthy versus diseased hearts and cardiomyocytes.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/overall_diagram.pdf}
-    \caption{Conceptual diagram to illustrate a data assimilation framework.}
-    \label{fig:framework_diagram}
+    \includegraphics{figs/ap_currents.pdf}
+    \caption{(a) Ventricular action potential in mouse cardiomyocytes and (b) diverse K\textsubscript{v} isoforms and their currents contributing to the AP shape.}
+    \label{fig:ap_currents}
 \end{figure}
 
-However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}.
-In contrast, there is a predominant isoform for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}.
-Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}.
-Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments.
-Typically, a form of curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}.
-Then the shape parameters are used to calibrate computer models \citep{kim2022simulation}, or kinetics information is added to the data assimilation process with additional curve fitting if applicable \citep{du2017silico}.
-However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately.
-These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}.
-Therefore, there is an urgent need to develop new data assimilation methods and tools that delineate the kinetics of K\textsubscript{v} isoforms for better understanding their roles in heart diseases.
+Transgenic mouse models are the most commonly used animal models that have provided insights into cardiac research despite the differences between rodents and humans; the isoforms expressed in both small and large mammals share many similarities \citep{milani2014small}.
+Most existing studies using mouse models separate individual K\textsuperscript{+} currents mathematically via a curve-fitting procedure \citep{costantini2005homeodomain,ednie2019reduced2,teng2022tmem65}, assuming the I\textsubscript{Ksum} current trace represents the summation of exponential decay functions (two or three, usually) and a constant term; each represents the shape of individual K\textsuperscript{+} currents.
+However, the ability to scrutinize the kinetics and dynamics of individual K\textsubscript{v} isoforms is limited by this traditional method because: 1) curve-fitting decomposition only relies on the shape of K\textsuperscript{+} current traces as discreet exponential functions which does not discern the underlying kinetics and interactions of K\textsubscript{v} isoforms; 2) I\textsubscript{Ksum} traces can only be decomposed for one clamp voltage at a time, and estimated information from each fitting is analyzed in a population-averaged way later.
+New methods are urgently needed to simultaneously delineate the underlying kinetics of multiple I\textsubscript{Ksum} traces from the same cell concurrently for a better understanding of K\textsubscript{v} channels and their roles as well as cellular variability in diseased cardiomyocytes.
 
-This paper presents a novel approach to data assimilation for analyzing whole-cell K\textsubscript{v} current traces (I\textsubscript{K}) and modeling the underlying kinetics of K\textsubscript{v} isoforms.
-A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}.
-First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting.
-Computer models of ion channels are a system of differential equations consisting of gating variables and kinetic functions that continuously simulate ion channel activities over time.
-Parameters in kinetic functions control the gating of ion channels and, in turn, determine current outputs.
-Second, a sensitivity analysis was performed to screen out parameters that marginally influence the model outputs.
-Third, we developed a model calibration that minimizes discrepancy between model outputs and experimental data generated from multiple protocols.
-Last, we made estimations and inferences of impacts of disease-related perturbations on potassium channel kinetics and used low-dimensional embedding to show cellular variability.
-Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
-\begin{figure}
-    \centering
-    \includegraphics{figs/flow_chart.pdf}
-    \caption{Flowcharts of (A) the current approach and (B) the proposed method.}
-    \label{fig:flow_chart}
-\end{figure}
+This paper presents a new approach of subject-specific concurrent data assimilation for model-guided learning of K\textsubscript{v} isoforms.
+First, computer models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents.
+It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism.
+Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant effects on the current generation.
+Several markers are defined that best represent the characteristics of the K\textsuperscript{+} currents in different perspectives.
+Third, a calibration routine is proposed to adjust parameters selected by the sensitivity analysis to couple \textit{in-silico} models with \textit{in-vitro} data of I\textsubscript{Ksum} recordings from the same cell concurrently.
+Last, low-dimensional embedding is utilized to visualize calibrated parameters in a collective way across the cells.
+The proposed methodology is evaluated and validated by comparing I\textsubscript{Ksum} recordings on healthy control versus chronically glycosylation-deficient cardiomyocyte.
+Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other VGICs \citep{ohtsubo2006glycosylation,ednie2012modulation}.
+Our in-vitro experiments showed that chronic reduction in cardiomyocyte N-glycosylation modulates VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}.
+In fact, these studies also showed that reduced cardiomyocyte complex N-glycosylation is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}.
 
-The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases.
+Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} currents with underlying kinetics and successfully captures cellular variability.
+For example, it verifies the experimental data that showed the significant reduction in the current magnitude and elongated decaying for the glycosylation-deficient cardiomyocyte group.
+Additionally, computational modeling estimates channel kinetics, which was not possible in the electrophysiological experiments alone.
+Cellular variability is visualized by low-dimensional embedding of calibration parameters into 3D space, which shows one of the powers of systematic computer models that allow one to predict how changes in VGICs impact cellular functions.
+Further, we develop a graphical-user-interface (GUI) application to make the proposed method accessible to biomedical scientists for investigating K\textsubscript{v}-related channelopathy that is available as online supplemental material.
+The proposed method shows strong potential for modeling the kinetics and gating mechanism of K\textsubscript{v} to study heart diseases.
 Our contributions are summarized as follows:
 \begin{itemize}
-    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
-    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
-    \item The proposed approach builds separate K\textsubscript{v} models and calibrates them for each cell, which allows analyzing the cellular variability. 
-    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
+    \item We model the underlying kinetics of K\textsubscript{v} isoforms from I\textsubscript{Ksum} recordings, which enables to simulate individual K\textsuperscript{+} currents based on biophysical principles.
+    \item We develop a subject-specific concurrent data assimilation routine that learns I\textsubscript{Ksum} recordings measured from the same cell at multiple membrane potentials simultaneously rather than a single I\textsubscript{Ksum} trace independently to model cellular-level dynamics.
+    \item The prediction uncertainty is quantified and the distribution of kinetic parameters is visualized using low-dimensional embedding.
+    \item A GUI application is provided to make the proposed method an enabling tool for biomedical scientists.
 \end{itemize}
-The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.
+The software package of the GUI application, tutorial, and reproducible MATLAB codes for modeling and analysis results are available as supplementary material.
 
 \section{Research Background}
-\subsection{Computational Modeling of the Heart}
-Computer models of the heart have been extensively studied and used in cardiology to understand the mechanism of the organ, investigate fatal diseases, and develop drugs.
-Models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
-Mathematical and computational implementations of the heart allow studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
+\subsection{Data Assimilation and Calibration of Cardiac Models}
+The functionality of the heart to pump blood through the body is controlled by electrical signals generated by the organ itself.
+Cardiac electrophysiology has contributed significantly to our understanding of the heart and disease-related modifications, such as congenital disorders of glycosylation (CDG).
+Advancements in laboratory techniques allow researchers to study molecular-level activities in cardiomyocytes via measuring ionic currents conducted through VGICs.
+For example, whole-cell current recordings are used to show the functions of a certain gene that encodes $\alpha$ subunits of Na\textsubscript{v} in human cardiomyocytes, leading to a small but inherent and chronic risk of acquired arrhythmia \citep{splawski2002variant}; pathophysiological roles of reduced sialylation impacting Na\textsubscript{v} and K\textsubscript{v} activities in mouse cardiomyocytes \citep{ednie2015sialicNav2,ednie2015sialicKv}; modulation of O-glycosylation causing aberrant K\textsubscript{v} activities \citep{schwetz2011sialic}.
+However, in-vitro experiments alone are limited to investigate detailed channel activities.
+There is a need to assimilate experimental data to estimate information that are not able to be observed directly.
+Traditionally, a curve-fitting method is used to make inferences, which assumes a functional form of information to estimate (e.g., individual currents of K\textsubscript{v} isoforms) \citep{liu2011dissection}.
+
+In contrast, computer models consist of mathematical equations describing biophysical properties and physiological functions compatible with experimental observations.
+Mathematical and computational modeling allows for studying the heart in a quantitative and predictive way \citep{whittaker2020calibration}.
+Computer models, coupled with experimental observations, provide integrative insights into the data by calibrating model parameters \citep{winslow2011integrative}.
 For example, models can be used to interpolate processes not directly observed in experiments and extrapolate to novel conditions such as disease-related perturbations \citep{rodriguez2010systems}.
-With the advancement of sensing technology and data analysis, computational cardiology is expanding its field.
-Personalized heart models pave a new way for monitoring and managing the cardiac health of individuals by characterizing individual heart signals from the large-scale IoT network and calculating dissimilarities to detect potential cardiac risk \citep{yang2020network}.
-Hence, cardiac computer models provide a great opportunity to mitigate heart disease, the number one cause of death in the U.S.
-\citep{murphy2021mortality}.
+From an engineering perspective, computer models of ion channels are dynamic system simulations that represent continuously changing gating kinetics.
+The dynamic nature of the models complicates the calibration process.
+Complex nonlinear structures of cardiac computer models also make calibration difficult.
+We developed statistical metamodeling and sequential design \citep{du2015statistical}, nonlinear optimization algorithms \citep{du2013silico,du2017silico}, and a heuristic optimization method \citep{kim2022simulation} in our previous in-silico studies to cope with the complexity of cardiac models.
+However, all these studies are based on current-shape parameters via curve fitting, or characteristic curves/statistics from additional curve fitting applied to estimated currents.
+In addition, only population-average models are derived and calibrated via learning the entire dataset for each healthy and diseased group.
+Little work has been done to leverage computer models to calibrate and delineate kinetic dynamics of K\textsubscript{v} isoforms with multiple voltage-clamp responses simultaneously.
 
-The functionality of the heart to pump blood through the body is controlled by electrical signals generated in the heart itself.
-Hence, the electrical activities and waves spreading through the whole heart have been studied considerably.
-For examle, \citet{yang2016whole} a new reaction-diffusion and pattern recognition method to model spatiotemporal dynamics of electrical waves propagating throughout the heart surface.
-Sensing methods, such as vectorcardiogram (VCG) and electrocardiogram (ECG), measure the electrical activities of the heart.
-Various analytics methods have been proposed to cope with their complex nonlinear dynamics in different heart regions over time.
-\citet{chen2019heterogeneous} developed a heterogeneous recurrence method that characterizes spatiotemporal patterns in disease-altered multi-channel cardiac signals.
-Nonlinear recurrence analysis captures and quantifies topological relationships of recurring patterns in signals, which has been proven to be an effective tool for cardiac signals that are naturally recurring over time \citep{yang2010multiscale,meyers2020cross}.
-
-However, to perform a fine-grained analysis, it is necessary to study the cellular-level activities of the heart as opposed to the organ-level signals.
-Voltage-gated ion channels (VGICs) are membrane proteins in cardiomyocytes that are the functional unit of the electrical activities of cardiac muscle cells.
-The whole-cell patch clamp recording is a powerful experimental technique for investigating the electrical properties of excitable cells such as cardiomyocytes and neurons.
-It allows measuring ionic currents conducted through designated VGICs and the action potential (AP).
-Although the laboratory technique reveals essential properties of VGICs and cardiomyocytes, the detailed gating mechanism remains ambiguous.
-For example, it is hard to determine rigorously how different ion channels collectively lead to changes in the AP.
-Computer models, coupled with experimental observations, provide integrative insights into the data, and they have become an essential part of studying the heart, as illustrated in Figure~\ref{fig:framework_diagram} \citep{winslow2011integrative}.
-
-Hence, modeling VGICs involves a data assimilation procedure that combines experimental data with computer models.
-Traditionally, computer models are calibrated to characteristic curves estimated from ionic-current recordings \citep{colman2018description}.
-These estimation processes include curve fitting, which assumes a functional form of the characteristic curve and finds shape parameters best fit data.
-However, curve fitting only considers a single current recording at each time independently, so it cannot capture cellular-level dynamics from the current recordings by changes in input stimuli, i.e., voltage-clamp protocols.
-It raises an additional challenge for analyzing I\textsubscript{K} recordings, because one-more curve fitting is used to separate individual current traces generated by various K\textsubscript{v} isoforms.
-I\textsubscript{Ksum} traces are decomposed only based on the assumption of the shapes of the component currents but not the underlying biophysical gating kinetics \citep{kim2022simulation}.
-Therefore, there is an urgent need to develop a novel data assimilation method for modeling and calibrating K\textsubscript{v} kinetics with considering cellular dynamics.
-
-\subsection{Altered Glycosylation and Heart Diseases}
+\subsection{Congenital Disorders of Glycosylation}
 Protein glycosylation is one of the most abundant and diverse forms of co/posttranslational modifications that impact essential protein functions, such as modulation of receptor or ion channel activities \citep{ohtsubo2006glycosylation,ednie2012modulation}.
-A growing number of studies has shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ohtsubo2006glycosylation,ednie2019reduced2}.
-It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly have modest reductions in protein glycosylation, present with cardiac deficits and idiopathic DCM \citep{marques2017cardiac}.
-However, the underlying pathological mechanism still remains elusive.
-
+A growing number of studies have shown the association between altered glycosylation and heart diseases, such as dilated cardiomyopathy (DCM) and hypertrophic cardiomyopathy \citep{ednie2019reduced2,ohtsubo2006glycosylation}.
+It is reported that up to 20\% of patients with congenital disorders of glycosylation (CDG), who commonly show modest reductions in protein glycosylation, present with cardiac deficits, including idiopathic DCM \citep{marques2017cardiac}.
+However, uncovering the underlying pathological mechanisms still remains elusive.
 We have investigated how regulated glycosylation contributes to heart failure in the context of electrophysiology.
 Electrical signaling is orchestrated activities of a variety of ion channels and transporters.
-Voltage-gated ion channels (VGICs) are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
+VGICs are heavily glycosylated, with ~30\% of the channel mass consisting of N-/O-linked glycans \citep{ednie2012modulation}.
 Glycosylation is a multi-step process and usually ends with sialic acid added.
-We also reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
-Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical, du2017silico}.
-However, these studies also rely on characteristic curves and statistics estimated from curve fitting.
+We reported that a saturating, electrostatic effect of negatively charged sialic attached to the terminal of N-/O-glycan branches significantly altered electrical signaling in Na\textsubscript{v} \citep{ednie2013sialicNav1,ednie2015sialicNav2} as well as K\textsubscript{v} \citep{ednie2015sialicKv}.
+Computational modeling has been used to further investigate the functional role of reduced sialylation in Na\textsubscript{v} and K\textsubscript{v} \citep{du2015statistical,du2017silico}.
 
-\section{Methods}
+\section{Research Methodology}
 \subsection{Computer Models of Potassium Channel Isoforms}
-Mouse models have been very useful methods for studying human cardiac electric signaling \citep{nerbonne2004studying,milani2014small,kim2022irx5}.
-Computer models of mouse cardiomyocytes simulate important biophysical properties of ion channels and molecular dynamics in producing the AP.
-In this paper, the AP is modeled as a differential equation of transmembrane currents and stimulus current I\textsubscript{stim} as below in Equation~\ref{eq:ap} where $C_{m}$ is the membrane capacitance, $t$ is time.
-\begin{equation}
-    \label{eq:ap}
-    \begin{split}
-    -C_{m}\frac{dV}{dt} &= \mathrm{I}_{\mathrm{CaL}}+\mathrm{I}_{\mathrm{p}(Ca)}+\mathrm{I}_{\mathrm{NaCa}}+\mathrm{I}_{\mathrm{Cab}}+\mathrm{I}_{\mathrm{Na}}+\mathrm{I}_{\mathrm{Nab}}+\mathrm{I}_{\mathrm{NaK}}+\mathrm{I}_{\mathrm{Cl},Ca} \\
-    &+\mathrm{I}_{\mathrm{Kto}}+\mathrm{I}_{\mathrm{Kslow1}}+\mathrm{I}_{\mathrm{Kslow2}}+\mathrm{I}_{\mathrm{Kss}}+\mathrm{I}_{\mathrm{Ks}}+\mathrm{I}_{\mathrm{Kr}}+\mathrm{I}_{\mathrm{K1}}+\mathrm{I}_{\mathrm{stim}}
-    \end{split}
-\end{equation}
-There are 15 transmembrane currents in this AP model: the L-type calcium current (I\textsubscript{CaL}), the calcium pump current (I\textsubscript{p(Ca)}), the Na\textsuperscript{+}/Ca\textsuperscript{2+} exchange current (I\textsubscript{NaCa}), the calcium background current (I\textsubscript{Cab}), the fast Na\textsuperscript{+} current (I\textsubscript{Na}), the background Na\textsuperscript{+} current (I\textsubscript{Nab}), the Na\textsuperscript{+}/K\textsuperscript{+} pump current (I\textsubscript{NaK}), the Ca\textsuperscript{2+}-activated Cl\textsuperscript{-} current (I\textsubscript{Cl,Ca}), the rapidly inactivating transient outward K\textsuperscript{+} current (I\textsubscript{Kto}), the two ultra-rapidly activating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}), the slow delayed rectifier K\textsuperscript{+} current (I\textsubscript{Ks}), the rapid delayed rectifier K\textsuperscript{+} current (I\textsubscript{Kr}), and the time independent K\textsuperscript{+} current (I\textsubscript{K1}).
-Detailed implementation of mouse AP and VGIC models can be found in \cite{bondarenko2014compartmentalized,asfaw2020compartmentalized}.
-
-The three types of K\textsuperscript{+} currents have larger magnitude compared to the other K\textsuperscript{+} currents: the rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), the delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}).
-The rapidly inactivating transient outward currents have two components I\textsubscript{Kto,f} and I\textsubscript{Kto,s}, but the later is essentially absent in apex cardiomyocytes.
-We only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because we focus on apex myocytes in this investigation.
-However, the transient outward current can be easily modified according to the region of ventricular cardiomyocytes.
-Therefore, we model K\textsubscript{v} isoforms that conduct these dominant currents: K\textsubscript{v}4.2 (I\textsubscript{Kto}), K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}), and K\textsubscript{2P} family (I\textsubscript{Kss}).
-\textbf{Figure~\ref{fig:kcurrent_example}B} illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the I\textsubscript{K} traces (I\textsubscript{Ksum}) given the protocol in \textbf{Figure~\ref{fig:kcurrent_example}A} that applies 0 mV voltage step from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
-\textbf{Figure~\ref{fig:kcurrent_example}C} and \textbf{D} show a range of protocols and consequent change in I\textsubscript{Ksum}.
+Because of the relative ease at which their genetics can be altered, mice have been useful models for studying cardiac electric signaling \citep{milani2014small,nerbonne2004studying}.
+In ventricular and atrial mouse cardiomyocytes, there are three major components of I\textsubscript{Ksum}: rapidly inactivating transient outward K\textsuperscript{+} currents (I\textsubscript{Kto,f} and/or I\textsubscript{Kto,s}), which are conducted through K\textsubscript{v}4.2 and K\textsubscript{v}1.4 respectively, slowly inactivating delayed rectifier K\textsuperscript{+} currents (I\textsubscript{Kslow1} and I\textsubscript{Kslow2}), which are conducted through K\textsubscript{v}1.5/K\textsubscript{v}2.1, and the non-inactivating steady-state K\textsuperscript{+} current (I\textsubscript{Kss}) \citep{xu1999four}.
+I\textsubscript{Kto,s} (K\textsubscript{v}1.4) is primarily found in septal cardiomyocytes \citep{bondarenko2004computer}.
+In this study, we only include I\textsubscript{Kto,f} in I\textsubscript{Kto}, because our in-vitro studies focused on the ventricular apex myocytes.
+However, I\textsubscript{Kto} can be easily modified according to the region of ventricular cardiomyocytes.
+We model these four dominant K\textsuperscript{+} currents: I\textsubscript{Kto}, I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}.
+Figure~\ref{fig:kcurrent_example}(b) illustrates the shape of the primary K\textsuperscript{+} currents and their contribution to the whole-cell I\textsubscript{K} trace (I\textsubscript{Ksum}) given the protocol in Figure~\ref{fig:kcurrent_example}(a) that applies 0 mV voltage pulse from holding potential -70 mV from time $t^{(h)}$ to $t^{(e)}$.
+Figure~\ref{fig:kcurrent_example}(c) and (d) show a range of voltage steps from -30 mV to 50 mV in 10 mV increments and consequent changes in I\textsubscript{Ksum}.
 Because of their shape, I\textsubscript{Kss} is assumed as a constant and the other currents an exponential function in the traditional data-driven curve-fitting approach.
-In this approach, K\textsuperscript{+} current can be defined by Equation~\ref{eq:expl_fitting} with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$, and a set of them is estimated that have the best fit with experimental I\textsubscript{Ksum} data.
+In this approach, if all four current models are included, I\textsubscript{Ksum} is defined by \eqref{eq:expl_fitting} (i.e., tri-exponential fitting) with the shape parameters such as amplitude $\mathrm{A}_{i}$ and time constant $\tau_{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
+In some cases, it is reduced to a bi-exponential function combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2}.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/exemplary_kv_currents.pdf}
-    \caption{Example of voltage-clamp K\textsuperscript{+} currents.
-(A) Voltage step of 0 mV from holding potential -70 mV.
-(B) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
-(C) Range of protocols from -30 to 50 mV from holding potential -70 mV and (D) consequent K\textsuperscript{+} current traces.}
+    \includegraphics{figs/kcurrent_example.pdf}
+    \caption{Example of voltage-clamp protocol and K\textsuperscript{+} current traces.
+(a) Clamp-voltage pulse of 0 mV from the holding potential -70 mV.
+(b) Dominant K\textsuperscript{+} currents and their contributions to I\textsubscript{Ksum}.
+(c) Series of clamp-voltage pulses (-30-50 mV in 10 mV increments) from the holding potential -70 mV and (d) consequent I\textsubscript{Ksum} traces.}
     \label{fig:kcurrent_example}
@@
     \label{eq:expl_fitting}
-    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}
+    \mathrm{I}_{\mathrm{Ksum}} = \mathrm{A}_{\mathrm{Kto}}e^{-t/\tau_{\mathrm{Kto}}} + \mathrm{A}_{\mathrm{Kslow1}}e^{-t/\tau_{\mathrm{Kslow1}}} + \mathrm{A}_{\mathrm{Kslow2}}e^{-t/\tau_{\mathrm{Kslow2}}} +  \mathrm{A}_{\mathrm{Kss}}.
 \end{equation}
 
-Unlike the simple current functions, computer models of K\textsubscript{v} isoforms can simulate detailed gating kinetics.
-We developed them with Hodgkin-Huxley modeling, which has two gating variables controlling conductance, a popular modeling scheme in various species \citep{ten2004model,mahajan2008rabbit}.
-For example, K\textsubscript{v} isoforms that conducts current I\textsubscript{K} can be defined by Equation~\ref{eq:hh_ex}
+We developed mouse K\textsubscript{v} models based on \citep{bondarenko2014compartmentalized,asfaw2020compartmentalized}, using the Hodgkin-Huxley modeling scheme that has been used for various species, such as humans \citep{ten2004model} and rabbits \citep{mahajan2008rabbit}.
+This type of model consists of two gating variables controlling the channel conductance, and its canonical form is defined by \eqref{eq:hh_ex}
 \begin{equation}
     \label{eq:hh_ex}
-    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}})
+    \mathrm{I}_{\mathrm{K}} = G_{\mathrm{K}}a^{n}i^{m}(V-E_{\mathrm{K}}),
 \end{equation}
 where $G_{\mathrm{K}}$ is the maximum conductance, $a^{n}$ and $i^{m}$ are the gating variables for $n,m \in \mathbb{N}$, $V$ is the transmembrane potential, and $E_{\mathrm{K}}$ is the K\textsuperscript{+} Nernst potential.
-$V-E_{\mathrm{K}}$ is the driving force of the ion movement.
-Important components in this equation are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$.
-These processes are governed by first-order kinetics and its voltage-dependent transition rates $\alpha$ and $\beta$.
-$\alpha$ is the rate at which a gate in closed state opens, whereas $\beta$ is the rates at which a gate in open state closes.
-Equation~\ref{eq:gv} shows a schematic relationship of this gating kinetics.
+$V-E_{\mathrm{K}}$ implies the driving force of the ion movement.
+Important components in \eqref{eq:hh_ex} are the gating variables $a$ and $i$, representing the fraction of activation and recovery from inactivation of the channel where $a,i \in [0,1]$.
+These processes are governed by first-order kinetics and voltage-dependent transition rates $\alpha$ and $\beta$.
+$\alpha$ is the rate at which a gate in a closed state opens, whereas $\beta$ is the rate at which a gate in an open state closes.
+Equation~\eqref{eq:gv} shows a schematic relationship of this gating kinetics.
+$n$ and $m$ represent the numbers of activation and inactivation gates, which are dependent on a particular K\textsubscript{v} isoform and determined by specific kinetics and properties of that channel.
+In general, the number of gates depends on the complexity of the channel behavior and the level of detail required in the model.
 \begin{align}
     \label{eq:gv}
-    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i
+    (1-a)&\xrightleftharpoons[\beta_{a}]{\alpha_{a}}a & (1-i)&\xrightleftharpoons[\beta_{i}]{\alpha_{i}}i.
 \end{align}
 
-The two biophysical processes can be modeled using differential equations in two ways as follows:
+These two biophysical processes $a$ and $i$ can be expressed using differential equations in two ways \eqref{eq:1stkinetic1} and \eqref{eq:1stkenetic2}
 \begin{align}
+    \label{eq:1stkinetic1}
     \frac{da}{dt} &=\alpha_{a}(1-a)-\beta_{a}a   &\frac{di}{dt} &=\alpha_{i}(1-i)-\beta_{i}i \\
-    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}}
+    \label{eq:1stkenetic2}
+    \frac{da}{dt} &= \frac{a_{\infty}-a}{\tau_{a}}  &\frac{di}{dt} &= \frac{i_{\infty}-i}{\tau_{i}},
 \end{align}
@@
     a_{\infty} &= \frac{\alpha_{a}}{\alpha_{a}+\beta_{a}} & i_{\infty} &=  \frac{\alpha_{i}}{\alpha_{i}+\beta_{i}} \\
-    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}
+    \tau_{a} &= \frac{1}{\alpha_{a}+\beta_{a}} & \tau_{i} &= \frac{1}{\alpha_{i}+\beta_{i}}.
 \end{align}
-The steady-state values and time constants can be defined directly by functions of voltage without transition rates in some cases.
-These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters that control the behavior of the kinetics of an ion channel.
+The steady-state values and time constants can be defined directly by functions of voltage $V$ without transition rates in some cases.
+These voltage-dependent functions, such as transition rates, steady states, or time constants, have parameters $p_{s}$ that control the behavior of the kinetics of an ion channel.
 
-\subsubsection{In-silico Modeling of K\textsubscript{v}4.2 (I\textsubscript{Kto})}
-The rapidly inactivating transient outward current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation.
+\subsubsection{In-silico Modeling of I\textsubscript{Kto}}
+The rapidly inactivating transient outward current I\textsubscript{Kto}, which is conducted through K\textsubscript{v}4.2, is characterized by a sharp upstroke during activation and subsequent rapid inactivation.
 It mainly contributes to the peak at the very beginning of activation in I\textsubscript{Ksum}.
@@
     &\alpha_{i} = \frac{p_{9}e^{-(V+p_{2})/p_{4}}}{1+p_{10}e^{-(V+p_{2}+p_{3})/p_{4}}} \\
-    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}} \label{eq:ikto_beta2}
+    & \beta_{i} = \frac{p_{11}e^{(V+p_{2}+p{3})/p_{4}}}{1+p_{12}e^{(V+p_{2}+p_{3})/p_{4}}}.
+    \label{eq:ikto_beta2}
 \end{align}
 There are two gating variables $a_{\mathrm{Kto}}$ and $i_{\mathrm{Kto}}$ responsible for activation and inactivation, respectively.
-Their kinetics are governed by transition-rate functions from Equation~\ref{eq:ikto_alpha1} to Equation~\ref{eq:ikto_beta2}.
-Parameters $p_{i}$ for $i=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
+Their kinetics are governed by transition-rate functions from \eqref{eq:ikto_alpha1} to \eqref{eq:ikto_beta2}.
+Parameters $p_{s}$ for $s=\{1, 2, \dots, 12\}$ in these equations act like knobs, allowing to control the behavior of the I\textsubscript{Kto} model.
 
-\subsubsection{In-silico Modeling of K\textsubscript{v}1.5 (I\textsubscript{Kslow1}), K\textsubscript{v}2.1 (I\textsubscript{Kslow2}) and K\textsubscript{2P} (I\textsubscript{Kss})} \label{s:method.model2}
-There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are very rapidly activating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
-As shown in \textbf{Figure~\ref{fig:kcurrent_example}}, both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, but I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1}.
-The non-inactivating steady-state current, which conducted through K\textsubscript{2P} family, remains constant during the voltage-clamp recording.
-These three currents collectively address the most part of the decaying phase of I\textsubscript{Ksum}.
-We assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and I\textsubscript{Kss} has similar activation behavior with a slightly different rate to keep the model as simple as possible to reduce the structural risk of overfitting.
+\subsubsection{In-silico Modeling of I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}}
+There are two major delayed rectifier currents in mouse ventricular cardiomyocytes that are rapidly activating and slowly inactivating: I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, which are conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1, respectively.
+As shown in Figure~\ref{fig:kcurrent_example}(b), both rectifier currents inactivate slower and have smaller magnitude than I\textsubscript{Kto}, and I\textsubscript{Kslow2} decays more gradually than I\textsubscript{Kslow1} \citep{liu2011dissection}.
+The non-inactivating steady-state current I\textsubscript{Kss}, which is likely conducted through K\textsubscript{v} isoforms K\textsubscript{2P} family \citep{feliciangeli2015family}, remains constant during the voltage-clamp recording.
+These three currents contribute to most part of the decaying portion of I\textsubscript{Ksum}.
+To keep the models as simple as possible to reduce the structural risk of overfitting, we assume that I\textsubscript{Kslow1} and I\textsubscript{Kslow2} have the same activation gating variable, and they have a similar inactivation pattern; I\textsubscript{Kss} has similar activation behavior with the two delayed rectifier currents but a slightly different rate.
 
-I\textsubscript{Kslow1} is modeled without transition-rate functions as opposed to I\textsubscript{Kto}.
-Its gating variables, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$ are defined directly by steady-state ($a_{ss}$ and $i_{ss}$) and time-constant functions ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) as follows.
+I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss} are modeled without transition-rate functions as opposed to I\textsubscript{Kto}, and their steady-state and time-constant functions are directly defined.
+First, the gating variables of I\textsubscript{Kslow1}, activation $a_{\mathrm{Kslow1}}$, and inactivation $i_{\mathrm{Kslow1}}$, are defined by \eqref{eq:akslow1} and \eqref{eq:ikslow1}.
+The steady-state functions ($a_{ss}$ and $i_{ss}$) in \eqref{eq:ass} and \eqref{eq:iss} will be shared with the other two current models.
+A full description of I\textsubscript{Kslow1} is given as follows:
 \begin{align}
     &\mathrm{I}_{\mathrm{Kslow1}} = G_{\mathrm{Kslow1}}a_{\mathrm{Kslow1}}i_{\mathrm{Kslow1}}(V-E_{\mathrm{K}}) \\
-    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(1)}} \\
-    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(1)}} \\
+    &\frac{da_{\mathrm{Kslow1}}}{dt} = \frac{a_{ss}-a_{\mathrm{Kslow1}}}{\tau_{a}^{(\mathrm{Kslow1})}} \label{eq:akslow1} \\
+    &\frac{di_{\mathrm{Kslow1}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow1}}}{\tau_{i}^{(\mathrm{Kslow1})}} \label{eq:ikslow1} \\
     &a_{ss} = \frac{1}{1+e^{-(V+p_{1})/p_{4}}} \label{eq:ass} \\
     &i_{ss} = \frac{1}{1+e^{(V+p_{2})/p_{5}}} \label{eq:iss} \\
-    &\tau_{a}^{(2)} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
-    &\tau_{i}^{(2)} = p_{10} - p_{8}i_{ss}
+    &\tau_{a}^{(\mathrm{Kslow1})} = \frac{p_{7}}{e^{p_{6}(V+p_{3})} + e^{-p_{6}(V+p_{3})}} + p_{9} \\
+    &\tau_{i}^{(\mathrm{Kslow1})} = p_{10} - p_{8}i_{ss}.
 \end{align}
 
-I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ that contains the same steady-state function $i_{ss}$ in I\textsubscript{Kslow1}.
-As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows.
+As a structural regularization, I\textsubscript{Kslow2} has the same activation variable with I\textsubscript{Kslow1} as in \eqref{eq:act_var}, and the time-constant function of the inactivation $i_{\mathrm{Kslow2}}$ shares the same steady-state function $i_{ss}$ \eqref{eq:iss} with I\textsubscript{Kslow1}.
+As a result of this modeling strategy, mathematical equations of I\textsubscript{Kslow} are given as follows:
 \begin{align}
     &\mathrm{I}_{\mathrm{Kslow2}} = G_{\mathrm{Kslow2}}a_{\mathrm{Kslow2}}i_{\mathrm{Kslow2}}(V-E_{\mathrm{K}}) \\
-    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \\
-    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(2)}} \\
-    &\tau_{i}^{(3)} = p_{2} - p_{1}i_{ss}
+    &a_{\mathrm{Kslow2}} = a_{\mathrm{Kslow1}} \label{eq:act_var} \\
+    &\frac{di_{\mathrm{Kslow2}}}{dt} = \frac{i_{ss}-i_{\mathrm{Kslow2}}}{\tau_{i}^{(\mathrm{Kslow2})}} \\
+    &\tau_{i}^{(\mathrm{Kslow2})} = p_{2} - p_{1}i_{ss}.
 \end{align}
 
-I\textsubscript{Kss} does not have an inactivation variable as it is non-inactivating current.
-It shares the same steady-state function for activation $a_{ss}$ with the rectifier currents but have a separate time-constant function to address the different activation rate.
+I\textsubscript{Kss} does not have an inactivation variable because it is non-inactivating current.
+It shares the same steady-state function for activation $a_{ss}$ \eqref{eq:ass} with the other two delayed rectifier currents but have a separate time-constant function \eqref{eq:time_ikss} to address the different activation rate.
 I\textsubscript{Kss} is modeled as
@@
 
-\subsection{Data Assimilation and Model Calibration}
-\textit{Data assimilation} is a method to find the optimal configuration and state of computational models by coupling them with experimental data.
+\subsection{Concurrent Assimilation of Functional Data}
+Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling them with experimental data.
 Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation.
 The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment.
-Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term.
+Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in \eqref{eq:rd_relation}, where $\epsilon$ is the error term.
 \begin{equation}
@@
     y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
-    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
+    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon.
 \end{align}
 Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints.
-By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.
+By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{D}$, which provides two complementary angles to study the real process $\mathcal{R}$.
 
-In this study, $\delta$ is defined by the sum of root-mean-square errors (RMSEs) as in Equation~\ref{eq:rmse}, which measures deviations between experimental data and model predictions from the end of a holding potential $t_i^{(h)}$ to the end of voltage step $t_i^{(e)}$ across a range of protocols, $i=1,2,\cdots,n$.
+From this perspective, bi-/tri-exponential function in \eqref{eq:expl_fitting} serves as $y^{\mathcal{M}}(t,v|\theta)$ in the curve-fitting approach, where $\theta=\{\mathrm{A}_{i}(v),\tau_{i}(v)\}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$.
+$v$ represents voltage.
+Note that $\mathrm{A}_{i}(v)$ and $\tau_{i}(v)$ are dependent on input data $v$, so for each voltage, we need to perform a data assimilation procedure separately.
+In general, multiple input voltage steps are applied to a cardiomyocyte producing a set of I\textsubscript{Ksum} recordings to study the voltage-dependent characteristics.
+Suppose the sum of in-silico models of K\textsuperscript{+} currents $\mathrm{I}_\mathrm{i}$ for $i \in \{\mathrm{Kto}, \mathrm{Kslow1}, \mathrm{Kslow2}, \mathrm{Kss}\}$ serve as a computer model for data assimilation as in \eqref{eq:comp_model_as_insilico}:
+\begin{equation}
+    \label{eq:comp_model_as_insilico}
+    y^{\mathcal{M}}(t,v|\theta) = \sum_{i} \mathrm{I}_{i},
+\end{equation}
+where $\theta$ is the union of kinetic parameters $p_{s}$ for each $\mathrm{I}_{i}$, which are constants.
+In contrast, this computer model generates I\textsubscript{Ksum} for different input voltages given one set of parameters, because the in-silico models are designed by biophysical principles encoding voltage dependence.
+Therefore, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in \eqref{eq:rmse}.
+RMSE evaluates the goodness-of-fit between experimental I\textsubscript{Ksum} and model prediction.
+RMSEs for a set of I\textsubscript{Ksum} traces are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltages $v=1,2,\cdots,n$.
 \begin{equation}
     \label{eq:rmse}
-    \delta = \sum_{i=1}^{n} \sqrt{\int_{t_i^{(h)}}^{t_i^{(e)}}\frac{(y_i^{\mathcal{D}}(t) - y_i^{\mathcal{M}}(t|\theta))^2}{t_i^{(e)}-t_i^{(h)}}dt}
+    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y^{\mathcal{M}}(t,v|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
 \end{equation}
-This approach differs from previous studies in two-fold.
-First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method includes multiple protocols simultaneously for comprehensive cellular-level modeling.
-Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}.
-Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.
 
-We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$.
+Cardiac models enable this concurrent data assimilation, because they generate multiple current traces with different input voltages by simulating underlying gating kinetics.
+Note that the proposed method calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data via curve-fitting \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}.
+We develop the box-constrained nonlinear optimization routine with multi-random initial points to minimize $\delta$.
 Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube.
 In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling.
-Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting.
-Global optimum is not guaranteed due to the complexity of our models.
-We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible.
+The multi-random-start scheme helps escape local optima and find the solution as close to the optimum as possible.
 Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden.
-This work is implemented in MATLAB R2022a, and the code is available at xyz.
+This work is implemented in MATLAB R2022a.
 
-\subsection{Model Regularization and Sensitivity Analysis}
-The principle of parsimony is critical in model calibration not only for enhancing fitting accuracy and preventing overfitting, but also for improving the interpretability of $\theta$.
-Excessive flexibility of a model has a risk of causing overfitting that occurs when the model fits data too closely, even including noise and random effects in the data.
-Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results of parameters.
-It is worth noting that the models presented in Section 3.1 are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simply the model structure.
-Specifically, the same parameters that control voltage thresholds, such as $p_{1}$, $p_{2}$, $p_{3}$, and $p_{4}$, are used in equations having a similar function.
-The steady-state functions (Equation~\ref{eq:ass} and \ref{eq:iss}) are reused in the three rectifier-type currents.
+\subsection{Sensitivity Analysis and Model Regularization}
+The principle of parsimony is critical in model calibration to enhance fitting accuracy, prevent overfitting, and improve interpretability of $\theta$.
+Excessive flexibility has a risk of overfitting that occurs when the model fits data too closely, even including noise and random effects in data.
+Besides, as the number of parameters increases, it becomes complicated to interpret the calibration results.
+It is worth emphasizing that the presented models are designed with structural regularization, in which some parameters and equations are recycled in multiple places to simplify the model structure.
 
-We also performed a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
-In order to do that, factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output.
-As illustrated in \textbf{Figure~\ref{fig:current_markers}}, six markers were defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
-Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
-Marker f will equal to the total recording time if current does not decline enough as in \textbf{Figure~\ref{fig:current_markers}C} and \textbf{~\ref{fig:current_markers}D}.
+We also perform a sensitivity analysis to identify a subset of the parameters that have significant impacts on the model output and only calibrate these sensitive parameters.
+Factorial designs are developed in which parameters vary at two levels contrasting their effect on the model output.
+As illustrated in Figure~\ref{fig:current_markers}, six markers are defined that capture characteristics of K\textsuperscript{+} current traces in voltage-clamp experiments.
+Each marker represents: a) the current magnitude of 10 ms after applying a voltage step, which measures the activation rate; b) 25\% of the total recording time has elapsed, c) 50\%, and d) 75\%, which collectively estimate the inactivation rate over time; e) the peak magnitude; and f) the time when current has decayed ($1-e^{-1}$)\% (almost 63\%) from the peak.
+Marker f will be equal to the total recording time if current does not decline enough as in Figure~\ref{fig:current_markers}(c) and (d).
 \begin{figure}[!ht]
@@
 
-We adopted a fractional factorial design of 1024 runs for I\textsubscript{Kto}, which results in resolution VIII.
+A fractional factorial design of 1024 runs is adopted for I\textsubscript{Kto}, which results in resolution VIII.
 The resolution ensures that the main effects and 2-/3-factor interactions are strongly clear.
 Full factorial designs are used for the other three currents.
-The marker points are evaluated at each design point and factorial effects are calculated using the linear model in Equation~\ref{eq:fact_effect} for each current.
+The marker points $\lambda^{k}$ defined as in Figure~\ref{fig:current_markers} are evaluated at each design sample and factorial effects are calculated via the linear model in \eqref{eq:fact_effect} for each current.
+The least squares method is used to estimate $\beta_{s}^{k}$.
+Then, half-normal plots are drawn to test significance of the estimated factorial effects.
 \begin{equation}
-    \delta^{k} = \beta_{o}^{k} + \sum\limits_{i}\beta_{i}^{k}p_{i}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\},
     \label{eq:fact_effect}
+    \lambda^{k} = \beta_{o}^{k} + \sum\limits_{s}\beta_{s}^{k}p_{s}^{k} + \epsilon, \ k \in \{a,b,c,d,e,f\}.
 \end{equation}
-The least squares method is used to estimate $\beta_{i}^{k}$. Then, half-normal plots are drawn to test significance of the estimated effects.
 
-\subsection{Variability Visualization and Low-dimensional Embedding}
+\subsection{Low-dimensional Embedding}
 Most cardiac electrophysiology studies utilize statistical tests for the mean to support hypotheses; in turn, the majority of ion channel models are based on population-averaged data.
-However, fine-grained analyses that consider the variability of cells can provide new insights into data.
-For example, it is possible there are differences between control and disease-altered groups but also within the groups according to characteristics of cells such as size or age.
-Hence, there is an urgent need to develop a tool to investigate cellular variability.
-Because the proposed approach calibrates models for the individual dataset from each cell, it allows us to quantify and visualize cellular variability in the tuned parameters.
-Histograms are used in this study to visualize distributions of calibration parameters and the effects of disease-related perturbations on the parameter distributions.
-Although a histogram is a simple yet powerful tool, one of the limitations of a histogram is that it cannot visualize multiple variables more than three on the same plot.
-It is desirable to inspect how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy control and disease groups.
-Low-dimensional embedding methods transform high-dimensional data into a plane or 3D space while preserving relative locations of data points as much as possible so that the structure of data can be visualized.
-\textbf{Figure~\ref{fig:embd}} illustrates how a low-dimensional embedding method is used to map the high-dimensional parameter space of data points from wild-type and disease groups into 3D space.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/embd.pdf}
-    \caption{Diagram illustrating the procedure of low-dimensional embedding to reveal cellular-level structure of the parameter space.}
-    \label{fig:embd}
-\end{figure}
+However, subject-specific analyses that consider the cellular-level variability can provide new insights into data.
+For example, it is possible there are differences between control and experimental groups but also within the groups.
+Hence, there is an urgent need to develop a tool to investigate cell-specific characteristics.
+Because the proposed approach calibrates models using the dataset for each cell, it allows us to quantify cellular variability in the tuned parameters.
+We adopt low-dimensional embedding that transforms high-dimensional data into a plane or 3D space while preserving relative locations of data points to visualize how distributional differences in calibration parameters collectively impact inter/intra-cell variability in healthy and diseased groups.
 
-For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \cite{van2008visualizing}.
+For this purpose, t-distributed stochastic neighbor embedding (t-SNE) is used \citep{van2008visualizing}.
 It has proven to be an effective method for visualizing high-dimensional data.
 t-SNE is a statistical method that constructs two sets of probability distributions $p_{ij}$ and $q_{ij}$ over pairs of data points $i$ and $j$ in a high-/low-dimensional space, respectively.
-These are probabilities of similarities of data points such that neighboring points have a higher probability while dissimilar points have a lower probability.
+These are probabilities of similarities such that neighboring points have a higher probability while dissimilar points have a lower probability.
 We first define the conditional probability of $j$ given $i$:
@@
 \end{equation}
-Typically, a gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.
+A gradient descent method is used for the minimization of $\mathrm{KL}(P||Q)$ with respect to $y$.
 
-\section{Experimental Design}
-More recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM using genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
-MGAT1KO mice developed DCM, deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
+\section{Experimental Design and Results}
+We apply the proposed framework to our data for investigating the pathophysiology of electrical signaling altered by reduced glycosylation.
+Recently, we showed that preventing hybrid/complex N-glycosylation in mouse cardiomyocytes was sufficient to cause DCM, achieved through genetic ablation of the \textit{MGAT1} gene (MGAT1KO model), which encodes a critical glycosyltransferase, GlcNAcT1 \citep{ednie2019reduced2}.
+MGAT1KO mice developed DCM that deteriorated into heart failure, and 100\% died early, presumably from ventricular arrhythmias leading to sudden cardiac death.
 To further investigate the role of altered glycosylation in pathogenesis and disease progression of the heart, we conducted whole-cell patch-clamp experiments that showed reductions in N-glycosylation significantly impact electrical signaling in mouse cardiomyocytes \citep{ednie2019reduced}.
-To be specific, whole-cell I\textsubscript{K} traces were measured in left ventricular apex cardiomyocytes of ~14 week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 - 50 mV) from holding potential of -70 mV.
-There were 32 observations for each WT and MGAT1KO.
+To be specific, whole-cell I\textsubscript{Ksum} traces were measured in left ventricular apex cardiomyocytes of $\sim$14-week-old control (wild type; WT) and MGAT1KO mice, elicited by 4.5 s 10mV voltage steps (-30 to +50 mV) from holding potential of -70 mV.
+There were 31 sets of whole-cell I\textsubscript{K} recordings from different WT cells and 30 from MGAT1KO cells.
+Animals were used and cared for as outlined by the NIH’s Guide for the Care and Use of Laboratory Animals.
+All animal protocols were reviewed and approved by the Wright State University Institutional Animal Use and Care Committee.
 
-We applied the proposed framework to our investigations in the pathophysiology of electrical signaling by altered glycosylation.
-It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1 ventricular myocytes.
-A curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in Equation~\ref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents.
-All three component currents were reduced, but the rectifier current was significantly reduced and slowed notably.
-Although the \textit{in-vitro} investigation discovered aberrant reductions in K\textsuperscript{+} currents with altered glycosylation, it was difficult to determine channel kinetics rigorously.
-For example, extra curve fitting was not available due to small current magnitude of the MGAT1KO model.
-Therefore, we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{K} recordings.
-This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics with considering the cellular-level dynamics of multiple protocols.
+It was observed in the experiment that K\textsuperscript{+} currents were reduced in MGAT1KO ventricular myocytes \citep{ednie2019reduced}.
+Because of the overlapping inactivation rates of the K\textsubscript{v} isoforms responsible for I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1), a curve-fitting method of bi-exponential, combining I\textsubscript{Kslow1} and I\textsubscript{Kslow2} of tri-exponential in \eqref{eq:expl_fitting}, was applied to decompose I\textsubscript{Ksum} traces into component currents.
+All three component currents were reduced, but the rectifier current I\textsubscript{Kslow} was significantly reduced and slowed notably.
+Although this in-vitro investigation discovered aberrant reductions in K\textsuperscript{+} currents with chronic glycosylation deficiency, it was difficult to determine channel kinetics rigorously.
+For example, curve fitting was not able to provide reliable results due to small current magnitude of the MGAT1KO model, particularly at the lower voltage steps.
+Therefore, here we leverage the suggested framework for modeling K\textsubscript{v} isoforms kinetics from our experimental data of I\textsubscript{Ksum} recordings.
+This new approach allows dissecting whole-cell K\textsuperscript{+} current traces into isoform components and modeling their underlying kinetics concurrently.
 
-\section{Results}
 \subsection{Parameter Screening}
-\textbf{Figure~\ref{fig:half-norm}} shows the half-normal plots of factorial effects of parameters on the six markers.
+Figure~\ref{fig:half-norm} shows the half-normal plots of factorial effects of parameters on the six markers.
 The red straight lines on the plots serve as a criterion to identify the parameters that are sensitive to the marker points.
@@
 We picked the parameters to be calibrated as a union of the sets of parameters falling above the straight lines on the six half-normal plots.
-For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1},\ p_{2}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
+For I\textsubscript{Kto} $\{p_{1},\ p_{2},\ p_{3},\ p_{4},\ p_{5},\ p_{7},\ p_{11}\}$ are selected, for I\textsubscript{Kslow1} $\{p_{1},\ p_{2},\ p_{4},\ p_{5},\ p_{9},\ p_{10}\}$, for I\textsubscript{Kslow2} $\{p_{1}\}$, and for I\textsubscript{Kss} $\{p_{1},\ p_{2},\ p_{3}\}$.
+In addition, all the maximum conductance variables are included in the calibration parameter set.
 \begin{figure}
@@
 \subsection{Model Fitness}
-We used the BFGS algorithm, one of the most popular nonlinear optimization methods, in the proposed model calibration routine because it supports box constraints.
-A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than utilizing the final objective function value of the calibration routine to report the fitness.
-The objective function, the sum of RMSEs over voltage steps, does not provide a straightforward interpretation that makes it difficult to measure the performance of the models.
-$R^{2}$, defined in Equation~\ref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model.
-However, this interpretation is not applicable, and its value can be negative for nonlinear models because the variance decomposition in Equation~\ref{eq:var_decomp} does not hold anymore.
+We tested various nonlinear optimization algorithms in the proposed model calibration routine that support box constraints, and the BFGS algorithm provided the most reliable results.
+A separate accuracy metric is used to measure the goodness-of-fit and potential prediction power of the calibration results rather than reporting the final objective function value.
+The objective function (i.e., sum of RMSEs) does not provide a straightforward interpretation that makes it difficult to evaluate the calibration performance.
+$R^{2}$, defined in \eqref{eq:r2}, is the most common prediction accuracy metric ranging from 0 to 1 for linear regression models that represents the proportion of variation in data that is explained by the model.
+However, this interpretation is not applicable to nonlinear models because the variance decomposition in \eqref{eq:var_decomp} does not hold anymore.
 \begin{align}
-    R^{2} &= 1 - \frac{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2}{\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2} \label{eq:r2} \\
-    \sum_{k=1}^{n}(y^{\mathcal{D}}(k)-\bar{y})^2 &= \sum_{k=1}^{n}(y^{\mathcal{M}}(k|\theta))-\bar{y})^2+\sum_{k=1}^{n}(y^{\mathcal{D}}(k)-y^{\mathcal{M}}(k|\theta))^2 \label{eq:var_decomp}
+    \label{eq:r2}
+    &R^{2} = 1 - \frac{\sum_{i=1}^{n}(y_{i}-\hat{y_{i}})^2}{\sum_{i=1}^{n}(y_{i}-\bar{y})^2} \\
+    \label{eq:var_decomp}
+    &\sum_{i=1}^{n}(y_{i}-\bar{y})^2 = \sum_{i=1}^{n}(\hat{y_{i}}-\bar{y})^2+\sum_{i=1}^{n}(y_{i}-\hat{y_{i}})^2
 \end{align}
 
-Therefore, a pseudo-R-squared measure, proposed in \citet{li2019prediction} was used to evaluate the model fitness in this study.
+Therefore, a pseudo-R-squared measure, proposed in \citep{li2019prediction} is used to evaluate the model fitness in this study.
 This nonlinear $R^{2}$ measure is based on the concept of linear correction of prediction function, allowing the variance decomposition for nonlinear functions, which guarantees straightforward interpretation and normalized scores between 0 and 1 as in the classical $R^{2}$.
-\textbf{Fig~\ref{fig:model_fitness}A} and \textbf{C} present averages of $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively.
-In most cases, more than 80 \% of variances on average in multiple current recordings are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
-The model fitness measures in MGAT1KO are less than WT in general.
+Figure~\ref{fig:model_fitness}(a) and (c) present averages of nonlinear $R^{2}$ values over voltage steps from -30 to 50 mV by 10 mV step size for WT and MGAT1KO, respectively.
+In most cases, more than 90 \% of variances on average are explained by models (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
+The model fitness measures in MGAT1KO are lower than WT in general.
 It is expected that the small current magnitude of MGAT1KO challenges the calibration process.
-We compared the actual model predictions and experimental I\textsubscript{K} traces in \textbf{Figure~\ref{fig:model_fitness}B} and \textbf{D} to validate the model calibration results further.
-As illustrated in the comparison plot of MGAT1KO at -30 mV clamp voltage, small and noisy current trace results in decreased $R^{2}$, 0.8335.
+We randomly select a cell from each group and compare the actual model predictions and experimental I\textsubscript{Ksum} traces in Figure~\ref{fig:model_fitness}(b) and (d) to validate the model calibration results further.
+As shown in Figure~\ref{fig:model_fitness}(d), MGAT1KO at -30 mV clamp voltage, small and noisy current trace, results in decreased $R^{2}$, 0.8335.
+The fitting was performed on 4-core Xeon E7-4830 processors.
+Without parallel computing on multiple cores, each cell took an average of $438.1 \pm 6.8$ to complete the data assimilation procedure for 5 random initial points.
+It took an average of $333.8 \pm 6.0$ s with multicore processing, which is ~31 \% efficient.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/model_fitness_4half.pdf}
-    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (A) WT and (C) MGAT1KO.
-Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{K} recordings (in pink) at -30 mV and 50 mV voltage steps of (B) cell 6 in WT and (D) 12 in MGAT1KO.}
+    \includegraphics{figs/model_fitness.pdf}
+    \caption{Bar graphs of the averages of nonlinear $R^{2}$s for nine voltage steps from -30 mV to 50 mV by the 10 mV step size for (a) WT and (c) MGAT1KO.
+Exemplary actual fitness between model predictions (in green) and experimental I\textsubscript{Ksum} recordings (in pink) at six voltage steps of (b) cell 6 in WT and (d) cell 12 in MGAT1KO.}
     \label{fig:model_fitness}
@@
 
-\subsection{Kinetics Modeling} \label{s:results.kinetics}
-The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior.
-\textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2.
-There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}).
-Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}.
-$\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV.
-$\tau_{i}^{(1)}$ for MGAT1KO shifted toward left.
-The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
+\subsection{Prediction of K\textsubscript{v} Activities}
+The in-silico modeling predicts that chronic reduction in cardiomyocyte N-glycosylation results in significant changes in channel steady states and kinetic behaviors.
+The most notable result in the in-vitro experimental data was the significant reduction in the current density in response to a 50-mV test potential.
+To verify this observation, Figure~\ref{fig:current_density} presents the prediction results of the current-density relationship of the four K\textsuperscript{+} currents.
+The model predicts the compatible result showing the I\textsubscript{Kto}, I\textsubscript{Kslow1}, and I\textsubscript{Kslow2} densities are significantly reduced in MGAT1KO cardiomyocytes, while there are no remarkable differences in I\textsubscript{Kss}.
+Note that in the in-vitro experiments, bi-exponential fitting was used, in which the two delayed rectifier currents are combined (I\textsubscript{Kslow}), because it is difficult to reliably fit a tri-exponential function with a relatively short voltage step (4.5 s here) \citep{liu2011dissection}.
+In contrast, although the proposed model-guided data assimilation distinguishes K\textsubscript{v}1.5 and K\textsubscript{v}2.1, it shows compatible prediction results with the experimental data as well as high fitness accuracy ($\sim$90 \% $R^{2}$) \citep{ednie2019reduced}.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/kinetics_ikto.pdf}
-    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
-    \label{fig:kinetics_ikto}
+    \includegraphics{figs/current_density.pdf}
+    \caption{Current-density relationships of the major K\textsubscript{+} currents: (a) I\textsubscript{Kto} (b) I\textsubscript{Kslow1}, I\textsubscript{Kslow2}, and I\textsubscript{Kss}.}
+    \label{fig:current_density}
 \end{figure}
 
-\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}.
-The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}).
-Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO.
-SSI is also shifted $\sim$ 10 mV to the right.
-The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO.
-Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}C}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
+Another aberrant activity of K\textsuperscript{+} currents was slower inactivation of the delayed rectifier currents in MGAT1KO cardiomyocytes.
+Figure~\ref{fig:time_constants} shows the prediction results of the inactivation time constants of I\textsubscript{Kslow1} and I\textsubscript{Kslow2}.
+The computer models successfully capture this trend, resulting in the time constant of inactivation in for I\textsubscript{Kslow1} ($\tau_{i}^{(\mathrm{Kslow1})}$) being estimated to be significantly slowed in MGAT1KO compared to WT, i.e., $\sim$348.9 ms, as shown in Figure~\ref{fig:time_constants} (a).
+I\textsubscript{Kslow2} is also slowed (see Figure~\ref{fig:time_constants} (b)).
+The inactivation time constants are important because they determine how slowly I\textsubscript{Kslow1} and I\textsubscript{Kslow2} inactivate and, in turn, affect the repolarization of the AP significantly.
+$\tau_{i}^{(Kslow1)}$ show higher variability than other currents to address most of the decaying portion of I\textsubscript{Ksum}.
+This is likely due to the fact that while K\textsubscript{v}4.2 and K\textsubscript{v}2.1 are O-glycosylated, K\textsubscript{v}1.5 possess a single occupied N-linked glycosylation.
+Hence, I\textsubscript{Kslow1} is more affected by reduced N-glycosylation.
+A plot is not provided here, but there is no significant effect on I\textsubscript{Kto} for membrane potential greater than 0 mV.
+It is worth mentioning that the proposed method was able to extrapolate $\tau_{i}^{(\mathrm{Kslow2})}$ that is greater than 5 s from the 4.5-s-pulse protocol, which shows the generalization power of the in-silico modeling.
 \begin{figure}[!ht]
     \centering
-    \includegraphics{figs/kinetics_rec.pdf}
-    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
-    \label{fig:kinetics_rec}
+    \includegraphics{figs/time_constants.pdf}
+    \caption{Inactivation time constants of (a) I\textsubscript{Kslow1}, and (b) I\textsubscript{Kslow2}.}
+    \label{fig:time_constants}
 \end{figure}
 
-\subsection{Parameter Distributions Between WT and MGAT1KO} \label{s:results.distribution}
-We estimated distributions of the calibrated kinetic parameters using histograms to analyze the impacts of reduced N-glycosylation further.
-\textbf{Figure~\ref{fig:hist_pkto}} shows the distributional differences of the kinetic parameters in the K\textsubscript{4.2} model between the two group.
-A notable fact about the figure is that MGAT1KO shows small variances for Parameters 1, 5, and 7 in transition $\alpha_{a}$ and $\beta_{a}$, which control kinetics activation of I\textsubscript{Kto}.
-Reduced glycosylation affects activation gating that plays a vital role in I\textsubscript{Kto}, as it is characterized by a high peak at the early depolarization and rapid inactivation.
-In addition, MGAT1KO causes a leftward shift in the distribution of the maximum conductance $G_{\mathrm{Kto}}$, resulting in smaller conduction than WT in general.
-It is likely due to the observed current reductions in MGAT1KO.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/hist_pkto.pdf}
-    \caption{Distribution estimations of calibration parameters in the K\textsubscript{Kto} model.}
-    \label{fig:hist_pkto}
-\end{figure}
+\subsection{Cellular Variability}
+The calibrated parameters are in a 61×20 matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters), and we apply t-SNE to encapsulate this high-dimensional data into three-dimensional space to visualize each data point.
+Figure~\ref{fig:embd_result} (a) presents the visualization of the t-SNE embedding of the calibration results.
+It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups across the entire data, indicating that MGAT1KO cells have higher variance in the aggregated kinetic parameter space.
+Determinants of the covariance matrices of the 3D embedding data to quantify the variability of WT and MGAT1KO in the kinetic parameter space.
+It turns out that WT has 38 \% higher variability than MGAT1KO.
 
-\textbf{Figure~\ref{fig:hist_prectifiers}} shows distribution estimations of the calibration parameters in the rectifier and steady-state currents models.
-\textbf{Figure~\ref{fig:hist_prectifiers}A} shows significant distributional differences in Parameters 9 and 10, which control time constants of I\textsubscript{Kslow1}.
-It is compatible with the previous observations of significant differences in time constants to voltage relationships of I\textsubscript{Kslow1} between WT and MGAT1KO.
-There are slight shifts in Parameters 1 and 4 of I\textsubscript{Kslow1} that affect steady-state activation.
-The voltage-threshold parameter for MGAT1KO in steady-state inactivation has higher densities at low values around 30 than WT, which has the peak mode around 48.
-It accounts for the shift at low voltages less than 0 mV in SSI.
-The maximum conductance $G_{\mathrm{Kslow1}}$ of MGAT1KO is centered around relatively small values compared to WT.
-This leftward shift in the maximum conductance is also observed in I\textsubscript{Kslow2} (\textbf{Figure~\ref{fig:kinetics_rec}B}).
-Although there is a difference in variance of $G_{\mathrm{Kss}}$, the two groups have similar means around 0.03 (i.e.
-WT: $0.0375\pm0.0027$ vs MGAT1KO: $0.0325\pm0.015$).
-It is compatible with the observation that I\textsubscript{Kss} does not show significant reductions.
-\begin{figure}[!ht]
-    \centering
-    \includegraphics{figs/hist_prectifiers.pdf}
-    \caption{Distribution estimations of calibration parameters in (A) the I\textsubscript{Kslow1} model, (B) the I\textsubscript{Kslow2} model, and (C) the I\textsubscript{Kss} model.}
-    \label{fig:hist_prectifiers}
-\end{figure}
-
-\subsection{Low-dimensional Embedding and Visualization}
-We aggregated calibration results into one to analyze the parameters in a collective way.
-It resulted in a $61\times20$ matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters).
-Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point.
-\textbf{Figure~\ref{fig:embd_result}} presents the visualization of the t-SNE embedding of the calibration results.
-It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
+To further investigate it, we generated the histograms of the maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant parameter determining the magnitude of the current (see Figure~\ref{fig:embd_result} (b)), and experimental I\textsubscript{K} density from the in-vitro data.
+There is clear skewness in MGAT1KO I\textsubscript{Ksum} density (small variance), while WT I\textsubscript{K} density is more disperse (high variance).
+Calibrated parameters, for example \textit{G}\textsubscript{Kslow1}, likely show smaller variations in MGAT1KO to address this trend in the experimental data.
+It is likely due to significant I\textsubscript{K} reduction in the diseased group combined with the elimination of complex N-glycosylation, thereby minimizing potential variability of N-glycan structures
 \begin{figure}[!ht]
@@
     \includegraphics{figs/embd_result.pdf}
-    \caption{t-SNE embedding of calibration parameters.}
+    \caption{(a) Low-dimensional embedding of calibration parameters into 3D space.
+Blue circles - WT myocytes ($n=31$); Red triangles - MGAT1KO myocytes ($n=30$).
+(b) Histogram of the calibrated kinetic parameter for maximum conductance of I\textsubscript{Kslow1} (\textit{G}\textsubscript{Kslow1}), which is the most significant factor determining the magnitude of I\textsubscript{Kslow1}.
+(c) Histogram of experimental I\textsubscript{Ksum} density.}
     \label{fig:embd_result}
@@
 
-\section{Conclusions}\label{s:conclusions}
-We propose a novel data assimilation framework of voltage-gated K\textsuperscript{+} channels that combines experimental data with computer models guided by simulated channel kinetics to investigate disease-altered cardiac electrical signaling.
-This approach provides a complementary tool to physical experiments that are limited in their ability to determine molecular-level channel kinetics rigorously and provide predictive insights.
-Traditional methods, based on curve-fitting, do not consider interactions of various gating variables across isoforms and dynamics of kinetics according to changes in protocols.
-The proposed method focuses on more comprehensive modeling of underlying channel kinetics of K\textsubscript{v} isoforms than traditional approaches.
-Experimental results show the strong potential of the proposed framework to discover channel kinetics from various angles.
-Predictions of kinetics variables to voltage relationships provide information about changes in the voltage dependence of channels.
-In addition, calibrated parameters of the computer models present a micro-level analysis of channel gating.
-The embedding method allows visualizing the variability of each cell.
-In summary, the proposed framework paves a new way for investigating the electrical signaling of potassium channels and heart diseases.
+\section{Discussion and Conclusions}
+Repolarization is a complex process that involves various K\textsubscript{v} isoforms.
+It is critical to understand the unique properties and functional roles of each K\textsubscript{v} to investigate pathological mechanisms of diseased cardiomyocytes that contribute to fatal heart diseases.
+However, current laboratory techniques, such as whole-cell patch-clamp recording, are not able to measure individual activities of K\textsubscript{v} isoforms reliably that activate, inactivate, and close at overlapping times during recordings, except through the attempt to remove K\textsubscript{v} isoform activity through less-than-fully-specific pharmacological intervention.
+Thus, only the sum of the different K\textsubscript{v} isoform activities can be measured as a single current trace, I\textsubscript{Ksum}.
+Hence, it is necessary to decompose I\textsubscript{Ksum} into individual K\textsuperscript{+} currents and estimate their channel activity via data assimilation.
+This paper presents a subject-specific concurrent data assimilation method for learning K\textsubscript{v} activities using multiple I\textsubscript{Ksum} recordings simultaneously for each cell.
+A case study is provided using our in-vitro experimental data of mouse cardiomyocyte I\textsubscript{K} in control conditions (WT) and under conditions of reduced complex N-glycosylation (MGAT1KO).
+We evaluate the calibration results using an adjusted $R^2$ measure for nonlinear models that preserves the interpretability of the classical $R^2$ based on the variance decomposition for linear models.
+Experimental results show the proposed method explains more than 90\% of variances by calibrated models in most cases (WT: $0.9624 \pm 0.0051$; MGAT1KO: $0.8938 \pm 0.0077$).
+A GUI application is provided to make an enabling tool for biomedical scientists without full expertise in modeling and computational analysis.
 
-\if0\blind{
-\section*{Acknowledgements}
-The authors acknowledge the generous support from the funding agency of XYZ and funding list} \fi
+The estimation of kinetics provides novel insights into potential mechanisms by which specific K\textsubscript{v} isoforms contribute to the overall reduction in I\textsubscript{K} observed in in-vitro experiments following chronic reductions in N-glycosylation.
+Thanks to the cell-specific approach, prediction uncertainty is quantified, and error bars are provided in the prediction results.
+Further, calibrated parameters are visualized via low-dimensional embedding that allows for encapsulating calibrated parameters in 3D space and, in turn, visualizing the variability across cells.
+WT cells show higher variability than MGAT1KO myocytes, which is likely due to significant I\textsubscript{K} reduction in the diseased group, combined with the elimination of complex N-glycosylation, thereby minimizing potential variability of N-glycan structures.
+The proposed method and pertinent software show strong potential for studying K\textsubscript{v} kinetics in various heart diseases.
+However, there are still some limitations and challenges that need to be addresses in future work.
+One of the main limitations of this study is that it does not address the interactions of K\textsubscript{v} with other major VGICs, namely Na\textsubscript{v} and Ca\textsubscript{v}, and cellular-level processes.
+Thus, causal inference of the effects of chronic glycosylation deficiency is limited.
+We plan to develop an integrative modeling method for learning activities of major VGICs together.
+
+\section*{Funding}
+This work is supported by the National Science Foundation (MCB-1856132 to HY and HK), (MCB-1856199 to EB and AE), (IOS-1146882 and IOS-1660926 to EB), and the American Heart Association (Postdoctoral Fellowship 15POST25710010 to AE).
+HY and HK would also like to thank the NSF I/UCRC Center for Health Organization Transformation (CHOT) award NSF IIP-1624727 for the support of their research work.
+Any opinions, findings, or conclusions found in this research are those of the authors and do not necessarily reflect the views of the sponsors.
+
+\section*{Disclosure of Interest}
+The authors report there are no competing interests to declare.
+
+\section*{Consent and Approval}
+This research does not involve any human participants or data. Therefore, no consent or approval was required for this study.
 
